


Quick comparison of the primate metadata from various sources
Looking at the subclass label mismatches to figure out what is going on.

data originating from the gillis labs FTP site. The confirmed_subclass, neighborhood, and confirmed_cluster all match the Subclass, Neighborhood, Cluster metadata from the same data deposited on CZI 
https://cellxgene.cziscience.com/collections/4dca242c-d302-4dba-a68f-4c61e7bad553


```{r}

human_10x = readRDS(file = '/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/primates/human_data_10x_gaba.rds')
human_ss = readRDS( file = '/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/primates/human_data_ss_gaba.rds')

chimp_10x = readRDS(file = '/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/primates/chimp_data_10x_gaba.rds')
chimp_ss = readRDS(file = '/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/primates/chimp_data_ss_gaba.rds')

gorilla_10x = readRDS(file = '/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/primates/gorilla_data_10x_gaba.rds')
gorilla_ss = readRDS(file = '/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/primates/gorilla_data_ss_gaba.rds')

macaca_10x = readRDS(file = '/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/primates/macaca_data_10x_gaba.rds')

marmoset_10x = readRDS(file = '/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/primates/marmoset_data_10x_gaba.rds')

```

This is supplemental table 6 from the Jorstad paper
https://pmc.ncbi.nlm.nih.gov/articles/PMC10659116/#ABS1

According to Hamsini, the 'metaneighbor_cluster' here maps to the 'cross-species_cluster' in the FTP data 
And the 'within_species_clusters' here maps to the 'confirmed_cluster' labels in the FTP data
```{r}

jorstad_supp6_meta = data.table::fread('/home/werner/projects/mouse_brain_consensus_taxonomy/data/jorstad_S6_cross_species_cluster_mapping.csv')

jorstad_supp6_meta



```


And this is the separate dendrogram table Hamsini shared where I first noticed the disagreements between these subclass labels and the confirmed subclass labels in the FTP data. I originally used the cross_species_cluster labels to map between the two

```{r}

primate_cluster_mappings = data.table::fread('../data/cross_species_mapping_cls86.csv')
primate_cluster_mappings = primate_cluster_mappings[1:86, ]

primate_cluster_mappings
```


```{r}
table(jorstad_supp6_meta$metaneighbor_cluster %in% primate_cluster_mappings$cross_species_cluster)


index = match(jorstad_supp6_meta$metaneighbor_cluster, primate_cluster_mappings$cross_species_cluster)

jorstad_supp6_meta
primate_cluster_mappings[index, ]

table(jorstad_supp6_meta$metaneighbor_cluster == primate_cluster_mappings$cross_species_cluster[index] )
table(jorstad_supp6_meta$subclass == primate_cluster_mappings$subclass_label[index] )




```

human_10x

```{r}
#human_10x
temp_ftp_data_meta = as.data.frame(colData(human_10x)) %>% group_by(cross_species_cluster, confirmed_subclass) %>% summarise(n = n())

index = match(temp_ftp_data_meta$cross_species_cluster, jorstad_supp6_meta$metaneighbor_cluster )
temp_ftp_data_meta$supp_table_6_subclass = jorstad_supp6_meta$subclass[index]


temp_ftp_data_meta %>% filter(confirmed_subclass != supp_table_6_subclass)

#human_ss

temp_ftp_data_meta = as.data.frame(colData(human_ss)) %>% group_by(cross_species_cluster, confirmed_subclass) %>% 
  summarise(n = n(), .groups = 'keep')
index = match(temp_ftp_data_meta$cross_species_cluster, jorstad_supp6_meta$metaneighbor_cluster )
temp_ftp_data_meta$supp_table_6_subclass = jorstad_supp6_meta$subclass[index]


temp_ftp_data_meta %>% filter(confirmed_subclass != supp_table_6_subclass)

```


```{r}

temp_ftp_data_meta = as.data.frame(colData(human_ss))

temp_ftp_data_meta %>% 
  group_by(cross_species_cluster, confirmed_subclass, cluster ) %>% 
  summarise(n = n()) %>% filter(cross_species_cluster %in% c('lamp5_sncg_vip|7','lamp5_sncg_vip|9'))



```

```{r}
jorstad_supp6_meta %>% filter(metaneighbor_cluster == 'lamp5_sncg_vip|9') %>% pull(within_species_clusters)
```











```{r}

temp_ftp_data_meta = as.data.frame(colData(chimp_10x)) %>% group_by(cross_species_cluster, confirmed_subclass) %>% summarise(n = n())

index = match(temp_ftp_data_meta$cross_species_cluster, jorstad_supp6_meta$metaneighbor_cluster )
temp_ftp_data_meta$supp_table_6_subclass = jorstad_supp6_meta$subclass[index]


temp_ftp_data_meta %>% filter(confirmed_subclass != supp_table_6_subclass)

temp_ftp_data_meta = as.data.frame(colData(chimp_ss)) %>% group_by(cross_species_cluster, confirmed_subclass) %>% summarise(n = n())

index = match(temp_ftp_data_meta$cross_species_cluster, jorstad_supp6_meta$metaneighbor_cluster )
temp_ftp_data_meta$supp_table_6_subclass = jorstad_supp6_meta$subclass[index]


temp_ftp_data_meta %>% filter(confirmed_subclass != supp_table_6_subclass)


```
























