


Putting the homologous labels back into the full mouse datasets, to explore the expression of genes that are not present in the 1-to-1 orthologs across the species



All the mouse data with homologous annotations

```{r}

library(Seurat)
library(SingleCellExperiment)
library(ggplot2)
library(dplyr)
```


```{r}
celltype_color_palette = c("Chandelier"='#F641A8', 
                           "Pvalb"='#D93137', 
                           "Sst"='#FF9900', 
                           "Sst Chodl"='gold', 
                           "Lamp5"='#DA808C',
                           "Sncg"='#DF70FF',
                           "Vip"='#A45FBF', 
                           "Lamp5 Lhx6"='#935F50',
                           "Pax6"='#71238C',
                           'Serpinf1' = 'grey20')

```

```{r}

load(file = '/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/mouse_primate_MN_integrated.Rdata')
interneuron.combined$cell_subclass[interneuron.combined$cell_subclass == 'Serpinf1'] = 'Vip'
```


Also get a summary plot for the datasets, number of cells and technology


```{r}


mouse_metadata = interneuron.combined[[]]


mouse_cell_num = mouse_metadata %>% group_by(study) %>% summarise(n = n()) %>% mutate(species = 'Mouse')


mouse_cell_num$study = c('Broad sN 10V3 2023','Allen sC Smart-seq V4 2018','Allen sC 10xV2 2021','Allen sN 10xV2 2021',
                         'Allen sC 10xV2 2023','Allen sC 10xV3 2023','Allen sC Smart-seq V4 2021','Allen sN Smart-seq V4 2021')


mouse_cell_num

```





```{r}

tasic18 = readRDS(file = '/vault/werner/rotation/jonathan/data/tasic18.rds')
table(as.character(tasic18$source.name), tasic18$characteristics..cell_subclass)

#Just grab the GABAergic cells
inhib_filt = colData(tasic18)$cell_class == 'GABAergic'
tasic18 = tasic18[ ,inhib_filt]

table(colData(tasic18)$cell_cluster)
ncol(tasic18)

```


```{r}

rownames(tasic18)[grepl('Cck', rownames(tasic18))]


current_study_ortho_meta = interneuron.combined[[]] %>% filter(study == 'tasic18')
index = match( colData(tasic18)$title,rownames(current_study_ortho_meta) )

tasic18$homologous_subclass = current_study_ortho_meta$primate_subclass[index]
tasic18$homologous_type = current_study_ortho_meta$orthology[index]
tasic18$mouse_sanity_check = current_study_ortho_meta$cell_subclass[index]


as.data.frame(colData(tasic18))

as.data.frame(colData(tasic18))[(tasic18$cell_subclass != tasic18$mouse_sanity_check), ]

```




```{r}
assay(tasic18, "cpm") = MetaMarkers::convert_to_cpm(assay(tasic18))

tasic18_seurat = as.Seurat(tasic18, counts = "counts", data = "cpm")

all.genes <- rownames(tasic18_seurat)
tasic18_seurat <- FindVariableFeatures(tasic18_seurat, nfeatures = 2000)
tasic18_seurat <- ScaleData(tasic18_seurat, features = all.genes)
tasic18_seurat <- RunPCA(tasic18_seurat, features = VariableFeatures(object = tasic18_seurat), verbose = F)
tasic18_seurat <- RunUMAP(tasic18_seurat , dims = 1:20, verbose = F)



DimPlot(tasic18_seurat, group.by = 'homologous_subclass',cols = celltype_color_palette)
DimPlot(tasic18_seurat, group.by = 'cell_subclass',cols = celltype_color_palette)

FeaturePlot(tasic18_seurat, features = 'Id2', slot = 'scale.data') + 
  scale_color_gradient2(low = '#313695', mid ='grey95', high = '#A50026', midpoint = 0 )

FeaturePlot(tasic18_seurat, features = 'Sncg', slot = 'scale.data') + 
  scale_color_gradient2(low = '#313695', mid ='grey95', high = '#A50026', midpoint = 0 )

FeaturePlot(tasic18_seurat, features = 'Pax6', slot = 'scale.data') + 
  scale_color_gradient2(low = '#313695', mid ='grey95', high = '#A50026', midpoint = 0 )

FeaturePlot(tasic18_seurat, features = 'Cck', slot = 'scale.data') + 
  scale_color_gradient2(low = '#313695', mid ='grey95', high = '#A50026', midpoint = 0 )

FeaturePlot(tasic18_seurat, features = 'Cnr1', slot = 'scale.data') + 
  scale_color_gradient2(low = '#313695', mid ='grey95', high = '#A50026', midpoint = 0 )

FeaturePlot(tasic18_seurat , features = 'Vip', slot = 'data') + 
  scale_color_gradient2(low = '#313695', mid ='grey95', high = '#A50026', midpoint = 0 )

```


```{r}
library(org.Mm.eg.db)
gene_map <- select(org.Mm.eg.db, keys(org.Mm.eg.db), c("SYMBOL", "ENSEMBL"), "ENTREZID")
gene_map


```

```{r}

zeng_10x_cells = readRDS("/vault/werner/rotation/jonathan/data/zeng_10x_cells/zeng_10x_cells.rds")
zeng_10x_cells


zeng_10x_nuclei = readRDS("/vault/werner/rotation/jonathan/data/zeng_10x_nuclei/zeng_10x_nuclei.rds")
zeng_10x_nuclei

zeng_smart_cells = readRDS("/vault/werner/rotation/jonathan/data/zeng_smart_cells/zeng_smart_cells.rds")
zeng_smart_nuclei = readRDS("/vault/werner/rotation/jonathan/data/zeng_smart_nuclei/zeng_smart_nuclei.rds")


inhib_filt = colData(zeng_10x_cells)$class_label == 'GABAergic'
zeng_10x_cells = zeng_10x_cells[ , inhib_filt]

inhib_filt = colData(zeng_10x_nuclei)$class_label == 'GABAergic'
zeng_10x_nuclei = zeng_10x_nuclei[ , inhib_filt]

inhib_filt = colData(zeng_smart_cells)$class_label == 'GABAergic'
zeng_smart_cells = zeng_smart_cells[ , inhib_filt]

inhib_filt = colData(zeng_smart_nuclei)$class_label == 'GABAergic'
zeng_smart_nuclei = zeng_smart_nuclei[ , inhib_filt]


ncol(zeng_10x_cells)
ncol(zeng_10x_nuclei)
ncol(zeng_smart_cells)
ncol(zeng_smart_nuclei)
```

```{r}
#10X datasets are already ensembl genes
zeng_10x_cells_genes = rownames(zeng_10x_cells)
index = match(zeng_10x_cells_genes, gene_map$ENSEMBL)
zeng_10x_cells_genes = gene_map[index, 'SYMBOL']
rownames(zeng_10x_cells) = paste(rownames(zeng_10x_cells), zeng_10x_cells_genes, sep = '_' )

zeng_10x_nuclei_genes = rownames(zeng_10x_nuclei)
index = match(zeng_10x_nuclei_genes, gene_map$ENSEMBL)
zeng_10x_nuclei_genes = gene_map[index, 'SYMBOL']
rowData(zeng_10x_nuclei)$full_gene_id = zeng_10x_nuclei_genes
zeng_10x_nuclei_genes[zeng_10x_nuclei_genes == 'NA'] = NA
zeng_10x_nuclei_genes = zeng_10x_nuclei_genes[!is.na(zeng_10x_nuclei_genes)]
rownames(zeng_10x_nuclei) = rowData(zeng_10x_nuclei)$full_gene_id


```




```{r}

rownames(zeng_smart_cells)[grepl('Cnr1', rownames(zeng_smart_cells))]

table( interneuron.combined$study)

current_study_ortho_meta = interneuron.combined[[]] %>% filter(study == 'zeng_10x_cells')
index = match( rownames(colData(zeng_10x_cells)),rownames(current_study_ortho_meta) )

zeng_10x_cells$homologous_subclass = current_study_ortho_meta$primate_subclass[index]
zeng_10x_cells$homologous_type = current_study_ortho_meta$orthology[index]
zeng_10x_cells$mouse_sanity_check = current_study_ortho_meta$cell_subclass[index]


as.data.frame(colData(zeng_10x_cells))
table(zeng_10x_cells$subclass_label == zeng_10x_cells$mouse_sanity_check)

as.data.frame(colData(zeng_10x_cells))[(zeng_10x_cells$cell_subclass != zeng_10x_cells$mouse_sanity_check), ]

```


```{r}
assay(zeng_10x_cells, "cpm") = MetaMarkers::convert_to_cpm(assay(zeng_10x_cells))
names(assays(zeng_10x_cells)) = c('counts','cpm')

dim(cpm(zeng_10x_cells))

zeng_10x_cells_seurat = as.Seurat(zeng_10x_cells, counts = "counts", data = "cpm")

all.genes <- rownames(zeng_10x_cells_seurat)
zeng_10x_cells_seurat <- FindVariableFeatures(zeng_10x_cells_seurat, nfeatures = 2000)
zeng_10x_cells_seurat <- ScaleData(zeng_10x_cells_seurat, features = all.genes)
zeng_10x_cells_seurat <- RunPCA(zeng_10x_cells_seurat, features = VariableFeatures(object = zeng_10x_cells_seurat), verbose = F)
zeng_10x_cells_seurat <- RunUMAP(zeng_10x_cells_seurat , dims = 1:20, verbose = F)

zeng_10x_cells_seurat[[]]

DimPlot(zeng_10x_cells_seurat, group.by = 'homologous_subclass',cols = celltype_color_palette)
DimPlot(zeng_10x_cells_seurat, group.by = 'subclass_label',cols = celltype_color_palette)


rownames(zeng_10x_cells)[grepl('Vip', rownames(zeng_10x_cells))]

FeaturePlot(zeng_10x_cells_seurat, features = 'ENSMUSG00000020644-Id2', slot = 'scale.data') + 
  scale_color_gradient2(low = '#313695', mid ='grey95', high = '#A50026', midpoint = 0 )

FeaturePlot(zeng_10x_cells_seurat, features = 'ENSMUSG00000023064-Sncg', slot = 'scale.data') + 
  scale_color_gradient2(low = '#313695', mid ='grey95', high = '#A50026', midpoint = 0 )

FeaturePlot(zeng_10x_cells_seurat, features = 'ENSMUSG00000027168-Pax6', slot = 'scale.data') + 
  scale_color_gradient2(low = '#313695', mid ='grey95', high = '#A50026', midpoint = 0 )

FeaturePlot(zeng_10x_cells_seurat, features = 'ENSMUSG00000032532-Cck', slot = 'scale.data') + 
  scale_color_gradient2(low = '#313695', mid ='grey95', high = '#A50026', midpoint = 0 )

FeaturePlot(zeng_10x_cells_seurat, features = 'ENSMUSG00000044288-Cnr1', slot = 'scale.data') + 
  scale_color_gradient2(low = '#313695', mid ='grey95', high = '#A50026', midpoint = 0 )

FeaturePlot(zeng_10x_cells_seurat, features = 'ENSMUSG00000019772-Vip', slot = 'data') + 
  scale_color_gradient2(low = '#313695', mid ='grey95', high = '#A50026', midpoint = 0 )



```



And check out the human data



```{r}

human_data = readRDS(file = '/inkwell03/werner/sc_data/primate_MTG/human_raw_UMI_expression_matrix.RDS')

table('ID2' %in% rownames(human_data))


```


```{r}

cell_index = human_data$confirmed_subclass %in% c('Vip','Sncg','Pax6','Lamp5','Lamp5_Lhx6','Pvalb','Chandelier','Sst','Sst Chodl')

dim(human_data)
human_data = human_data[, cell_index ]
dim(human_data)

#Convert to singlecell experiment
human_data = as.SingleCellExperiment(human_data)
#CPM normalize
assay(human_data, "cpm") = MetaMarkers::convert_to_cpm(assay(human_data))
#Split by smartseq and 10x
ss_index = colData(human_data)$species_tech == 'human_ss'
human_data_ss = human_data[ , ss_index]
human_data_10x = human_data[ , !ss_index]

dim(human_data_ss)
dim(human_data_10x)


```



```{r}

primate_cluster_mappings = data.table::fread('../../data/cross_species_mapping_cls86.csv')
table(primate_cluster_mappings$consensus_cluster_label)
```

```{r}
celltype_color_palette = c("Chandelier"='#F641A8', 
                           "Pvalb"='#D93137', 
                           "Sst"='#FF9900', 
                           "Sst Chodl"='gold', 
                           "Lamp5"='#DA808C',
                           "Sncg"='#DF70FF',
                           "Vip"='#A45FBF', 
                           "Lamp5_Lhx6"='#935F50',
                           "Pax6"='#71238C',
                           'Serpinf1' = 'grey20')

```

```{r}

colData(human_data_ss)

index = match(human_data_ss$cross_species_cluster, primate_cluster_mappings$cross_species_cluster)
human_data_ss$consensus_cluster_label = primate_cluster_mappings$consensus_cluster_label[index]
human_data_ss$subclass_label = primate_cluster_mappings$subclass_label[index]
#Note the cells without cross-species consensus
human_data_ss$subclass_label[is.na(human_data_ss$consensus_cluster_label)] = NA

human_data_ss_seurat = as.Seurat(human_data_ss, counts = "counts", data = "cpm")

all.genes <- rownames(human_data_ss_seurat)
human_data_ss_seurat <- FindVariableFeatures(human_data_ss_seurat, nfeatures = 2000)
human_data_ss_seurat <- ScaleData(human_data_ss_seurat, features = all.genes)
human_data_ss_seurat <- RunPCA(human_data_ss_seurat, features = VariableFeatures(object = human_data_ss_seurat), verbose = F)
human_data_ss_seurat <- RunUMAP(human_data_ss_seurat , dims = 1:20, verbose = F)

human_data_ss_seurat[[]]


DimPlot(human_data_ss_seurat, group.by = 'subclass_label',cols = celltype_color_palette)


FeaturePlot(human_data_ss_seurat, features = 'ID2', slot = 'scale.data') + 
  scale_color_gradient2(low = '#313695', mid ='grey95', high = '#A50026', midpoint = 0 )

FeaturePlot(human_data_ss_seurat, features = 'SNCG', slot = 'scale.data') + 
  scale_color_gradient2(low = '#313695', mid ='grey95', high = '#A50026', midpoint = 0 )

FeaturePlot(human_data_ss_seurat, features = 'PAX6', slot = 'scale.data') + 
  scale_color_gradient2(low = '#313695', mid ='grey95', high = '#A50026', midpoint = 0 )

FeaturePlot(human_data_ss_seurat, features = 'CCK', slot = 'scale.data') + 
  scale_color_gradient2(low = '#313695', mid ='grey95', high = '#A50026', midpoint = 0 )

FeaturePlot(human_data_ss_seurat, features = 'CNR1', slot = 'scale.data') + 
  scale_color_gradient2(low = '#313695', mid ='grey95', high = '#A50026', midpoint = 0 )

FeaturePlot(human_data_ss_seurat, features = 'VIP', slot = 'data') + 
  scale_color_gradient2(low = '#313695', mid ='grey95', high = '#A50026', midpoint = 0 )




```


```{r}

colData(human_data_10x)

index = match(human_data_10x$cross_species_cluster, primate_cluster_mappings$cross_species_cluster)
human_data_10x$consensus_cluster_label = primate_cluster_mappings$consensus_cluster_label[index]
human_data_10x$subclass_label = primate_cluster_mappings$subclass_label[index]
#Note the cells without cross-species consensus
human_data_10x$subclass_label[is.na(human_data_10x$consensus_cluster_label)] = NA

human_data_10x_seurat = as.Seurat(human_data_10x, counts = "counts", data = "cpm")

all.genes <- rownames(human_data_10x_seurat)
human_data_10x_seurat <- FindVariableFeatures(human_data_10x_seurat, nfeatures = 2000)
human_data_10x_seurat <- ScaleData(human_data_10x_seurat, features = all.genes)
human_data_10x_seurat <- RunPCA(human_data_10x_seurat, features = VariableFeatures(object = human_data_10x_seurat), verbose = F)
human_data_10x_seurat <- RunUMAP(human_data_10x_seurat , dims = 1:20, verbose = F)

human_data_10x_seurat[[]]


DimPlot(human_data_10x_seurat, group.by = 'subclass_label',cols = celltype_color_palette)


FeaturePlot(human_data_10x_seurat, features = 'ID2', slot = 'scale.data') + 
  scale_color_gradient2(low = '#313695', mid ='grey95', high = '#A50026', midpoint = 0 )

FeaturePlot(human_data_10x_seurat, features = 'SNCG', slot = 'scale.data') + 
  scale_color_gradient2(low = '#313695', mid ='grey95', high = '#A50026', midpoint = 0 )

FeaturePlot(human_data_10x_seurat, features = 'PAX6', slot = 'scale.data') + 
  scale_color_gradient2(low = '#313695', mid ='grey95', high = '#A50026', midpoint = 0 )

FeaturePlot(human_data_10x_seurat, features = 'CCK', slot = 'scale.data') + 
  scale_color_gradient2(low = '#313695', mid ='grey95', high = '#A50026', midpoint = 0 )

FeaturePlot(human_data_10x_seurat, features = 'CNR1', slot = 'scale.data') + 
  scale_color_gradient2(low = '#313695', mid ='grey95', high = '#A50026', midpoint = 0 )

FeaturePlot(human_data_10x_seurat, features = 'VIP', slot = 'data') + 
  scale_color_gradient2(low = '#313695', mid ='grey95', high = '#A50026', midpoint = 0 )




```

Get all the Allen primate data

```{r}

human_data_10x  = readRDS(file = '/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/primates/human_data_10x_gaba.rds')
human_data_ss = readRDS( file = '/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/primates/human_data_ss_gaba.rds')

chimp_data_10x  = readRDS(file = '/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/primates/chimp_data_10x_gaba.rds')
chimp_data_ss = readRDS( file = '/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/primates/chimp_data_ss_gaba.rds')

gorilla_data_10x = readRDS(file = '/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/primates/gorilla_data_10x_gaba.rds')
gorilla_data_ss = readRDS(file = '/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/primates/gorilla_data_ss_gaba.rds')

macaca_data = readRDS(file = '/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/primates/macaca_data_10x_gaba.rds')

marmoset_data = readRDS(file = '/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/primates/marmoset_data_10x_gaba.rds')

```

```{r}
mouse_cell_num


allen_primate_cell_num = data.frame(study = c('Allen Human sN 10xV3 2023', 'Allen Human sN Smart-seq V4 2023', 'Allen Chimpanzee sN 10xV3 2023', 'Allen Chimpanzee sN Smart-seq V4 2023',
                                              'Allen Gorilla sN 10xV3 2023', 'Allen Gorilla sN Smart-seq V4 2023', 'Allen Macaque sN 10xV3 2023', 'Allen Marmoset sN 10xV3 2023'), 
                                    n = c(ncol(human_data_10x), ncol(human_data_ss), ncol(chimp_data_10x), ncol(chimp_data_ss), ncol(gorilla_data_10x), ncol(gorilla_data_ss),
                                          ncol(macaca_data), ncol(marmoset_data) ),
                                    species = c('Human','Human','Chimpanzee','Chimpanzee','Gorilla','Gorilla','Macaque','Marmoset'))

allen_primate_cell_num 

```


And the Yale and UT southwestern primate datasets

```{r}


caglayan_utSouthwestern = readRDS(file = '/inkwell03/werner/sc_data/primate_caglayan_konopka/caglayan_gaba_processed.rds')
utSouth_cell_num = caglayan_utSouthwestern[[]] %>% group_by(Species) %>% summarise(n = n())

utSouth_cell_num$study = c('UT SouthWestern Human sN 10xV3 2023','UT SouthWestern Chimpanzee sN 10xV3 2023','UT SouthWestern Macaque sN 10xV3 2023')
colnames(utSouth_cell_num) = c('species','n','study')
utSouth_cell_num$species = c('Human','Chimpanzee','Macaque')



utSouth_cell_num
```


```{r}

sesten_gaba_data = readRDS('/inkwell03/werner/sc_data/primate_ma_sestan/sestan_integrated_gaba.rds')

yale_cell_num = sesten_gaba_data[[]] %>% group_by(species) %>% summarise(n = n())

yale_cell_num$species[yale_cell_num$species == 'Rhesus'] = 'Macaque'


yale_cell_num$study = c('Yale Chimpanzee sN 10xV3 2022', 'Yale Human sN 10xV3 2022','Yale Marmoset sN 10xV3 2022','Yale Macaque sN 10xV3 2022' )


yale_cell_num

```


```{r}


mouse_cell_num

allen_primate_cell_num

utSouth_cell_num

yale_cell_num



all_cell_num_df = do.call('rbind', list(mouse_cell_num,allen_primate_cell_num, utSouth_cell_num, yale_cell_num ))

all_cell_num_df$species = factor(all_cell_num_df$species, levels = c('Mouse','Marmoset','Macaque','Gorilla','Chimpanzee','Human'))


all_cell_num_df$study = factor(all_cell_num_df$study , levels = c("Allen Human sN 10xV3 2023","Allen Human sN Smart-seq V4 2023", 
                                                                  "Allen Chimpanzee sN 10xV3 2023","Allen Chimpanzee sN Smart-seq V4 2023",
                                                                  "Allen Gorilla sN 10xV3 2023","Allen Gorilla sN Smart-seq V4 2023",
                                                                  "Allen Macaque sN 10xV3 2023","Allen Marmoset sN 10xV3 2023",
                                                                  "UT SouthWestern Human sN 10xV3 2023" , "UT SouthWestern Chimpanzee sN 10xV3 2023" ,
                                                                  "UT SouthWestern Macaque sN 10xV3 2023" , "Yale Human sN 10xV3 2022", 
                                                                  "Yale Chimpanzee sN 10xV3 2022", "Yale Macaque sN 10xV3 2022", "Yale Marmoset sN 10xV3 2022",
                                                                  'Allen sC Smart-seq V4 2018', 'Allen sC 10xV2 2021', 'Allen sN 10xV2 2021', 
                                                                  'Allen sC Smart-seq V4 2021', 'Allen sN Smart-seq V4 2021',
                                                                  'Allen sC 10xV2 2023','Allen sC 10xV3 2023','Broad sN 10V3 2023'))



all_cell_num_df$technology = '10x'
all_cell_num_df$technology[all_cell_num_df$study %in% c('Allen sC Smart-seq V4 2018', 'Allen sC Smart-seq V4 2021','Allen sN Smart-seq V4 2021',
                                                        'Allen Human sN Smart-seq V4 2023','Allen Chimpanzee sN Smart-seq V4 2023', 'Allen Gorilla sN Smart-seq V4 2023')] = 'Smart-seq'

ggplot(all_cell_num_df, aes(x = study, y = species, size = log10(n), color = technology )) + geom_point() +
  guides(size=guide_legend(title="log10(Cell num.)")) + scale_color_manual(values = c('10x' = 'slateblue1', 'Smart-seq' = 'goldenrod')) +
  theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0))

ggsave(filename = 'cell_num_dataset_summary.pdf', path = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/', 
       device = 'pdf', useDingbats = F, height = 5, width = 8)



```






