

Non-allen institute cross primate data

Initial idea, cluster the data de novo using standard Seurat approaches. 
Using just the subclass name-sake genes (VIP, Pvalb, ) etc, annotate the clusters. 

The idea is to quantify 
1.) whether SNCG is DE in the primate data
2.) whether PAX6 is DE in the primate data
3.) if the Vip cells have a consistent outgroup with marker gene expression consistent with the orthologous SNCG subclass.

```{r}
library(SingleCellExperiment)
library(Seurat)
library(rhdf5)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(stats)
library(data.table)
library(MetaMarkers)
setwd('/home/werner/projects/mouse_brain_consensus_taxonomy/code')

```


```{r}
celltype_color_palette = c("Chandelier"='#F641A8', 
                           "Pvalb"='#D93137', 
                           "Sst"='#FF9900', 
                           "Sst Chodl"='gold', 
                           "Lamp5"='#DA808C',
                           "Sncg"='#DF70FF',
                           "Vip"='#A45FBF', 
                           "Lamp5 Lhx6"='#935F50',
                           "Pax6"='#71238C')


```

Caglayan_konopka primate

downloaded from https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE192773
Paper: https://pmc.ncbi.nlm.nih.gov/articles/PMC11161302/#S32

snRNA-seq of Brodmann area 23, part of the posterior cingulate cortex

```{r}

caglayan_inhib = readRDS('/inkwell03/werner/sc_data/primate_caglayan_konopka/GSE192773_Seurat_Inhibitory_RNA.RDS')

caglayan_inhib

```

```{r}

rownames(caglayan_inhib)[1:10]

caglayan_inhib[[]]

table(caglayan_inhib$Species, caglayan_inhib$CellType )

table(caglayan_inhib$Sample, caglayan_inhib$CellType)


cge_mge_label = rep('CGE', length = ncol(caglayan_inhib))
cge_mge_label[caglayan_inhib$CellType %in% c('PVALB_Basket','PVALB_Chandelier','SST_CALB1-','SST_CALB1+')] = 'MGE'

caglayan_inhib$cge_mge = cge_mge_label

```


```{r}

final_present_1to1s = readRDS('/home/werner/projects/mouse_brain_consensus_taxonomy/data/final_present_1to1s.rds')

final_present_1to1s

#mouse_present_1to1s= paste(final_present_1to1s$`Mouse gene stable ID`, final_present_1to1s$`Mouse gene name`, sep = '_')
#length(mouse_present_1to1s)

caglayan_genes = rownames(caglayan_inhib)

length(caglayan_genes)
table(caglayan_genes %in% final_present_1to1s$`Gene name`)



```


Try clustering the data like normal
CPM, standard clustering, do it species specific first


```{r}
caglayan_human_inhib = subset(caglayan_inhib, subset = Species == 'human' )


#CPM normalization
caglayan_human_inhib <- NormalizeData(caglayan_human_inhib, normalization.method = 'RC', scale.factor = 1e6, verbose = F)
caglayan_human_inhib <- FindVariableFeatures(caglayan_human_inhib, selection.method = "vst", nfeatures = 2000, verbose = F)
all.genes <- rownames(caglayan_human_inhib)
caglayan_human_inhib <- ScaleData(caglayan_human_inhib, features = all.genes, verbose = F)


caglayan_human_inhib <- RunPCA(caglayan_human_inhib, features = VariableFeatures(object =caglayan_human_inhib), verbose = F)

#caglayan_human_inhib <- RunPCA(caglayan_human_inhib, features = c('VIP','SNCG','PAX6','LAMP5','LHX6', 'SST', 'CHODL', 'PVALB', 'TRPS1'), 
#                               npcs = 5, verbose = F)


caglayan_human_inhib <- FindNeighbors(caglayan_human_inhib, dims = 1:25, verbose = F)
caglayan_human_inhib <- FindClusters(caglayan_human_inhib, verbose = F)

caglayan_human_inhib <- RunUMAP(caglayan_human_inhib, dims = 1:25, verbose = F)

DimPlot(caglayan_human_inhib, reduction = "pca", label = T)
DimPlot(caglayan_human_inhib, reduction = "umap", label = T)
DimPlot(caglayan_human_inhib, reduction = "umap", group.by = 'CellType')
DimPlot(caglayan_human_inhib, reduction = "umap", group.by = 'cge_mge')

```


```{r}
caglayan_human_inhib_markers = compute_markers(caglayan_human_inhib@assays$RNA@data, caglayan_human_inhib$seurat_clusters)
gc()

caglayan_human_inhib_class_markers = compute_markers(caglayan_human_inhib@assays$RNA@data, caglayan_human_inhib$cge_mge)
gc()

caglayan_human_inhib_markers %>% filter(gene == 'PAX6') %>% arrange(desc(fold_change))
caglayan_human_inhib_markers %>% filter(gene == 'SNCG') %>% arrange(desc(fold_change))
caglayan_human_inhib_markers %>% filter(gene == 'RXRG') %>% arrange(desc(fold_change))
caglayan_human_inhib_markers %>% filter(gene == 'LAMP5') %>% arrange(desc(fold_change))
caglayan_human_inhib_markers %>% filter(gene == 'PALMD') %>% arrange(desc(fold_change))
caglayan_human_inhib_markers %>% filter(gene == 'CXCL14') %>% arrange(desc(fold_change))


caglayan_human_inhib_class_markers %>% filter(gene == 'PAX6') %>% arrange(desc(fold_change))
caglayan_human_inhib_class_markers %>% filter(gene == 'SNCG') %>% arrange(desc(fold_change))
caglayan_human_inhib_class_markers %>% filter(gene == 'RXRG') %>% arrange(desc(fold_change))
caglayan_human_inhib_class_markers %>% filter(gene == 'LAMP5') %>% arrange(desc(fold_change))
caglayan_human_inhib_class_markers %>% filter(gene == 'PALMD') %>% arrange(desc(fold_change))
caglayan_human_inhib_class_markers %>% filter(gene == 'CXCL14') %>% arrange(desc(fold_change))



features = c('ADARB2','VIP','PALMD','SNCG','RXRG','CXCL14','LAMP5','LHX6','SST', 'CHODL','PVALB','TRPS1')
VlnPlot(caglayan_human_inhib, group.by = 'seurat_clusters',
        features, stack = TRUE, sort = F, flip = T, slot = 'data',) +
        theme(legend.position = "none") + ggtitle("Caglayan Human GABA")

```




```{r}
caglayan_chimp_inhib = subset(caglayan_inhib, subset = Species == 'chimp' )

#CPM normalization
caglayan_chimp_inhib <- NormalizeData(caglayan_chimp_inhib, normalization.method = 'RC', scale.factor = 1e6, verbose = F)
caglayan_chimp_inhib <- FindVariableFeatures(caglayan_chimp_inhib, selection.method = "vst", nfeatures = 2000, verbose = F)
all.genes <- rownames(caglayan_chimp_inhib)
caglayan_chimp_inhib <- ScaleData(caglayan_chimp_inhib, features = all.genes, verbose = F)

caglayan_chimp_inhib <- RunPCA(caglayan_chimp_inhib, features = VariableFeatures(object =caglayan_chimp_inhib), verbose = F)

caglayan_chimp_inhib <- FindNeighbors(caglayan_chimp_inhib, dims = 1:25, verbose = F)
caglayan_chimp_inhib <- FindClusters(caglayan_chimp_inhib, verbose = F, resolution = .25)

caglayan_chimp_inhib <- RunUMAP(caglayan_chimp_inhib, dims = 1:25, verbose = F)

DimPlot(caglayan_chimp_inhib, reduction = "pca", label = T)
DimPlot(caglayan_chimp_inhib, reduction = "umap", label = T)
DimPlot(caglayan_chimp_inhib, reduction = "umap", group.by = 'CellType')
DimPlot(caglayan_chimp_inhib, reduction = "umap", group.by = 'cge_mge')


FeaturePlot(caglayan_chimp_inhib, features = 'VIP', slot = 'scale.data')
FeaturePlot(caglayan_chimp_inhib, features = 'CBLN2', slot = 'scale.data')
FeaturePlot(caglayan_chimp_inhib, features = 'CXCL14', slot = 'scale.data')
FeaturePlot(caglayan_chimp_inhib, features = 'PAX6', slot = 'scale.data')
FeaturePlot(caglayan_chimp_inhib, features = 'PALMD', slot = 'scale.data')
```






```{r}

caglayan_chimp_inhib_markers = compute_markers(caglayan_chimp_inhib@assays$RNA@data, caglayan_chimp_inhib$seurat_clusters)
gc()

caglayan_chimp_inhib_class_markers = compute_markers(caglayan_chimp_inhib@assays$RNA@data, caglayan_chimp_inhib$cge_mge)
gc()


caglayan_chimp_inhib_class_markers %>% filter(gene == 'PAX6') %>% arrange(desc(fold_change))
caglayan_chimp_inhib_class_markers %>% filter(gene == 'SNCG') %>% arrange(desc(fold_change))
caglayan_chimp_inhib_class_markers %>% filter(gene == 'RXRG') %>% arrange(desc(fold_change))
caglayan_chimp_inhib_class_markers %>% filter(gene == 'LAMP5') %>% arrange(desc(fold_change))
caglayan_chimp_inhib_class_markers %>% filter(gene == 'PALMD') %>% arrange(desc(fold_change))
caglayan_chimp_inhib_class_markers %>% filter(gene == 'CXCL14') %>% arrange(desc(fold_change))


```


```{r}
caglayan_macaque_inhib = subset(caglayan_inhib, subset = Species == 'macaque' )


#CPM normalization
caglayan_macaque_inhib <- NormalizeData(caglayan_macaque_inhib, normalization.method = 'RC', scale.factor = 1e6, verbose = F)
caglayan_macaque_inhib <- FindVariableFeatures(caglayan_macaque_inhib, selection.method = "vst", nfeatures = 2000, verbose = F)
all.genes <- rownames(caglayan_macaque_inhib)
caglayan_macaque_inhib <- ScaleData(caglayan_macaque_inhib, features = all.genes, verbose = F)

caglayan_macaque_inhib <- RunPCA(caglayan_macaque_inhib, features = VariableFeatures(object =caglayan_macaque_inhib), verbose = F)

caglayan_macaque_inhib <- FindNeighbors(caglayan_macaque_inhib, dims = 1:25, verbose = F)
caglayan_macaque_inhib <- FindClusters(caglayan_macaque_inhib, verbose = F, resolution = .25)

caglayan_macaque_inhib <- RunUMAP(caglayan_macaque_inhib, dims = 1:25, verbose = F)

DimPlot(caglayan_macaque_inhib, reduction = "pca", label = T)
DimPlot(caglayan_macaque_inhib, reduction = "umap", label = T)
DimPlot(caglayan_macaque_inhib, reduction = "umap", group.by = 'CellType')
DimPlot(caglayan_macaque_inhib, reduction = "umap", group.by = 'cge_mge')


FeaturePlot(caglayan_macaque_inhib, features = 'VIP', slot = 'scale.data')
FeaturePlot(caglayan_macaque_inhib, features = 'CBLN2', slot = 'scale.data')
FeaturePlot(caglayan_macaque_inhib, features = 'CXCL14', slot = 'scale.data')
FeaturePlot(caglayan_macaque_inhib, features = 'PAX6', slot = 'scale.data')
FeaturePlot(caglayan_macaque_inhib, features = 'PALMD', slot = 'scale.data')
```

```{r}
caglayan_macaque_inhib_markers = compute_markers(caglayan_macaque_inhib@assays$RNA@data, caglayan_macaque_inhib$seurat_clusters)
gc()

caglayan_macaque_inhib_class_markers = compute_markers(caglayan_macaque_inhib@assays$RNA@data, caglayan_macaque_inhib$cge_mge)
gc()


caglayan_macaque_inhib_class_markers %>% filter(gene == 'PAX6') %>% arrange(desc(fold_change))
caglayan_macaque_inhib_class_markers %>% filter(gene == 'SNCG') %>% arrange(desc(fold_change))
caglayan_macaque_inhib_class_markers %>% filter(gene == 'RXRG') %>% arrange(desc(fold_change))
caglayan_macaque_inhib_class_markers %>% filter(gene == 'LAMP5') %>% arrange(desc(fold_change))
caglayan_macaque_inhib_class_markers %>% filter(gene == 'PALMD') %>% arrange(desc(fold_change))
caglayan_macaque_inhib_class_markers %>% filter(gene == 'CXCL14') %>% arrange(desc(fold_change))


```


```{r}


export_markers(caglayan_macaque_inhib_markers, '/inkwell03/werner/sc_data/primate_caglayan_konopka/species_meta_markers/macaque_gaba_celltype_markers.csv' )
export_markers(caglayan_chimp_inhib_markers, '/inkwell03/werner/sc_data/primate_caglayan_konopka/species_meta_markers/chimp_gaba_celltype_markers.csv' )
export_markers(caglayan_human_inhib_markers, '/inkwell03/werner/sc_data/primate_caglayan_konopka/species_meta_markers/human_gaba_celltype_markers.csv' )


export_markers(caglayan_macaque_inhib_class_markers, '/inkwell03/werner/sc_data/primate_caglayan_konopka/species_meta_markers/macaque_gaba_class_markers.csv' )
export_markers(caglayan_chimp_inhib_class_markers, '/inkwell03/werner/sc_data/primate_caglayan_konopka/species_meta_markers/chimp_gaba_class_markers.csv' )
export_markers(caglayan_human_inhib_class_markers, '/inkwell03/werner/sc_data/primate_caglayan_konopka/species_meta_markers/human_gaba_class_markers.csv' )

```


```{r}


caglayan_macaque_inhib_class_markers = read_markers('/inkwell03/werner/sc_data/primate_caglayan_konopka/species_meta_markers/macaque_gaba_class_markers.csv.gz' )
caglayan_chimp_inhib_class_markers = read_markers( '/inkwell03/werner/sc_data/primate_caglayan_konopka/species_meta_markers/chimp_gaba_class_markers.csv.gz' )
caglayan_human_inhib_class_markers = read_markers( '/inkwell03/werner/sc_data/primate_caglayan_konopka/species_meta_markers/human_gaba_class_markers.csv.gz' )

```



```{r}
#PAX6
pax6_sp_1 = caglayan_human_inhib_class_markers %>% filter(gene == 'PAX6') %>% 
  select(cell_type, gene, fold_change) %>% 
  mutate(species = 'human')

pax6_sp_2 = caglayan_chimp_inhib_class_markers %>% filter(gene == 'PAX6') %>% 
  select(cell_type, gene, fold_change) %>% 
  mutate(species = 'chimp')

pax6_sp_3 = caglayan_macaque_inhib_class_markers %>% filter(gene == 'PAX6') %>% 
  select(cell_type, gene, fold_change) %>% 
  mutate(species = 'macaque')


pax6_DE_df = do.call('rbind', list(pax6_sp_1 , pax6_sp_2 , pax6_sp_3))
pax6_DE_df = pax6_DE_df %>% tidyr::pivot_wider(names_from = cell_type, values_from = fold_change)

#LAMP5
lamp5_sp_1 = caglayan_human_inhib_class_markers %>% filter(gene == 'LAMP5') %>% 
  select(cell_type, gene, fold_change) %>% 
  mutate(species = 'human')

lamp5_sp_2 = caglayan_chimp_inhib_class_markers %>% filter(gene == 'LAMP5') %>% 
  select(cell_type, gene, fold_change) %>% 
  mutate(species = 'chimp')

lamp5_sp_3 = caglayan_macaque_inhib_class_markers %>% filter(gene == 'LAMP5') %>% 
  select(cell_type, gene, fold_change) %>% 
  mutate(species = 'macaque')


lamp5_DE_df = do.call('rbind', list(lamp5_sp_1 , lamp5_sp_2 , lamp5_sp_3))
lamp5_DE_df = lamp5_DE_df %>% tidyr::pivot_wider(names_from = cell_type, values_from = fold_change)

#VIP
vip_sp_1 = caglayan_human_inhib_class_markers %>% filter(gene == 'VIP') %>% 
  select(cell_type, gene, fold_change) %>% 
  mutate(species = 'human')

vip_sp_2 = caglayan_chimp_inhib_class_markers %>% filter(gene == 'VIP') %>% 
  select(cell_type, gene, fold_change) %>% 
  mutate(species = 'chimp')

vip_sp_3 = caglayan_macaque_inhib_class_markers %>% filter(gene == 'VIP') %>% 
  select(cell_type, gene, fold_change) %>% 
  mutate(species = 'macaque')


vip_DE_df = do.call('rbind', list(vip_sp_1 , vip_sp_2 , vip_sp_3))
vip_DE_df = vip_DE_df %>% tidyr::pivot_wider(names_from = cell_type, values_from = fold_change)


#SNCG
sncg_sp_1 = caglayan_human_inhib_class_markers %>% filter(gene == 'SNCG') %>% 
  select(cell_type, gene, fold_change) %>% 
  mutate(species = 'human')

sncg_sp_2 = caglayan_chimp_inhib_class_markers %>% filter(gene == 'SNCG') %>% 
  select(cell_type, gene, fold_change) %>% 
  mutate(species = 'chimp')

sncg_sp_3 = caglayan_macaque_inhib_class_markers %>% filter(gene == 'SNCG') %>% 
  select(cell_type, gene, fold_change) %>% 
  mutate(species = 'macaque')


sncg_DE_df = do.call('rbind', list(sncg_sp_1 , sncg_sp_2 , sncg_sp_3))
sncg_DE_df = sncg_DE_df %>% tidyr::pivot_wider(names_from = cell_type, values_from = fold_change)


marker_gene_df = rbind(pax6_DE_df, lamp5_DE_df, vip_DE_df, sncg_DE_df)

marker_gene_df = marker_gene_df %>% mutate(log2_FC_CGE = log2(CGE))
marker_gene_df$species = factor(marker_gene_df$species , levels = c('macaque','chimp','human'))

ggplot(marker_gene_df  , aes(x = log2_FC_CGE, y = species , color = gene, shape = species)) + geom_point(size = 5,alpha = .5) +
  xlab('log2(FC) CGE/MGE') + ylab('Species') +
  scale_color_manual(values = c('SNCG' = celltype_color_palette[['Sncg']], 
                                'PAX6' = celltype_color_palette[['Pax6']],
                                'LAMP5' = celltype_color_palette[['Lamp5']],
                                'VIP' = celltype_color_palette[['Vip']])) + 
  geom_hline(yintercept = 0, linetype = 'dashed', color = 'red') +
  geom_vline(xintercept = 0, linetype = 'dashed', color = 'red') +
  theme_bw() + ggtitle('Caglayan cross-primate data')

ggsave(filename = 'caglayan_pax6_sncg_vip_lamp5_MGE_CGE_DE_scatter.pdf', path = '../../graphs/non_allen_data/', device = 'pdf',
       useDingbats = F, height = 3, width = 5)


```

Just cluster the data taxonomically like the allen, see what it looks like.
Show the VIP values across the centroids, is there a consistent  out group to VIP that expresses the orthologous SNCG markers?
CXCL14?


```{r}
all_species_markers = list(
    human_10x = read_markers("metamarkers/markers_human_10x.csv.gz"),
    human_ss = read_markers("metamarkers/markers_human_ss.csv.gz"),
    chimp_10x = read_markers("metamarkers/markers_chimp_10x.csv.gz"),
    chimp_ss = read_markers("metamarkers/markers_chimp_ss.csv.gz"),
    gorilla_10x = read_markers("metamarkers/markers_gorilla_10x.csv.gz"),
    gorilla_ss = read_markers("metamarkers/markers_gorilla_ss.csv.gz"),
    macaca_10x = read_markers("metamarkers/markers_macaca_10x.csv.gz"),
    marmoset_10x= read_markers("metamarkers/markers_marmoset_10x.csv.gz"),
    tasic18 = read_markers("metamarkers/ortho_markers_tasic18.csv.gz"),
    zeng_10x_cells = read_markers("metamarkers/ortho_markers_zeng_10x_cells.csv.gz"),
    zeng_10x_nuclei = read_markers("metamarkers/ortho_markers_zeng_10x_nuclei.csv.gz"),
    zeng_smart_nuclei = read_markers("metamarkers/ortho_markers_zeng_smart_nuclei.csv.gz"),
    zeng_smart_cells = read_markers("metamarkers/ortho_markers_zeng_smart_cells.csv.gz"),
    zeng_10xv2_wba = read_markers("metamarkers/ortho_markers_zeng_10xv2_wba.csv.gz"),
    zeng_10xv3_wba = read_markers("metamarkers/ortho_markers_zeng_10xv3_wba.csv.gz"),
    macosko_wba = read_markers("metamarkers/ortho_markers_macosko_wba.csv.gz")
)


all_species_markers$human_10x$cell_type[all_species_markers$human_10x$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
all_species_markers$human_ss$cell_type[all_species_markers$human_ss$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
all_species_markers$chimp_10x$cell_type[all_species_markers$chimp_10x$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
all_species_markers$chimp_ss$cell_type[all_species_markers$chimp_ss$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
all_species_markers$gorilla_10x$cell_type[all_species_markers$gorilla_10x$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
all_species_markers$gorilla_ss$cell_type[all_species_markers$gorilla_ss$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
all_species_markers$macaca_10x$cell_type[all_species_markers$macaca_10x$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
all_species_markers$marmoset_10x$cell_type[all_species_markers$marmoset_10x$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'

conserved_subclass_meta_markers = make_meta_markers(all_species_markers, detailed_stats = TRUE)

conserved_subclass_meta_markers %>% group_by(cell_type) %>% filter(rank <= 10)


all_primate_markers = all_species_markers[c('human_10x', 'human_ss', 'chimp_10x','chimp_ss','gorilla_10x','gorilla_ss','macaca_10x','marmoset_10x')]

primate_subclass_meta_markers = make_meta_markers(all_primate_markers, detailed_stats = TRUE)

primate_subclass_meta_markers %>% group_by(cell_type) %>% filter(rank <= 10)



```

Cluster all the species using the integrated space
And then compute markers from the normalized data

```{r}

caglayan_inhib_copy = caglayan_inhib
caglayan_inhib_copy <- FindNeighbors(caglayan_inhib_copy , dims = 1:25, verbose = F, assay = 'integrated')
caglayan_inhib_copy <- FindClusters(caglayan_inhib_copy , verbose = F, graph.name = 'integrated_snn')



caglayan_inhib_copy<- NormalizeData(caglayan_inhib_copy, normalization.method = 'RC', scale.factor = 1e6, verbose = F)


```



```{r}


saveRDS(caglayan_inhib_copy, file = '/inkwell03/werner/sc_data/primate_caglayan_konopka/caglayan_gaba_processed.rds')

```


Get UMAPs of the integrated species, original author labels, the CGE and MGE annots, and the de novo clusters


```{r}

caglayan_inhib_copy = readRDS('/inkwell03/werner/sc_data/primate_caglayan_konopka/caglayan_gaba_processed.rds')

cge_mge_label = rep('CGE', length = ncol(caglayan_inhib_copy))
cge_mge_label[caglayan_inhib_copy$CellType %in% c('PVALB_Basket','PVALB_Chandelier','SST_CALB1-','SST_CALB1+')] = 'MGE'

caglayan_inhib_copy$cge_mge = cge_mge_label


DimPlot(caglayan_inhib_copy, group.by = 'Species') + ggtitle('UT Southwestern primate species')
ggsave(filename = 'caglayan_integrated_species_umap.pdf', path = '../../graphs/non_allen_data/umaps/', device = 'pdf',
       useDingbats = F, height = 4, width = 6)

DimPlot(caglayan_inhib_copy, group.by = 'CellType') + ggtitle('UT Southwestern original author annotations')
ggsave(filename = 'caglayan_integrated_celltype_umap.pdf', path = '../../graphs/non_allen_data/umaps/', device = 'pdf',
       useDingbats = F, height = 4, width = 6)

DimPlot(caglayan_inhib_copy, group.by = 'cge_mge') + ggtitle('UT Southwestern CGE-MGE labels')
ggsave(filename = 'caglayan_integrated_cge_mge_umap.pdf', path = '../../graphs/non_allen_data/umaps/', device = 'pdf',
       useDingbats = F, height = 4, width = 6)

DimPlot(caglayan_inhib_copy, group.by = 'seurat_clusters', label = T) + ggtitle('UT Southwestern de novo clusters')
ggsave(filename = 'caglayan_integrated_de_novo_clust_umap.pdf', path = '../../graphs/non_allen_data/umaps/', device = 'pdf',
       useDingbats = F, height = 4, width = 6)

```



SingleCellExperiment for MetaMarkers

```{r}

caglayan_inhib_copy = readRDS('/inkwell03/werner/sc_data/primate_caglayan_konopka/caglayan_gaba_processed.rds')

DefaultAssay(caglayan_inhib_copy) = 'RNA'

caglayan_inhib_copy = as.SingleCellExperiment(caglayan_inhib_copy)

names(assays(caglayan_inhib_copy)) = c('counts','cpm')

caglayan_integ_inhib_markers = compute_markers(cpm(caglayan_inhib_copy), caglayan_inhib_copy$seurat_clusters)
gc()


```



##########################
Checking out hierarchical clustering of the primate data plus the marker expression of the Allen primate vs Mouse metamarkers

###########################

orthologous mouse markers

```{r}
all_mouse_markers = list(
    tasic18 = read_markers("metamarkers/ortho_markers_tasic18.csv.gz"),
    zeng_10x_cells = read_markers("metamarkers/ortho_markers_zeng_10x_cells.csv.gz"),
    zeng_10x_nuclei = read_markers("metamarkers/ortho_markers_zeng_10x_nuclei.csv.gz"),
    zeng_smart_nuclei = read_markers("metamarkers/ortho_markers_zeng_smart_nuclei.csv.gz"),
    zeng_smart_cells = read_markers("metamarkers/ortho_markers_zeng_smart_cells.csv.gz"),
    zeng_10xv2_wba = read_markers("metamarkers/ortho_markers_zeng_10xv2_wba.csv.gz"),
    zeng_10xv3_wba = read_markers("metamarkers/ortho_markers_zeng_10xv3_wba.csv.gz"),
    macosko_wba = read_markers("metamarkers/ortho_markers_macosko_wba.csv.gz")
)


mouse_subclass_meta_markers = make_meta_markers(all_mouse_markers, detailed_stats = TRUE)

mouse_subclass_meta_markers %>% group_by(cell_type) %>% filter(rank <= 10)


```

And the original mouse markers

MetaMarkers
```{r}
orig_mouse_markers = list(
    tasic18 = read_markers("metamarkers/markers_tasic18.csv.gz"),
    zeng_10x_cells = read_markers("metamarkers/markers_zeng_10x_cells.csv.gz"),
    zeng_10x_nuclei = read_markers("metamarkers/markers_zeng_10x_nuclei.csv.gz"),
    zeng_smart_nuclei = read_markers("metamarkers/markers_zeng_smart_nuclei.csv.gz"),
    zeng_smart_cells = read_markers("metamarkers/markers_zeng_smart_cells.csv.gz"),
    zeng_10xv2_wba = read_markers("metamarkers/markers_zeng_10xv2_wba.csv.gz"),
    zeng_10xv3_wba = read_markers("metamarkers/markers_zeng_10xv3_wba.csv.gz"),
    macosko_wba = read_markers("metamarkers/markers_macosko_wba.csv.gz")
)


original_mouse_subclass_meta_markers = make_meta_markers(orig_mouse_markers, detailed_stats = TRUE)

original_mouse_subclass_meta_markers %>% group_by(cell_type) %>% filter(rank <= 10)


```




Functions for pulling together hierarchical clustering plus looking at marker gene expression for each cluster


```{r}

get_max_norm_agg_col_annotation = function(sce_obj, clust_for_markers, celltype_color_palette){
  #Get the average of the aggregated subclass MetaMarker expression per cluster as column annotations
  col_annot_data = as.data.frame(colData(sce_obj)) %>% select(!! sym(clust_for_markers), agg_vip, agg_sncg, agg_pax6, 
                                                      agg_lamp5, agg_lamp5_lhx6,
                                                      agg_pvalb, agg_chandelier, 
                                                      agg_sst, agg_sst_chodl) %>%
    group_by(!! sym(clust_for_markers)) %>% summarise(agg_vip = mean(agg_vip), agg_sncg = mean(agg_sncg), 
                                            agg_pax6 = mean(agg_pax6), agg_lamp5 = mean(agg_lamp5),
                                            agg_lamp5_lhx6 = mean(agg_lamp5_lhx6),
                                            agg_pvalb = mean(agg_pvalb), agg_chandelier = mean(agg_chandelier),
                                            agg_sst = mean(agg_sst), agg_sst_chodl = mean(agg_sst_chodl))
  
  col_annot_data = col_annot_data %>% select(-c(!! sym(clust_for_markers)))
  #Get the mean value for the agg expression within each cluster
  celltype_cent_means = rowMeans(col_annot_data)
  #Dividing each cluster values by the cluster mean, produces enrichment with respect to that cluster
  #Pretending a cluster is a cell, this gives the enrichment for that marker expression assuming equal marker expression across all the marker sets
  col_annot_data = col_annot_data / celltype_cent_means
  
  #Then max scale so each cell-type is 0-1
  #Value of 1 indicates that cluster was the cluster with the highest enrichment for that marker set compared to all the other clusters
  col_annot_data = as.data.frame(apply(col_annot_data, 2, function(x) x/max(x) ))

  
  vip_col = circlize::colorRamp2(c(0, .4, 1), c("white","white", celltype_color_palette[['Vip']]))
  sncg_col = circlize::colorRamp2(c(0, .4, 1), c("white","white", celltype_color_palette[['Sncg']]))
  pax6_col = circlize::colorRamp2(c(0, .4, 1), c("white","white", celltype_color_palette[['Pax6']]))
  lamp5_col = circlize::colorRamp2(c(0, .4, 1), c("white","white", celltype_color_palette[['Lamp5']]))
  lamp5_lhx6_col = circlize::colorRamp2(c(0, .4, 1), c("white","white", celltype_color_palette[['Lamp5 Lhx6']]))
  pvalb_col = circlize::colorRamp2(c(0, .4, 1), c("white","white", celltype_color_palette[['Pvalb']]))
  chandelier_col = circlize::colorRamp2(c(0, .4, 1), c("white","white", celltype_color_palette[['Chandelier']]))
  sst_col = circlize::colorRamp2(c(0, .4, 1), c("white","white", celltype_color_palette[['Sst']]))
  sst_chodl_col = circlize::colorRamp2(c(0, .4, 1), c("white","white", celltype_color_palette[['Sst Chodl']]))
  
  col_annot = ComplexHeatmap::HeatmapAnnotation(agg_vip = col_annot_data$agg_vip,
                                                agg_sncg = col_annot_data$agg_sncg,
                                                agg_pax6 = col_annot_data$agg_pax6,
                                                agg_lamp5 = col_annot_data$agg_lamp5,
                                                agg_lamp5_lhx6 = col_annot_data$agg_lamp5_lhx6,
                                                agg_pvalb = col_annot_data$agg_pvalb,
                                                agg_chandelier = col_annot_data$agg_chandelier,
                                                agg_sst = col_annot_data$agg_sst,
                                                agg_sst_chodl = col_annot_data$agg_sst_chodl,
                                                col = list(agg_vip = vip_col,
                                                           agg_sncg = sncg_col, 
                                                           agg_pax6 = pax6_col,
                                                           agg_lamp5 = lamp5_col,
                                                           agg_lamp5_lhx6 = lamp5_lhx6_col,
                                                           agg_pvalb = pvalb_col,
                                                           agg_chandelier = chandelier_col,
                                                           agg_sst = sst_col,
                                                           agg_sst_chodl = sst_chodl_col),
                                                show_legend = c(T, F, F, F, F, F, F, F, F))

  return(col_annot)
}


get_scaled_enrichment_col_annotation = function(sce_obj, clust_for_markers, celltype_color_palette){
  
  #Get the average of the aggregated subclass MetaMarker expression per cluster as column annotations
  col_annot_data = as.data.frame(colData(sce_obj)) %>% select(!! sym(clust_for_markers), `enrich.Vip`, `enrich.Sncg`, `enrich.Pax6`, 
                                                      `enrich.Lamp5`, `enrich.Lamp5.Lhx6`,
                                                      `enrich.Pvalb`, `enrich.Chandelier`, 
                                                      `enrich.Sst`, `enrich.Sst.Chodl`) %>%
    group_by(!! sym(clust_for_markers)) %>% summarise(`enrich.Vip` = mean(`enrich.Vip`),  `enrich.Sncg` = mean( `enrich.Sncg`), 
                                            `enrich.Pax6` = mean(`enrich.Pax6`), `enrich.Lamp5` = mean(`enrich.Lamp5`),
                                            `enrich.Lamp5.Lhx6` = mean(`enrich.Lamp5.Lhx6`),
                                            `enrich.Pvalb` = mean(`enrich.Pvalb`), `enrich.Chandelier` = mean(`enrich.Chandelier`),
                                            `enrich.Sst` = mean(`enrich.Sst`), `enrich.Sst.Chodl` = mean(`enrich.Sst.Chodl`))
  
  #Scale the average enrichments per cluster
  col_annot_data = col_annot_data %>% select(-c(!! sym(clust_for_markers)))
  col_annot_data = as.data.frame(scale(col_annot_data))
  col_mins = matrixStats::colMins(as.matrix(col_annot_data))
  col_maxs = matrixStats::colMaxs(as.matrix(col_annot_data))
  
  vip_col = circlize::colorRamp2(c(col_mins['enrich.Vip'], 0, col_maxs['enrich.Vip']), c("white","white", celltype_color_palette[['Vip']]))
  sncg_col = circlize::colorRamp2(c(col_mins['enrich.Sncg'], 0, col_maxs['enrich.Sncg']), c("white","white", celltype_color_palette[['Sncg']]))
  pax6_col = circlize::colorRamp2(c(col_mins['enrich.Pax6'], 0, col_maxs['enrich.Pax6']), c("white","white", celltype_color_palette[['Pax6']]))
  lamp5_col = circlize::colorRamp2(c(col_mins['enrich.Lamp5'], 0, col_maxs['enrich.Lamp5']), c("white","white", celltype_color_palette[['Lamp5']]))
  lamp5_lhx6_col = circlize::colorRamp2(c(col_mins['enrich.Lamp5.Lhx6'], 0, col_maxs['enrich.Lamp5.Lhx6']), 
                                        c("white","white", celltype_color_palette[['Lamp5 Lhx6']]))
  pvalb_col = circlize::colorRamp2(c(col_mins['enrich.Pvalb'], 0, col_maxs['enrich.Pvalb']), c("white","white", celltype_color_palette[['Pvalb']]))
  chandelier_col = circlize::colorRamp2(c(col_mins['enrich.Chandelier'], 0, col_maxs['enrich.Chandelier']), 
                                        c("white","white", celltype_color_palette[['Chandelier']]))
  sst_col = circlize::colorRamp2(c(col_mins['enrich.Sst'], 0, col_maxs['enrich.Sst']), c("white","white", celltype_color_palette[['Sst']]))
  sst_chodl_col = circlize::colorRamp2(c(col_mins['enrich.Sst.Chodl'], 0, col_maxs['enrich.Sst.Chodl']), 
                                       c("white","white", celltype_color_palette[['Sst Chodl']]))
  
  col_annot = ComplexHeatmap::HeatmapAnnotation(enrich.Vip = col_annot_data$enrich.Vip,
                                                enrich.Sncg = col_annot_data$enrich.Sncg,
                                                enrich.Pax6 = col_annot_data$enrich.Pax6,
                                                enrich.Lamp5 = col_annot_data$enrich.Lamp5,
                                                enrich.Lamp5.Lhx6 = col_annot_data$enrich.Lamp5.Lhx6,
                                                enrich.Pvalb = col_annot_data$enrich.Pvalb,
                                                enrich.Chandelier = col_annot_data$enrich.Chandelier,
                                                enrich.Sst = col_annot_data$enrich.Sst,
                                                enrich.Sst.Chodl = col_annot_data$enrich.Sst.Chodl,
                                                col = list(enrich.Vip = vip_col,
                                                           enrich.Sncg  = sncg_col, 
                                                           enrich.Pax6 = pax6_col,
                                                           enrich.Lamp5 = lamp5_col,
                                                           enrich.Lamp5.Lhx6 = lamp5_lhx6_col,
                                                           enrich.Pvalb = pvalb_col,
                                                           enrich.Chandelier = chandelier_col,
                                                           enrich.Sst = sst_col,
                                                           enrich.Sst.Chodl = sst_chodl_col),
                                                show_legend = c(T, F, F, F, F, F, F, F, F))

  return(col_annot)
}


get_hierarch_clust = function(sce_obj, cluster_markers, num_markers, plot_title, centroid_clust_name, want_scaled = F){
  
  top_clust_markers = cluster_markers %>% group_by(cell_type) %>% arrange(desc(auroc), .by_group = T) %>% slice_max(auroc, n = num_markers)
  top_clust_markers = unique(top_clust_markers$gene)
  
  #Expression data for that list
  gene_filt = rownames(sce_obj) %in% top_clust_markers
  exp_data = cpm(sce_obj)[gene_filt, ]

  #Compute centroids of gene expression per cluster
  exp_data = as.data.frame(t(as.matrix(exp_data)))
  exp_data = exp_data %>% mutate(cluster = colData(sce_obj)[[centroid_clust_name]])
  
  centroids = exp_data %>% group_by(cluster) %>% summarize(across(which(colnames(exp_data)!= 'cluster'), median)) %>% as.data.frame()
  #Get back into genes on rows and subclass on columns
  rownames(centroids) = centroids$cluster
  centroids = t(centroids[ ,2:ncol(centroids)])
  
  #Compute a distance matrix from the correlations of the centroids
  centroid_corr = cor(centroids, method = 'spearman')

  #Hierarchical clusting on the correlation distance matrix
  hclust_avg <- hclust(as.dist(1-centroid_corr), method = 'average')

  if(want_scaled){
    column_annotation = get_scaled_enrichment_col_annotation(sce_obj, centroid_clust_name, celltype_color_palette)
  }else{
    column_annotation = get_max_norm_agg_col_annotation(sce_obj, celltype_color_palette)
  }
  
  
  corr_cols <- rev(grDevices::colorRampPalette(RColorBrewer::brewer.pal(11,"RdYlBu"))(100))
  cent_hm = ComplexHeatmap::Heatmap(centroid_corr, col = corr_cols, name = 'spearman' ,
                                    cluster_columns = hclust_avg,
                                    cluster_rows = hclust_avg,
                                    column_title = plot_title,
                                    top_annotation = column_annotation)
  cent_hm = ComplexHeatmap::draw(cent_hm)
  return(cent_hm)
}


```



Get aggregate expression of the top conserved marker genes
```{r}
num_primate_markers = 50
num_mouse_markers = 50

get_agg_exp = function(meta_markers, cellType, num_markers, sce_obj){
  grab_genes = meta_markers %>% filter(cell_type == cellType & rank <= num_markers) %>% pull(gene)
  gene_index = rownames(sce_obj) %in% grab_genes
  agg_exp = colSums(cpm(sce_obj)[gene_index, ])
  return(agg_exp)
}


get_enrich_exp_meta = function(meta_markers,num_markers, sce_obj){
  #Drop any metdata columns with enrich in the name
  current_metadata = colData(sce_obj)
  enrich_skip_index = !grepl('enrich', colnames(current_metadata))
  current_metadata  = current_metadata[ , enrich_skip_index]
  
  top_markers = meta_markers %>% filter(rank <= num_markers)
  ct_scores = score_cells(log1p(cpm(sce_obj)), top_markers)
  ct_enrichment = compute_marker_enrichment(ct_scores)
  rownames(ct_enrichment) = gsub( 'all|', 'enrich|', rownames(ct_enrichment), fixed = T)
  current_metadata = cbind(current_metadata, t(ct_enrichment))
  
  return(current_metadata)
}


#Enrichment of Allen primate marker expression
colData(caglayan_inhib_copy) = get_enrich_exp_meta(primate_subclass_meta_markers, num_primate_markers, caglayan_inhib_copy)

#Alltogether in a hierarchical clustered plot
caglayan_hierarch_plot = get_hierarch_clust(caglayan_inhib_copy, cluster_markers = caglayan_integ_inhib_markers , 
                                            num_markers = 50,plot_title = 'Caglayan Primate clusters: Allen primate MetaMarkers', 
                                            centroid_clust_name = 'seurat_clusters', want_scale = T)
caglayan_hierarch_plot

pdf(file = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/non_allen_data/caglayan_cross_primate_primate_MM_cluster_tax_heat.pdf',
    useDingbats = F, height = 8, width = 8)
caglayan_hierarch_plot
dev.off()


#Enrichment of Mouse marker expression
colData(caglayan_inhib_copy) = get_enrich_exp_meta(mouse_subclass_meta_markers, num_mouse_markers, caglayan_inhib_copy)

#Alltogether in a hierarchical clustered plot
caglayan_hierarch_plot = get_hierarch_clust(caglayan_inhib_copy, cluster_markers = caglayan_integ_inhib_markers , 
                                            num_markers = 50,plot_title = 'Caglayan Primate clusters: Mouse MetaMarkers', 
                                            centroid_clust_name = 'seurat_clusters', want_scale = T)
caglayan_hierarch_plot

pdf(file = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/non_allen_data/caglayan_cross_primate_mouse_MM_cluster_tax_heat.pdf',
    useDingbats = F, height = 8, width = 8)
caglayan_hierarch_plot
dev.off()


```







################################
Other cross-primate dataset

Sestan dataset

downloaded the seurat rds object from : http://resources.sestanlab.org/PFC/
paper:https://pmc.ncbi.nlm.nih.gov/articles/PMC9614553

snRNA-seq

dorsolateral preFrontal cortex

101, 845 inhibitory cells

################################




```{r}

sestan_data = readRDS('/inkwell03/werner/sc_data/primate_ma_sestan/PFC_snRNAseq_liftover.rds')
sestan_data

```


```{r}

sestan_data@assays$RNA@meta.features


table(sestan_data$species, sestan_data$subclass)


sestan_inhib = subset(sestan_data, subset = class == 'InN')
rm(sestan_data)
gc()
```


UMAP with the object looks to be integrated, but they don't include the integrated values in the seurat object

```{r}

sestan_inhib[[]]

DimPlot(sestan_inhib, reduction = "umap", group.by = 'species')
DimPlot(sestan_inhib, reduction = "umap", group.by = 'subclass')


cge_mge_label = rep('CGE', length = ncol(sestan_inhib))
cge_mge_label[sestan_inhib$subclass %in% c('PVALB','PVALB ChC','SST','SST HGF','SST NPY')] = 'MGE'

table(cge_mge_label)
table(sestan_inhib$subclass)


sestan_inhib$cge_mge = cge_mge_label

```


Get the same SNCG and PAX6 DE plot as from the Caglayan data, split by species, normalize, get markers, etc.


```{r}

sestan_human_data = subset(sestan_inhib, subset = species == 'Human')
sestan_chimp_data = subset(sestan_inhib, subset = species == 'Chimpanzee')
sestan_macaque_data = subset(sestan_inhib, subset = species == 'Rhesus')
sestan_marmoset_data = subset(sestan_inhib, subset = species == 'Marmoset')


sestan_human_data = NormalizeData(sestan_human_data, normalization.method = 'RC', scale.factor = 1e6, verbose = F)
sestan_human_inhib_class_markers = compute_markers(sestan_human_data@assays$RNA@data, sestan_human_data$cge_mge)
gc()
sestan_chimp_data = NormalizeData(sestan_chimp_data, normalization.method = 'RC', scale.factor = 1e6, verbose = F)
sestan_chimp_inhib_class_markers = compute_markers(sestan_chimp_data@assays$RNA@data, sestan_chimp_data$cge_mge)
gc()
sestan_macaque_data = NormalizeData(sestan_macaque_data, normalization.method = 'RC', scale.factor = 1e6, verbose = F)
sestan_macaque_inhib_class_markers = compute_markers(sestan_macaque_data@assays$RNA@data, sestan_macaque_data$cge_mge)
gc()
sestan_marmoset_data = NormalizeData(sestan_marmoset_data, normalization.method = 'RC', scale.factor = 1e6, verbose = F)
sestan_marmoset_inhib_class_markers = compute_markers(sestan_marmoset_data@assays$RNA@data, sestan_marmoset_data$cge_mge)
gc()

```



```{r}
#pax6
pax6_sp_1 = sestan_human_inhib_class_markers %>% filter(gene == 'PAX6') %>% 
  select(cell_type, gene, fold_change) %>% 
  mutate(species = 'human')

pax6_sp_2 = sestan_chimp_inhib_class_markers %>% filter(gene == 'PAX6') %>% 
  select(cell_type, gene, fold_change) %>% 
  mutate(species = 'chimp')

pax6_sp_3 = sestan_macaque_inhib_class_markers %>% filter(gene == 'PAX6') %>% 
  select(cell_type, gene, fold_change) %>% 
  mutate(species = 'macaque')

pax6_sp_4 = sestan_marmoset_inhib_class_markers %>% filter(gene == 'PAX6') %>% 
  select(cell_type, gene, fold_change) %>% 
  mutate(species = 'marmoset')

pax6_DE_df = do.call('rbind', list(pax6_sp_1 , pax6_sp_2 , pax6_sp_3, pax6_sp_4))
pax6_DE_df = pax6_DE_df %>% tidyr::pivot_wider(names_from = cell_type, values_from = fold_change)

#Lamp5
lamp5_sp_1 = sestan_human_inhib_class_markers %>% filter(gene == 'LAMP5') %>% 
  select(cell_type, gene, fold_change) %>% 
  mutate(species = 'human')

lamp5_sp_2 = sestan_chimp_inhib_class_markers %>% filter(gene == 'LAMP5') %>% 
  select(cell_type, gene, fold_change) %>% 
  mutate(species = 'chimp')

lamp5_sp_3 = sestan_macaque_inhib_class_markers %>% filter(gene == 'LAMP5') %>% 
  select(cell_type, gene, fold_change) %>% 
  mutate(species = 'macaque')

lamp5_sp_4 = sestan_marmoset_inhib_class_markers %>% filter(gene == 'LAMP5') %>% 
  select(cell_type, gene, fold_change) %>% 
  mutate(species = 'marmoset')

lamp5_DE_df = do.call('rbind', list(lamp5_sp_1 , lamp5_sp_2 , lamp5_sp_3, lamp5_sp_4))
lamp5_DE_df = lamp5_DE_df %>% tidyr::pivot_wider(names_from = cell_type, values_from = fold_change)

#Vip
vip_sp_1 = sestan_human_inhib_class_markers %>% filter(gene == 'VIP') %>% 
  select(cell_type, gene, fold_change) %>% 
  mutate(species = 'human')

vip_sp_2 = sestan_chimp_inhib_class_markers %>% filter(gene == 'VIP') %>% 
  select(cell_type, gene, fold_change) %>% 
  mutate(species = 'chimp')

vip_sp_3 = sestan_macaque_inhib_class_markers %>% filter(gene == 'VIP') %>% 
  select(cell_type, gene, fold_change) %>% 
  mutate(species = 'macaque')

vip_sp_4 = sestan_marmoset_inhib_class_markers %>% filter(gene == 'VIP') %>% 
  select(cell_type, gene, fold_change) %>% 
  mutate(species = 'marmoset')

vip_DE_df = do.call('rbind', list(vip_sp_1 , vip_sp_2 , vip_sp_3, vip_sp_4))
vip_DE_df = vip_DE_df %>% tidyr::pivot_wider(names_from = cell_type, values_from = fold_change)


#sncg
sncg_sp_1 = sestan_human_inhib_class_markers %>% filter(gene == 'SNCG') %>% 
  select(cell_type, gene, fold_change) %>% 
  mutate(species = 'human')

sncg_sp_2 = sestan_chimp_inhib_class_markers %>% filter(gene == 'SNCG') %>% 
  select(cell_type, gene, fold_change) %>% 
  mutate(species = 'chimp')

sncg_sp_3 = sestan_macaque_inhib_class_markers %>% filter(gene == 'SNCG') %>% 
  select(cell_type, gene, fold_change) %>% 
  mutate(species = 'macaque')

sncg_sp_4 = sestan_marmoset_inhib_class_markers %>% filter(gene == 'SNCG') %>% 
  select(cell_type, gene, fold_change) %>% 
  mutate(species = 'marmoset')


sncg_DE_df = do.call('rbind', list(sncg_sp_1 , sncg_sp_2 , sncg_sp_3, sncg_sp_4))
sncg_DE_df = sncg_DE_df %>% tidyr::pivot_wider(names_from = cell_type, values_from = fold_change)


marker_gene_df = rbind(pax6_DE_df, sncg_DE_df, vip_DE_df, lamp5_DE_df)

marker_gene_df = marker_gene_df %>% mutate(log2_FC_CGE = log2(CGE))
marker_gene_df$species = factor(marker_gene_df$species , levels = c('marmoset','macaque','chimp','human'))

ggplot(marker_gene_df  , aes(x = log2_FC_CGE, y = species , color = gene, shape = species)) + geom_point(size = 5,alpha = .5) +
  xlab('log2(FC) CGE/MGE') + ylab('Species') +
  scale_color_manual(values = c('SNCG' = celltype_color_palette[['Sncg']], 
                                'PAX6' = celltype_color_palette[['Pax6']],
                                'LAMP5' = celltype_color_palette[['Lamp5']],
                                'VIP' = celltype_color_palette[['Vip']])) + 
  geom_hline(yintercept = 0, linetype = 'dashed', color = 'red') +
  geom_vline(xintercept = 0, linetype = 'dashed', color = 'red') +
  scale_shape_manual(values = c('human' = 'square', 'chimp' = 'triangle', 'macaque' = 'circle', 'marmoset' = 'cross')) +
  theme_bw() + ggtitle('Sestan cross-primate data')


ggsave(filename = 'sesten_pax6_sncg_vip_lamp5_MGE_CGE_DE_scatter.pdf', path = '../../graphs/non_allen_data/', device = 'pdf',
       useDingbats = F, height = 3, width = 5)


```









```{r}

#CPM normalization
sestan_inhib <- NormalizeData(sestan_inhib, normalization.method = 'RC', scale.factor = 1e6, verbose = F)
sestan_inhib <- FindVariableFeatures(sestan_inhib, selection.method = "vst", nfeatures = 2000, verbose = F)
all.genes <- rownames(sestan_inhib)
sestan_inhib <- ScaleData(sestan_inhib, features = all.genes, verbose = F)

sestan_inhib <- RunPCA(sestan_inhib, features = VariableFeatures(object =sestan_inhib), verbose = F)

sestan_inhib <- FindNeighbors(sestan_inhib, dims = 1:25, verbose = F)
sestan_inhib <- FindClusters(sestan_inhib, verbose = F)

sestan_inhib <- RunUMAP(sestan_inhib, dims = 1:25, verbose = F)


```



```{r}

DimPlot(sestan_inhib, reduction = "umap", group.by = 'species')
DimPlot(sestan_inhib, reduction = "umap", group.by = 'subclass')

```

Split by species and integrate

```{r}
options(future.globals.maxSize = 1000 * 1024^2)

```


```{r}
table(sestan_inhib$species)

sestan_human = subset(sestan_inhib, subset = species == 'Human')
sestan_chimp = subset(sestan_inhib, subset = species == 'Chimpanzee')
sestan_marmoset = subset(sestan_inhib, subset = species == 'Marmoset')
sestan_macaque = subset(sestan_inhib, subset = species == 'Rhesus')

rm(sestan_inhib)
gc()

dim(sestan_human)
dim(sestan_chimp)
dim(sestan_marmoset)
dim(sestan_macaque)
```


```{r}

#Make the list of objects
seurat.list = list('sestan_human' = sestan_human , 'sestan_chimp' = sestan_chimp, 'sestan_macaque' = sestan_macaque, 'sestan_marmoset' = sestan_marmoset)
#Rm individual datasets to save space, doesn't take that long for the above chuncks to rerun
rm(sestan_human , sestan_chimp, sestan_macaque,  sestan_marmoset)
gc()


integration_features <- SelectIntegrationFeatures(object.list = seurat.list)
length(integration_features)

start = Sys.time()
integration_anchors <- FindIntegrationAnchors(object.list = seurat.list, anchor.features = integration_features)
start - Sys.time()

```


```{r}
rm(seurat.list)
gc()

start = Sys.time()

#Make a single integrated object
interneuron.combined <- IntegrateData(anchorset = integration_anchors)

start - Sys.time()

rm(integration_anchors)
gc()
```



```{r}
start = Sys.time()

DefaultAssay(interneuron.combined) <- "integrated"
# Run the standard workflow for visualization and clustering
interneuron.combined <- ScaleData(interneuron.combined, verbose = FALSE)
interneuron.combined <- RunPCA(interneuron.combined, verbose = FALSE)
interneuron.combined <- RunUMAP(interneuron.combined, reduction = "pca", dims = 1:25)

Sys.time() - start

```


```{r}
DimPlot(interneuron.combined, reduction = "umap", group.by = 'species')
DimPlot(interneuron.combined, reduction = "umap", group.by = 'subclass')

```




Cluster in the integrated space

```{r}

interneuron.combined <- FindNeighbors(interneuron.combined , dims = 1:25, verbose = F, assay = 'integrated')
interneuron.combined <- FindClusters(interneuron.combined , verbose = F, graph.name = 'integrated_snn')


```


```{r}
DimPlot(interneuron.combined, reduction = "umap", group.by = 'species')
DimPlot(interneuron.combined, reduction = "umap", group.by = 'subclass')
DimPlot(interneuron.combined, reduction = "umap", group.by = 'seurat_clusters')
```



Save the integrated dataset

```{r}

saveRDS(interneuron.combined, file = '/inkwell03/werner/sc_data/primate_ma_sestan/sestan_integrated_gaba.rds')

```





```{r}

sesten_gaba_data = readRDS('/inkwell03/werner/sc_data/primate_ma_sestan/sestan_integrated_gaba.rds')

DefaultAssay(sesten_gaba_data) <- "RNA"
```


```{r}

#Add the cge and MGE labels of the original authors annotations
sesten_gaba_data[[]]
table(sesten_gaba_data$subclass)

cge_mge_label = rep('CGE', length = ncol(sesten_gaba_data))
cge_mge_label[sesten_gaba_data$subclass %in% c('PVALB','PVALB ChC','SST','SST HGF','SST NPY')] = 'MGE'
sesten_gaba_data$cge_mge = cge_mge_label


DimPlot(sesten_gaba_data , reduction = "umap", group.by = 'species')
ggsave(filename = 'sestan_integrated_species_umap.pdf', path = '../../graphs/non_allen_data/umaps/', device = 'pdf',
       useDingbats = F, height = 4, width = 6)


DimPlot(sesten_gaba_data , reduction = "umap", group.by = 'subclass')
ggsave(filename = 'sestan_integrated_author_annot_umap.pdf', path = '../../graphs/non_allen_data/umaps/', device = 'pdf',
       useDingbats = F, height = 4, width = 6)

DimPlot(sesten_gaba_data , reduction = "umap", group.by = 'cge_mge')
ggsave(filename = 'sestan_integrated_author_cge_mge_umap.pdf', path = '../../graphs/non_allen_data/umaps/', device = 'pdf',
       useDingbats = F, height = 4, width = 6)


DimPlot(sesten_gaba_data , reduction = "umap", group.by = 'seurat_clusters', label = T)
ggsave(filename = 'sestan_integrated_de_novo_clusters_umap.pdf', path = '../../graphs/non_allen_data/umaps/', device = 'pdf',
       useDingbats = F, height = 4, width = 6)
```


Also cluster each species independently in their non-integrated data and save as species specific clusters




```{r}

#human
temp_sestan_human = subset(sesten_gaba_data, subset = species == 'Human')
temp_sestan_human  <- FindVariableFeatures(temp_sestan_human , selection.method = "vst", nfeatures = 2000, verbose = F)
temp_sestan_human <- ScaleData(temp_sestan_human, verbose = FALSE)
temp_sestan_human <- RunPCA(temp_sestan_human, verbose = FALSE)
temp_sestan_human <- RunUMAP(temp_sestan_human, reduction = "pca", dims = 1:25, verbose = F)
temp_sestan_human <- FindNeighbors(temp_sestan_human , dims = 1:25, verbose = F, assay = 'RNA')
temp_sestan_human <- FindClusters(temp_sestan_human , verbose = F, graph.name = 'RNA_snn')

temp_sestan_human$species_specific_clusters = temp_sestan_human$RNA_snn_res.0.8

DimPlot(temp_sestan_human, reduction = 'pca', group.by = 'species_specific_clusters')
DimPlot(temp_sestan_human, reduction = 'umap', group.by = 'species_specific_clusters')
DimPlot(temp_sestan_human, reduction = 'umap', group.by = 'subclass')


#chimp
temp_sestan_chimp = subset(sesten_gaba_data, subset = species == 'Chimpanzee')
temp_sestan_chimp  <- FindVariableFeatures(temp_sestan_chimp , selection.method = "vst", nfeatures = 2000, verbose = F)
temp_sestan_chimp <- ScaleData(temp_sestan_chimp, verbose = FALSE)
temp_sestan_chimp <- RunPCA(temp_sestan_chimp, verbose = FALSE)
temp_sestan_chimp <- RunUMAP(temp_sestan_chimp, reduction = "pca", dims = 1:25, verbose = F)
temp_sestan_chimp <- FindNeighbors(temp_sestan_chimp , dims = 1:25, verbose = F, assay = 'RNA')
temp_sestan_chimp <- FindClusters(temp_sestan_chimp , verbose = F, graph.name = 'RNA_snn')

temp_sestan_chimp$species_specific_clusters = temp_sestan_chimp$RNA_snn_res.0.8

DimPlot(temp_sestan_chimp, reduction = 'pca', group.by = 'species_specific_clusters')
DimPlot(temp_sestan_chimp, reduction = 'umap', group.by = 'species_specific_clusters')
DimPlot(temp_sestan_chimp, reduction = 'umap', group.by = 'subclass')

#macaque
temp_sestan_macaque = subset(sesten_gaba_data, subset = species == 'Rhesus')
temp_sestan_macaque  <- FindVariableFeatures(temp_sestan_macaque , selection.method = "vst", nfeatures = 2000, verbose = F)
temp_sestan_macaque <- ScaleData(temp_sestan_macaque, verbose = FALSE)
temp_sestan_macaque <- RunPCA(temp_sestan_macaque, verbose = FALSE)
temp_sestan_macaque <- RunUMAP(temp_sestan_macaque, reduction = "pca", dims = 1:25, verbose = F)
temp_sestan_macaque <- FindNeighbors(temp_sestan_macaque , dims = 1:25, verbose = F, assay = 'RNA')
temp_sestan_macaque <- FindClusters(temp_sestan_macaque , verbose = F, graph.name = 'RNA_snn')

temp_sestan_macaque$species_specific_clusters = temp_sestan_macaque$RNA_snn_res.0.8

DimPlot(temp_sestan_macaque, reduction = 'pca', group.by = 'species_specific_clusters')
DimPlot(temp_sestan_macaque, reduction = 'umap', group.by = 'species_specific_clusters')
DimPlot(temp_sestan_macaque, reduction = 'umap', group.by = 'subclass')


#marmoset
temp_sestan_marmoset = subset(sesten_gaba_data, subset = species == 'Marmoset')
temp_sestan_marmoset  <- FindVariableFeatures(temp_sestan_marmoset , selection.method = "vst", nfeatures = 2000, verbose = F)
temp_sestan_marmoset <- ScaleData(temp_sestan_marmoset, verbose = FALSE)
temp_sestan_marmoset <- RunPCA(temp_sestan_marmoset, verbose = FALSE)
temp_sestan_marmoset <- RunUMAP(temp_sestan_marmoset, reduction = "pca", dims = 1:25, verbose = F)
temp_sestan_marmoset <- FindNeighbors(temp_sestan_marmoset , dims = 1:25, verbose = F, assay = 'RNA')
temp_sestan_marmoset <- FindClusters(temp_sestan_marmoset , verbose = F, graph.name = 'RNA_snn')

temp_sestan_marmoset$species_specific_clusters = temp_sestan_marmoset$RNA_snn_res.0.8

DimPlot(temp_sestan_marmoset, reduction = 'pca', group.by = 'species_specific_clusters')
DimPlot(temp_sestan_marmoset, reduction = 'umap', group.by = 'species_specific_clusters')
DimPlot(temp_sestan_marmoset, reduction = 'umap', group.by = 'subclass')

DimPlot(temp_sestan_human, reduction = 'umap', group.by = 'species_specific_clusters', label = T)
DimPlot(temp_sestan_chimp, reduction = 'umap', group.by = 'species_specific_clusters', label = T)
DimPlot(temp_sestan_macaque, reduction = 'umap', group.by = 'species_specific_clusters', label = T)
DimPlot(temp_sestan_marmoset, reduction = 'umap', group.by = 'species_specific_clusters', label = T)

```


```{r}

sesten_gaba_data$species_specific_clusters = rep(NA, length = ncol(sesten_gaba_data))
table(is.na(sesten_gaba_data $species_specifc_clusters))

all_species_barcodes = c(temp_sestan_human$cell.name, temp_sestan_chimp$cell.name, temp_sestan_macaque$cell.name, temp_sestan_marmoset$cell.name)

all_species_clusters = c(temp_sestan_human$species_specific_clusters, temp_sestan_chimp$species_specific_clusters, 
                         temp_sestan_macaque$species_specific_clusters, temp_sestan_marmoset$species_specific_clusters)

index = match(sesten_gaba_data$cell.name, all_species_barcodes)
sesten_gaba_data$species_specific_clusters = all_species_clusters[index]
sesten_gaba_data[[]] %>% filter(integrated_snn_res.0.8 == 17) %>% group_by(species, species_specific_clusters) %>% summarise(n = n())

sesten_gaba_data[[]] = sesten_gaba_data[[]]  %>% select(-c(species_specifc_clusters))



saveRDS(sesten_gaba_data, file = '/inkwell03/werner/sc_data/primate_ma_sestan/sestan_integrated_gaba.rds')

```





```{r}

sesten_gaba_data = readRDS('/inkwell03/werner/sc_data/primate_ma_sestan/sestan_integrated_gaba.rds')
length(rownames(sesten_gaba_data))
table(rownames(sesten_gaba_data) %in% final_present_1to1s$`Gene name`)

DefaultAssay(sesten_gaba_data) <- "RNA"

sesten_gaba_data = as.SingleCellExperiment(sesten_gaba_data)
names(assays(sesten_gaba_data)) = c('counts','cpm')

colnames(colData(sesten_gaba_data))


```




######################
Looking at the cluster markers from clusters derived via the integrated data

#####################


```{r}
sesten_integ_inhib_markers = compute_markers(cpm(sesten_gaba_data), sesten_gaba_data$seurat_clusters)
gc()
```



Get aggregate expression of the top conserved marker genes
```{r}

#Enrichment of Allen primate marker expression
colData(sesten_gaba_data) = get_enrich_exp_meta(primate_subclass_meta_markers, num_primate_markers, sesten_gaba_data)

#Alltogether in a hierarchical clustered plot
sesten_hierarch_plot = get_hierarch_clust(sesten_gaba_data, cluster_markers = sesten_integ_inhib_markers  , 
                                            num_markers = 50,plot_title = 'Sestan Primate clusters: Allen primate MetaMarkers', 
                                            centroid_clust_name = 'seurat_clusters', want_scale = T)
sesten_hierarch_plot 

pdf(file = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/non_allen_data/sesten_cross_primate_primate_MM_cluster_tax_heat.pdf',
    useDingbats = F, height = 8, width = 8)
sesten_hierarch_plot 
dev.off()


#Enrichment of Mouse marker expression
colData(sesten_gaba_data) = get_enrich_exp_meta(mouse_subclass_meta_markers, num_mouse_markers, sesten_gaba_data)

#Alltogether in a hierarchical clustered plot
sesten_hierarch_plot = get_hierarch_clust(sesten_gaba_data, cluster_markers = sesten_integ_inhib_markers  , 
                                            num_markers = 50,plot_title = 'Sestan Primate clusters: Mouse MetaMarkers', 
                                            centroid_clust_name = 'seurat_clusters', want_scale = T)
sesten_hierarch_plot 


pdf(file = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/non_allen_data/sesten_cross_primate_mouse_MM_cluster_tax_heat.pdf',
    useDingbats = F, height = 8, width = 8)
sesten_hierarch_plot 
dev.off()



```


Everything mostly lines up, with mouse markers and primate markers segregating across the de novo clusters into subclasses

Except with one primate Vip cluster. It clusters with Vip via the centroids, but it has the highest Mouse Sncg expression. It also looks to have the weakest primate VIP marker expression.

As a quick check, use the Allen Primate MetaMarkers to annotate the cells and look at the subclass annotation distributions across the clusters.
Anticipating that most of the clusters should have a clear subclass winner, what about this seemingly intermediate Vip/Sncg primate cluster?

It is cluster 20 in Caglayan and cluster 17 in Sestan.


```{r}

top_allen_markers = primate_subclass_meta_markers %>% filter(rank <= 50)
top_mouse_markers = mouse_subclass_meta_markers %>% filter(rank <= 50)

ct_scores = score_cells(log1p(cpm(sesten_gaba_data)), top_allen_markers)
ct_enrichment = compute_marker_enrichment(ct_scores)
ct_pred = assign_cells(ct_scores)
sesten_gaba_data$primate_metaMarker_annot = ct_pred$predicted

confuse_mat_sestan = table(sesten_gaba_data$primate_metaMarker_annot, sesten_gaba_data$seurat_clusters)
cluster_totals_sestan = colSums(confuse_mat_sestan )
confuse_mat_sestan = t(t(confuse_mat_sestan)/cluster_totals_sestan)

#Ordering by the clusters with the most almost perfect matches
row_order = order(rowSums(round(confuse_mat_sestan)), decreasing = T)
round(confuse_mat_sestan)[row_order, ]

col_order = c(2,5,7,13,14,18,21,4,9,10,12,17,19,0,3,6,23,1,15,22,24,16,20,8,11,25) +1

col_fun = circlize::colorRamp2(breaks = c(0,.5, 1) , colors = c('white','orange','red'))
hm = ComplexHeatmap::Heatmap(confuse_mat_sestan,name = 'cluster proportion',
                        col = col_fun,
                        cluster_rows = F, row_order = row_order,
                        cluster_columns = F, column_order = col_order,
                        column_title = 'Sestan cluster predictions with Allen Primate MetaMarkers')
hm = ComplexHeatmap::draw(hm)
pdf(file = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/non_allen_data/primate_MM_predict_sestan_clusters.pdf', 
    useDingbats = F, height = 4, width = 8)
hm
dev.off()


#Mouse subclass markers
ct_scores = score_cells(log1p(cpm(sesten_gaba_data)), top_mouse_markers)
ct_enrichment = compute_marker_enrichment(ct_scores)
ct_pred = assign_cells(ct_scores)
sesten_gaba_data$mouse_metaMarker_annot = ct_pred$predicted

confuse_mat_sestan = table(sesten_gaba_data$mouse_metaMarker_annot, sesten_gaba_data$seurat_clusters)
cluster_totals_sestan = colSums(confuse_mat_sestan )
confuse_mat_sestan = t(t(confuse_mat_sestan)/cluster_totals_sestan)

#Keep the same order as the primate confusion matrix
col_fun = circlize::colorRamp2(breaks = c(0,.5, 1) , colors = c('white','orange','red'))
hm = ComplexHeatmap::Heatmap(confuse_mat_sestan, name = 'cluster proportion',
                        col = col_fun,
                        cluster_rows = F, row_order = row_order,
                        cluster_columns = F, column_order = col_order,
                        column_title = 'Sestan cluster predictions with Mouse MetaMarkers')
hm = ComplexHeatmap::draw(hm)

pdf(file = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/non_allen_data/mouse_MM_predict_sestan_clusters.pdf', 
    useDingbats = F, height = 4, width = 8)
hm
dev.off()




ct_scores = score_cells(log1p(cpm(caglayan_inhib_copy)), top_allen_markers)
ct_enrichment = compute_marker_enrichment(ct_scores)
ct_pred = assign_cells(ct_scores)
caglayan_inhib_copy$primate_metaMarker_annot = ct_pred$predicted

confuse_mat_caglayan = table(caglayan_inhib_copy$primate_metaMarker_annot, caglayan_inhib_copy$seurat_clusters)
cluster_totals_caglayan = colSums(confuse_mat_caglayan )
confuse_mat_caglayan = t(t(confuse_mat_caglayan)/cluster_totals_caglayan)

#Ordering by the clusters with the most almost perfect matches
row_order = order(rowSums(round(confuse_mat_caglayan)), decreasing = T)
round(confuse_mat_caglayan)[row_order, ]

col_order = c(5,7,11,13,17,18,20,26,1,4,6,8,10,21,22,0,3,15,16,23,2,14,27,24,25,12,9,19,28) +1

col_fun = circlize::colorRamp2(breaks = c(0,.5, 1) , colors = c('white','orange','red'))
hm = ComplexHeatmap::Heatmap(confuse_mat_caglayan,name = 'cluster proportion',
                        col = col_fun,
                        cluster_rows = F, row_order = row_order,
                        cluster_columns = F, column_order = col_order,
                        column_title = 'Caglayan cluster predictions with Allen Primate MetaMarkers')
hm = ComplexHeatmap::draw(hm)
pdf(file = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/non_allen_data/primate_MM_predict_caglayan_clusters.pdf', 
    useDingbats = F, height = 4, width = 8)
hm
dev.off()


#Mouse subclass markers
ct_scores = score_cells(log1p(cpm(caglayan_inhib_copy)), top_mouse_markers)
ct_enrichment = compute_marker_enrichment(ct_scores)
ct_pred = assign_cells(ct_scores)
caglayan_inhib_copy$mouse_metaMarker_annot = ct_pred$predicted

confuse_mat_caglayan = table(caglayan_inhib_copy$mouse_metaMarker_annot, caglayan_inhib_copy$seurat_clusters)
cluster_totals_caglayan = colSums(confuse_mat_caglayan )
confuse_mat_caglayan = t(t(confuse_mat_caglayan)/cluster_totals_caglayan)

#Keep the same order as the primate confusion matrix
col_fun = circlize::colorRamp2(breaks = c(0,.5, 1) , colors = c('white','orange','red'))
hm = ComplexHeatmap::Heatmap(confuse_mat_caglayan, name = 'cluster proportion',
                        col = col_fun,
                        cluster_rows = F, row_order = row_order,
                        cluster_columns = F, column_order = col_order,
                        column_title = 'Caglayan cluster predictions with Mouse MetaMarkers')
hm = ComplexHeatmap::draw(hm)

pdf(file = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/non_allen_data/mouse_MM_predict_caglayan_clusters.pdf', 
    useDingbats = F, height = 4, width = 8)
hm
dev.off()

```

Save the non-allen primate cluster markers and send to Hamsini
From the above and the MetaNeighbor results between Allen primate and Mouse data for the non-Allen primate datasets,
Sestan cluster 17 looks Vip when you cluster on centroids and is also large majority Vip when you predict cells based on Allen Primate MetaMarkers.
But, it is majority Sncg when you use the Mouse MetaMarkers and clusters with Mouse Sncg in MetaNeighbor.
Leading idea is the Mouse Sncg is actually an ancestral Sncg subclass, which was derived from Vip. The Mouse Sncg and the Sestan Vip 17 are reflective of this ancestral state, with the Primate Sncg subclass being diversified from them.

Hamsini will take a look at the Mouse Subclass Markers and Sestan cluster markers and see if there's anything interesting with her paralog approach.

So I need the Mouse Subclass MetaMarkers and the Sestan Cluster Markers. Just send all of that over with the plots to show which clusters are which


```{r}

export_meta_markers(mouse_subclass_meta_markers, 'markers_to_share/mouse_subclass_metaMarkers.csv',names(all_mouse_markers) )


export_markers(sesten_integ_inhib_markers, 'markers_to_share/sestan_de_novo_cluster_markers.csv' )

export_markers(caglayan_integ_inhib_markers, 'markers_to_share/caglayan_de_novo_cluster_markers.csv' )

```

```{r}

mouse_subclass_meta_markers = read_meta_markers('markers_to_share/mouse_subclass_metaMarkers.csv.gz')

```




For a specific comparison, plot the primate vip against the primate sncg marker expression for the de novo Vip and Sncg clusters
Then plot mouse Vip against mouse Sncg for those clusters


```{r}

sesten_gaba_data = readRDS('/inkwell03/werner/sc_data/primate_ma_sestan/sestan_integrated_gaba.rds')
DefaultAssay(sesten_gaba_data) <- "RNA"

sesten_gaba_data = as.SingleCellExperiment(sesten_gaba_data)
names(assays(sesten_gaba_data)) = c('counts','cpm')

#get the top primate and mouse MetaMarkers

top_primate_markers = primate_subclass_meta_markers %>% filter(rank <= 50)
top_mouse_markers = mouse_subclass_meta_markers %>% filter(rank <= 50)

#Get the primate enrichments, summarise by the mean enrichment per de novo cluster
ct_scores_primate = score_cells(log1p(cpm(sesten_gaba_data)), top_primate_markers)
ct_enrichment_primate = compute_marker_enrichment(ct_scores_primate)

colnames(ct_enrichment_primate) = sesten_gaba_data$seurat_clusters
ct_enrichment_primate_df = reshape2::melt(t(ct_enrichment_primate), varnames = c('de_novo_cluster','primate_markers'), value.name = c('enrichment'))
ct_enrichment_primate_df = ct_enrichment_primate_df %>% group_by(de_novo_cluster, primate_markers) %>% 
  summarise(mean_enrichment = mean(enrichment), sd_enrichment = sd(enrichment))

#Get the Sncg and Vip clusters
sestan_primate_vip_sncg_df = ct_enrichment_primate_df %>% 
  filter(de_novo_cluster %in% c(16, 20, 4, 9, 10, 12, 17, 19) & primate_markers %in% c('all|Vip','all|Sncg')) %>%
  tidyr::pivot_wider(names_from = primate_markers, values_from = mean_enrichment)

sestan_primate_vip_sncg_df$subclass = rep('Vip', length = nrow(sestan_primate_vip_sncg_df))
sestan_primate_vip_sncg_df$subclass[sestan_primate_vip_sncg_df$de_novo_cluster %in% c(16, 20)] = 'Sncg'

p_yale_primate = ggplot(sestan_primate_vip_sncg_df, aes(x = `all|Vip`, y = `all|Sncg`, color = subclass, label = de_novo_cluster)) + 
  geom_point(size = 4, show.legend = F) +
  geom_label_repel(min.segment.length = 0, show.legend = F, force_pull = .1, box.padding = .5) +
  geom_abline(intercept = 0, slope = 1, linetype = 'dashed', color = 'black') +
  scale_color_manual(values = celltype_color_palette) +
  ylab('Primate Sncg marker enrichment') + xlab('Primate Vip marker enrichment') + ggtitle('Yale cross-primate data') +
  theme_bw() + theme(aspect.ratio = 1)


#Get the mouse enrichments, summarise by the mean enrichment per de novo cluster
ct_scores_mouse = score_cells(log1p(cpm(sesten_gaba_data)), top_mouse_markers)
ct_enrichment_mouse = compute_marker_enrichment(ct_scores_mouse)

colnames(ct_enrichment_mouse) = sesten_gaba_data$seurat_clusters
ct_enrichment_mouse_df = reshape2::melt(t(ct_enrichment_mouse), varnames = c('de_novo_cluster','mouse_markers'), value.name = c('enrichment'))
ct_enrichment_mouse_df = ct_enrichment_mouse_df %>% group_by(de_novo_cluster, mouse_markers) %>% summarise(mean_enrichment = mean(enrichment))

#Get the Sncg and Vip clusters
sestan_mouse_vip_sncg_df = ct_enrichment_mouse_df %>% 
  filter(de_novo_cluster %in% c(16, 20, 4, 9, 10, 12, 17, 19) & mouse_markers %in% c('all|Vip','all|Sncg')) %>%
  tidyr::pivot_wider(names_from = mouse_markers, values_from = mean_enrichment)

sestan_mouse_vip_sncg_df$subclass = rep('Vip', length = nrow(sestan_mouse_vip_sncg_df))
sestan_mouse_vip_sncg_df$subclass[sestan_mouse_vip_sncg_df$de_novo_cluster %in% c(16, 20)] = 'Sncg'

p_yale_mouse = ggplot(sestan_mouse_vip_sncg_df, aes(x = `all|Vip`, y = `all|Sncg`, color = subclass, label = de_novo_cluster)) + 
  geom_point(size = 4, show.legend = F) +
  geom_label_repel(min.segment.length = 0, show.legend = F, force_pull = .1, box.padding = .5) +
  geom_abline(intercept = 0, slope = 1, linetype = 'dashed', color = 'black') +
  scale_color_manual(values = celltype_color_palette) +
  ylab('Mouse Sncg marker enrichment') + xlab('Mouse Vip marker enrichment') + ggtitle('Yale cross-primate data') +
  theme_bw() + theme(aspect.ratio = 1)



p_yale_primate
p_yale_mouse

ggsave(plot = p_yale_primate, filename = 'sestan_vip_sncg_clust_mean_primate_enrich.pdf', path = '../graphs/non_allen_data/', device = 'pdf',
       useDingbats = F, height = 4, width = 4)
ggsave(plot = p_yale_mouse, filename = 'sestan_vip_sncg_clust_mean_mouse_enrich.pdf', path = '../graphs/non_allen_data/', device = 'pdf',
       useDingbats = F, height = 4, width = 4)


```


Plot the mean enrichment +- SD per cluster, group clusters by primate subclass, specifically just Vip, Sncg, and maybe Pax6
```{r}

sestan_clust_to_subclass = rep(NA, length = nrow(ct_enrichment_primate_df))
sestan_clust_to_subclass[ct_enrichment_primate_df$de_novo_cluster %in% c(25)] = 'Sst Chodl'
sestan_clust_to_subclass[ct_enrichment_primate_df$de_novo_cluster %in% c(2, 5, 7, 13, 14, 18, 21)] = 'Sst'
sestan_clust_to_subclass[ct_enrichment_primate_df$de_novo_cluster %in% c(8)] = 'Chandelier'
sestan_clust_to_subclass[ct_enrichment_primate_df$de_novo_cluster %in% c(0,3,6,23)] = 'Pvalb'
sestan_clust_to_subclass[ct_enrichment_primate_df$de_novo_cluster %in% c(15, 1)] = 'Lamp5'
sestan_clust_to_subclass[ct_enrichment_primate_df$de_novo_cluster %in% c(11)] = 'Lamp5 Lhx6'
sestan_clust_to_subclass[ct_enrichment_primate_df$de_novo_cluster %in% c(22, 24)] = 'Pax6'
sestan_clust_to_subclass[ct_enrichment_primate_df$de_novo_cluster %in% c(16, 20)] = 'Sncg'
sestan_clust_to_subclass[ct_enrichment_primate_df$de_novo_cluster %in% c(4, 9, 10, 12, 17, 19)] = 'Vip'
table(is.na(sestan_clust_to_subclass))


ct_enrichment_primate_df$mapped_primate_subclass = sestan_clust_to_subclass
ct_enrichment_primate_df$mapped_primate_subclass = factor(ct_enrichment_primate_df$mapped_primate_subclass, 
                                                          levels = c('Sst','Sst Chodl','Pvalb','Chandelier','Lamp5 Lhx6','Lamp5','Vip','Pax6','Sncg'))


plotting_df = ct_enrichment_primate_df %>% filter(primate_markers == 'all|Vip')

order_vec = plotting_df %>% arrange(match(mapped_primate_subclass, c('Sst','Sst Chodl','Pvalb','Chandelier','Lamp5 Lhx6','Lamp5','Vip','Pax6','Sncg'))) %>%
  pull(de_novo_cluster)

plotting_df$de_novo_cluster = factor(plotting_df$de_novo_cluster, levels = order_vec)

yale_vip_enrich_dot_plot = ggplot(plotting_df, aes(x = de_novo_cluster, y = mean_enrichment, color = mapped_primate_subclass)) +
  geom_point(size = 3, show.legend = F) + 
  geom_linerange(aes(ymin = mean_enrichment - sd_enrichment, ymax = mean_enrichment + sd_enrichment),
                 show.legend = F) +
  geom_hline(yintercept = 1, linetype = 'dashed', color = 'black') +
  ylab('Mean Primate Vip enrichment') + xlab('Yale cross-primate de novo clusters') + ylim(0,3) +
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +
  scale_color_manual(values = celltype_color_palette) +
  theme_bw() + theme(panel.border = element_blank(), axis.line = element_line(),
                     panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank())


plotting_df = ct_enrichment_primate_df %>% filter(primate_markers == 'all|Pax6')

order_vec = plotting_df %>% arrange(match(mapped_primate_subclass, c('Sst','Sst Chodl','Pvalb','Chandelier','Lamp5 Lhx6','Lamp5','Vip','Pax6','Sncg'))) %>%
  pull(de_novo_cluster)

plotting_df$de_novo_cluster = factor(plotting_df$de_novo_cluster, levels = order_vec)

yale_pax6_enrich_dot_plot = ggplot(plotting_df, aes(x = de_novo_cluster, y = mean_enrichment, color = mapped_primate_subclass)) +
  geom_point(size = 3, show.legend = F) + 
  geom_linerange(aes(ymin = mean_enrichment - sd_enrichment, ymax = mean_enrichment + sd_enrichment),
                 show.legend = F) +
  geom_hline(yintercept = 1, linetype = 'dashed', color = 'black') +
  ylab('Mean Primate Pax6 enrichment') + xlab('Yale cross-primate de novo clusters') +  ylim(0,3) +
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +
  scale_color_manual(values = celltype_color_palette) +
  theme_bw() + theme(panel.border = element_blank(), axis.line = element_line(),
                     panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank())


plotting_df = ct_enrichment_primate_df %>% filter(primate_markers == 'all|Sncg')

order_vec = plotting_df %>% arrange(match(mapped_primate_subclass, c('Sst','Sst Chodl','Pvalb','Chandelier','Lamp5 Lhx6','Lamp5','Vip','Pax6','Sncg'))) %>%
  pull(de_novo_cluster)

plotting_df$de_novo_cluster = factor(plotting_df$de_novo_cluster, levels = order_vec)

yale_sncg_enrich_dot_plot = ggplot(plotting_df, aes(x = de_novo_cluster, y = mean_enrichment, color = mapped_primate_subclass)) +
  geom_point(size = 3, show.legend = F) + 
  geom_linerange(aes(ymin = mean_enrichment - sd_enrichment, ymax = mean_enrichment + sd_enrichment),
                 show.legend = F) +
  geom_hline(yintercept = 1, linetype = 'dashed', color = 'black') +
  ylab('Mean Primate Sncg enrichment') + xlab('Yale cross-primate de novo clusters') +  ylim(0,3) +
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +
  scale_color_manual(values = celltype_color_palette) +
  theme_bw() + theme(panel.border = element_blank(), axis.line = element_line(),
                     panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank())
  
yale_sncg_enrich_dot_plot
yale_pax6_enrich_dot_plot
yale_vip_enrich_dot_plot



ggsave(plot = yale_sncg_enrich_dot_plot, filename = 'dot_sestan_sncg_mean_primate_enrich.pdf', path = '../graphs/non_allen_data/', device = 'pdf',
       useDingbats = F, height = 2, width = 4)
ggsave(plot = yale_pax6_enrich_dot_plot, filename = 'dot_sestan_pax6_mean_primate_enrich.pdf', path = '../graphs/non_allen_data/', device = 'pdf',
       useDingbats = F, height = 2, width = 4)
ggsave(plot = yale_vip_enrich_dot_plot, filename = 'dot_sestan_vip_mean_primate_enrich.pdf', path = '../graphs/non_allen_data/', device = 'pdf',
       useDingbats = F, height = 2, width = 4)


#And the mouse Sncg marker enrichment
top_mouse_markers = mouse_subclass_meta_markers %>% filter(rank <= 50)

#Get the primate enrichments, summarise by the mean enrichment per de novo cluster
ct_scores_mouse = score_cells(log1p(cpm(sesten_gaba_data)), top_mouse_markers)
ct_enrichment_mouse = compute_marker_enrichment(ct_scores_mouse)

colnames(ct_enrichment_mouse) = sesten_gaba_data$seurat_clusters
ct_enrichment_mouse_df = reshape2::melt(t(ct_enrichment_mouse), varnames = c('de_novo_cluster','mouse_markers'), value.name = c('enrichment'))
ct_enrichment_mouse_df = ct_enrichment_mouse_df %>% group_by(de_novo_cluster, mouse_markers) %>% 
  summarise(mean_enrichment = mean(enrichment), sd_enrichment = sd(enrichment))

sestan_clust_to_subclass = rep(NA, length = nrow(ct_enrichment_mouse_df))
sestan_clust_to_subclass[ct_enrichment_mouse_df$de_novo_cluster %in% c(25)] = 'Sst Chodl'
sestan_clust_to_subclass[ct_enrichment_mouse_df$de_novo_cluster %in% c(2, 5, 7, 13, 14, 18, 21)] = 'Sst'
sestan_clust_to_subclass[ct_enrichment_mouse_df$de_novo_cluster %in% c(8)] = 'Chandelier'
sestan_clust_to_subclass[ct_enrichment_mouse_df$de_novo_cluster %in% c(0,3,6,23)] = 'Pvalb'
sestan_clust_to_subclass[ct_enrichment_mouse_df$de_novo_cluster %in% c(15, 1)] = 'Lamp5'
sestan_clust_to_subclass[ct_enrichment_mouse_df$de_novo_cluster %in% c(11)] = 'Lamp5 Lhx6'
sestan_clust_to_subclass[ct_enrichment_mouse_df$de_novo_cluster %in% c(22, 24)] = 'Pax6'
sestan_clust_to_subclass[ct_enrichment_mouse_df$de_novo_cluster %in% c(16, 20)] = 'Sncg'
sestan_clust_to_subclass[ct_enrichment_mouse_df$de_novo_cluster %in% c(4, 9, 10, 12, 17, 19)] = 'Vip'
table(is.na(sestan_clust_to_subclass))


ct_enrichment_mouse_df$mapped_primate_subclass = sestan_clust_to_subclass
ct_enrichment_mouse_df$mapped_primate_subclass = factor(ct_enrichment_mouse_df$mapped_primate_subclass, 
                                                          levels = c('Sst','Sst Chodl','Pvalb','Chandelier','Lamp5 Lhx6','Lamp5','Vip','Pax6','Sncg'))

plotting_df = ct_enrichment_mouse_df %>% filter(mouse_markers == 'all|Sncg')

order_vec = plotting_df %>% arrange(match(mapped_primate_subclass, c('Sst','Sst Chodl','Pvalb','Chandelier','Lamp5 Lhx6','Lamp5','Vip','Pax6','Sncg'))) %>%
  pull(de_novo_cluster)

plotting_df$de_novo_cluster = factor(plotting_df$de_novo_cluster, levels = order_vec)

yale_mouse_sncg_enrich_dot_plot = ggplot(plotting_df, aes(x = de_novo_cluster, y = mean_enrichment, color = mapped_primate_subclass)) +
  geom_point(size = 3, show.legend = F) + 
  geom_linerange(aes(ymin = mean_enrichment - sd_enrichment, ymax = mean_enrichment + sd_enrichment),
                 show.legend = F) +
  geom_hline(yintercept = 1, linetype = 'dashed', color = 'black') +
  ylab('Mean Mouse Sncg enrichment') + xlab('Yale cross-primate de novo clusters') +  ylim(0,3) +
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +
  scale_color_manual(values = celltype_color_palette) +
  theme_bw() + theme(panel.border = element_blank(), axis.line = element_line(),
                     panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank())
  
yale_mouse_sncg_enrich_dot_plot
ggsave(plot = yale_mouse_sncg_enrich_dot_plot, filename = 'dot_sestan_sncg_mean_mouse_enrich.pdf', path = '../graphs/non_allen_data/', device = 'pdf',
       useDingbats = F, height = 2, width = 4)
```





```{r}

caglayan_gaba_data = readRDS('/inkwell03/werner/sc_data/primate_caglayan_konopka/caglayan_gaba_processed.rds')
DefaultAssay(caglayan_gaba_data ) = 'RNA'

caglayan_gaba_data = as.SingleCellExperiment(caglayan_gaba_data)
names(assays(caglayan_gaba_data)) = c('counts','cpm')

#get the top primate and mouse MetaMarkers
top_primate_markers = primate_subclass_meta_markers %>% filter(rank <= 50)
top_mouse_markers = mouse_subclass_meta_markers %>% filter(rank <= 50)

#Get the primate enrichments, summarise by the mean enrichment per de novo cluster
ct_scores_primate = score_cells(log1p(cpm(caglayan_gaba_data)), top_primate_markers)
ct_enrichment_primate = compute_marker_enrichment(ct_scores_primate)

colnames(ct_enrichment_primate) = caglayan_gaba_data$seurat_clusters
ct_enrichment_primate_df = reshape2::melt(t(ct_enrichment_primate), varnames = c('de_novo_cluster','primate_markers'), value.name = c('enrichment'))
ct_enrichment_primate_df = ct_enrichment_primate_df %>% group_by(de_novo_cluster, primate_markers) %>% 
  summarise(mean_enrichment = mean(enrichment), sd_enrichment = sd(enrichment))

#Get the Sncg and Vip clusters
caglayan_primate_vip_sncg_df = ct_enrichment_primate_df %>% 
  filter(de_novo_cluster %in% c(19, 5, 7, 11, 13, 17, 18, 20, 26) & primate_markers %in% c('all|Vip','all|Sncg')) %>%
  tidyr::pivot_wider(names_from = primate_markers, values_from = mean_enrichment)

caglayan_primate_vip_sncg_df$subclass = rep('Vip', length = nrow(caglayan_primate_vip_sncg_df))
caglayan_primate_vip_sncg_df$subclass[caglayan_primate_vip_sncg_df$de_novo_cluster %in% c(19)] = 'Sncg'

p_texas_primate = ggplot(caglayan_primate_vip_sncg_df, aes(x = `all|Vip`, y = `all|Sncg`, color = subclass, label = de_novo_cluster)) + 
  geom_point(size = 4, show.legend = F) +
  geom_label_repel(min.segment.length = 0, show.legend = F, force_pull = .1, box.padding = .5) +
  geom_abline(intercept = 0, slope = 1, linetype = 'dashed', color = 'black') +
  scale_color_manual(values = celltype_color_palette) +
  ylab('Primate Sncg marker enrichment') + xlab('Primate Vip marker enrichment') + ggtitle('UT Southwestern cross-primate data') +
  theme_bw() + theme(aspect.ratio = 1)


#Get the mouse enrichments, summarise by the mean enrichment per de novo cluster
ct_scores_mouse = score_cells(log1p(cpm(caglayan_gaba_data)), top_mouse_markers)
ct_enrichment_mouse = compute_marker_enrichment(ct_scores_mouse)

colnames(ct_enrichment_mouse) = caglayan_gaba_data$seurat_clusters
ct_enrichment_mouse_df = reshape2::melt(t(ct_enrichment_mouse), varnames = c('de_novo_cluster','mouse_markers'), value.name = c('enrichment'))
ct_enrichment_mouse_df = ct_enrichment_mouse_df %>% group_by(de_novo_cluster, mouse_markers) %>% summarise(mean_enrichment = mean(enrichment))

#Get the Sncg and Vip clusters
caglayan_mouse_vip_sncg_df = ct_enrichment_mouse_df %>% 
  filter(de_novo_cluster %in% c(19, 5, 7, 11, 13, 17, 18, 20, 26) & mouse_markers %in% c('all|Vip','all|Sncg')) %>%
  tidyr::pivot_wider(names_from = mouse_markers, values_from = mean_enrichment)

caglayan_mouse_vip_sncg_df$subclass = rep('Vip', length = nrow(caglayan_mouse_vip_sncg_df))
caglayan_mouse_vip_sncg_df$subclass[caglayan_mouse_vip_sncg_df$de_novo_cluster %in% c(19)] = 'Sncg'

p_texas_mouse = ggplot(caglayan_mouse_vip_sncg_df, aes(x = `all|Vip`, y = `all|Sncg`, color = subclass, label = de_novo_cluster)) + 
  geom_point(size = 4, show.legend = F) +
  geom_label_repel(min.segment.length = 0, show.legend = F, force_pull = .1, box.padding = .5) +
  geom_abline(intercept = 0, slope = 1, linetype = 'dashed', color = 'black') +
  scale_color_manual(values = celltype_color_palette) +
  ylab('Mouse Sncg marker enrichment') + xlab('Mouse Vip marker enrichment') + ggtitle('UT Southwestern cross-primate data') +
  theme_bw() + theme(aspect.ratio = 1)



p_texas_primate
p_texas_mouse

ggsave(plot = p_texas_primate, filename = 'caglayan_vip_sncg_clust_mean_primate_enrich.pdf', path = '../graphs/non_allen_data/', device = 'pdf',
       useDingbats = F, height = 4, width = 4)
ggsave(plot = p_texas_mouse, filename = 'caglayan_vip_sncg_clust_mean_mouse_enrich.pdf', path = '../graphs/non_allen_data/', device = 'pdf',
       useDingbats = F, height = 4, width = 4)


```


```{r}

caglayan_clust_to_subclass = rep(NA, length = nrow(ct_enrichment_primate_df))
caglayan_clust_to_subclass[ct_enrichment_primate_df$de_novo_cluster %in% c(28)] = 'Sst Chodl'
caglayan_clust_to_subclass[ct_enrichment_primate_df$de_novo_cluster %in% c(1,4,6,8,10,21,22)] = 'Sst'
caglayan_clust_to_subclass[ct_enrichment_primate_df$de_novo_cluster %in% c(12)] = 'Chandelier'
caglayan_clust_to_subclass[ct_enrichment_primate_df$de_novo_cluster %in% c(0,3,15,16,23)] = 'Pvalb'
caglayan_clust_to_subclass[ct_enrichment_primate_df$de_novo_cluster %in% c(2,14,27)] = 'Lamp5'
caglayan_clust_to_subclass[ct_enrichment_primate_df$de_novo_cluster %in% c(9)] = 'Lamp5 Lhx6'
caglayan_clust_to_subclass[ct_enrichment_primate_df$de_novo_cluster %in% c(24,25)] = 'Pax6'
caglayan_clust_to_subclass[ct_enrichment_primate_df$de_novo_cluster %in% c(19)] = 'Sncg'
caglayan_clust_to_subclass[ct_enrichment_primate_df$de_novo_cluster %in% c(5,7,11,13,17,18,20,26)] = 'Vip'
table(is.na(caglayan_clust_to_subclass))


ct_enrichment_primate_df$mapped_primate_subclass = caglayan_clust_to_subclass
ct_enrichment_primate_df$mapped_primate_subclass = factor(ct_enrichment_primate_df$mapped_primate_subclass, 
                                                          levels = c('Sst','Sst Chodl','Pvalb','Chandelier','Lamp5 Lhx6','Lamp5','Vip','Pax6','Sncg'))


plotting_df = ct_enrichment_primate_df %>% filter(primate_markers == 'all|Vip')

order_vec = plotting_df %>% arrange(match(mapped_primate_subclass, c('Sst','Sst Chodl','Pvalb','Chandelier','Lamp5 Lhx6','Lamp5','Vip','Pax6','Sncg'))) %>%
  pull(de_novo_cluster)

plotting_df$de_novo_cluster = factor(plotting_df$de_novo_cluster, levels = order_vec)

texas_vip_enrich_dot_plot = ggplot(plotting_df, aes(x = de_novo_cluster, y = mean_enrichment, color = mapped_primate_subclass)) +
  geom_point(size = 3, show.legend = F) + 
  geom_linerange(aes(ymin = mean_enrichment - sd_enrichment, ymax = mean_enrichment + sd_enrichment),
                 show.legend = F) +
  geom_hline(yintercept = 1, linetype = 'dashed', color = 'black') +
  ylab('Mean Primate Vip enrichment') + xlab('UT Southwestern cross-primate de novo clusters') + ylim(0,3) +
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +
  scale_color_manual(values = celltype_color_palette) +
  theme_bw() + theme(panel.border = element_blank(), axis.line = element_line(),
                     panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank())


plotting_df = ct_enrichment_primate_df %>% filter(primate_markers == 'all|Pax6')

order_vec = plotting_df %>% arrange(match(mapped_primate_subclass, c('Sst','Sst Chodl','Pvalb','Chandelier','Lamp5 Lhx6','Lamp5','Vip','Pax6','Sncg'))) %>%
  pull(de_novo_cluster)

plotting_df$de_novo_cluster = factor(plotting_df$de_novo_cluster, levels = order_vec)

texas_pax6_enrich_dot_plot = ggplot(plotting_df, aes(x = de_novo_cluster, y = mean_enrichment, color = mapped_primate_subclass)) +
  geom_point(size = 3, show.legend = F) + 
  geom_linerange(aes(ymin = mean_enrichment - sd_enrichment, ymax = mean_enrichment + sd_enrichment),
                 show.legend = F) +
  geom_hline(yintercept = 1, linetype = 'dashed', color = 'black') +
  ylab('Mean Primate Pax6 enrichment') + xlab('UT Southwestern cross-primate de novo clusters') +  ylim(0,3) +
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +
  scale_color_manual(values = celltype_color_palette) +
  theme_bw() + theme(panel.border = element_blank(), axis.line = element_line(),
                     panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank())


plotting_df = ct_enrichment_primate_df %>% filter(primate_markers == 'all|Sncg')

order_vec = plotting_df %>% arrange(match(mapped_primate_subclass, c('Sst','Sst Chodl','Pvalb','Chandelier','Lamp5 Lhx6','Lamp5','Vip','Pax6','Sncg'))) %>%
  pull(de_novo_cluster)

plotting_df$de_novo_cluster = factor(plotting_df$de_novo_cluster, levels = order_vec)

texas_sncg_enrich_dot_plot = ggplot(plotting_df, aes(x = de_novo_cluster, y = mean_enrichment, color = mapped_primate_subclass)) +
  geom_point(size = 3, show.legend = F) + 
  geom_linerange(aes(ymin = mean_enrichment - sd_enrichment, ymax = mean_enrichment + sd_enrichment),
                 show.legend = F) +
  geom_hline(yintercept = 1, linetype = 'dashed', color = 'black') +
  ylab('Mean Primate Sncg enrichment') + xlab('UT Southwester cross-primate de novo clusters') +  ylim(0,3) +
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +
  scale_color_manual(values = celltype_color_palette) +
  theme_bw() + theme(panel.border = element_blank(), axis.line = element_line(),
                     panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank())
  
texas_sncg_enrich_dot_plot
texas_pax6_enrich_dot_plot
texas_vip_enrich_dot_plot



ggsave(plot = texas_sncg_enrich_dot_plot, filename = 'dot_caglayan_sncg_mean_primate_enrich.pdf', path = '../graphs/non_allen_data/', device = 'pdf',
       useDingbats = F, height = 2, width = 4)
ggsave(plot = texas_pax6_enrich_dot_plot, filename = 'dot_caglayan_pax6_mean_primate_enrich.pdf', path = '../graphs/non_allen_data/', device = 'pdf',
       useDingbats = F, height = 2, width = 4)
ggsave(plot = texas_vip_enrich_dot_plot, filename = 'dot_caglayan_vip_mean_primate_enrich.pdf', path = '../graphs/non_allen_data/', device = 'pdf',
       useDingbats = F, height = 2, width = 4)


#And the mouse Sncg marker enrichment
top_mouse_markers = mouse_subclass_meta_markers %>% filter(rank <= 50)

#Get the primate enrichments, summarise by the mean enrichment per de novo cluster
ct_scores_mouse = score_cells(log1p(cpm(caglayan_gaba_data)), top_mouse_markers)
ct_enrichment_mouse = compute_marker_enrichment(ct_scores_mouse)

colnames(ct_enrichment_mouse) = caglayan_gaba_data$seurat_clusters
ct_enrichment_mouse_df = reshape2::melt(t(ct_enrichment_mouse), varnames = c('de_novo_cluster','mouse_markers'), value.name = c('enrichment'))
ct_enrichment_mouse_df = ct_enrichment_mouse_df %>% group_by(de_novo_cluster, mouse_markers) %>% 
  summarise(mean_enrichment = mean(enrichment), sd_enrichment = sd(enrichment))

caglayan_clust_to_subclass = rep(NA, length = nrow(ct_enrichment_mouse_df))
caglayan_clust_to_subclass[ct_enrichment_mouse_df$de_novo_cluster %in% c(28)] = 'Sst Chodl'
caglayan_clust_to_subclass[ct_enrichment_mouse_df$de_novo_cluster %in% c(1,4,6,8,10,21,22)] = 'Sst'
caglayan_clust_to_subclass[ct_enrichment_mouse_df$de_novo_cluster %in% c(12)] = 'Chandelier'
caglayan_clust_to_subclass[ct_enrichment_mouse_df$de_novo_cluster %in% c(0,3,15,16,23)] = 'Pvalb'
caglayan_clust_to_subclass[ct_enrichment_mouse_df$de_novo_cluster %in% c(2,14,27)] = 'Lamp5'
caglayan_clust_to_subclass[ct_enrichment_mouse_df$de_novo_cluster %in% c(9)] = 'Lamp5 Lhx6'
caglayan_clust_to_subclass[ct_enrichment_mouse_df$de_novo_cluster %in% c(24,25)] = 'Pax6'
caglayan_clust_to_subclass[ct_enrichment_mouse_df$de_novo_cluster %in% c(19)] = 'Sncg'
caglayan_clust_to_subclass[ct_enrichment_mouse_df$de_novo_cluster %in% c(5,7,11,13,17,18,20,26)] = 'Vip'
table(is.na(caglayan_clust_to_subclass))


ct_enrichment_mouse_df$mapped_primate_subclass = caglayan_clust_to_subclass
ct_enrichment_mouse_df$mapped_primate_subclass = factor(ct_enrichment_mouse_df$mapped_primate_subclass, 
                                                          levels = c('Sst','Sst Chodl','Pvalb','Chandelier','Lamp5 Lhx6','Lamp5','Vip','Pax6','Sncg'))

plotting_df = ct_enrichment_mouse_df %>% filter(mouse_markers == 'all|Sncg')

order_vec = plotting_df %>% arrange(match(mapped_primate_subclass, c('Sst','Sst Chodl','Pvalb','Chandelier','Lamp5 Lhx6','Lamp5','Vip','Pax6','Sncg'))) %>%
  pull(de_novo_cluster)

plotting_df$de_novo_cluster = factor(plotting_df$de_novo_cluster, levels = order_vec)

texas_mouse_sncg_enrich_dot_plot = ggplot(plotting_df, aes(x = de_novo_cluster, y = mean_enrichment, color = mapped_primate_subclass)) +
  geom_point(size = 3, show.legend = F) + 
  geom_linerange(aes(ymin = mean_enrichment - sd_enrichment, ymax = mean_enrichment + sd_enrichment),
                 show.legend = F) +
  geom_hline(yintercept = 1, linetype = 'dashed', color = 'black') +
  ylab('Mean Mouse Sncg enrichment') + xlab('UT SouthWestern cross-primate de novo clusters') +  ylim(0,3) +
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +
  scale_color_manual(values = celltype_color_palette) +
  theme_bw() + theme(panel.border = element_blank(), axis.line = element_line(),
                     panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank())
  
texas_mouse_sncg_enrich_dot_plot
ggsave(plot = texas_mouse_sncg_enrich_dot_plot, filename = 'dot_caglayan_sncg_mean_mouse_enrich.pdf', path = '../graphs/non_allen_data/', device = 'pdf',
       useDingbats = F, height = 2, width = 4)




```








################################################
Now in the non_allen primate datasets, look at the correlations in DE between the primate clusters and the mouse subclasses

But split by species. Is there variability across the species? 

If Mouse Sncg is the ancestral Primate Sncg and there remains an Ancestral Sncg in the form of a Vip cell-type in primates,
is there also phylogenetic variability there as well across the primates?

You could imagine the marmoset Ancestral Sncg(Vip) is closer to the Mouse Sncg than Human

################################################

First I need the marker genes for the de novo clusters

```{r}


human_index = sesten_gaba_data$species == 'Human'
sestan_human = sesten_gaba_data[ , human_index]

chimp_index = sesten_gaba_data$species == 'Chimpanzee'
sestan_chimp = sesten_gaba_data[ , chimp_index]

macaque_index = sesten_gaba_data$species == 'Rhesus'
sestan_macaque = sesten_gaba_data[ , macaque_index]

marmoset_index = sesten_gaba_data$species == 'Marmoset'
sestan_marmoset = sesten_gaba_data[ , marmoset_index]


sestan_human_cluster_markers = compute_markers(cpm(sestan_human),sestan_human$seurat_clusters )
sestan_chimp_cluster_markers = compute_markers(cpm(sestan_chimp),sestan_chimp$seurat_clusters )
sestan_macaque_cluster_markers = compute_markers(cpm(sestan_macaque),sestan_macaque$seurat_clusters )
sestan_marmoset_cluster_markers = compute_markers(cpm(sestan_marmoset),sestan_marmoset$seurat_clusters )

sestan_human_cluster_markers
sestan_chimp_cluster_markers
sestan_macaque_cluster_markers
sestan_marmoset_cluster_markers


export_markers(sestan_human_cluster_markers, 'markers_to_share/sestan_human_de_novo_cluster_markers.csv')
export_markers(sestan_chimp_cluster_markers, 'markers_to_share/sestan_chimp_de_novo_cluster_markers.csv')
export_markers(sestan_macaque_cluster_markers, 'markers_to_share/sestan_macaque_de_novo_cluster_markers.csv')
export_markers(sestan_marmoset_cluster_markers, 'markers_to_share/sestan_marmoset_de_novo_cluster_markers.csv')

```



```{r}

human_index = caglayan_inhib_copy$Species == 'human'
caglayan_human = caglayan_inhib_copy[ , human_index]

chimp_index = caglayan_inhib_copy$Species == 'chimp'
caglayan_chimp = caglayan_inhib_copy[ , chimp_index]

macaque_index = caglayan_inhib_copy$Species == 'macaque'
caglayan_macaque = caglayan_inhib_copy[ , macaque_index]


caglayan_human_cluster_markers = compute_markers(cpm(caglayan_human),caglayan_human$seurat_clusters )
caglayan_chimp_cluster_markers = compute_markers(cpm(caglayan_chimp),caglayan_chimp$seurat_clusters )
caglayan_macaque_cluster_markers = compute_markers(cpm(caglayan_macaque),caglayan_macaque$seurat_clusters )

caglayan_human_cluster_markers
caglayan_chimp_cluster_markers
caglayan_macaque_cluster_markers


export_markers(caglayan_human_cluster_markers, 'markers_to_share/caglayan_human_de_novo_cluster_markers.csv')
export_markers(caglayan_chimp_cluster_markers, 'markers_to_share/caglayan_chimp_de_novo_cluster_markers.csv')
export_markers(caglayan_macaque_cluster_markers, 'markers_to_share/caglayan_macaque_de_novo_cluster_markers.csv')

```






Compare DE correlations to the mouse Subclass markers
Really just want the focus on the primate 17 cluster
Primate Vip clusters: 4, 9, 10, 12, 17, 19
Primate Sncg clusters: 16, 20
Primate Pax6 clusters: 22, 24

```{r}

intersect_genes = intersect(sestan_human_cluster_markers$gene, mouse_subclass_meta_markers$gene)
length(intersect_genes)

filt_mouse_subclass_markers = mouse_subclass_meta_markers %>% filter(gene %in% intersect_genes)
filt_primate_subclass_markers = primate_subclass_meta_markers %>% filter(gene %in% intersect_genes)


filt_sestan_human_markers = sestan_human_cluster_markers %>% filter(gene %in% intersect_genes)
filt_sestan_chimp_markers = sestan_chimp_cluster_markers %>% filter(gene %in% intersect_genes)
filt_sestan_macaque_markers = sestan_macaque_cluster_markers %>% filter(gene %in% intersect_genes)
filt_sestan_marmoset_markers = sestan_marmoset_cluster_markers %>% filter(gene %in% intersect_genes)




```

```{r}

all_markers = seq(10, 500, 10)

all_corr_summary_dfs = list()

for(i in 1:length(all_markers)){

  num_top_markers = all_markers[i]
  
  num_top_markers = 100
  mouse_subclass = 'Sncg'
  de_metric = 'auroc'
  primate_clusters =  c('17', '16', '20', '4', '9', '10', '12', '19')
  
  
  test_mouse_subclass_markers = filt_mouse_subclass_markers %>% filter(cell_type == mouse_subclass & rank <= num_top_markers) %>% pull(gene)
  
  mouse_subclass_de_column = filt_mouse_subclass_markers %>% filter(cell_type == mouse_subclass & rank <= num_top_markers) %>% 
    select(cell_type, gene, !! sym(de_metric)) %>%
    tidyr::pivot_wider(names_from = cell_type, values_from = !! sym(de_metric))
  mouse_subclass_de_column = tibble::column_to_rownames(mouse_subclass_de_column, var = 'gene')
  colnames(mouse_subclass_de_column) = paste0('Mouse_', colnames(mouse_subclass_de_column))
  
  #Human cluster
  sestan_human_cluster_mat = filt_sestan_human_markers %>% filter(cell_type %in% primate_clusters & 
                                                                    gene %in% test_mouse_subclass_markers ) %>%
    select(cell_type, gene, !! sym(de_metric)) %>% tidyr::pivot_wider(names_from = cell_type, values_from = !! sym(de_metric))
  sestan_human_cluster_mat  = tibble::column_to_rownames(sestan_human_cluster_mat , var = 'gene')
  colnames(sestan_human_cluster_mat) = paste0('human_', colnames(sestan_human_cluster_mat) )
  gene_index = match(rownames(mouse_subclass_de_column), rownames(sestan_human_cluster_mat))
  sestan_human_cluster_mat = sestan_human_cluster_mat[gene_index, , drop = F]
  #chimp cluster
  sestan_chimp_cluster_mat = filt_sestan_chimp_markers %>% filter(cell_type %in% primate_clusters & 
                                                                    gene %in% test_mouse_subclass_markers ) %>%
    select(cell_type, gene, !! sym(de_metric)) %>% tidyr::pivot_wider(names_from = cell_type, values_from = !! sym(de_metric))
  sestan_chimp_cluster_mat  = tibble::column_to_rownames(sestan_chimp_cluster_mat , var = 'gene')
  colnames(sestan_chimp_cluster_mat) = paste0('chimp_', colnames(sestan_chimp_cluster_mat) )
  gene_index = match(rownames(mouse_subclass_de_column), rownames(sestan_chimp_cluster_mat))
  sestan_chimp_cluster_mat = sestan_chimp_cluster_mat[gene_index, , drop = F]
  #macaque cluster
  sestan_macaque_cluster_mat = filt_sestan_macaque_markers %>% filter(cell_type %in% primate_clusters & 
                                                                    gene %in% test_mouse_subclass_markers ) %>%
    select(cell_type, gene, !! sym(de_metric)) %>% tidyr::pivot_wider(names_from = cell_type, values_from = !! sym(de_metric))
  sestan_macaque_cluster_mat  = tibble::column_to_rownames(sestan_macaque_cluster_mat , var = 'gene')
  colnames(sestan_macaque_cluster_mat) = paste0('macaque_', colnames(sestan_macaque_cluster_mat) )
  gene_index = match(rownames(mouse_subclass_de_column), rownames(sestan_macaque_cluster_mat))
  sestan_macaque_cluster_mat = sestan_macaque_cluster_mat[gene_index, , drop = F]
  #marmoset cluster
  sestan_marmoset_cluster_mat = filt_sestan_marmoset_markers %>% filter(cell_type %in% primate_clusters & 
                                                                    gene %in% test_mouse_subclass_markers ) %>%
    select(cell_type, gene, !! sym(de_metric)) %>% tidyr::pivot_wider(names_from = cell_type, values_from = !! sym(de_metric))
  sestan_marmoset_cluster_mat  = tibble::column_to_rownames(sestan_marmoset_cluster_mat , var = 'gene')
  colnames(sestan_marmoset_cluster_mat) = paste0('marmoset_', colnames(sestan_marmoset_cluster_mat) )
  gene_index = match(rownames(mouse_subclass_de_column), rownames(sestan_marmoset_cluster_mat))
  sestan_marmoset_cluster_mat = sestan_marmoset_cluster_mat[gene_index, , drop = F]
  
  
  all_species_de = do.call('cbind', list(sestan_human_cluster_mat, sestan_chimp_cluster_mat, 
                                             sestan_macaque_cluster_mat, sestan_marmoset_cluster_mat, mouse_subclass_de_column ))
  
  cor_mat = cor(all_species_de, method = 'spearman')
  
  ComplexHeatmap::Heatmap(cor_mat, name = 'spearman',
                         clustering_distance_rows = function(m) dist(1-m), clustering_distance_columns = function(m) dist(1-m),
                         clustering_method_rows = 'ward.D2', clustering_method_columns = 'ward.D2',
                         column_title = 'Similarity in Mouse Sncg DE')
  
  
  long_cor_df = reshape2::melt(cor_mat, value.name = 'spearman', varnames = c('cluster_1','cluster_2'))
  long_cor_df$cluster_name = sapply(strsplit(as.character(long_cor_df$cluster_1), split = '_', fixed = T ), '[[', 2 )
  
  corr_summary_df = long_cor_df %>% filter(cluster_2 == sprintf('Mouse_%s', mouse_subclass)) %>% 
    group_by(cluster_name) %>% summarise(mean_spearman = mean(spearman)) %>%
    mutate(num_markers = num_top_markers)
  
  all_corr_summary_dfs[[i]] = corr_summary_df 
  
}


all_corr_summary_dfs = do.call('rbind', all_corr_summary_dfs)

primate_cluster_colors = c('red','#DF70FF','#DF70FF','#A45FBF','#A45FBF','#A45FBF','#A45FBF','#A45FBF')
names(primate_cluster_colors ) = primate_clusters
ggplot(all_corr_summary_dfs , aes(x = num_markers, y = mean_spearman, color = cluster_name)) + geom_line() +
  ggtitle(sprintf('Similarity in DE to Mouse %s', mouse_subclass)) + 
  ylab('Correlation in marker DE') +
  xlab('Number of top markers') +
  scale_color_manual(values = primate_cluster_colors)

```



```{r}

all_markers = seq(10, 500, 10)

all_corr_summary_dfs = list()

for(i in 1:length(all_markers)){

  num_top_markers = all_markers[i]
  
  mouse_subclass = 'Vip'
  de_metric = 'auroc'
  primate_clusters =  c('9')
  
  
  test_mouse_subclass_markers = filt_mouse_subclass_markers %>% filter(cell_type == mouse_subclass & rank <= num_top_markers) %>% pull(gene)
  
  mouse_subclass_de_column = filt_mouse_subclass_markers %>% filter(cell_type == mouse_subclass & rank <= num_top_markers) %>% 
    select(cell_type, gene, !! sym(de_metric)) %>%
    tidyr::pivot_wider(names_from = cell_type, values_from = !! sym(de_metric))
  mouse_subclass_de_column = tibble::column_to_rownames(mouse_subclass_de_column, var = 'gene')
  colnames(mouse_subclass_de_column) = paste0('Mouse_', colnames(mouse_subclass_de_column))
  
  #Human cluster
  sestan_human_cluster_mat = filt_sestan_human_markers %>% filter(cell_type %in% primate_clusters & 
                                                                    gene %in% test_mouse_subclass_markers ) %>%
    select(cell_type, gene, !! sym(de_metric)) %>% tidyr::pivot_wider(names_from = cell_type, values_from = !! sym(de_metric))
  sestan_human_cluster_mat  = tibble::column_to_rownames(sestan_human_cluster_mat , var = 'gene')
  colnames(sestan_human_cluster_mat) = paste0('human_', colnames(sestan_human_cluster_mat) )
  gene_index = match(rownames(mouse_subclass_de_column), rownames(sestan_human_cluster_mat))
  sestan_human_cluster_mat = sestan_human_cluster_mat[gene_index, , drop = F]
  #chimp cluster
  sestan_chimp_cluster_mat = filt_sestan_chimp_markers %>% filter(cell_type %in% primate_clusters & 
                                                                    gene %in% test_mouse_subclass_markers ) %>%
    select(cell_type, gene, !! sym(de_metric)) %>% tidyr::pivot_wider(names_from = cell_type, values_from = !! sym(de_metric))
  sestan_chimp_cluster_mat  = tibble::column_to_rownames(sestan_chimp_cluster_mat , var = 'gene')
  colnames(sestan_chimp_cluster_mat) = paste0('chimp_', colnames(sestan_chimp_cluster_mat) )
  gene_index = match(rownames(mouse_subclass_de_column), rownames(sestan_chimp_cluster_mat))
  sestan_chimp_cluster_mat = sestan_chimp_cluster_mat[gene_index, , drop = F]
  #macaque cluster
  sestan_macaque_cluster_mat = filt_sestan_macaque_markers %>% filter(cell_type %in% primate_clusters & 
                                                                    gene %in% test_mouse_subclass_markers ) %>%
    select(cell_type, gene, !! sym(de_metric)) %>% tidyr::pivot_wider(names_from = cell_type, values_from = !! sym(de_metric))
  sestan_macaque_cluster_mat  = tibble::column_to_rownames(sestan_macaque_cluster_mat , var = 'gene')
  colnames(sestan_macaque_cluster_mat) = paste0('macaque_', colnames(sestan_macaque_cluster_mat) )
  gene_index = match(rownames(mouse_subclass_de_column), rownames(sestan_macaque_cluster_mat))
  sestan_macaque_cluster_mat = sestan_macaque_cluster_mat[gene_index, , drop = F]
  #marmoset cluster
  sestan_marmoset_cluster_mat = filt_sestan_marmoset_markers %>% filter(cell_type %in% primate_clusters & 
                                                                    gene %in% test_mouse_subclass_markers ) %>%
    select(cell_type, gene, !! sym(de_metric)) %>% tidyr::pivot_wider(names_from = cell_type, values_from = !! sym(de_metric))
  sestan_marmoset_cluster_mat  = tibble::column_to_rownames(sestan_marmoset_cluster_mat , var = 'gene')
  colnames(sestan_marmoset_cluster_mat) = paste0('marmoset_', colnames(sestan_marmoset_cluster_mat) )
  gene_index = match(rownames(mouse_subclass_de_column), rownames(sestan_marmoset_cluster_mat))
  sestan_marmoset_cluster_mat = sestan_marmoset_cluster_mat[gene_index, , drop = F]
  
  
  all_species_de = do.call('cbind', list(sestan_human_cluster_mat, sestan_chimp_cluster_mat, 
                                             sestan_macaque_cluster_mat, sestan_marmoset_cluster_mat, mouse_subclass_de_column ))
  
  cor_mat = cor(all_species_de, method = 'spearman')
  
  #ComplexHeatmap::Heatmap(cor_mat, name = 'spearman',
  #                       clustering_distance_rows = function(m) dist(1-m), clustering_distance_columns = function(m) dist(1-m),
  #                       clustering_method_rows = 'average', clustering_method_columns = 'average')
  
  
  long_cor_df = reshape2::melt(cor_mat, value.name = 'spearman', varnames = c('cluster_1','cluster_2'))
  #long_cor_df$cluster_name = sapply(strsplit(as.character(long_cor_df$cluster_1), split = '_', fixed = T ), '[[', 2 )
  
  corr_summary_df = long_cor_df %>% filter(cluster_2 == sprintf('Mouse_%s', mouse_subclass)) %>% 
    mutate(num_markers = num_top_markers)
  
  all_corr_summary_dfs[[i]] = corr_summary_df 
  
}


all_corr_summary_dfs = do.call('rbind', all_corr_summary_dfs)

primate_17_species_colors = c('darkblue',  'lightblue',  'purple', 'orange')
names(primate_17_species_colors) = paste0(c('human_','chimp_','macaque_','marmoset_'), primate_clusters)
ggplot(all_corr_summary_dfs , aes(x = num_markers, y = spearman, color = cluster_1)) + geom_line() +
  ggtitle(sprintf('Similarity in DE to Mouse %s', mouse_subclass)) + 
  ylab('Correlation in marker DE') +
  xlab('Number of top markers') +
  scale_color_manual(values = primate_17_species_colors)

```


And the primate markers

```{r}
all_markers = seq(10, 500, 10)

all_corr_summary_dfs = list()

for(i in 1:length(all_markers)){

  num_top_markers = all_markers[i]
  
  num_top_markers = 100
  primate_subclass = 'Sncg'
  de_metric = 'auroc'
  primate_clusters =  c('17', '16', '20', '4', '9', '10', '12', '19')
  
  
  test_primate_subclass_markers = filt_primate_subclass_markers %>% filter(cell_type == primate_subclass & rank <= num_top_markers) %>% pull(gene)
  
  primate_subclass_de_column = filt_primate_subclass_markers %>% filter(cell_type == primate_subclass & rank <= num_top_markers) %>% 
    select(cell_type, gene, !! sym(de_metric)) %>%
    tidyr::pivot_wider(names_from = cell_type, values_from = !! sym(de_metric))
  primate_subclass_de_column = tibble::column_to_rownames(primate_subclass_de_column, var = 'gene')
  colnames(primate_subclass_de_column) = paste0('Primate_', colnames(primate_subclass_de_column))
  
  #Human cluster
  sestan_human_cluster_mat = filt_sestan_human_markers %>% filter(cell_type %in% primate_clusters & 
                                                                    gene %in% test_primate_subclass_markers ) %>%
    select(cell_type, gene, !! sym(de_metric)) %>% tidyr::pivot_wider(names_from = cell_type, values_from = !! sym(de_metric))
  sestan_human_cluster_mat  = tibble::column_to_rownames(sestan_human_cluster_mat , var = 'gene')
  colnames(sestan_human_cluster_mat) = paste0('human_', colnames(sestan_human_cluster_mat) )
  gene_index = match(rownames(primate_subclass_de_column), rownames(sestan_human_cluster_mat))
  sestan_human_cluster_mat = sestan_human_cluster_mat[gene_index, , drop = F]
  #chimp cluster
  sestan_chimp_cluster_mat = filt_sestan_chimp_markers %>% filter(cell_type %in% primate_clusters & 
                                                                    gene %in% test_primate_subclass_markers ) %>%
    select(cell_type, gene, !! sym(de_metric)) %>% tidyr::pivot_wider(names_from = cell_type, values_from = !! sym(de_metric))
  sestan_chimp_cluster_mat  = tibble::column_to_rownames(sestan_chimp_cluster_mat , var = 'gene')
  colnames(sestan_chimp_cluster_mat) = paste0('chimp_', colnames(sestan_chimp_cluster_mat) )
  gene_index = match(rownames(primate_subclass_de_column), rownames(sestan_chimp_cluster_mat))
  sestan_chimp_cluster_mat = sestan_chimp_cluster_mat[gene_index, , drop = F]
  #macaque cluster
  sestan_macaque_cluster_mat = filt_sestan_macaque_markers %>% filter(cell_type %in% primate_clusters & 
                                                                    gene %in% test_primate_subclass_markers ) %>%
    select(cell_type, gene, !! sym(de_metric)) %>% tidyr::pivot_wider(names_from = cell_type, values_from = !! sym(de_metric))
  sestan_macaque_cluster_mat  = tibble::column_to_rownames(sestan_macaque_cluster_mat , var = 'gene')
  colnames(sestan_macaque_cluster_mat) = paste0('macaque_', colnames(sestan_macaque_cluster_mat) )
  gene_index = match(rownames(primate_subclass_de_column), rownames(sestan_macaque_cluster_mat))
  sestan_macaque_cluster_mat = sestan_macaque_cluster_mat[gene_index, , drop = F]
  #marmoset cluster
  sestan_marmoset_cluster_mat = filt_sestan_marmoset_markers %>% filter(cell_type %in% primate_clusters & 
                                                                    gene %in% test_primate_subclass_markers ) %>%
    select(cell_type, gene, !! sym(de_metric)) %>% tidyr::pivot_wider(names_from = cell_type, values_from = !! sym(de_metric))
  sestan_marmoset_cluster_mat  = tibble::column_to_rownames(sestan_marmoset_cluster_mat , var = 'gene')
  colnames(sestan_marmoset_cluster_mat) = paste0('marmoset_', colnames(sestan_marmoset_cluster_mat) )
  gene_index = match(rownames(primate_subclass_de_column), rownames(sestan_marmoset_cluster_mat))
  sestan_marmoset_cluster_mat = sestan_marmoset_cluster_mat[gene_index, , drop = F]
  
  
  all_species_de = do.call('cbind', list(sestan_human_cluster_mat, sestan_chimp_cluster_mat, 
                                             sestan_macaque_cluster_mat, sestan_marmoset_cluster_mat, primate_subclass_de_column ))
  
  cor_mat = cor(all_species_de, method = 'spearman')
  
  ComplexHeatmap::Heatmap(cor_mat, name = 'spearman',
                         clustering_distance_rows = function(m) dist(1-m), clustering_distance_columns = function(m) dist(1-m),
                         clustering_method_rows = 'average', clustering_method_columns = 'average',
                         column_title = 'Similarity in Primate Sncg DE')

  long_cor_df = reshape2::melt(cor_mat, value.name = 'spearman', varnames = c('cluster_1','cluster_2'))
  long_cor_df$cluster_name = sapply(strsplit(as.character(long_cor_df$cluster_1), split = '_', fixed = T ), '[[', 2 )
  
  corr_summary_df = long_cor_df %>% filter(cluster_2 == sprintf('Primate_%s', primate_subclass)) %>% 
    group_by(cluster_name) %>% summarise(mean_spearman = mean(spearman)) %>%
    mutate(num_markers = num_top_markers)
  
  all_corr_summary_dfs[[i]] = corr_summary_df 
  
}

all_corr_summary_dfs = do.call('rbind', all_corr_summary_dfs)

primate_cluster_colors = c('red','#DF70FF','#DF70FF','#A45FBF','#A45FBF','#A45FBF','#A45FBF','#A45FBF')
names(primate_cluster_colors ) = primate_clusters
ggplot(all_corr_summary_dfs , aes(x = num_markers, y = mean_spearman, color = cluster_name)) + geom_line() +
  ggtitle(sprintf('Similarity in DE to Primate %s', primate_subclass)) + 
  ylab('Correlation in marker DE') +
  xlab('Number of top markers') +
  scale_color_manual(values = primate_cluster_colors)


```





And the primate markers

```{r}
all_markers = seq(10, 500, 10)

all_corr_summary_dfs = list()

for(i in 1:length(all_markers)){

  num_top_markers = all_markers[i]
  
  primate_subclass = 'Sncg'
  de_metric = 'auroc'
  primate_clusters =  c('16')
  
  
  test_primate_subclass_markers = filt_primate_subclass_markers %>% filter(cell_type == primate_subclass & rank <= num_top_markers) %>% pull(gene)
  
  primate_subclass_de_column = filt_primate_subclass_markers %>% filter(cell_type == primate_subclass & rank <= num_top_markers) %>% 
    select(cell_type, gene, !! sym(de_metric)) %>%
    tidyr::pivot_wider(names_from = cell_type, values_from = !! sym(de_metric))
  primate_subclass_de_column = tibble::column_to_rownames(primate_subclass_de_column, var = 'gene')
  colnames(primate_subclass_de_column) = paste0('Primate_', colnames(primate_subclass_de_column))
  
  #Human cluster
  sestan_human_cluster_mat = filt_sestan_human_markers %>% filter(cell_type %in% primate_clusters & 
                                                                    gene %in% test_primate_subclass_markers ) %>%
    select(cell_type, gene, !! sym(de_metric)) %>% tidyr::pivot_wider(names_from = cell_type, values_from = !! sym(de_metric))
  sestan_human_cluster_mat  = tibble::column_to_rownames(sestan_human_cluster_mat , var = 'gene')
  colnames(sestan_human_cluster_mat) = paste0('human_', colnames(sestan_human_cluster_mat) )
  gene_index = match(rownames(primate_subclass_de_column), rownames(sestan_human_cluster_mat))
  sestan_human_cluster_mat = sestan_human_cluster_mat[gene_index, , drop = F]
  #chimp cluster
  sestan_chimp_cluster_mat = filt_sestan_chimp_markers %>% filter(cell_type %in% primate_clusters & 
                                                                    gene %in% test_primate_subclass_markers ) %>%
    select(cell_type, gene, !! sym(de_metric)) %>% tidyr::pivot_wider(names_from = cell_type, values_from = !! sym(de_metric))
  sestan_chimp_cluster_mat  = tibble::column_to_rownames(sestan_chimp_cluster_mat , var = 'gene')
  colnames(sestan_chimp_cluster_mat) = paste0('chimp_', colnames(sestan_chimp_cluster_mat) )
  gene_index = match(rownames(primate_subclass_de_column), rownames(sestan_chimp_cluster_mat))
  sestan_chimp_cluster_mat = sestan_chimp_cluster_mat[gene_index, , drop = F]
  #macaque cluster
  sestan_macaque_cluster_mat = filt_sestan_macaque_markers %>% filter(cell_type %in% primate_clusters & 
                                                                    gene %in% test_primate_subclass_markers ) %>%
    select(cell_type, gene, !! sym(de_metric)) %>% tidyr::pivot_wider(names_from = cell_type, values_from = !! sym(de_metric))
  sestan_macaque_cluster_mat  = tibble::column_to_rownames(sestan_macaque_cluster_mat , var = 'gene')
  colnames(sestan_macaque_cluster_mat) = paste0('macaque_', colnames(sestan_macaque_cluster_mat) )
  gene_index = match(rownames(primate_subclass_de_column), rownames(sestan_macaque_cluster_mat))
  sestan_macaque_cluster_mat = sestan_macaque_cluster_mat[gene_index, , drop = F]
  #marmoset cluster
  sestan_marmoset_cluster_mat = filt_sestan_marmoset_markers %>% filter(cell_type %in% primate_clusters & 
                                                                    gene %in% test_primate_subclass_markers ) %>%
    select(cell_type, gene, !! sym(de_metric)) %>% tidyr::pivot_wider(names_from = cell_type, values_from = !! sym(de_metric))
  sestan_marmoset_cluster_mat  = tibble::column_to_rownames(sestan_marmoset_cluster_mat , var = 'gene')
  colnames(sestan_marmoset_cluster_mat) = paste0('marmoset_', colnames(sestan_marmoset_cluster_mat) )
  gene_index = match(rownames(primate_subclass_de_column), rownames(sestan_marmoset_cluster_mat))
  sestan_marmoset_cluster_mat = sestan_marmoset_cluster_mat[gene_index, , drop = F]
  
  
  all_species_de = do.call('cbind', list(sestan_human_cluster_mat, sestan_chimp_cluster_mat, 
                                             sestan_macaque_cluster_mat, sestan_marmoset_cluster_mat, primate_subclass_de_column ))
  
  cor_mat = cor(all_species_de, method = 'spearman')
  
  #ComplexHeatmap::Heatmap(cor_mat, name = 'spearman',
  #                       clustering_distance_rows = function(m) dist(1-m), clustering_distance_columns = function(m) dist(1-m),
  #                       clustering_method_rows = 'average', clustering_method_columns = 'average')

  long_cor_df = reshape2::melt(cor_mat, value.name = 'spearman', varnames = c('cluster_1','cluster_2'))

  corr_summary_df = long_cor_df %>% filter(cluster_2 == sprintf('Primate_%s', primate_subclass)) %>% 
    mutate(num_markers = num_top_markers)
  
  all_corr_summary_dfs[[i]] = corr_summary_df 
  
}

all_corr_summary_dfs = do.call('rbind', all_corr_summary_dfs)

primate_17_species_colors = c('darkblue',  'lightblue',  'purple', 'orange')
names(primate_17_species_colors) = paste0(c('human_','chimp_','macaque_','marmoset_'), primate_clusters)
ggplot(all_corr_summary_dfs , aes(x = num_markers, y = spearman, color = cluster_1)) + geom_line() +
  ggtitle(sprintf('Similarity in DE to Primate %s', primate_subclass)) + 
  ylab('Correlation in marker DE') +
  xlab('Number of top markers') +
  scale_color_manual(values = primate_17_species_colors)


```


