

Exploring some human and mouse published patch-seq data

At least for some mouse data, the cells are already mapped to the Tasic18 clusters, can assign them to the homologous primate subclasses.


```{r}
library(ggplot2)
library(dplyr)
library(SingleCellExperiment)

```

```{r}
celltype_color_palette = c("Chandelier"='#F641A8', 
                           "Pvalb"='#D93137', 
                           "Sst"='#FF9900', 
                           "Sst Chodl"='gold', 
                           "Lamp5"='#DA808C',
                           "Sncg"='#DF70FF',
                           "Vip"='#A45FBF', 
                           "Lamp5 Lhx6"='#935F50',
                           "Pax6"='#71238C')


```


```{r}

tasic18 = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/tasic18_gaba_primate_MN.rds')
tasic_metadata = as.data.frame(colData(tasic18))

tasic_clust_to_subclass_df = tasic_metadata %>% group_by(cell_cluster, cell_subclass, primate_subclass, orthology) %>% summarise(n = n()) %>% arrange(orthology)


clust_to_primate_subclass = c(tasic_clust_to_subclass_df$primate_subclass)
names(clust_to_primate_subclass) = tasic_clust_to_subclass_df$cell_cluster
clust_to_primate_subclass

clust_to_mouse_subclass = c(tasic_clust_to_subclass_df$cell_subclass)
names(clust_to_mouse_subclass) = tasic_clust_to_subclass_df$cell_cluster
clust_to_mouse_subclass


```





Tasic18 pax6 clusters
c('Lamp5 Fam19a1 Pax6','Lamp5 Krt73','Serpinf1 Clrn1','Sncg Slc17a8')

Tasic18 sncg cluster
c('Vip Igfbp6 Pltp')


```{r}

mouse_file_path = '/home/werner/projects/mouse_brain_consensus_taxonomy/data/patch_seq/mouse/'

#Has cell IDs and their cluster annotation
mouse_patch_meta = data.table::fread(paste0(mouse_file_path, '20200625_patchseq_metadata_mouse.csv'))

table(mouse_patch_meta$corresponding_AIT2.3.1_alias)


mouse_patch_meta %>% filter(corresponding_AIT2.3.1_alias %in% c('Lamp5 Fam19a1 Pax6','Lamp5 Krt73','Serpinf1 Clrn1','Sncg Slc17a8') )

#Add the mouse and primate subclass labels to the clusters
mouse_patch_meta$primate_subclass = clust_to_primate_subclass[mouse_patch_meta$corresponding_AIT2.3.1_alias]
mouse_patch_meta$mouse_subclass = clust_to_mouse_subclass[mouse_patch_meta$corresponding_AIT2.3.1_alias]


```



Add the cluster annotations from the metadata and then add the mouse subclass and homologous subclass annotations from the Tasic data



```{r}

mouse_patch_data = data.table::fread(paste0(mouse_file_path, 'ephys_features_gouwens_w_QC.csv'))

index = match(mouse_patch_data$cell_specimen_id,mouse_patch_meta$cell_specimen_id )
table(mouse_patch_data$cell_specimen_id == mouse_patch_meta$cell_specimen_id[index] )

mouse_patch_data$cell_cluster = mouse_patch_meta$corresponding_AIT2.3.1_alias[index]
mouse_patch_data$mouse_subclass = mouse_patch_meta$mouse_subclass[index]
mouse_patch_data$primate_subclass = mouse_patch_meta$primate_subclass[index]





```


```{r}

ggplot(filter(mouse_patch_data, mouse_subclass %in% c('Vip','Sncg','Lamp5')), 
       aes(x = qc_initial_access_resistance_mohm , y = input_resistance, color = mouse_subclass )) + 
  geom_point(alpha = .5) +
  scale_color_manual(values = celltype_color_palette) 


ggplot(filter(mouse_patch_data, mouse_subclass %in% c('Vip','Sncg','Lamp5')), 
       aes(x = rheo_fast_trough_t , y = input_resistance, color = primate_subclass )) + 
  geom_point(alpha = .5) +
  scale_color_manual(values = celltype_color_palette) 


```


Trying PCA just to see 
Filtering out the variables that have mostly NAs
filt_vars = c('file_id','file_name','cell_cluster','mouse_subclass','primate_subclass','qc_blowout_mv','rheo_slow_trough_t','rheo_adp_v','rheo_slow_trough_v',
'hero_adapt','hero_first_isi','hero_isi_cv','hearo_mean_isi')

```{r}

#Denote the Lamp5 Lhx6 as a subclass in the original mouse annotations
mouse_patch_data$mouse_subclass[mouse_patch_data$cell_cluster == 'Lamp5 Lhx6'] = 'Lamp5 Lhx6'
#Denote the Serpin mouse subclass as Vip
mouse_patch_data$mouse_subclass[mouse_patch_data$mouse_subclass == 'Serpinf1'] = 'Vip'



filt_vars = c('file_id','file_name','cell_cluster','mouse_subclass','primate_subclass','qc_blowout_mv','rheo_slow_trough_t','rheo_adp_v','rheo_slow_trough_v',
'hero_adapt','hero_first_isi','hero_isi_cv','hero_mean_isi', 'qc_recording_date')

filt_mouse_patch_data = mouse_patch_data %>% filter(mouse_subclass %in% c('Vip','Lamp5','Sncg', 'Lamp5 Lhx6')) %>% select(-filt_vars)

#Filter out cells with NA values
row_filt = rowSums(is.na(filt_mouse_patch_data)) == 0 
filt_mouse_patch_data = filt_mouse_patch_data[row_filt, ]


filt_mouse_patch_data = as.matrix(filt_mouse_patch_data %>% tibble::column_to_rownames(var = 'cell_specimen_id'))


data.pca <- princomp(filt_mouse_patch_data)
summary(data.pca)

pca_loadings_df = as.data.frame(data.pca$scores)


index = match(rownames(pca_loadings_df), mouse_patch_data$cell_specimen_id )
table(rownames(pca_loadings_df) == mouse_patch_data$cell_specimen_id[index] )

pca_loadings_df$cell_cluster = mouse_patch_data$cell_cluster[index]
pca_loadings_df$mouse_subclass = mouse_patch_data$mouse_subclass[index]
pca_loadings_df$primate_subclass = mouse_patch_data$primate_subclass[index]


pca_loadings_df

```


Generally very subtle differences across the ephys differences for the different subtypes, largely looks like a Lamp5 vs Vip spectrum, Sncg and Pax6 are inbetween, but very continuous.


```{r}

ggplot(pca_loadings_df, aes(x = Comp.1, y = Comp.2, color = mouse_subclass)) + geom_point(alpha = .75) +
  scale_color_manual(values = celltype_color_palette)


ggplot(pca_loadings_df, aes(x = Comp.1, y = Comp.2, color = primate_subclass)) + geom_point(alpha = .75) +
  scale_color_manual(values = celltype_color_palette)

```


PUlling together a standard variance decomposition approach to compare across the different annotations.
Starting simply with a one-way ANOVA, univariate decomposition

The goal is to compare the variance explained by the original annotations to the variance explained by the homologous annotations.
Use ANOVA decomposition to get the coefficient of determination for labelsA (mouse subclasses) and then labelsB (homologous subclasses). 
Can get a p-value for each coefficient/feature and can also compare the distribution of coefficients across all the features for the two labels.

Since the homologous labels have 1 more annotation, need to use either adjustedR^2 or omega^2 as the coefficient. Empirical p-values will also be useful due to the varying sample size of each annotation group.

Z-score the ephys features to start, since all in different units.

Don't necessarily have to z-score the genes, since all CPM (may try logCPM), but if I want to compare the omegas across the modalities, will need to zscore anyhow.


Code is from chatGPT, double checking implementation and output

```{r}

library(parallel)

# ---------- omega^2 estimator ----------
omega2 <- function(y, grp) {
  grp <- factor(grp)
  n <- length(y)
  k <- nlevels(grp)
  if (k < 2) return(0)
  grand_mean <- mean(y)
  group_means <- tapply(y, grp, mean)
  n_k <- as.numeric(table(grp))
  ss_b <- sum(n_k * (group_means - grand_mean)^2)
  ss_t <- sum((y - grand_mean)^2)
  ms_w <- (ss_t - ss_b) / (n - k)
  num <- ss_b - (k - 1) * ms_w
  denom <- ss_t + ms_w
  if (denom <= 0 || num <= 0) return(0)
  num / denom
}

# ---------- permutation test per feature ----------
omega_perm_test <- function(y, grp, n_perm = 1000) {
  obs <- omega2(y, grp)
  # permute while preserving group sizes
  n_k <- table(grp)
  perm_fun <- function() {
    shuffled <- sample(y)
    shuffled_grp <- rep(names(n_k), n_k)
    omega2(shuffled, shuffled_grp)
  }
  perms <- replicate(n_perm, perm_fun())
  pval <- (1 + sum(perms >= obs)) / (1 + n_perm)
  list(obs = obs, pval = pval)
}

# ---------- wrapper across many features ----------
run_omega_perm <- function(mat, grp, n_perm = 1000, n_cores = 4) {
  stopifnot(is.matrix(mat))
  cl <- makeCluster(n_cores)
  clusterExport(cl, c("omega2", "omega_perm_test", "grp", "n_perm"), envir = environment())
  res <- parApply(cl, mat, 1, function(y) omega_perm_test(y, grp, n_perm))
  stopCluster(cl)
  
  # Convert list output to data.frame
  obs <- sapply(res, function(x) x$obs)
  pval <- sapply(res, function(x) x$pval)
  data.frame(feature = rownames(mat), omega2 = obs, pval = pval)
}

# ---------- example usage ----------
# expr: matrix of features (genes/ephys vars) x samples (cells)
# labels_A and labels_B: factor vectors of length ncol(expr)

#res_A <- run_omega_perm(expr, labels_A, n_perm = 1000, n_cores = 8)
#res_B <- run_omega_perm(expr, labels_B, n_perm = 1000, n_cores = 8)

# Add FDR correction
#res_A$padj <- p.adjust(res_A$pval, method = "BH")
#res_B$padj <- p.adjust(res_B$pval, method = "BH")

```





```{r}
#Rows are cells, columns are ephys features
filt_mouse_patch_data 

scaled_filt_mouse_patch_data = scale(filt_mouse_patch_data)

#Get the two different annotations
index = match(rownames(filt_mouse_patch_data ), mouse_patch_data$cell_specimen_id )
mouse_subclass = mouse_patch_data$mouse_subclass[index]
primate_subclass = mouse_patch_data$primate_subclass[index]

table(mouse_subclass)
table(primate_subclass)


```

```{r}

# ---------- example usage ----------
# expr: matrix of features (genes/ephys vars) x samples (cells)
# labels_A and labels_B: factor vectors of length ncol(expr)

ephys_feature_mat = t(scaled_filt_mouse_patch_data )

res_mouse <- run_omega_perm(ephys_feature_mat, mouse_subclass, n_perm = 1000, n_cores = 8)
res_primate <- run_omega_perm(ephys_feature_mat, primate_subclass, n_perm = 1000, n_cores = 8)

# Add FDR correction
res_mouse$padj <- p.adjust(res_mouse$pval, method = "BH")
res_primate$padj <- p.adjust(res_primate$pval, method = "BH")

par(pty= 's')
plot(res_mouse$omega2, res_primate$omega2, main = 'Ephys features', 
     ylab = 'Variance explained by homologous subclasses', 
     xlab = 'Variance explained by original subclasses', xlim = c(0,1), ylim = c(0,1) )
abline(a = 0, b = 1, col = 'red', lty = 'dashed')


res_mouse
res_primate


combined_res = res_mouse
combined_res$primate_omega2 = res_primate$omega2
combined_res$primate_pval = res_primate$pval
combined_res$primate_padj = res_primate$padj

combined_res = combined_res %>% mutate(omega_diff = abs(omega2 - primate_omega2)) %>% arrange(desc(omega_diff))
combined_res

```


good features for separability
tau
rheo_width
rheobase_i
qc_input_resistance_mohm



```{r}

library(ggExtra)

```


```{r}

p_mouse = ggplot(filter(mouse_patch_data, primate_subclass %in% c('Vip','Sncg','Lamp5', 'Lamp5 Lhx6','Pax6')), 
       aes(x = tau, y = rheo_width, color = mouse_subclass )) + 
  geom_point(alpha = .75, size = 2) +
  scale_color_manual(values = celltype_color_palette) + theme_bw()

ggMarginal(p_mouse, type="density",  groupColour = TRUE, groupFill = TRUE)

p_primate = ggplot(filter(mouse_patch_data, primate_subclass %in% c('Vip','Sncg','Lamp5', 'Lamp5 Lhx6','Pax6')), 
       aes(x = tau , y = rheo_width, color = primate_subclass )) + 
  geom_point(alpha = .75, size = 2) +
  scale_color_manual(values = celltype_color_palette) + theme_bw()

ggMarginal(p_primate, type="density",  groupColour = TRUE, groupFill = TRUE)

```

The features with the biggest differn

```{r}

p_mouse = ggplot(filter(mouse_patch_data, mouse_subclass %in% c('Lamp5')), 
       aes(x = sag, y = rheo_fast_trough_t, color = mouse_subclass )) + 
  geom_point(alpha = .75, size = 2) +
  scale_color_manual(values = celltype_color_palette) + theme_bw()

ggMarginal(p_mouse, type="density",  groupColour = TRUE, groupFill = TRUE)

p_primate = ggplot(filter(mouse_patch_data, mouse_subclass %in% c('Lamp5')), 
       aes(x = sag , y = rheo_fast_trough_t, color = primate_subclass )) + 
  geom_point(alpha = .75, size = 2) +
  scale_color_manual(values = celltype_color_palette) + theme_bw()

ggMarginal(p_primate, type="density",  groupColour = TRUE, groupFill = TRUE)



p_mouse = ggplot(filter(mouse_patch_data, mouse_subclass %in% c('Vip')), 
       aes(x = sag, y = rheo_fast_trough_t, color = mouse_subclass )) + 
  geom_point(alpha = .75, size = 2) +
  scale_color_manual(values = celltype_color_palette) + theme_bw()

ggMarginal(p_mouse, type="density",  groupColour = TRUE, groupFill = TRUE)

p_primate = ggplot(filter(mouse_patch_data, mouse_subclass %in% c('Vip')), 
       aes(x = sag , y = rheo_fast_trough_t, color = primate_subclass )) + 
  geom_point(alpha = .75, size = 2) +
  scale_color_manual(values = celltype_color_palette) + theme_bw()

ggMarginal(p_primate, type="density",  groupColour = TRUE, groupFill = TRUE)



p_mouse = ggplot(filter(mouse_patch_data, primate_subclass %in% c('Pax6')), 
       aes(x = sag, y = rheo_fast_trough_t, color = mouse_subclass )) + 
  geom_point(alpha = .75, size = 2) +
  scale_color_manual(values = celltype_color_palette) + theme_bw()

ggMarginal(p_mouse, type="density",  groupColour = TRUE, groupFill = TRUE)

p_primate = ggplot(filter(mouse_patch_data, primate_subclass %in% c('Pax6')), 
       aes(x = sag , y = rheo_fast_trough_t, color = primate_subclass )) + 
  geom_point(alpha = .75, size = 2) +
  scale_color_manual(values = celltype_color_palette) + theme_bw()

ggMarginal(p_primate, type="density",  groupColour = TRUE, groupFill = TRUE)




```


And do the same univariate decomp with the tasic expression data
Only use highly variable genes


```{r}

tasic18$cell_subclass[tasic18$cell_cluster == 'Lamp5 Lhx6'] = 'Lamp5 Lhx6'
tasic18$cell_subclass[tasic18$cell_subclass == 'Serpinf1'] = 'Vip'

cell_filt = tasic18$cell_subclass %in% c('Lamp5','Lamp5 Lhx6','Vip','Sncg')
tasic18_cge = tasic18[ , cell_filt]

table(tasic18_cge$cell_subclass)
table(tasic18_cge$primate_subclass)




#Get highly variable genes
logcounts(tasic18_cge) = log1p(cpm(tasic18_cge))
stats = scran::modelGeneVar(tasic18_cge)
dataset_hvg = scran::getTopHVGs(stats, n = 2000)
dataset_hvg

#Get the logcounts of the Hvgs
gene_filt = rownames(tasic18_cge) %in% dataset_hvg
feature_exp = logcounts(tasic18_cge)[gene_filt, ]
scale_feature_exp = t(scale(t(feature_exp)))

#And the subclass annotations
exp_mouse_subclass = tasic18_cge$cell_subclass
exp_primate_subclass = tasic18_cge$primate_subclass

table(exp_mouse_subclass)
table(exp_primate_subclass)
```


```{r}
res_exp_mouse <- run_omega_perm(scale_feature_exp, exp_mouse_subclass, n_perm = 1000, n_cores = 8)
res_exp_primate <- run_omega_perm(scale_feature_exp, exp_primate_subclass, n_perm = 1000, n_cores = 8)

# Add FDR correction
res_exp_mouse$padj <- p.adjust(res_exp_mouse$pval, method = "BH")
res_exp_primate$padj <- p.adjust(res_exp_primate$pval, method = "BH")

par(pty= 's')
plot(res_exp_mouse$omega2, res_exp_primate$omega2, main = 'Transcriptomic features', 
     ylab = 'Variance explained by homologous subclasses', 
     xlab = 'Variance explained by original subclasses', xlim = c(0,1), ylim = c(0,1) )
abline(a = 0, b = 1, col = 'red', lty = 'dashed')

combined_res = res_exp_mouse
combined_res$primate_omega2 = res_exp_primate$omega2
combined_res$primate_pval = res_exp_primate$pval
combined_res$primate_padj = res_exp_primate$padj

combined_res = combined_res %>% mutate(omega_diff = abs(omega2 - primate_omega2)) %>% arrange(desc(omega_diff))
combined_res


```

```{r}

feature_exp


current_metadata = as.data.frame(colData(tasic18_cge))


current_metadata$gene_1_exp = feature_exp['CALN1' ,]
current_metadata$gene_2_exp = feature_exp['VSNL1' ,]

ggplot(current_metadata, aes(x = gene_1_exp, y = gene_2_exp, color = cell_subclass)) +
  geom_point() + scale_color_manual(values = celltype_color_palette) + theme_bw()

ggplot(current_metadata, aes(x = gene_1_exp, y = gene_2_exp, color = primate_subclass)) +
  geom_point() + scale_color_manual(values = celltype_color_palette) + theme_bw()



```
















