

Processing rat brain data for the consensus interneuron subclass labels

Not too many rat brain single cell datasets easily available
Attempting a similar analysis as with the non-allen primate data

Pull the data all together from each study
De novo cluster and than predict cell-type annotations using the original mouse subclass markers, homologous markers, and the probably the primate markers too


Aydin dataset
https://pubmed.ncbi.nlm.nih.gov/39455606/

Aydin data seems non-conventional


Wang dataset
https://pubmed.ncbi.nlm.nih.gov/38495123/

Wang data is in the cellRanger barcodes, genes, matrix output

```{r}
library(Seurat)
library(SingleCellExperiment)
library(MetaMarkers)
library(dplyr)
library(ggplot2)
```

```{r}
celltype_color_palette = c("Chandelier"='#F641A8', 
                           "Pvalb"='#D93137', 
                           "Sst"='#FF9900', 
                           "Sst Chodl"='gold', 
                           "Lamp5"='#DA808C',
                           "Sncg"='#DF70FF',
                           "Vip"='#A45FBF', 
                           "Lamp5 Lhx6"='#935F50',
                           "Pax6"='#71238C')


```

```{r}

final_present_1to1s = readRDS('/home/werner/projects/mouse_brain_consensus_taxonomy/data/final_present_1to1s.rds')

final_present_1to1s


```


Aydin data

Data files are ParseBiosciences output, Seurat has a wrapper function




```{r}

aydin_path = '/inkwell03/werner/sc_data/rattus_brain/aydin_2024'

aydin_counts = Seurat::ReadParseBio(aydin_path)

aydin_cell_metadata = data.table::fread(paste0(aydin_path, '/GSM8182337_cell_metadata.csv'))


#Only keep the genes in the final 1to1s, match to the mouse gene name and swap in the human gene name

index = match(rownames(aydin_counts), final_present_1to1s$`Mouse gene name`)
final_present_1to1s[index,]$`Gene name`

rownames(aydin_counts) = final_present_1to1s[index,]$`Gene name`

gene_filt = !is.na(rownames(aydin_counts))
aydin_counts = aydin_counts[gene_filt, ]
dim(aydin_counts)



aydin_seurat <- CreateSeuratObject(counts = aydin_counts, project = "aydin_rat", min.features = 500, meta.data = aydin_cell_metadata )

aydin_seurat[["percent.mt"]] <- PercentageFeatureSet(aydin_seurat, pattern = "^Mt-")
aydin_seurat$region = 'Auditory_cortex'

Idents(aydin_seurat) = aydin_seurat$region

VlnPlot(aydin_seurat, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```

Keep the genes that are in common with the final mouse orthologs in the final cross-species 1to1s that we have 
And swap in the human name for the gene

```{r}

table(rownames(aydin_seurat) %in% final_present_1to1s$`Gene name`)


aydin_seurat@assays$RNA@features[[]]
```







```{r}
#CPM normalization
aydin_seurat <- NormalizeData(aydin_seurat, normalization.method = 'RC', scale.factor = 1e6, verbose = F)
aydin_seurat <- FindVariableFeatures(aydin_seurat, selection.method = "vst", nfeatures = 2000, verbose = F)
all.genes <- rownames(aydin_seurat)
aydin_seurat <- ScaleData(aydin_seurat, features = all.genes, verbose = F)


aydin_seurat <- RunPCA(aydin_seurat, features = VariableFeatures(object =aydin_seurat), verbose = F)


aydin_seurat <- FindNeighbors(aydin_seurat, dims = 1:25, verbose = F)
aydin_seurat <- FindClusters(aydin_seurat, verbose = F)

aydin_seurat <- RunUMAP(aydin_seurat, dims = 1:25, verbose = F)

DimPlot(aydin_seurat, reduction = "pca", label = T)
DimPlot(aydin_seurat, reduction = "umap", label = T)


```

```{r}

cross_primate_markers = data.table::fread('/inkwell03/werner/sc_data/primate_MTG/primate_MTG_consensus_markers.csv', skip = 4)

primate_class_markers = cross_primate_markers[ , c('gene','InN','NonN','ExN')]


table(primate_class_markers$gene %in% rownames(aydin_seurat))


top_markers = list('InN' = primate_class_markers$gene[as.logical(primate_class_markers$InN)],
     'NonN' = primate_class_markers$gene[as.logical(primate_class_markers$NonN)],
     'ExN' = primate_class_markers$gene[as.logical(primate_class_markers$ExN)])



aydin_sce = as.SingleCellExperiment(aydin_seurat)
names(assays(aydin_sce)) = c('counts','cpm','scaledata')



ct_scores = score_cells(log1p(cpm(aydin_sce)), top_markers)
ct_enrichment = compute_marker_enrichment(ct_scores)
ct_pred = assign_cells(ct_scores)

ct_pred[1:10, ]



aydin_seurat$glutamatergic_enrichment = ct_enrichment['all|ExN', ] / max(ct_enrichment['all|ExN', ])
aydin_seurat$GABAergic_enrichment = ct_enrichment['all|InN', ] / max(ct_enrichment['all|InN', ])
aydin_seurat$non_neuronal_enrichment = ct_enrichment['all|NonN', ] / max(ct_enrichment['all|NonN', ])

aydin_seurat$predicted_primate_MM_class = ct_pred$predicted

```


```{r}

DimPlot(aydin_seurat, reduction = "umap", label = T, group.by = 'predicted_primate_MM_class')

```


```{r}

cell_index = aydin_seurat$predicted_primate_MM_class == 'InN'
aydin_seurat_inhibitory = aydin_seurat[ , cell_index]

aydin_seurat_inhibitory

saveRDS(aydin_seurat, file = '/inkwell03/werner/sc_data/rattus_brain/aydin_2024/aydin_seurat_full_data.rds')

saveRDS(aydin_seurat_inhibitory, file = '/inkwell03/werner/sc_data/rattus_brain/aydin_2024/aydin_seurat_GABA.rds')


```



```{r}


aydin_seurat_inhibitory = readRDS(file = '/inkwell03/werner/sc_data/rattus_brain/aydin_2024/aydin_seurat_GABA.rds')

aydin_seurat_inhibitory
```





New de novo clusters for just the inhibitory neurons


```{r}
#CPM normalization
aydin_seurat_inhibitory <- FindVariableFeatures(aydin_seurat_inhibitory, selection.method = "vst", nfeatures = 2000, verbose = F)
all.genes <- rownames(aydin_seurat_inhibitory)
aydin_seurat_inhibitory <- ScaleData(aydin_seurat_inhibitory, features = all.genes, verbose = F)


aydin_seurat_inhibitory <- RunPCA(aydin_seurat_inhibitory, features = VariableFeatures(object =aydin_seurat_inhibitory), verbose = F)


aydin_seurat_inhibitory <- FindNeighbors(aydin_seurat_inhibitory, dims = 1:25, verbose = F)
aydin_seurat_inhibitory <- FindClusters(aydin_seurat_inhibitory, verbose = F)

aydin_seurat_inhibitory <- RunUMAP(aydin_seurat_inhibitory, dims = 1:25, verbose = F)

DimPlot(aydin_seurat_inhibitory, reduction = "pca", label = T)
DimPlot(aydin_seurat_inhibitory, reduction = "umap", label = T)


```

Load up the mouse and primate CGE/MGE metamarkers

```{r}
mouse_class_meta_markers = read_meta_markers("../metamarkers/MGE_CGE/mouse_meta_markers.csv.gz")
primate_class_meta_markers = read_meta_markers("../metamarkers/MGE_CGE/primate_meta_markers.csv.gz")

mammal_class_meta_markers = read_meta_markers("../metamarkers/MGE_CGE/mammal_CGE_MGE_meta_markers.csv.gz")
```




Check out the CGE and MGE predictions for the tree shrew data

```{r}

rownames(aydin_seurat_inhibitory@assays$RNA@layers$data) = rownames(aydin_seurat_inhibitory)


#mouse class markers
top_markers = mouse_class_meta_markers %>% filter(rank <= 100)
ct_scores = score_cells(log1p(aydin_seurat_inhibitory@assays$RNA@layers$data), top_markers)
ct_enrichment = compute_marker_enrichment(ct_scores)
ct_pred = assign_cells(ct_scores)

aydin_seurat_inhibitory$mouse_cge_mge = ct_pred$predicted

#Primate class markers
top_markers = primate_class_meta_markers %>% filter(rank <= 100)
ct_scores = score_cells(log1p(aydin_seurat_inhibitory@assays$RNA@layers$data), top_markers)
ct_enrichment = compute_marker_enrichment(ct_scores)
ct_pred = assign_cells(ct_scores)

aydin_seurat_inhibitory$primate_cge_mge = ct_pred$predicted

#Cross_mammal class markers
top_markers = mammal_class_meta_markers %>% filter(rank <= 100)
ct_scores = score_cells(log1p(aydin_seurat_inhibitory@assays$RNA@layers$data), top_markers)
ct_enrichment = compute_marker_enrichment(ct_scores)
ct_pred = assign_cells(ct_scores)

aydin_seurat_inhibitory$mammal_cge_mge = ct_pred$predicted

#Add the cross-mammal CGE-MGE enrichments for plotting
aydin_seurat_inhibitory$mammal_cge_enrich = ct_enrichment['all|CGE', ]
aydin_seurat_inhibitory$mammal_mge_enrich = ct_enrichment['all|MGE', ]


DimPlot(aydin_seurat_inhibitory, group.by = 'seurat_clusters', label = T) + ggtitle('Initial Aydin-Rat de novo GABA clusters')

DimPlot(aydin_seurat_inhibitory, group.by = 'mouse_cge_mge', label = T) + ggtitle('CGE-MGE annotations using Mouse markers')

DimPlot(aydin_seurat_inhibitory, group.by = 'primate_cge_mge', label = T)+ ggtitle('CGE-MGE annotations using Primate markers')

DimPlot(aydin_seurat_inhibitory, group.by = 'mammal_cge_mge', label = T)+ 
  ggtitle('CGE-MGE annotations using Mammal markers') +
  scale_color_manual(values = c('CGE' = 'darkorchid', 'MGE' = 'goldenrod1', 'unassigned' = 'red'))

FeaturePlot(aydin_seurat_inhibitory, features = c('mammal_cge_enrich'))+ 
  ggtitle('Mammalian CGE enrichment: Aydin-Rat') +
  scale_colour_gradient2(low = "white", mid = "white", high = "black", midpoint = 1)

FeaturePlot(aydin_seurat_inhibitory, features = c('mammal_mge_enrich'))+ 
  ggtitle('Mammalian MGE enrichment: Aydin-Rat') +
  scale_colour_gradient2(low = "white", mid = "white", high = "black", midpoint = 1)

```

14 had clear enrichment of chandelier markers

```{r}
mouse_class_nums = table(aydin_seurat_inhibitory$mouse_cge_mge, aydin_seurat_inhibitory$seurat_clusters)
total_clust_nums = colSums(mouse_class_nums)
mouse_class_ratios = t(t(mouse_class_nums) / total_clust_nums)

primate_class_nums = table(aydin_seurat_inhibitory$primate_cge_mge, aydin_seurat_inhibitory$seurat_clusters)
total_clust_nums = colSums(primate_class_nums)
primate_class_ratios = t(t(primate_class_nums) / total_clust_nums)


primate_class_ratios = t(primate_class_ratios)
for_ggplot_df = data.frame(CGE = primate_class_ratios[ ,'CGE'], MGE = primate_class_ratios[ ,'MGE'], cluster = rownames(primate_class_ratios))
for_ggplot_df$cluster[for_ggplot_df$CGE >= .80 | for_ggplot_df$CGE <= .20] = NA

ggplot(for_ggplot_df, aes(x = CGE, y = MGE, label = cluster)) + geom_point() +
  ggrepel::geom_label_repel(max.overlaps = Inf, min.segment.length = 0) + 
  theme_bw() + coord_fixed() + ggtitle('CGE-MGE cluster proportions: Aydin-Rat')


mouse_class_ratios = t(mouse_class_ratios)
for_ggplot_df = data.frame(CGE = mouse_class_ratios[ ,'CGE'], MGE = mouse_class_ratios[ ,'MGE'], cluster = rownames(mouse_class_ratios))
for_ggplot_df$cluster[for_ggplot_df$CGE >= .80 | for_ggplot_df$CGE <= .20] = NA

ggplot(for_ggplot_df, aes(x = CGE, y = MGE, label = cluster)) + geom_point() +
  ggrepel::geom_label_repel(max.overlaps = Inf, min.segment.length = 0) + 
  theme_bw() + coord_fixed() + ggtitle('Mouse CGE-MGE cluster proportions: Aydin-Rat')



ggsave(filename = 'initial_aydin_rat_mouse_CGEMGE_clust_ratios_ggplot.pdf', path = '../../graphs/non_allen_data/', useDingbats = F,
       device = 'pdf', height = 4, width = 4)


```

Filter out cluster 10 and re cluster

```{r}

#filter out cluster 10
cell_index = aydin_seurat_inhibitory$seurat_clusters != '10'
aydin_seurat_inhibitory = aydin_seurat_inhibitory[, cell_index]


aydin_seurat_inhibitory <- RunPCA(aydin_seurat_inhibitory, features = VariableFeatures(object =aydin_seurat_inhibitory), verbose = F)

aydin_seurat_inhibitory <- FindNeighbors(aydin_seurat_inhibitory, dims = 1:25, verbose = F)
aydin_seurat_inhibitory <- FindClusters(aydin_seurat_inhibitory, verbose = F)

aydin_seurat_inhibitory <- RunUMAP(aydin_seurat_inhibitory, dims = 1:25, verbose = F)

DimPlot(aydin_seurat_inhibitory, reduction = "pca", label = T)
DimPlot(aydin_seurat_inhibitory, reduction = "umap", label = T)


```



Add in the primate and mouse subclass predictions

```{r}
all_mouse_markers = list(
    tasic18 = read_markers("../metamarkers/ortho_markers_tasic18.csv.gz"),
    zeng_10x_cells = read_markers("../metamarkers/ortho_markers_zeng_10x_cells.csv.gz"),
    zeng_10x_nuclei = read_markers("../metamarkers/ortho_markers_zeng_10x_nuclei.csv.gz"),
    zeng_smart_nuclei = read_markers("../metamarkers/ortho_markers_zeng_smart_nuclei.csv.gz"),
    zeng_smart_cells = read_markers("../metamarkers/ortho_markers_zeng_smart_cells.csv.gz"),
    zeng_10xv2_wba = read_markers("../metamarkers/ortho_markers_zeng_10xv2_wba.csv.gz"),
    zeng_10xv3_wba = read_markers("../metamarkers/ortho_markers_zeng_10xv3_wba.csv.gz"),
    macosko_wba = read_markers("../metamarkers/ortho_markers_macosko_wba.csv.gz")
)


mouse_subclass_meta_markers = make_meta_markers(all_mouse_markers, detailed_stats = TRUE)

mouse_subclass_meta_markers %>% group_by(cell_type) %>% filter(rank <= 10)


```


```{r}
all_species_markers = list(
    human_10x = read_markers("../metamarkers/markers_human_10x.csv.gz"),
    human_ss = read_markers("../metamarkers/markers_human_ss.csv.gz"),
    chimp_10x = read_markers("../metamarkers/markers_chimp_10x.csv.gz"),
    chimp_ss = read_markers("../metamarkers/markers_chimp_ss.csv.gz"),
    gorilla_10x = read_markers("../metamarkers/markers_gorilla_10x.csv.gz"),
    gorilla_ss = read_markers("../metamarkers/markers_gorilla_ss.csv.gz"),
    macaca_10x = read_markers("../metamarkers/markers_macaca_10x.csv.gz"),
    marmoset_10x= read_markers("../metamarkers/markers_marmoset_10x.csv.gz"),
    tasic18 = read_markers("../metamarkers/ortho_markers_tasic18.csv.gz"),
    zeng_10x_cells = read_markers("../metamarkers/ortho_markers_zeng_10x_cells.csv.gz"),
    zeng_10x_nuclei = read_markers("../metamarkers/ortho_markers_zeng_10x_nuclei.csv.gz"),
    zeng_smart_nuclei = read_markers("../metamarkers/ortho_markers_zeng_smart_nuclei.csv.gz"),
    zeng_smart_cells = read_markers("../metamarkers/ortho_markers_zeng_smart_cells.csv.gz"),
    zeng_10xv2_wba = read_markers("../metamarkers/ortho_markers_zeng_10xv2_wba.csv.gz"),
    zeng_10xv3_wba = read_markers("../metamarkers/ortho_markers_zeng_10xv3_wba.csv.gz"),
    macosko_wba = read_markers("../metamarkers/ortho_markers_macosko_wba.csv.gz")
)


all_species_markers$human_10x$cell_type[all_species_markers$human_10x$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
all_species_markers$human_ss$cell_type[all_species_markers$human_ss$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
all_species_markers$chimp_10x$cell_type[all_species_markers$chimp_10x$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
all_species_markers$chimp_ss$cell_type[all_species_markers$chimp_ss$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
all_species_markers$gorilla_10x$cell_type[all_species_markers$gorilla_10x$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
all_species_markers$gorilla_ss$cell_type[all_species_markers$gorilla_ss$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
all_species_markers$macaca_10x$cell_type[all_species_markers$macaca_10x$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
all_species_markers$marmoset_10x$cell_type[all_species_markers$marmoset_10x$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'

conserved_subclass_meta_markers = make_meta_markers(all_species_markers, detailed_stats = TRUE)

conserved_subclass_meta_markers %>% group_by(cell_type) %>% filter(rank <= 10)


all_primate_markers = all_species_markers[c('human_10x', 'human_ss', 'chimp_10x','chimp_ss','gorilla_10x','gorilla_ss','macaca_10x','marmoset_10x')]

primate_subclass_meta_markers = make_meta_markers(all_primate_markers, detailed_stats = TRUE)

primate_subclass_meta_markers %>% group_by(cell_type) %>% filter(rank <= 10)



```

```{r}
orig_mouse_markers = list(
    tasic18 = read_markers("../metamarkers/markers_tasic18.csv.gz"),
    zeng_10x_cells = read_markers("../metamarkers/markers_zeng_10x_cells.csv.gz"),
    zeng_10x_nuclei = read_markers("../metamarkers/markers_zeng_10x_nuclei.csv.gz"),
    zeng_smart_nuclei = read_markers("../metamarkers/markers_zeng_smart_nuclei.csv.gz"),
    zeng_smart_cells = read_markers("../metamarkers/markers_zeng_smart_cells.csv.gz"),
    zeng_10xv2_wba = read_markers("../metamarkers/markers_zeng_10xv2_wba.csv.gz"),
    zeng_10xv3_wba = read_markers("../metamarkers/markers_zeng_10xv3_wba.csv.gz"),
    macosko_wba = read_markers("../metamarkers/markers_macosko_wba.csv.gz")
)


original_mouse_subclass_meta_markers = make_meta_markers(orig_mouse_markers, detailed_stats = TRUE)

original_mouse_subclass_meta_markers %>% group_by(cell_type) %>% filter(rank <= 10)


```


```{r}

top_allen_markers = primate_subclass_meta_markers %>% filter(rank <= 50)
top_mouse_markers = mouse_subclass_meta_markers %>% filter(rank <= 50)
top_ori_mouse_markers = original_mouse_subclass_meta_markers %>% filter(rank <= 50)

rownames(aydin_seurat_inhibitory@assays$RNA@layers$data) = rownames(aydin_seurat_inhibitory)

ct_scores = score_cells(log1p(aydin_seurat_inhibitory@assays$RNA@layers$data), top_allen_markers)
ct_enrichment = compute_marker_enrichment(ct_scores)
ct_pred = assign_cells(ct_scores)
aydin_seurat_inhibitory$primate_subclass_annot = ct_pred$predicted

ct_scores = score_cells(log1p(aydin_seurat_inhibitory@assays$RNA@layers$data), top_mouse_markers)
ct_enrichment = compute_marker_enrichment(ct_scores)
ct_pred = assign_cells(ct_scores)
aydin_seurat_inhibitory$mouse_subclass_annot = ct_pred$predicted

ct_scores = score_cells(log1p(aydin_seurat_inhibitory@assays$RNA@layers$data), top_ori_mouse_markers)
ct_enrichment = compute_marker_enrichment(ct_scores)
ct_pred = assign_cells(ct_scores)
aydin_seurat_inhibitory$ori_mouse_subclass_annot = ct_pred$predicted

DimPlot(aydin_seurat_inhibitory, group.by = 'seurat_clusters', label = T) + ggtitle('Initial Aydin-Rat de novo GABA clusters')

DimPlot(aydin_seurat_inhibitory, group.by = 'mouse_cge_mge', label = T) + ggtitle('CGE-MGE annotations using Mouse markers')

DimPlot(aydin_seurat_inhibitory, group.by = 'primate_cge_mge', label = T)+ ggtitle('CGE-MGE annotations using Primate markers')

DimPlot(aydin_seurat_inhibitory, group.by = 'primate_subclass_annot', label = T) + ggtitle('subclass annotations using Primate markers')

DimPlot(aydin_seurat_inhibitory, group.by = 'mouse_subclass_annot', label = T)+ ggtitle('subclass annotations using Mouse markers')

DimPlot(aydin_seurat_inhibitory, group.by = 'ori_mouse_subclass_annot', label = T)+ ggtitle('subclass annotations using Mouse markers')


```


```{r}

DimPlot(aydin_seurat_inhibitory, group.by = 'seurat_clusters', label = T) + ggtitle('Aydin-Rat de novo GABA clusters') 
ggsave(filename = 'aydin_rat_de_novo_clusters_umap.pdf', path = '../../graphs/non_allen_data/umaps/', device = 'pdf',
       useDingbats = F, height = 4, width = 6)

DimPlot(aydin_seurat_inhibitory, group.by = 'mouse_subclass_annot')+ ggtitle('subclass annotations using homologous Mouse markers') + scale_color_manual(values = celltype_color_palette)
ggsave(filename = 'aydin_rat_homo_mouse_predict_umap.pdf', path = '../../graphs/non_allen_data/umaps/', device = 'pdf',
       useDingbats = F, height = 4, width = 6)

DimPlot(aydin_seurat_inhibitory, group.by = 'ori_mouse_subclass_annot')+ ggtitle('subclass annotations using original Mouse markers') + scale_color_manual(values = celltype_color_palette)
ggsave(filename = 'aydin_rat_ori_mouse_predict_umap.pdf', path = '../../graphs/non_allen_data/umaps/', device = 'pdf',
       useDingbats = F, height = 4, width = 6)


```


Save with the de novo clusters and all the annotations
```{r}
saveRDS(aydin_seurat_inhibitory, file = '/inkwell03/werner/sc_data/rattus_brain/aydin_2024/aydin_seurat_GABA_annotated.rds')

```




And get the confusion matrix per de novo cluster

```{r}


confuse_mat_aydin_rat = table(aydin_seurat_inhibitory$ori_mouse_subclass_annot, aydin_seurat_inhibitory$seurat_clusters)
cluster_totals_aydin_rat = colSums(confuse_mat_aydin_rat )
confuse_mat_aydin_rat = t(t(confuse_mat_aydin_rat)/cluster_totals_aydin_rat)

#Ordering by CGE and MGE
row_order = match(c('Sst','Sst Chodl','Pvalb','Chandelier','Lamp5','Lamp5 Lhx6','Vip','Sncg'), rownames(confuse_mat_aydin_rat))
round(confuse_mat_aydin_rat)[row_order, ]

col_order = c(2,4,5,6,14,13,0,1,7,11,8,10,3,12, 9) +1

col_fun = circlize::colorRamp2(breaks = c(0,.5, 1) , colors = c('white','orange','red'))
hm = ComplexHeatmap::Heatmap(confuse_mat_aydin_rat,name = 'cluster proportion',
                        col = col_fun,
                        cluster_rows = F, row_order = row_order,
                        cluster_columns = F, column_order = col_order,
                        column_title = 'Aydin-rat cluster predictions with original mouse markers')
hm = ComplexHeatmap::draw(hm)

pdf(file = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/non_allen_data/ori_mouse_MM_predict_aydin_rat_clusters.pdf', 
    useDingbats = F, height = 4, width = 8)
hm
dev.off()


confuse_mat_aydin_rat = table(aydin_seurat_inhibitory$mouse_subclass_annot, aydin_seurat_inhibitory$seurat_clusters)
cluster_totals_aydin_rat = colSums(confuse_mat_aydin_rat )
confuse_mat_aydin_rat = t(t(confuse_mat_aydin_rat)/cluster_totals_aydin_rat)

#Ordering by CGE and MGE
row_order = match(c('Sst','Sst Chodl','Pvalb','Chandelier','Lamp5','Lamp5 Lhx6','Vip','Sncg', 'Pax6', 'unassigned'), rownames(confuse_mat_aydin_rat))
#round(confuse_mat_aydin_rat)[row_order, ]
col_order = c(2,4,5,6,14,13,0,1,7,11,8,10,3,12, 9) +1

col_fun = circlize::colorRamp2(breaks = c(0,.5, 1) , colors = c('white','orange','red'))
hm = ComplexHeatmap::Heatmap(confuse_mat_aydin_rat,name = 'cluster proportion',
                        col = col_fun,
                        cluster_rows = F, row_order = row_order,
                        cluster_columns = F, column_order = col_order,
                        column_title = 'Aydin-rat cluster predictions with homologous mouse markers')
hm = ComplexHeatmap::draw(hm)

pdf(file = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/non_allen_data/homo_mouse_MM_predict_aydin_rat_clusters.pdf', 
    useDingbats = F, height = 4, width = 8)
hm
dev.off()


```

#########################################

And now the Wang dataset

Each sample has separate files

And the number of genes per sample don't match, which is odd considering they make the 10X files available, not sure what's going on there

###########################################


```{r}

wang_path = '/inkwell03/werner/sc_data/rattus_brain/wang_2024'
wang_directories = list.dirs(wang_path, recursive = F)

cortex_directories = wang_directories[grepl('CT', wang_directories)]

cortex_samples = sapply(strsplit(cortex_directories, split = '/', fixed = T), '[[', 7)
cortex_samples = sapply(strsplit(cortex_samples, split = '_', fixed = T), '[[', 1)


all_cortex_list = vector(mode = 'list', length = length(cortex_samples))
for(i in 1:length(cortex_samples)){

  seurat_counts = Read10X(cortex_directories[[i]])
  
  #Filter for the 1to1s and swap in the human gene names
  index = match(rownames(seurat_counts), final_present_1to1s$`Mouse gene name`)
  rownames(seurat_counts) = final_present_1to1s[index,]$`Gene name`
  
  gene_filt = !is.na(rownames(seurat_counts))
  seurat_counts = seurat_counts[gene_filt, ]
  
  current_seurat <- CreateSeuratObject(counts = seurat_counts, project = "current_rat", min.features = 500)
  current_seurat[["percent.mt"]] <- PercentageFeatureSet(current_seurat, pattern = "^MT-")
  current_seurat$region = 'Cortex'
  current_seurat$sample = cortex_samples[i]
  all_cortex_list[[i]] = current_seurat

}

all_cortex_list

#Go through and get the intersect genes
all_genes_list = vector(mode = 'list', length = length(all_cortex_list))
for(i in 1:length(all_cortex_list)){
  all_genes_list[[i]] = rownames(all_cortex_list[[i]])
}
final_genes = Reduce(intersect, all_genes_list)

#And go through and filter to the same gene set

for(i in 1:length(all_cortex_list)){
 
  current_seurat = all_cortex_list[[i]]
  gene_index = rownames(current_seurat) %in% final_genes
  current_seurat = current_seurat[gene_index, ]
  
  gene_index = match(final_genes, rownames(current_seurat))
  current_seurat = current_seurat[gene_index, ]
  
  all_cortex_list[[i]] = as.SingleCellExperiment(current_seurat)

}

wang_sce = do.call('cbind', all_cortex_list)
table(colData(wang_sce)$sample)



wang_seurat = as.Seurat(wang_sce, counts = "counts", data = NULL)


Idents(wang_seurat) = wang_seurat$region

VlnPlot(wang_seurat, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

```



```{r}
#CPM normalization
wang_seurat <- NormalizeData(wang_seurat, normalization.method = 'RC', scale.factor = 1e6, verbose = F)
wang_seurat <- FindVariableFeatures(wang_seurat, selection.method = "vst", nfeatures = 2000, verbose = F)
all.genes <- rownames(wang_seurat)
wang_seurat <- ScaleData(wang_seurat, features = all.genes, verbose = F)


wang_seurat <- RunPCA(wang_seurat, features = VariableFeatures(object =wang_seurat), verbose = F)


wang_seurat <- FindNeighbors(wang_seurat, dims = 1:25, verbose = F)
wang_seurat <- FindClusters(wang_seurat, verbose = F)

wang_seurat <- RunUMAP(wang_seurat, dims = 1:25, verbose = F)



DimPlot(wang_seurat, reduction = "pca", label = T)
DimPlot(wang_seurat, reduction = "umap", label = T)
DimPlot(wang_seurat, reduction = 'umap', group.by = 'sample'  )

```


```{r}

cross_primate_markers = data.table::fread('/inkwell03/werner/sc_data/primate_MTG/primate_MTG_consensus_markers.csv', skip = 4)

primate_class_markers = cross_primate_markers[ , c('gene','InN','NonN','ExN')]


table(primate_class_markers$gene %in% rownames(wang_seurat))


top_markers = list('InN' = primate_class_markers$gene[as.logical(primate_class_markers$InN)],
     'NonN' = primate_class_markers$gene[as.logical(primate_class_markers$NonN)],
     'ExN' = primate_class_markers$gene[as.logical(primate_class_markers$ExN)])



wang_sce = as.SingleCellExperiment(wang_seurat)
names(assays(wang_sce)) = c('counts','cpm','scaledata')


ct_scores = score_cells(log1p(cpm(wang_sce)), top_markers)
ct_enrichment = compute_marker_enrichment(ct_scores)
ct_pred = assign_cells(ct_scores)

ct_pred[1:10, ]


wang_seurat$glutamatergic_enrichment = ct_enrichment['all|ExN', ] / max(ct_enrichment['all|ExN', ])
wang_seurat$GABAergic_enrichment = ct_enrichment['all|InN', ] / max(ct_enrichment['all|InN', ])
wang_seurat$non_neuronal_enrichment = ct_enrichment['all|NonN', ] / max(ct_enrichment['all|NonN', ])

wang_seurat$predicted_primate_MM_class = ct_pred$predicted

```


```{r}

DimPlot(wang_seurat, reduction = "umap", label = T, group.by = 'predicted_primate_MM_class')

```

```{r}

cell_index = wang_seurat$predicted_primate_MM_class == 'InN'
wang_seurat_inhibitory = wang_seurat[ , cell_index]

wang_seurat_inhibitory

saveRDS(wang_seurat, file = '/inkwell03/werner/sc_data/rattus_brain/wang_2024/wang_seurat_full_data.rds')

saveRDS(wang_seurat_inhibitory, file = '/inkwell03/werner/sc_data/rattus_brain/wang_2024/wang_seurat_GABA.rds')


```

```{r}

wang_seurat_inhibitory = readRDS(file = '/inkwell03/werner/sc_data/rattus_brain/wang_2024/wang_seurat_GABA.rds')

wang_seurat_inhibitory 

```




New de novo clusters for just the inhibitory neurons


```{r}
#CPM normalization
wang_seurat_inhibitory <- FindVariableFeatures(wang_seurat_inhibitory, selection.method = "vst", nfeatures = 2000, verbose = F)
all.genes <- rownames(wang_seurat_inhibitory)
wang_seurat_inhibitory <- ScaleData(wang_seurat_inhibitory, features = all.genes, verbose = F)


wang_seurat_inhibitory <- RunPCA(wang_seurat_inhibitory, features = VariableFeatures(object =wang_seurat_inhibitory), verbose = F)


wang_seurat_inhibitory <- FindNeighbors(wang_seurat_inhibitory, dims = 1:25, verbose = F)
wang_seurat_inhibitory <- FindClusters(wang_seurat_inhibitory, verbose = F)

wang_seurat_inhibitory <- RunUMAP(wang_seurat_inhibitory, dims = 1:25, verbose = F)

DimPlot(wang_seurat_inhibitory, reduction = "pca", label = T)
DimPlot(wang_seurat_inhibitory, reduction = "umap", label = T, group.by = 'seurat_clusters')


```



Load up the mouse and primate CGE/MGE metamarkers

```{r}
mouse_class_meta_markers = read_meta_markers("../metamarkers/MGE_CGE/mouse_meta_markers.csv.gz")
primate_class_meta_markers = read_meta_markers("../metamarkers/MGE_CGE/primate_meta_markers.csv.gz")

mammal_class_meta_markers = read_meta_markers("../metamarkers/MGE_CGE/mammal_CGE_MGE_meta_markers.csv.gz")
```


Check out the CGE and MGE predictions

```{r}

#rownames(wang_seurat_inhibitory@assays$RNA@data) = rownames(wang_seurat_inhibitory)


#mouse class markers
top_markers = mouse_class_meta_markers %>% filter(rank <= 100)
ct_scores = score_cells(log1p(wang_seurat_inhibitory@assays$RNA@data), top_markers)
ct_enrichment = compute_marker_enrichment(ct_scores)
ct_pred = assign_cells(ct_scores)

wang_seurat_inhibitory$mouse_cge_mge = ct_pred$predicted

#Primate class markers
top_markers = primate_class_meta_markers %>% filter(rank <= 100)
ct_scores = score_cells(log1p(wang_seurat_inhibitory@assays$RNA@data), top_markers)
ct_enrichment = compute_marker_enrichment(ct_scores)
ct_pred = assign_cells(ct_scores)

wang_seurat_inhibitory$primate_cge_mge = ct_pred$predicted

#Cross_mammal class markers
top_markers = mammal_class_meta_markers %>% filter(rank <= 100)
ct_scores = score_cells(log1p(wang_seurat_inhibitory@assays$RNA@data), top_markers)
ct_enrichment = compute_marker_enrichment(ct_scores)
ct_pred = assign_cells(ct_scores)

wang_seurat_inhibitory$mammal_cge_mge = ct_pred$predicted

#Add the cross-mammal CGE-MGE enrichments for plotting
wang_seurat_inhibitory$mammal_cge_enrich = ct_enrichment['all|CGE', ]
wang_seurat_inhibitory$mammal_mge_enrich = ct_enrichment['all|MGE', ]


DimPlot(wang_seurat_inhibitory, group.by = 'seurat_clusters', label = T) + ggtitle('Initial Wang-Rat de novo GABA clusters')

DimPlot(wang_seurat_inhibitory, group.by = 'mouse_cge_mge', label = T) + ggtitle('CGE-MGE annotations using Mouse markers')

DimPlot(wang_seurat_inhibitory, group.by = 'primate_cge_mge', label = T)+ ggtitle('CGE-MGE annotations using Primate markers')

DimPlot(wang_seurat_inhibitory, group.by = 'mammal_cge_mge', label = T)+ 
  ggtitle('CGE-MGE annotations using Mammal markers') +
  scale_color_manual(values = c('CGE' = 'darkorchid', 'MGE' = 'goldenrod1', 'unassigned' = 'red'))

FeaturePlot(wang_seurat_inhibitory, features = c('mammal_cge_enrich'))+ 
  ggtitle('Mammalian CGE enrichment: wang-Rat') +
  scale_colour_gradient2(low = "white", mid = "white", high = "black", midpoint = 1)

FeaturePlot(wang_seurat_inhibitory, features = c('mammal_mge_enrich'))+ 
  ggtitle('Mammalian MGE enrichment: wang-Rat') +
  scale_colour_gradient2(low = "white", mid = "white", high = "black", midpoint = 1)

```

```{r}
mouse_class_nums = table(wang_seurat_inhibitory$mouse_cge_mge, wang_seurat_inhibitory$seurat_clusters)
total_clust_nums = colSums(mouse_class_nums)
mouse_class_ratios = t(t(mouse_class_nums) / total_clust_nums)

primate_class_nums = table(wang_seurat_inhibitory$primate_cge_mge, wang_seurat_inhibitory$seurat_clusters)
total_clust_nums = colSums(primate_class_nums)
primate_class_ratios = t(t(primate_class_nums) / total_clust_nums)


primate_class_ratios = t(primate_class_ratios)
for_ggplot_df = data.frame(CGE = primate_class_ratios[ ,'CGE'], MGE = primate_class_ratios[ ,'MGE'], cluster = rownames(primate_class_ratios))
for_ggplot_df$cluster[for_ggplot_df$CGE >= .80 | for_ggplot_df$CGE <= .20] = NA

ggplot(for_ggplot_df, aes(x = CGE, y = MGE, label = cluster)) + geom_point() +
  ggrepel::geom_label_repel(max.overlaps = Inf, min.segment.length = 0) + 
  theme_bw() + coord_fixed() + ggtitle('CGE-MGE cluster proportions: Wang-Rat')


mouse_class_ratios = t(mouse_class_ratios)
for_ggplot_df = data.frame(CGE = mouse_class_ratios[ ,'CGE'], MGE = mouse_class_ratios[ ,'MGE'], cluster = rownames(mouse_class_ratios))
for_ggplot_df$cluster[for_ggplot_df$CGE >= .80 | for_ggplot_df$CGE <= .20] = NA

ggplot(for_ggplot_df, aes(x = CGE, y = MGE, label = cluster)) + geom_point() +
  ggrepel::geom_label_repel(max.overlaps = Inf, min.segment.length = 0) + 
  theme_bw() + coord_fixed() + ggtitle('Mouse CGE-MGE cluster proportions: Wang-Rat')



ggsave(filename = 'initial_wang_rat_mouse_CGEMGE_clust_ratios_ggplot.pdf', path = '../../graphs/non_allen_data/', useDingbats = F,
       device = 'pdf', height = 4, width = 4)


```


Filter out the hybrid clusters, chandelier cells are cluster 11 so keep

```{r}

cell_index = !wang_seurat_inhibitory$seurat_clusters %in% c('1','10','12','13','16','17','18')
wang_seurat_inhibitory = wang_seurat_inhibitory[, cell_index]


wang_seurat_inhibitory <- RunPCA(wang_seurat_inhibitory, features = VariableFeatures(object =wang_seurat_inhibitory), verbose = F)

wang_seurat_inhibitory <- FindNeighbors(wang_seurat_inhibitory, dims = 1:25, verbose = F)
wang_seurat_inhibitory <- FindClusters(wang_seurat_inhibitory, verbose = F, resolution = .9)

wang_seurat_inhibitory <- RunUMAP(wang_seurat_inhibitory, dims = 1:25, verbose = F)

DimPlot(wang_seurat_inhibitory, reduction = "pca", label = T)
DimPlot(wang_seurat_inhibitory, reduction = "umap", label = T, group.by = 'seurat_clusters')


```



```{r}

top_allen_markers = primate_subclass_meta_markers %>% filter(rank <= 50)
top_mouse_markers = mouse_subclass_meta_markers %>% filter(rank <= 50)
top_ori_mouse_markers = original_mouse_subclass_meta_markers %>% filter(rank <= 50)

#rownames(wang_seurat_inhibitory@assays$RNA@layers$data) = rownames(wang_seurat_inhibitory)

ct_scores = score_cells(log1p(wang_seurat_inhibitory@assays$RNA@data), top_allen_markers)
ct_enrichment = compute_marker_enrichment(ct_scores)
ct_pred = assign_cells(ct_scores)
wang_seurat_inhibitory$primate_subclass_annot = ct_pred$predicted

ct_scores = score_cells(log1p(wang_seurat_inhibitory@assays$RNA@data), top_mouse_markers)
ct_enrichment = compute_marker_enrichment(ct_scores)
ct_pred = assign_cells(ct_scores)
wang_seurat_inhibitory$mouse_subclass_annot = ct_pred$predicted

ct_scores = score_cells(log1p(wang_seurat_inhibitory@assays$RNA@data), top_ori_mouse_markers)
ct_enrichment = compute_marker_enrichment(ct_scores)
ct_pred = assign_cells(ct_scores)
wang_seurat_inhibitory$ori_mouse_subclass_annot = ct_pred$predicted

DimPlot(wang_seurat_inhibitory, group.by = 'seurat_clusters', label = T) + ggtitle('Initial Wang-Rat de novo GABA clusters')

DimPlot(wang_seurat_inhibitory, group.by = 'mouse_cge_mge', label = T) + ggtitle('CGE-MGE annotations using Mouse markers')

DimPlot(wang_seurat_inhibitory, group.by = 'primate_cge_mge', label = T)+ ggtitle('CGE-MGE annotations using Primate markers')

DimPlot(wang_seurat_inhibitory, group.by = 'primate_subclass_annot', label = T) + ggtitle('subclass annotations using Primate markers')

DimPlot(wang_seurat_inhibitory, group.by = 'mouse_subclass_annot', label = T)+ ggtitle('subclass annotations using Mouse markers')

DimPlot(wang_seurat_inhibitory, group.by = 'ori_mouse_subclass_annot', label = T)+ ggtitle('subclass annotations using Mouse markers')


```

```{r}

DimPlot(wang_seurat_inhibitory, group.by = 'seurat_clusters', label = T) + ggtitle('Wang-Rat de novo GABA clusters') 
ggsave(filename = 'wang_rat_de_novo_clusters_umap.pdf', path = '../../graphs/non_allen_data/umaps/', device = 'pdf',
       useDingbats = F, height = 4, width = 6)

DimPlot(wang_seurat_inhibitory, group.by = 'mouse_subclass_annot')+ ggtitle('subclass annotations using homologous Mouse markers') + scale_color_manual(values = celltype_color_palette)
ggsave(filename = 'Wang_rat_homo_mouse_predict_umap.pdf', path = '../../graphs/non_allen_data/umaps/', device = 'pdf',
       useDingbats = F, height = 4, width = 6)

DimPlot(wang_seurat_inhibitory, group.by = 'ori_mouse_subclass_annot')+ ggtitle('subclass annotations using original Mouse markers') + scale_color_manual(values = celltype_color_palette)
ggsave(filename = 'Wang_rat_ori_mouse_predict_umap.pdf', path = '../../graphs/non_allen_data/umaps/', device = 'pdf',
       useDingbats = F, height = 4, width = 6)


```


Save with the de novo clusters and all the annotations
```{r}
saveRDS(wang_seurat_inhibitory, file = '/inkwell03/werner/sc_data/rattus_brain/wang_2024/wang_seurat_GABA_cortex_annotated.rds')

```




```{r}


confuse_mat_wang_rat = table(wang_seurat_inhibitory$ori_mouse_subclass_annot, wang_seurat_inhibitory$seurat_clusters)
cluster_totals_wang_rat = colSums(confuse_mat_wang_rat )
confuse_mat_wang_rat = t(t(confuse_mat_wang_rat)/cluster_totals_wang_rat)

#Ordering by CGE and MGE
row_order = match(c('Sst','Sst Chodl','Pvalb','Chandelier','Lamp5','Lamp5 Lhx6','Vip','Sncg', 'unassigned'), rownames(confuse_mat_wang_rat))
#round(confuse_mat_wang_rat)[row_order, ]

col_order = c(1,5,11,0,2,7,10,9,4,6,3,12,8) +1

col_fun = circlize::colorRamp2(breaks = c(0,.5, 1) , colors = c('white','orange','red'))
hm = ComplexHeatmap::Heatmap(confuse_mat_wang_rat,name = 'cluster proportion',
                        col = col_fun,
                        cluster_rows = F, row_order = row_order,
                        cluster_columns = F, column_order = col_order,
                        column_title = 'Wang-rat cluster predictions with original mouse markers')
hm = ComplexHeatmap::draw(hm)

pdf(file = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/non_allen_data/ori_mouse_MM_predict_wang_rat_clusters.pdf', 
    useDingbats = F, height = 4, width = 8)
hm
dev.off()


confuse_mat_wang_rat = table(wang_seurat_inhibitory$mouse_subclass_annot, wang_seurat_inhibitory$seurat_clusters)
cluster_totals_wang_rat = colSums(confuse_mat_wang_rat )
confuse_mat_wang_rat = t(t(confuse_mat_wang_rat)/cluster_totals_wang_rat)

#Ordering by CGE and MGE
row_order = match(c('Sst','Sst Chodl','Pvalb','Chandelier','Lamp5','Lamp5 Lhx6','Vip','Sncg', 'Pax6', 'unassigned'), rownames(confuse_mat_wang_rat))
#round(confuse_mat_wang_rat)[row_order, ]
col_order = c(1,5,11,0,2,7,10,9,4,6,3,12,8) +1

col_fun = circlize::colorRamp2(breaks = c(0,.5, 1) , colors = c('white','orange','red'))
hm = ComplexHeatmap::Heatmap(confuse_mat_wang_rat,name = 'cluster proportion',
                        col = col_fun,
                        cluster_rows = F, row_order = row_order,
                        cluster_columns = F, column_order = col_order,
                        column_title = 'wang-rat cluster predictions with homologous mouse markers')
hm = ComplexHeatmap::draw(hm)

pdf(file = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/non_allen_data/homo_mouse_MM_predict_wang_rat_clusters.pdf', 
    useDingbats = F, height = 4, width = 8)
hm
dev.off()


```



And just see what the hippocampal samples look like

```{r}

wang_path = '/inkwell03/werner/sc_data/rattus_brain/wang_2024'
wang_directories = list.dirs(wang_path, recursive = F)

hippo_directories = wang_directories[grepl('HP', wang_directories)]

hippo_samples = sapply(strsplit(hippo_directories, split = '/', fixed = T), '[[', 7)
hippo_samples = sapply(strsplit(hippo_samples, split = '_', fixed = T), '[[', 1)


all_hippo_list = vector(mode = 'list', length = length(hippo_samples))
for(i in 1:length(hippo_samples)){

  seurat_counts = Read10X(hippo_directories[[i]])
  
  #Filter for the 1to1s and swap in the human gene names
  index = match(rownames(seurat_counts), final_present_1to1s$`Mouse gene name`)
  rownames(seurat_counts) = final_present_1to1s[index,]$`Gene name`
  
  gene_filt = !is.na(rownames(seurat_counts))
  seurat_counts = seurat_counts[gene_filt, ]
  
  current_seurat <- CreateSeuratObject(counts = seurat_counts, project = "current_rat", min.features = 500)
  current_seurat[["percent.mt"]] <- PercentageFeatureSet(current_seurat, pattern = "^MT-")
  current_seurat$region = 'hippo'
  current_seurat$sample = hippo_samples[i]
  all_hippo_list[[i]] = current_seurat

}

all_hippo_list

#Go through and get the intersect genes
all_genes_list = vector(mode = 'list', length = length(all_hippo_list))
for(i in 1:length(all_hippo_list)){
  all_genes_list[[i]] = rownames(all_hippo_list[[i]])
}
final_genes = Reduce(intersect, all_genes_list)

#And go through and filter to the same gene set

for(i in 1:length(all_hippo_list)){
 
  current_seurat = all_hippo_list[[i]]
  gene_index = rownames(current_seurat) %in% final_genes
  current_seurat = current_seurat[gene_index, ]
  
  gene_index = match(final_genes, rownames(current_seurat))
  current_seurat = current_seurat[gene_index, ]
  
  all_hippo_list[[i]] = as.SingleCellExperiment(current_seurat)

}

wang_sce = do.call('cbind', all_hippo_list)
table(colData(wang_sce)$sample)



wang_seurat = as.Seurat(wang_sce, counts = "counts", data = NULL)


Idents(wang_seurat) = wang_seurat$region

VlnPlot(wang_seurat, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

```

```{r}
#CPM normalization
wang_seurat <- NormalizeData(wang_seurat, normalization.method = 'RC', scale.factor = 1e6, verbose = F)
wang_seurat <- FindVariableFeatures(wang_seurat, selection.method = "vst", nfeatures = 2000, verbose = F)
all.genes <- rownames(wang_seurat)
wang_seurat <- ScaleData(wang_seurat, features = all.genes, verbose = F)


wang_seurat <- RunPCA(wang_seurat, features = VariableFeatures(object =wang_seurat), verbose = F)


wang_seurat <- FindNeighbors(wang_seurat, dims = 1:25, verbose = F)
wang_seurat <- FindClusters(wang_seurat, verbose = F)

wang_seurat <- RunUMAP(wang_seurat, dims = 1:25, verbose = F)



DimPlot(wang_seurat, reduction = "pca", label = T)
DimPlot(wang_seurat, reduction = "umap", label = T)
DimPlot(wang_seurat, reduction = 'umap', group.by = 'sample'  )

```




```{r}
top_markers = list('InN' = primate_class_markers$gene[as.logical(primate_class_markers$InN)],
     'NonN' = primate_class_markers$gene[as.logical(primate_class_markers$NonN)],
     'ExN' = primate_class_markers$gene[as.logical(primate_class_markers$ExN)])

wang_sce = as.SingleCellExperiment(wang_seurat)
names(assays(wang_sce)) = c('counts','cpm','scaledata')

ct_scores = score_cells(log1p(cpm(wang_sce)), top_markers)
ct_enrichment = compute_marker_enrichment(ct_scores)
ct_pred = assign_cells(ct_scores)

dim(wang_seurat)
ct_enrichment[ , 1:10]

wang_seurat$glutamatergic_enrichment = ct_enrichment['all|ExN', ] / max(ct_enrichment['all|ExN', ])
wang_seurat$GABAergic_enrichment = ct_enrichment['all|InN', ] / max(ct_enrichment['all|InN', ])
wang_seurat$non_neuronal_enrichment = ct_enrichment['all|NonN', ] / max(ct_enrichment['all|NonN', ])

wang_seurat$predicted_primate_MM_class = ct_pred$predicted


DimPlot(wang_seurat, reduction = "umap", label = T, group.by = 'predicted_primate_MM_class')

```




```{r}

cell_index = wang_seurat$predicted_primate_MM_class == 'InN'
wang_seurat_inhibitory = wang_seurat[ , cell_index]

wang_seurat_inhibitory

saveRDS(wang_seurat, file = '/inkwell03/werner/sc_data/rattus_brain/wang_2024/wang_seurat_full_data_hippo.rds')

saveRDS(wang_seurat_inhibitory, file = '/inkwell03/werner/sc_data/rattus_brain/wang_2024/wang_seurat_GABA_hippo.rds')


```

New de novo clusters for just the inhibitory neurons


```{r}
#CPM normalization
wang_seurat_inhibitory <- FindVariableFeatures(wang_seurat_inhibitory, selection.method = "vst", nfeatures = 2000, verbose = F)
all.genes <- rownames(wang_seurat_inhibitory)
wang_seurat_inhibitory <- ScaleData(wang_seurat_inhibitory, features = all.genes, verbose = F)


wang_seurat_inhibitory <- RunPCA(wang_seurat_inhibitory, features = VariableFeatures(object =wang_seurat_inhibitory), verbose = F)


wang_seurat_inhibitory <- FindNeighbors(wang_seurat_inhibitory, dims = 1:25, verbose = F)
wang_seurat_inhibitory <- FindClusters(wang_seurat_inhibitory, verbose = F)

wang_seurat_inhibitory <- RunUMAP(wang_seurat_inhibitory, dims = 1:25, verbose = F)

DimPlot(wang_seurat_inhibitory, reduction = "pca", label = T)
DimPlot(wang_seurat_inhibitory, reduction = "umap", label = T, group.by = 'seurat_clusters')


```


Check out the CGE and MGE predictions

```{r}

#rownames(wang_seurat_inhibitory@assays$RNA@data) = rownames(wang_seurat_inhibitory)


#mouse class markers
top_markers = mouse_class_meta_markers %>% filter(rank <= 100)
ct_scores = score_cells(log1p(wang_seurat_inhibitory@assays$RNA@data), top_markers)
ct_enrichment = compute_marker_enrichment(ct_scores)
ct_pred = assign_cells(ct_scores)

wang_seurat_inhibitory$mouse_cge_mge = ct_pred$predicted

#Primate class markers
top_markers = primate_class_meta_markers %>% filter(rank <= 100)
ct_scores = score_cells(log1p(wang_seurat_inhibitory@assays$RNA@data), top_markers)
ct_enrichment = compute_marker_enrichment(ct_scores)
ct_pred = assign_cells(ct_scores)

wang_seurat_inhibitory$primate_cge_mge = ct_pred$predicted

#Cross_mammal class markers
top_markers = mammal_class_meta_markers %>% filter(rank <= 100)
ct_scores = score_cells(log1p(wang_seurat_inhibitory@assays$RNA@data), top_markers)
ct_enrichment = compute_marker_enrichment(ct_scores)
ct_pred = assign_cells(ct_scores)

wang_seurat_inhibitory$mammal_cge_mge = ct_pred$predicted

#Add the cross-mammal CGE-MGE enrichments for plotting
wang_seurat_inhibitory$mammal_cge_enrich = ct_enrichment['all|CGE', ]
wang_seurat_inhibitory$mammal_mge_enrich = ct_enrichment['all|MGE', ]


DimPlot(wang_seurat_inhibitory, group.by = 'seurat_clusters', label = T) + ggtitle('Initial Wang-Rat de novo GABA clusters')

DimPlot(wang_seurat_inhibitory, group.by = 'mouse_cge_mge', label = T) + ggtitle('CGE-MGE annotations using Mouse markers')

DimPlot(wang_seurat_inhibitory, group.by = 'primate_cge_mge', label = T)+ ggtitle('CGE-MGE annotations using Primate markers')

DimPlot(wang_seurat_inhibitory, group.by = 'mammal_cge_mge', label = T)+ 
  ggtitle('CGE-MGE annotations using Mammal markers') +
  scale_color_manual(values = c('CGE' = 'darkorchid', 'MGE' = 'goldenrod1', 'unassigned' = 'red'))

FeaturePlot(wang_seurat_inhibitory, features = c('mammal_cge_enrich'))+ 
  ggtitle('Mammalian CGE enrichment: wang-Rat') +
  scale_colour_gradient2(low = "white", mid = "white", high = "black", midpoint = 1)

FeaturePlot(wang_seurat_inhibitory, features = c('mammal_mge_enrich'))+ 
  ggtitle('Mammalian MGE enrichment: wang-Rat') +
  scale_colour_gradient2(low = "white", mid = "white", high = "black", midpoint = 1)

```




```{r}
mouse_class_nums = table(wang_seurat_inhibitory$mouse_cge_mge, wang_seurat_inhibitory$seurat_clusters)
total_clust_nums = colSums(mouse_class_nums)
mouse_class_ratios = t(t(mouse_class_nums) / total_clust_nums)

primate_class_nums = table(wang_seurat_inhibitory$primate_cge_mge, wang_seurat_inhibitory$seurat_clusters)
total_clust_nums = colSums(primate_class_nums)
primate_class_ratios = t(t(primate_class_nums) / total_clust_nums)


primate_class_ratios = t(primate_class_ratios)
for_ggplot_df = data.frame(CGE = primate_class_ratios[ ,'CGE'], MGE = primate_class_ratios[ ,'MGE'], cluster = rownames(primate_class_ratios))
for_ggplot_df$cluster[for_ggplot_df$CGE >= .8 | for_ggplot_df$CGE <= .2] = NA

ggplot(for_ggplot_df, aes(x = CGE, y = MGE, label = cluster)) + geom_point() +
  ggrepel::geom_label_repel(max.overlaps = Inf, min.segment.length = 0) + 
  theme_bw() + coord_fixed() + ggtitle('CGE-MGE cluster proportions: Wang-Rat')


mouse_class_ratios = t(mouse_class_ratios)
for_ggplot_df = data.frame(CGE = mouse_class_ratios[ ,'CGE'], MGE = mouse_class_ratios[ ,'MGE'], cluster = rownames(mouse_class_ratios))
for_ggplot_df$cluster[for_ggplot_df$CGE >= .8 | for_ggplot_df$CGE <= .2] = NA

ggplot(for_ggplot_df, aes(x = CGE, y = MGE, label = cluster)) + geom_point() +
  ggrepel::geom_label_repel(max.overlaps = Inf, min.segment.length = 0) + 
  theme_bw() + coord_fixed() + ggtitle('Mouse CGE-MGE cluster proportions: Wang-Rat')



ggsave(filename = 'initial_wang_rat_mouse_CGEMGE_hippo_clust_ratios_ggplot.pdf', path = '../../graphs/non_allen_data/', useDingbats = F,
       device = 'pdf', height = 4, width = 4)


```
cluster 11 is chandelier cells
```{r}

cell_index = !wang_seurat_inhibitory$seurat_clusters %in% c('1','3','9','13','14','15','16', '17')
wang_seurat_inhibitory = wang_seurat_inhibitory[, cell_index]


wang_seurat_inhibitory <- RunPCA(wang_seurat_inhibitory, features = VariableFeatures(object =wang_seurat_inhibitory), verbose = F)

wang_seurat_inhibitory <- FindNeighbors(wang_seurat_inhibitory, dims = 1:25, verbose = F)
wang_seurat_inhibitory <- FindClusters(wang_seurat_inhibitory, verbose = F)

wang_seurat_inhibitory <- RunUMAP(wang_seurat_inhibitory, dims = 1:25, verbose = F)

DimPlot(wang_seurat_inhibitory, reduction = "pca", label = T)
DimPlot(wang_seurat_inhibitory, reduction = "umap", label = T, group.by = 'seurat_clusters')


```



```{r}

top_allen_markers = primate_subclass_meta_markers %>% filter(rank <= 50)
top_mouse_markers = mouse_subclass_meta_markers %>% filter(rank <= 50)
top_ori_mouse_markers = original_mouse_subclass_meta_markers %>% filter(rank <= 50)

#rownames(wang_seurat_inhibitory@assays$RNA@layers$data) = rownames(wang_seurat_inhibitory)

ct_scores = score_cells(log1p(wang_seurat_inhibitory@assays$RNA@data), top_allen_markers)
ct_enrichment = compute_marker_enrichment(ct_scores)
ct_pred = assign_cells(ct_scores)
wang_seurat_inhibitory$primate_subclass_annot = ct_pred$predicted

ct_scores = score_cells(log1p(wang_seurat_inhibitory@assays$RNA@data), top_mouse_markers)
ct_enrichment = compute_marker_enrichment(ct_scores)
ct_pred = assign_cells(ct_scores)
wang_seurat_inhibitory$mouse_subclass_annot = ct_pred$predicted

ct_scores = score_cells(log1p(wang_seurat_inhibitory@assays$RNA@data), top_ori_mouse_markers)
ct_enrichment = compute_marker_enrichment(ct_scores)
ct_pred = assign_cells(ct_scores)
wang_seurat_inhibitory$ori_mouse_subclass_annot = ct_pred$predicted

DimPlot(wang_seurat_inhibitory, group.by = 'seurat_clusters', label = T) + ggtitle('Initial Wang-Rat de novo GABA clusters')

DimPlot(wang_seurat_inhibitory, group.by = 'mouse_cge_mge', label = T) + ggtitle('CGE-MGE annotations using Mouse markers')

DimPlot(wang_seurat_inhibitory, group.by = 'primate_cge_mge', label = T)+ ggtitle('CGE-MGE annotations using Primate markers')

DimPlot(wang_seurat_inhibitory, group.by = 'primate_subclass_annot', label = T) + ggtitle('subclass annotations using Primate markers')

DimPlot(wang_seurat_inhibitory, group.by = 'mouse_subclass_annot', label = T)+ ggtitle('subclass annotations using Mouse markers')

DimPlot(wang_seurat_inhibitory, group.by = 'ori_mouse_subclass_annot', label = T)+ ggtitle('subclass annotations using Mouse markers')


```


```{r}

DimPlot(wang_seurat_inhibitory, group.by = 'seurat_clusters', label = T) + ggtitle('Wang-Rat de novo GABA clusters') 
ggsave(filename = 'wang_rat_de_novo_clusters_hippo_umap.pdf', path = '../../graphs/non_allen_data/umaps/', device = 'pdf',
       useDingbats = F, height = 4, width = 6)

DimPlot(wang_seurat_inhibitory, group.by = 'mouse_subclass_annot')+ ggtitle('subclass annotations using homologous Mouse markers') + scale_color_manual(values = celltype_color_palette)
ggsave(filename = 'Wang_rat_homo_mouse_predict_hippo_umap.pdf', path = '../../graphs/non_allen_data/umaps/', device = 'pdf',
       useDingbats = F, height = 4, width = 6)

DimPlot(wang_seurat_inhibitory, group.by = 'ori_mouse_subclass_annot')+ ggtitle('subclass annotations using original Mouse markers') + scale_color_manual(values = celltype_color_palette)
ggsave(filename = 'Wang_rat_ori_mouse_predict_hippo_umap.pdf', path = '../../graphs/non_allen_data/umaps/', device = 'pdf',
       useDingbats = F, height = 4, width = 6)

```





```{r}

confuse_mat_wang_rat = table(wang_seurat_inhibitory$ori_mouse_subclass_annot, wang_seurat_inhibitory$seurat_clusters)
cluster_totals_wang_rat = colSums(confuse_mat_wang_rat )
confuse_mat_wang_rat = t(t(confuse_mat_wang_rat)/cluster_totals_wang_rat)

#Ordering by CGE and MGE
row_order = match(c('Sst','Sst Chodl','Pvalb','Chandelier','Lamp5','Lamp5 Lhx6','Vip','Sncg', 'unassigned'), rownames(confuse_mat_wang_rat))
#round(confuse_mat_wang_rat)[row_order, ]
col_order = c(0,1,9,10,8,6,4,7,5,3,2) +1

col_fun = circlize::colorRamp2(breaks = c(0,.5, 1) , colors = c('white','orange','red'))
hm = ComplexHeatmap::Heatmap(confuse_mat_wang_rat,name = 'cluster proportion',
                        col = col_fun,
                        cluster_rows = F, row_order = row_order,
                        cluster_columns = F, column_order = col_order,
                        column_title = 'Wang-rat cluster predictions with original mouse markers')
hm = ComplexHeatmap::draw(hm)

pdf(file = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/non_allen_data/ori_mouse_MM_predict_wang_rat_hippo_clusters.pdf', 
    useDingbats = F, height = 4, width = 8)
hm
dev.off()


confuse_mat_wang_rat = table(wang_seurat_inhibitory$mouse_subclass_annot, wang_seurat_inhibitory$seurat_clusters)
cluster_totals_wang_rat = colSums(confuse_mat_wang_rat )
confuse_mat_wang_rat = t(t(confuse_mat_wang_rat)/cluster_totals_wang_rat)

#Ordering by CGE and MGE
row_order = match(c('Sst','Sst Chodl','Pvalb','Chandelier','Lamp5','Lamp5 Lhx6','Vip','Sncg', 'Pax6', 'unassigned'), rownames(confuse_mat_wang_rat))
#round(confuse_mat_wang_rat)[row_order, ]
col_order = c(0,1,9,10,8,6,4,7,5,3,2) +1

col_fun = circlize::colorRamp2(breaks = c(0,.5, 1) , colors = c('white','orange','red'))
hm = ComplexHeatmap::Heatmap(confuse_mat_wang_rat,name = 'cluster proportion',
                        col = col_fun,
                        cluster_rows = F, row_order = row_order,
                        cluster_columns = F, column_order = col_order,
                        column_title = 'wang-rat cluster predictions with homologous mouse markers')
hm = ComplexHeatmap::draw(hm)

pdf(file = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/non_allen_data/homo_mouse_MM_predict_wang_rat_hippo_clusters.pdf', 
    useDingbats = F, height = 4, width = 8)
hm
dev.off()


```



###################################

Probably should do a mouse MetaNeighbor, focus on the Rat clusters that replicate in the mouse datasets

##################################

```{r}
aydin_seurat_inhibitory = readRDS(file = '/inkwell03/werner/sc_data/rattus_brain/aydin_2024/aydin_seurat_GABA_annotated.rds')

wang_seurat_inhibitory = readRDS( file = '/inkwell03/werner/sc_data/rattus_brain/wang_2024/wang_seurat_GABA_cortex_annotated.rds')


aydin_sce_inhibitory = as.SingleCellExperiment(aydin_seurat_inhibitory)
names(assays(aydin_sce_inhibitory)) = c('counts','cpm','scaledata')

aydin_sce_inhibitory$cluster_label = aydin_sce_inhibitory$seurat_clusters
aydin_sce_inhibitory$subclass_label = NA
aydin_sce_inhibitory$study = 'rat_aydin'

wang_sce_inhibitory = as.SingleCellExperiment(wang_seurat_inhibitory)
names(assays(wang_sce_inhibitory)) = c('counts','cpm','scaledata')

wang_sce_inhibitory$cluster_label = wang_sce_inhibitory$seurat_clusters
wang_sce_inhibitory$subclass_label = NA
wang_sce_inhibitory$study = 'rat_wang'

```
These are the mouse clusters that are replicable across the mouse datasets


```{r}

load('/home/werner/projects/mouse_brain_consensus_taxonomy/data/metaneighbor/mouse_MN_results.Rdata')

keep_mouse_clusters = unname(unlist(mouse_mclusters))

```


Load up the full mouse datasets and filter down just to the replicable mouse clusters
Add in the ratdata, using the de novo clusters, for metaNeighbor


```{r}

tasic18 = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/tasic18_gaba.rds')

#Change metdata columns to match other datasets
tasic18$subclass_label = tasic18$cell_subclass
tasic18$cluster_label = tasic18$cell_cluster

keep_meta = colData(tasic18)[ , c('subclass_label','cluster_label')]
colData(tasic18)= keep_meta

#keep the replicable mouse clusters
tasic18$study = 'tasic18'
dataset_cluster_labels = paste(tasic18$study, tasic18$cluster_label, sep = '|')
tasic18 = tasic18[, dataset_cluster_labels %in% keep_mouse_clusters]
tasic18

```


```{r}

zeng_10x_cells = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/zeng_10x_cells_gaba.rds')

keep_meta = colData(zeng_10x_cells)[ , c('subclass_label','cluster_label')]
colData(zeng_10x_cells)= keep_meta

#keep the replicable mouse clusters
zeng_10x_cells$study = 'zeng_10x_cells'
dataset_cluster_labels = paste(zeng_10x_cells$study, zeng_10x_cells$cluster_label, sep = '|')
zeng_10x_cells = zeng_10x_cells[, dataset_cluster_labels %in% keep_mouse_clusters]
zeng_10x_cells

```


```{r}

zeng_10x_nuclei = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/zeng_10x_nuclei_gaba.rds')

keep_meta = colData(zeng_10x_nuclei)[ , c('subclass_label','cluster_label')]
colData(zeng_10x_nuclei)= keep_meta

#keep the replicable mouse clusters
zeng_10x_nuclei$study = 'zeng_10x_nuclei'
dataset_cluster_labels = paste(zeng_10x_nuclei$study, zeng_10x_nuclei$cluster_label, sep = '|')
zeng_10x_nuclei = zeng_10x_nuclei[, dataset_cluster_labels %in% keep_mouse_clusters]
zeng_10x_nuclei



```

```{r}

zeng_smart_cells = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/zeng_smart_cells_gaba.rds')

keep_meta = colData(zeng_smart_cells)[ , c('subclass_label','cluster_label')]
colData(zeng_smart_cells)= keep_meta

#keep the replicable mouse clusters
zeng_smart_cells$study = 'zeng_smart_cells'
dataset_cluster_labels = paste(zeng_smart_cells$study, zeng_smart_cells$cluster_label, sep = '|')
zeng_smart_cells = zeng_smart_cells[, dataset_cluster_labels %in% keep_mouse_clusters]
zeng_smart_cells



```

```{r}

zeng_smart_nuclei = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/zeng_smart_nuclei_gaba.rds')

keep_meta = colData(zeng_smart_nuclei)[ , c('subclass_label','cluster_label')]
colData(zeng_smart_nuclei)= keep_meta

#keep the replicable mouse clusters
zeng_smart_nuclei$study = 'zeng_smart_nuclei'
dataset_cluster_labels = paste(zeng_smart_nuclei$study, zeng_smart_nuclei$cluster_label, sep = '|')
zeng_smart_nuclei = zeng_smart_nuclei[, dataset_cluster_labels %in% keep_mouse_clusters]
zeng_smart_nuclei



```

Whole brain atlases
```{r}

zeng_10xv2_wba = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/zeng_wb_atlas_10xv2_gaba.rds')

zeng_10xv2_wba$cluster_label = zeng_10xv2_wba$supertype_label

keep_meta = colData(zeng_10xv2_wba)[ , c('subclass_label','cluster_label')]
colData(zeng_10xv2_wba)= keep_meta

#keep the replicable mouse clusters
zeng_10xv2_wba$study = 'zeng_10xv2_wba'
dataset_cluster_labels = paste(zeng_10xv2_wba$study, zeng_10xv2_wba$cluster_label, sep = '|')
zeng_10xv2_wba = zeng_10xv2_wba[, dataset_cluster_labels %in% keep_mouse_clusters]
zeng_10xv2_wba


```

```{r}

zeng_10xv3_wba = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/zeng_wb_atlas_10xv3_gaba.rds')

zeng_10xv3_wba$cluster_label = zeng_10xv3_wba$supertype_label

keep_meta = colData(zeng_10xv3_wba)[ , c('subclass_label','cluster_label')]
colData(zeng_10xv3_wba)= keep_meta

#keep the replicable mouse clusters
zeng_10xv3_wba$study = 'zeng_10xv3_wba'
dataset_cluster_labels = paste(zeng_10xv3_wba$study, zeng_10xv3_wba$cluster_label, sep = '|')
zeng_10xv3_wba = zeng_10xv3_wba[, dataset_cluster_labels %in% keep_mouse_clusters]
zeng_10xv3_wba


```


```{r}

macosko_wba = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/macosko_wb_atlas_gaba.rds')

#No subclass labels
macosko_wba$cluster_label = macosko_wba$cell.type
macosko_wba$subclass_label = rep(NA, length = ncol(macosko_wba))

#keep the replicable mouse clusters
macosko_wba$study = 'macosko_wba'
dataset_cluster_labels = paste(macosko_wba$study, macosko_wba$cluster_label, sep = '|')
macosko_wba = macosko_wba[, dataset_cluster_labels %in% keep_mouse_clusters]
macosko_wba


```


```{r}

names(assays(zeng_10x_cells)) = c('counts','cpm')
names(assays(zeng_10x_nuclei)) = c('counts','cpm')
names(assays(zeng_smart_cells)) = c('counts','cpm')
names(assays(zeng_smart_nuclei)) = c('counts','cpm')
names(assays(zeng_10xv2_wba)) = c('counts','cpm')
names(assays(zeng_10xv3_wba)) = c('counts','cpm')


all_rodent_sce = list(
  tasic18 = tasic18,
  zeng_10x_cells = zeng_10x_cells,
  zeng_10x_nuclei = zeng_10x_nuclei,
  zeng_smart_cells = zeng_smart_cells,
  zeng_smart_nuclei = zeng_smart_nuclei,
  zeng_10xv2_wba = zeng_10xv2_wba,
  zeng_10xv3_wba = zeng_10xv3_wba,
  macosko_wba = macosko_wba,
  rat_aydin = aydin_sce_inhibitory,
  rat_wang = wang_sce_inhibitory
)



```


```{r}
lapply(all_rodent_sce, function(x) head(rownames(x), 3))

```

```{r}
lapply(all_rodent_sce, function(x) colnames(colData(x)))

```

```{r}
library(MetaNeighbor)

```


```{r}

merged_rodent_sce = mergeSCE(all_rodent_sce)
dim(merged_rodent_sce)

```


Get rid of the list of sce objects for space
```{r}

rm(all_rodent_sce)
gc()
```


```{r}

head(colData(merged_rodent_sce))

```

Run MetaNeighbor

Get highly variable genes
```{r}
start = Sys.time()
global_hvgs = variableGenes(dat = merged_rodent_sce, exp_labels = merged_rodent_sce$study_id, min_recurrence = 5)
length(global_hvgs)
Sys.time() - start
```


Just use the top 1000 hvgs
```{r}
global_hvgs = global_hvgs[1:1000]

```

And now the one-vs-best cell-types

```{r}
start = Sys.time()
rodent_best_hits = MetaNeighborUS(var_genes = global_hvgs, 
                              dat = merged_rodent_sce, 
                              study_id = merged_rodent_sce$study_id,
                              cell_type = merged_rodent_sce$cluster_label,
                              fast_version = T,
                              one_vs_best = T, symmetric_output = F)


Sys.time() - start
gc()



pdf(file = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/metaneighbor_plots/all_within_rodent_data_1vsbest.pdf', 
    useDingbats = F, height = 25, width = 25)
plotHeatmap(rodent_best_hits, cex = .2)
dev.off()



```

```{r}
rodent_mclusters = extractMetaClusters(rodent_best_hits, threshold = .7)
rodent_mclusters

```


save the rodent metaneighbor results
```{r}

save(rodent_best_hits , rodent_mclusters,
     file = '/home/werner/projects/mouse_brain_consensus_taxonomy/data/metaneighbor/rodent_MN_results.Rdata')


```



Non-replicable rat clusters
"rat_aydin|1"  "rat_aydin|14" "rat_aydin|4"  "rat_wang|0"   "rat_wang|10"  "rat_wang|7" 
```{r}
rodent_mclusters$outliers[grepl('rat', rodent_mclusters$outliers)]

```


```{r}

table(aydin_sce_inhibitory$cluster_label, aydin_sce_inhibitory$mouse_subclass_annot)

```












