
Checking out the differential expression of specific GO terms across the subclasses

Focus on the Ion channels, as proxies for electrophysiological properties of the subclasses



GO database
SynGO database
KEGG database


```{r}

library(dplyr)
library(tidyr)
library(SingleCellExperiment)
library(ggplot2)
library(org.Hs.eg.db)
library(GO.db)
library(ComplexHeatmap)

```


```{r}
celltype_color_palette = c("Chandelier"='#F641A8', 
                           "Pvalb"='#D93137', 
                           "Sst"='#FF9900', 
                           "Sst Chodl"='gold', 
                           "Lamp5"='#DA808C',
                           "Sncg"='#DF70FF',
                           "Vip"='#A45FBF', 
                           "Lamp5 Lhx6"='#935F50',
                           "Pax6"='#71238C')


```


SynGo Terms of interest

synapse adhesion between pre- and post-synapse
presynaptic modulation of chemical synaptic transmission
postsynaptic modulation of chemical synaptic transmission
transmitter-gated ion channel activity involved in regulation of postsynaptic membrane potential 
ligand-gated ion channel activity involved in regulation of presynaptic membrane potential
voltage-gated ion channel activity involved in regulation of presynaptic membrane potential 

```{r}
#synGo_path = '/home/werner/tools/syn_go/SynGO_bulk_download_release_20231201/'

synGo_annots_df = data.table::fread('/home/werner/tools/syn_go/SynGO_bulk_download_release_20231201/syngo_annotations.csv')

synGo_annots_df %>% filter(grepl('presynaptic membrane', synGo_annots_df$go_name)) %>% group_by(go_name) %>% summarise(n = n())

```


Go terms of interest


```{r}

org.Hs.eg.db #running this will print out the GO database version information
go_table <- as.data.frame(org.Hs.egGO2ALLEGS) #gene mapping to the direct GO term and that term's children
go_table$gene_symbol = mapIds(org.Hs.eg.db, go_table$gene_id, "SYMBOL","ENTREZID") #Add gene symbol 

go_table %>% filter(go_id == 'GO:0000009') %>% filter(!duplicated(gene_symbol))

```

```{r}


# Get all GO IDs in the database
all_ids <- keys(GO.db, keytype = "GOID")

# Get their term names (descriptions)
all_terms <- Term(all_ids)

# Get ontology type (BP, MF, CC)
all_ont <- Ontology(all_ids)

# Optional: get definitions
all_defs <- Definition(all_ids)


# 2. Map GO terms to genes
go2gene <- AnnotationDbi::as.list(org.Hs.egGO2ALLEGS)  # GOID -> list of Entrez IDs

# 3. Count how many genes per GO term
gene_counts <- sapply(all_ids, function(go) {
  genes <- go2gene[[go]]
  if (is.null(genes)) return(0)
  length(unique(genes))
})



# Combine into a data.frame
go_table <- data.frame(
  GOID = all_ids,
  Term = all_terms,
  N_genes = gene_counts,
  Ontology = all_ont,
  Definition = all_defs,
  stringsAsFactors = FALSE
)

head(go_table, 10)

```

```{r}

go_table %>% filter(grepl('calcium', Term)) %>% arrange(desc(N_genes))

```



Initially just visualize average expression of these gene sets, focus on just the CGE celltypes

```{r}

zeng_10xv2_wba = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/zeng_wb_atlas_10xv2_gaba_primate_MN.rds')
zeng_10xv2_wba = zeng_10xv2_wba[ , zeng_10xv2_wba$orthology == 'high_orthology']

```



```{r}

cge_index = zeng_10xv2_wba$cell_subclass %in% c('Lamp5','Lamp5 Lhx6','Vip','Sncg')
mouse_clusters = unique(as.character(zeng_10xv2_wba$cell_cluster[cge_index]))


clust_to_mouse_subclass = c('Lamp5', 'Lamp5', 'Lamp5 Lhx6', 'Sncg', 'Vip', 'Vip', 'Vip', 'Vip', 'Vip', 'Sncg','Vip','Vip','Vip','Vip','Vip','Sncg','Sncg','Vip', 'Sncg', 'Lamp5')
names(clust_to_mouse_subclass) = mouse_clusters

clust_to_primate_subclass = c('Lamp5', 'Lamp5', 'Lamp5 Lhx6', 'Pax6', 'Vip', 'Vip', 'Vip', 'Vip', 'Sncg', 'Pax6','Vip','Vip','Vip','Vip','Sncg','Pax6','Pax6','Vip', 'Pax6', 'Pax6')
names(clust_to_primate_subclass) = mouse_clusters


```



```{r}

get_go_genes = function(go_ID){
  
  genes <- org.Hs.egGO2ALLEGS[[go_ID]]
  # Add gene symbols
  symbols <- mapIds(org.Hs.eg.db,
                    keys = genes,
                    column = "SYMBOL",
                    keytype = "ENTREZID",
                    multiVals = "first")
  symbols = unique(symbols)
  return(symbols)

}
```


Just get the average expression of the genes in a gene set, at their cluster level


```{r}

gene_set = 'GO:0005891'
go_term_name = Term(gene_set)
genes <- org.Hs.egGO2ALLEGS[[gene_set]]
# Add gene symbols
symbols <- mapIds(org.Hs.eg.db,
                  keys = genes,
                  column = "SYMBOL",
                  keytype = "ENTREZID",
                  multiVals = "first")
symbols = unique(symbols)


cell_index = colData(zeng_10xv2_wba)$cell_subclass %in% c('Lamp5','Lamp5 Lhx6','Vip','Sncg') 
gene_index = rownames(zeng_10xv2_wba) %in% symbols

#Subset the data
go_set_exp_sce = zeng_10xv2_wba[gene_index, cell_index]

#Compute the mean CPM per gene for each cluster
exp_cpm = assay(go_set_exp_sce, 'cpm')
colnames(exp_cpm) = go_set_exp_sce$cell_cluster

exp_cpm = reshape2::melt(as.matrix(exp_cpm), varnames = c('Gene','Cluster'), value.name = 'CPM')

mean_exp_cpm_df = exp_cpm %>% group_by(Cluster, Gene) %>% summarise(mean_CPM = mean(CPM)) %>%
  pivot_wider(names_from = Cluster, values_from = mean_CPM) %>% tibble::column_to_rownames(var = 'Gene')

#Scale the data
mean_exp_cpm_df = as.matrix(mean_exp_cpm_df)
mean_exp_cpm_df = t(scale(t(mean_exp_cpm_df)))

column_annot = HeatmapAnnotation(mouse_subclass = clust_to_mouse_subclass[colnames(mean_exp_cpm_df)],
                                 homologous_subclass = clust_to_primate_subclass[colnames(mean_exp_cpm_df)],
                                 col = list(mouse_subclass = celltype_color_palette,
                                            homologous_subclass = celltype_color_palette))

Heatmap(mean_exp_cpm_df, name = 'z-scored Mean CPM',
        show_row_names = F, show_row_dend = F, 
        clustering_method_rows = 'ward.D2', clustering_method_columns = 'ward.D2',
        top_annotation = column_annot,
        column_title = go_term_name)



```


Try a PCA just on the CGE clusters and the ion channels


If the signal is clear enough, just use that


Get the set of ion channels to use, potassium, sodium, calcium, chloride

GO:0034704	calcium channel complex
GO:0005891	voltage-gated calcium channel complex
GO:0034705	potassium channel complex
GO:0008076	voltage-gated potassium channel complex
GO:0034706	sodium channel complex
GO:0001518	voltage-gated sodium channel complex
GO:0034707	chloride channel complex

```{r}

go_table %>% filter(grepl('chloride', Term)) %>% arrange(desc(N_genes))


ion_channels = unique(c(get_go_genes('GO:0034704'), get_go_genes('GO:0005891'), get_go_genes('GO:0034705'),
  get_go_genes('GO:0008076'), get_go_genes('GO:0034706'), get_go_genes('GO:0001518'),
  get_go_genes('GO:0034707')))

length(ion_channels)

```



```{r}

cell_index = colData(zeng_10xv2_wba)$cell_subclass %in% c('Lamp5','Lamp5 Lhx6','Vip','Sncg') 
gene_index = rownames(zeng_10xv2_wba) %in% ion_channels

#Subset the data
go_set_exp_sce = zeng_10xv2_wba[gene_index, cell_index]

mouse_clust_and_subclass = colData(go_set_exp_sce)$cell_subclass
mouse_clust_and_subclass[go_set_exp_sce$cell_cluster == 'Lamp5 Gaba_4'] = 'Lamp5 Gaba_4'
mouse_clust_and_subclass[go_set_exp_sce$cell_cluster == 'Vip Gaba_1'] = 'Vip Gaba_1'
mouse_clust_and_subclass[go_set_exp_sce$cell_cluster == 'Vip Gaba_9'] = 'Vip Gaba_9'

go_set_exp_sce$mouse_clust_and_subclass = mouse_clust_and_subclass


go_set_exp_sce 
```


```{r}
celltype_and_clust_palette = c("Chandelier"='grey80', 
                           "Pvalb"='grey80', 
                           "Sst"='grey80', 
                           "Sst Chodl"='grey80', 
                           "Lamp5"='grey80',
                           "Sncg"='grey80',
                           "Vip"='grey80', 
                           "Lamp5 Lhx6"='grey80',
                           "Pax6"='grey80',
                           'Lamp5 Gaba_4' = '#DA808C', 'Vip Gaba_1' = '#A45FBF', 'Vip Gaba_9' = '#A45FBF')


```




```{r}

#dataset_hvg = scran::getTopHVGs(stats, n = 2000)
logcounts(go_set_exp_sce) = log1p(cpm(go_set_exp_sce))
#Run PCA on the HVG
go_set_exp_sce = scater::runPCA(go_set_exp_sce , scale = T)
#Run UMAP
go_set_exp_sce = scater::runUMAP(go_set_exp_sce , dimred = 'PCA', n_dimred = 10)



p_mouse_subclass = scater::plotReducedDim(go_set_exp_sce , dimred = 'UMAP', color_by = 'mouse_clust_and_subclass') + 
  scale_color_manual(values = celltype_and_clust_palette)
#Randomly shuffle point order and then manually plot to add an alpha option
p_mouse_subclass$data <- p_mouse_subclass$data[sample(nrow(p_mouse_subclass$data)), ]
p_mouse_subclass$layers <- list()

#Add centroid labels of the clusters we want to highlight
coords <- as.data.frame(reducedDim(go_set_exp_sce, "UMAP"))
coords$cell_type <- colData(go_set_exp_sce)$mouse_clust_and_subclass

centroids <- coords %>%
  group_by(cell_type) %>%
  summarise(Dim1 = mean(UMAP1), Dim2 = mean(UMAP2)) %>%
  filter(cell_type %in% c('Lamp5 Gaba_4','Vip Gaba_1','Vip Gaba_9'))


scater::plotReducedDim(go_set_exp_sce , dimred = 'UMAP', color_by = 'cell_subclass') + 
  scale_color_manual(values = celltype_color_palette) + ggtitle('Ion channel UMAP: original mouse subclasses') +
  ggrepel::geom_label_repel(
  data = centroids,
  aes(x = Dim1, y = Dim2, label = cell_type),
  inherit.aes = FALSE,
  size = 4,
  box.padding = 1.25,
  segment.color = "black") 

scater::plotReducedDim(go_set_exp_sce , dimred = 'UMAP', color_by = 'primate_subclass') + 
  scale_color_manual(values = celltype_color_palette) + ggtitle('Ion channel UMAP: homologous mouse subclasses') +
  ggrepel::geom_label_repel(
  data = centroids,
  aes(x = Dim1, y = Dim2, label = cell_type),
  inherit.aes = FALSE,
  size = 4,
  box.padding = 1.25,
  segment.color = "black") 

p_mouse_subclass + geom_point(data = p_mouse_subclass$data, 
                              aes(x = p_mouse_subclass$data$X, y = p_mouse_subclass$data$Y, colour = p_mouse_subclass$data$colour_by),
                              alpha = 0.5) +
  ggrepel::geom_label_repel(
  data = centroids,
  aes(x = Dim1, y = Dim2, label = cell_type),
  inherit.aes = FALSE,
  size = 4,
  box.padding = 1.25,
  segment.color = "black") +
  ggtitle('Ion channel UMAP: original mouse clusters')

```


```{r}

dim(go_set_exp_sce)
```


Perhaps try integrating all the mouse data on just the ion channel genes to get a summary UMAP and summary gene expression matrix
Average expression at the GO term level.

















