
Checking out the differential expression of specific GO terms across the subclasses

Focus on the Ion channels, as proxies for electrophysiological properties of the subclasses



GO database
SynGO database
KEGG database


```{r}

library(dplyr)
library(tidyr)
library(SingleCellExperiment)
library(MetaMarkers)
library(Seurat)
library(ggplot2)
library(org.Hs.eg.db)
library(GO.db)
library(ComplexHeatmap)

```


```{r}
celltype_color_palette = c("Chandelier"='#F641A8', 
                           "Pvalb"='#D93137', 
                           "Sst"='#FF9900', 
                           "Sst Chodl"='gold', 
                           "Lamp5"='#DA808C',
                           "Sncg"='#DF70FF',
                           "Vip"='#A45FBF', 
                           "Lamp5 Lhx6"='#935F50',
                           "Pax6"='#71238C')


```


SynGo Terms of interest

synapse adhesion between pre- and post-synapse
presynaptic modulation of chemical synaptic transmission
postsynaptic modulation of chemical synaptic transmission
transmitter-gated ion channel activity involved in regulation of postsynaptic membrane potential 
ligand-gated ion channel activity involved in regulation of presynaptic membrane potential
voltage-gated ion channel activity involved in regulation of presynaptic membrane potential 

```{r}
#synGo_path = '/home/werner/tools/syn_go/SynGO_bulk_download_release_20231201/'

synGo_annots_df = data.table::fread('/home/werner/tools/syn_go/SynGO_bulk_download_release_20231201/syngo_annotations.csv')

synGo_annots_df %>% filter(grepl('presynaptic membrane', synGo_annots_df$go_name)) %>% group_by(go_name) %>% summarise(n = n())

```


Go terms of interest


```{r}

org.Hs.eg.db #running this will print out the GO database version information
go_table <- as.data.frame(org.Hs.egGO2ALLEGS) #gene mapping to the direct GO term and that term's children
go_table$gene_symbol = mapIds(org.Hs.eg.db, go_table$gene_id, "SYMBOL","ENTREZID") #Add gene symbol 

go_table %>% filter(go_id == 'GO:0000009') %>% filter(!duplicated(gene_symbol))

```

```{r}


# Get all GO IDs in the database
all_ids <- keys(GO.db, keytype = "GOID")

# Get their term names (descriptions)
all_terms <- Term(all_ids)

# Get ontology type (BP, MF, CC)
all_ont <- Ontology(all_ids)

# Optional: get definitions
all_defs <- Definition(all_ids)


# 2. Map GO terms to genes
go2gene <- AnnotationDbi::as.list(org.Hs.egGO2ALLEGS)  # GOID -> list of Entrez IDs

# 3. Count how many genes per GO term
gene_counts <- sapply(all_ids, function(go) {
  genes <- go2gene[[go]]
  if (is.null(genes)) return(0)
  length(unique(genes))
})



# Combine into a data.frame
go_table <- data.frame(
  GOID = all_ids,
  Term = all_terms,
  N_genes = gene_counts,
  Ontology = all_ont,
  Definition = all_defs,
  stringsAsFactors = FALSE
)

head(go_table, 10)

```

```{r}

go_table %>% filter(grepl('calcium', Term)) %>% arrange(desc(N_genes))

```



Load up and integrate the data just using the ion channel genes. Can then also compute the CPMs across all the datasets for a summary plot.


Get the set of ion channels to use, potassium, sodium, calcium, chloride

GO:0034704	calcium channel complex
GO:0005891	voltage-gated calcium channel complex
GO:0034705	potassium channel complex
GO:0008076	voltage-gated potassium channel complex
GO:0034706	sodium channel complex
GO:0001518	voltage-gated sodium channel complex
GO:0034707	chloride channel complex

```{r}

get_go_genes = function(go_ID){
  
  genes <- org.Hs.egGO2ALLEGS[[go_ID]]
  # Add gene symbols
  symbols <- mapIds(org.Hs.eg.db,
                    keys = genes,
                    column = "SYMBOL",
                    keytype = "ENTREZID",
                    multiVals = "first")
  symbols = unique(symbols)
  return(symbols)

}
```

```{r}

go_table %>% filter(grepl('chloride', Term)) %>% arrange(desc(N_genes))


ion_channels = unique(c(get_go_genes('GO:0034704'), get_go_genes('GO:0005891'), get_go_genes('GO:0034705'),
  get_go_genes('GO:0008076'), get_go_genes('GO:0034706'), get_go_genes('GO:0001518'),
  get_go_genes('GO:0034707')))

length(ion_channels)

```



```{r}
tasic18 = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/tasic18_gaba_primate_MN.rds')
zeng_10x_cells = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/zeng_10x_cells_gaba_primate_MN.rds')
zeng_10x_nuclei = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/zeng_10x_nuclei_gaba_primate_MN.rds')
zeng_smart_cells = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/zeng_smart_cells_gaba_primate_MN.rds')
zeng_smart_nuclei = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/zeng_smart_nuclei_gaba_primate_MN.rds')
zeng_10xv2_wba = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/zeng_wb_atlas_10xv2_gaba_primate_MN.rds')
zeng_10xv3_wba = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/zeng_wb_atlas_10xv3_gaba_primate_MN.rds')
macosko_wba = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/macosko_wb_atlas_gaba_primate_MN.rds')
```

filter out the low-orthologous clusters

```{r}
tasic18 = tasic18[ , tasic18$orthology == 'high_orthology']
zeng_10x_cells = zeng_10x_cells[ , zeng_10x_cells$orthology == 'high_orthology']
zeng_10x_nuclei = zeng_10x_nuclei[ , zeng_10x_nuclei$orthology == 'high_orthology']
zeng_smart_cells = zeng_smart_cells[ , zeng_smart_cells$orthology == 'high_orthology']
zeng_smart_nuclei = zeng_smart_nuclei[ , zeng_smart_nuclei$orthology == 'high_orthology']
zeng_10xv2_wba = zeng_10xv2_wba[ , zeng_10xv2_wba$orthology == 'high_orthology']
zeng_10xv3_wba = zeng_10xv3_wba[ , zeng_10xv3_wba$orthology == 'high_orthology']
macosko_wba = macosko_wba[ , macosko_wba$orthology == 'high_orthology']

gc()
```


Subset to just the CGE cells and the ion channel genes


```{r}

cell_index = tasic18$primate_subclass %in% c('Lamp5','Lamp5 Lhx6','Vip','Sncg','Pax6')
gene_index = rownames(tasic18) %in% ion_channels
tasic18 = tasic18[gene_index, cell_index]


cell_index = zeng_10x_cells$primate_subclass %in% c('Lamp5','Lamp5 Lhx6','Vip','Sncg','Pax6')
gene_index = rownames(zeng_10x_cells) %in% ion_channels
zeng_10x_cells = zeng_10x_cells[gene_index, cell_index]

cell_index = zeng_10x_nuclei$primate_subclass %in% c('Lamp5','Lamp5 Lhx6','Vip','Sncg','Pax6')
gene_index = rownames(zeng_10x_nuclei) %in% ion_channels
zeng_10x_nuclei = zeng_10x_nuclei[gene_index, cell_index]


cell_index = zeng_smart_cells$primate_subclass %in% c('Lamp5','Lamp5 Lhx6','Vip','Sncg','Pax6')
gene_index = rownames(zeng_smart_cells) %in% ion_channels
zeng_smart_cells = zeng_smart_cells[gene_index, cell_index]

cell_index = zeng_smart_nuclei$primate_subclass %in% c('Lamp5','Lamp5 Lhx6','Vip','Sncg','Pax6')
gene_index = rownames(zeng_smart_nuclei) %in% ion_channels
zeng_smart_nuclei = zeng_smart_nuclei[gene_index, cell_index]


cell_index = zeng_10xv2_wba$primate_subclass %in% c('Lamp5','Lamp5 Lhx6','Vip','Sncg','Pax6')
gene_index = rownames(zeng_10xv2_wba) %in% ion_channels
zeng_10xv2_wba = zeng_10xv2_wba[gene_index, cell_index]


cell_index = zeng_10xv3_wba$primate_subclass %in% c('Lamp5','Lamp5 Lhx6','Vip','Sncg','Pax6')
gene_index = rownames(zeng_10xv3_wba) %in% ion_channels
zeng_10xv3_wba = zeng_10xv3_wba[gene_index, cell_index]


cell_index = macosko_wba$primate_subclass %in% c('Lamp5','Lamp5 Lhx6','Vip','Sncg','Pax6')
gene_index = rownames(macosko_wba) %in% ion_channels
macosko_wba = macosko_wba[gene_index, cell_index]


dim(tasic18)
dim(zeng_10x_cells)
dim(zeng_10x_nuclei)
dim(zeng_smart_cells)
dim(zeng_smart_nuclei)
dim(zeng_10xv2_wba)
dim(zeng_10xv3_wba)
dim(macosko_wba)

```


Convert to seurat and integrate


```{r}

tasic18 = as.Seurat(tasic18, counts = "counts", data = "cpm")
zeng_10x_cells = as.Seurat(zeng_10x_cells, counts = "counts", data = "cpm")
zeng_10x_nuclei = as.Seurat(zeng_10x_nuclei, counts = "counts", data = "cpm")
zeng_smart_cells = as.Seurat(zeng_smart_cells, counts = "counts", data = "cpm")
zeng_smart_nuclei = as.Seurat(zeng_smart_nuclei, counts = "counts", data = "cpm")

zeng_10xv2_wba = as.Seurat(zeng_10xv2_wba, counts = "counts", data = "cpm")
zeng_10xv3_wba = as.Seurat(zeng_10xv3_wba, counts = "counts", data = "cpm")

macosko_wba = as.Seurat(macosko_wba, counts = "counts", data = "cpm")
gc()
```

FindIntegrationAnchors took 1.5 hours

```{r}

#Make the list of objects
seurat.list = list('tasic18' = tasic18, 'zeng_10x_cells' = zeng_10x_cells, 'zeng_smart_cells' = zeng_smart_cells ,  
                   'zeng_10x_nuclei' = zeng_10x_nuclei, 'zeng_smart_nuclei' =  zeng_smart_nuclei,
                   'zeng_10xv2_wba' = zeng_10xv2_wba, 'zeng_10xv3_wba' = zeng_10xv3_wba , 'macosko_wba' = macosko_wba )
#Rm individual datasets to save space, doesn't take that long for the above chuncks to rerun
rm(tasic18, zeng_10x_cells, zeng_smart_cells ,zeng_10x_nuclei, zeng_smart_nuclei, zeng_10xv2_wba, zeng_10xv3_wba ,macosko_wba )
gc()


#integration_features <- SelectIntegrationFeatures(object.list = seurat.list)
#length(integration_features)

#Integrate on all of the ion channel genes
integration_features = rownames(seurat.list$tasic18)

start = Sys.time()
integration_anchors <- FindIntegrationAnchors(object.list = seurat.list, anchor.features = integration_features)
start - Sys.time()

```

```{r}
rm(seurat.list)
gc()
```


Takes awhile and a lot of memory
```{r}
start = Sys.time()

#Make a single integrated object
interneuron.ion_channels <- IntegrateData(anchorset = integration_anchors)

start - Sys.time()
```

```{r}
rm(integration_anchors)
gc()
```

```{r}
start = Sys.time()

DefaultAssay(interneuron.ion_channels) <- "integrated"
# Run the standard workflow for visualization and clustering
interneuron.ion_channels <- ScaleData(interneuron.ion_channels, verbose = FALSE)
interneuron.ion_channels <- RunPCA(interneuron.ion_channels, verbose = FALSE)
interneuron.ion_channels <- RunUMAP(interneuron.ion_channels, reduction = "pca", dims = 1:20)

Sys.time() - start

```


```{r}

DimPlot(interneuron.ion_channels, group.by = 'study')

DimPlot(interneuron.ion_channels, group.by = 'cell_subclass', cols = celltype_color_palette) + ggtitle('Ion channel UMAP: Original mouse subclass')
ggsave(filename = 'mouse_ion_integrated_umap_mouse_labels.pdf', path = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/integrated_interneurons/ion_channel_plots/',
       device = 'pdf', useDingbats = F, height = 4, width = 6)


DimPlot(interneuron.ion_channels, group.by = 'primate_subclass', cols = celltype_color_palette) + ggtitle('Ion channel UMAP: Homologous primate subclass')
ggsave(filename = 'mouse_ion_integrated_umap_primate_labels.pdf', 
       path = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/integrated_interneurons/ion_channel_plots/',
       device = 'pdf', useDingbats = F, height = 4, width = 6)
```


```{r}

DimPlot(interneuron.ion_channels, group.by = 'study', reduction = 'pca')

DimPlot(interneuron.ion_channels, group.by = 'cell_subclass', cols = celltype_color_palette, reduction = 'pca') + 
  ggtitle('Ion channel UMAP: Original mouse subclass')

DimPlot(interneuron.ion_channels, group.by = 'primate_subclass', cols = celltype_color_palette, reduction = 'pca') + 
  ggtitle('Ion channel UMAP: Homologous primate subclass')

```



```{r}

save(interneuron.ion_channels, file = '/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/mouse_primate_ionChannels_integrated.Rdata')

```



Visualize heatmaps of some ion channel gene CPMs

GO:0034704	calcium channel complex
GO:0005891	voltage-gated calcium channel complex
GO:0034705	potassium channel complex
GO:0008076	voltage-gated potassium channel complex
GO:0034706	sodium channel complex
GO:0001518	voltage-gated sodium channel complex
GO:0034707	chloride channel complex

```{r}

potassium_genes = unique(c(get_go_genes('GO:0034705'), get_go_genes('GO:0008076')))
sodium_genes = unique(c(get_go_genes('GO:0034706'), get_go_genes('GO:0001518')))
calcium_genes = unique(c(get_go_genes('GO:0034704'), get_go_genes('GO:0005891')))
chloride_genes = get_go_genes('GO:0034707')
gene_index = rownames(interneuron.ion_channels) %in% potassium_genes


interneuron.ion_channels$cell_subclass[interneuron.ion_channels$cell_cluster == 'Lamp5 Lhx6'] = 'Lamp5 Lhx6'
interneuron.ion_channels$cell_subclass[interneuron.ion_channels$cell_subclass == 'Serpinf1'] = 'Vip'

cell_index = interneuron.ion_channels$cell_subclass %in% c('Lamp5', 'Sncg' )

#Subset the data
go_set_exp_sce = interneuron.ion_channels
#The smart seq cells and nuclei have an identically named cluster, separate the two
go_set_exp_sce$cell_cluster[go_set_exp_sce$study == 'zeng_smart_cells' & go_set_exp_sce$cell_cluster == 'Lamp5 Egln3_2'] = 'Lamp5 Egln3_2_2'

#Compute the mean CPM per gene for each cluster
exp_cpm = go_set_exp_sce@assays$integrated@data
colnames(exp_cpm) = go_set_exp_sce$cell_cluster

exp_cpm = reshape2::melt(as.matrix(exp_cpm), varnames = c('Gene','Cluster'), value.name = 'CPM')

mean_exp_cpm_df = exp_cpm %>% group_by(Cluster, Gene) %>% summarise(mean_CPM = mean(CPM)) %>%
  pivot_wider(names_from = Cluster, values_from = mean_CPM) %>% tibble::column_to_rownames(var = 'Gene')

#Scale the data
mean_exp_cpm_df = as.matrix(mean_exp_cpm_df)
mean_exp_cpm_df = t(scale(t(mean_exp_cpm_df)))


clust_to_subclass_df = go_set_exp_sce[[]] %>% group_by(cell_cluster, cell_subclass, primate_subclass) %>% summarise(n = n())
#clust_to_subclass_df$cell_subclass[clust_to_subclass_df$cell_subclass == 'Serpinf1'] = 'Vip'
#clust_to_subclass_df$cell_subclass[clust_to_subclass_df$cell_cluster == 'Lamp5 Lhx6'] = 'Lamp5 Lhx6'

index = match(colnames(mean_exp_cpm_df) , clust_to_subclass_df$cell_cluster)


column_annot = HeatmapAnnotation(mouse_subclass = clust_to_subclass_df$cell_subclass[index],
                                 homologous_subclass = clust_to_subclass_df$primate_subclass[index],
                                 col = list(mouse_subclass = celltype_color_palette,
                                            homologous_subclass = celltype_color_palette))


h1 = Heatmap(mean_exp_cpm_df, name = 'z-scored Mean CPM',
        show_row_names = F, show_row_dend = F, show_column_names = F,
        clustering_method_rows = 'ward.D2', clustering_method_columns = 'ward.D2',
        top_annotation = column_annot, 
        column_title = 'Ion channel GO term genes')
h1 = draw(h1)


pdf(file = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/integrated_interneurons/ion_channel_plots/all_ion_genes_avg_cpm_heatmap.pdf',
    useDingbats = F, height = 8, width = 6)
h1
dev.off()

```


```{r}


mouse_subclass_ion_markers = compute_markers(interneuron.ion_channels@assays$integrated@data, interneuron.ion_channels$cell_subclass)
mouse_marker_genes = mouse_subclass_ion_markers %>% group_by(cell_type) %>% slice_max(auroc, n = 10) %>% pull(gene)

primate_subclass_ion_markers = compute_markers(interneuron.ion_channels@assays$integrated@data, interneuron.ion_channels$primate_subclass)
primate_marker_genes = primate_subclass_ion_markers %>% group_by(cell_type) %>% slice_max(auroc, n = 10) %>% pull(gene)



gene_index = rownames(interneuron.ion_channels) %in% mouse_marker_genes

#Subset the data
go_set_exp_sce = interneuron.ion_channels[gene_index, ]
#The smart seq cells and nuclei have an identically named cluster, separate the two
go_set_exp_sce$cell_cluster[go_set_exp_sce$study == 'zeng_smart_cells' & go_set_exp_sce$cell_cluster == 'Lamp5 Egln3_2'] = 'Lamp5 Egln3_2_2'

#Compute the mean CPM per gene for each cluster
exp_cpm = go_set_exp_sce@assays$integrated@data
colnames(exp_cpm) = go_set_exp_sce$cell_cluster

exp_cpm = reshape2::melt(as.matrix(exp_cpm), varnames = c('Gene','Cluster'), value.name = 'CPM')

mean_exp_cpm_df = exp_cpm %>% group_by(Cluster, Gene) %>% summarise(mean_CPM = mean(CPM)) %>%
  pivot_wider(names_from = Cluster, values_from = mean_CPM) %>% tibble::column_to_rownames(var = 'Gene')

#Scale the data
mean_exp_cpm_df = as.matrix(mean_exp_cpm_df)
mean_exp_cpm_df = t(scale(t(mean_exp_cpm_df)))


clust_to_subclass_df = go_set_exp_sce[[]] %>% group_by(cell_cluster, cell_subclass, primate_subclass) %>% summarise(n = n())
#clust_to_subclass_df$cell_subclass[clust_to_subclass_df$cell_subclass == 'Serpinf1'] = 'Vip'
#clust_to_subclass_df$cell_subclass[clust_to_subclass_df$cell_cluster == 'Lamp5 Lhx6'] = 'Lamp5 Lhx6'

index = match(colnames(mean_exp_cpm_df) , clust_to_subclass_df$cell_cluster)


column_annot = HeatmapAnnotation(mouse_subclass = clust_to_subclass_df$cell_subclass[index],
                                 homologous_subclass = clust_to_subclass_df$primate_subclass[index],
                                 col = list(mouse_subclass = celltype_color_palette,
                                            homologous_subclass = celltype_color_palette))

set.seed(123)
h2 = Heatmap(mean_exp_cpm_df, name = 'z-scored Mean CPM',
        show_row_names = T, show_row_dend = F, show_column_names = F,
        clustering_method_rows = 'ward.D2', clustering_method_columns = 'ward.D2',
        top_annotation = column_annot, row_names_gp = grid::gpar(fontsize = 8),
        column_title = 'Ion channel GO term genes')
h2 = draw(h2)



pdf(file = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/integrated_interneurons/ion_channel_plots/top_mouse_de_ion_genes_avg_cpm_heatmap.pdf',
    useDingbats = F, height = 5, width = 7)
h2
dev.off()



```


















