
Checking out the differential expression of specific GO terms across the subclasses

Focus on the Ion channels, as proxies for electrophysiological properties of the subclasses



GO database
SynGO database
KEGG database


```{r}

library(dplyr)
library(tidyr)
library(SingleCellExperiment)
library(ggplot2)
library(org.Hs.eg.db)
library(GO.db)
library(ComplexHeatmap)

```


```{r}
celltype_color_palette = c("Chandelier"='#F641A8', 
                           "Pvalb"='#D93137', 
                           "Sst"='#FF9900', 
                           "Sst Chodl"='gold', 
                           "Lamp5"='#DA808C',
                           "Sncg"='#DF70FF',
                           "Vip"='#A45FBF', 
                           "Lamp5 Lhx6"='#935F50',
                           "Pax6"='#71238C')


```


SynGo Terms of interest

synapse adhesion between pre- and post-synapse
presynaptic modulation of chemical synaptic transmission
postsynaptic modulation of chemical synaptic transmission
transmitter-gated ion channel activity involved in regulation of postsynaptic membrane potential 
ligand-gated ion channel activity involved in regulation of presynaptic membrane potential
voltage-gated ion channel activity involved in regulation of presynaptic membrane potential 

```{r}
#synGo_path = '/home/werner/tools/syn_go/SynGO_bulk_download_release_20231201/'

synGo_annots_df = data.table::fread('/home/werner/tools/syn_go/SynGO_bulk_download_release_20231201/syngo_annotations.csv')

synGo_annots_df %>% filter(grepl('presynaptic membrane', synGo_annots_df$go_name)) %>% group_by(go_name) %>% summarise(n = n())

```


Go terms of interest


```{r}

org.Hs.eg.db #running this will print out the GO database version information
go_table <- as.data.frame(org.Hs.egGO2ALLEGS) #gene mapping to the direct GO term and that term's children
go_table$gene_symbol = mapIds(org.Hs.eg.db, go_table$gene_id, "SYMBOL","ENTREZID") #Add gene symbol 

go_table %>% filter(go_id == 'GO:0000009') %>% filter(!duplicated(gene_symbol))

```

```{r}


# Get all GO IDs in the database
all_ids <- keys(GO.db, keytype = "GOID")

# Get their term names (descriptions)
all_terms <- Term(all_ids)

# Get ontology type (BP, MF, CC)
all_ont <- Ontology(all_ids)

# Optional: get definitions
all_defs <- Definition(all_ids)


# 2. Map GO terms to genes
go2gene <- AnnotationDbi::as.list(org.Hs.egGO2ALLEGS)  # GOID -> list of Entrez IDs

# 3. Count how many genes per GO term
gene_counts <- sapply(all_ids, function(go) {
  genes <- go2gene[[go]]
  if (is.null(genes)) return(0)
  length(unique(genes))
})



# Combine into a data.frame
go_table <- data.frame(
  GOID = all_ids,
  Term = all_terms,
  N_genes = gene_counts,
  Ontology = all_ont,
  Definition = all_defs,
  stringsAsFactors = FALSE
)

head(go_table, 10)

```

```{r}

go_table %>% filter(grepl('calcium', Term)) %>% arrange(desc(N_genes))

```



Initially just visualize average expression of these gene sets, focus on just the CGE celltypes

```{r}

zeng_10xv2_wba = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/zeng_wb_atlas_10xv2_gaba_primate_MN.rds')
zeng_10xv2_wba = zeng_10xv2_wba[ , zeng_10xv2_wba$orthology == 'high_orthology']

```



```{r}

cge_index = zeng_10xv2_wba$cell_subclass %in% c('Lamp5','Lamp5 Lhx6','Vip','Sncg')
mouse_clusters = unique(as.character(zeng_10xv2_wba$cell_cluster[cge_index]))


clust_to_mouse_subclass = c('Lamp5', 'Lamp5', 'Lamp5 Lhx6', 'Sncg', 'Vip', 'Vip', 'Vip', 'Vip', 'Vip', 'Sncg','Vip','Vip','Vip','Vip','Vip','Sncg','Sncg','Vip', 'Sncg', 'Lamp5')
names(clust_to_mouse_subclass) = mouse_clusters

clust_to_primate_subclass = c('Lamp5', 'Lamp5', 'Lamp5 Lhx6', 'Pax6', 'Vip', 'Vip', 'Vip', 'Vip', 'Sncg', 'Pax6','Vip','Vip','Vip','Vip','Sncg','Pax6','Pax6','Vip', 'Pax6', 'Pax6')
names(clust_to_primate_subclass) = mouse_clusters


```





Just get the average expression of the genes in a gene set, at their cluster level


```{r}

gene_set = 'GO:0005891'
go_term_name = Term(gene_set)
genes <- org.Hs.egGO2ALLEGS[[gene_set]]
# Add gene symbols
symbols <- mapIds(org.Hs.eg.db,
                  keys = genes,
                  column = "SYMBOL",
                  keytype = "ENTREZID",
                  multiVals = "first")
symbols = unique(symbols)


cell_index = colData(zeng_10xv2_wba)$cell_subclass %in% c('Lamp5','Lamp5 Lhx6','Vip','Sncg') 
gene_index = rownames(zeng_10xv2_wba) %in% symbols

#Subset the data
go_set_exp_sce = zeng_10xv2_wba[gene_index, cell_index]

#Compute the mean CPM per gene for each cluster
exp_cpm = assay(go_set_exp_sce, 'cpm')
colnames(exp_cpm) = go_set_exp_sce$cell_cluster

exp_cpm = reshape2::melt(as.matrix(exp_cpm), varnames = c('Gene','Cluster'), value.name = 'CPM')

mean_exp_cpm_df = exp_cpm %>% group_by(Cluster, Gene) %>% summarise(mean_CPM = mean(CPM)) %>%
  pivot_wider(names_from = Cluster, values_from = mean_CPM) %>% tibble::column_to_rownames(var = 'Gene')

#Scale the data
mean_exp_cpm_df = as.matrix(mean_exp_cpm_df)
mean_exp_cpm_df = t(scale(t(mean_exp_cpm_df)))

column_annot = HeatmapAnnotation(mouse_subclass = clust_to_mouse_subclass[colnames(mean_exp_cpm_df)],
                                 homologous_subclass = clust_to_primate_subclass[colnames(mean_exp_cpm_df)],
                                 col = list(mouse_subclass = celltype_color_palette,
                                            homologous_subclass = celltype_color_palette))

Heatmap(mean_exp_cpm_df, name = 'z-scored Mean CPM',
        show_row_names = F, show_row_dend = F, 
        clustering_method_rows = 'ward.D2', clustering_method_columns = 'ward.D2',
        top_annotation = column_annot,
        column_title = go_term_name)



```




























