


Exploring marker gene variability across the allen+macosko mouse data and the allen primate data

Using the original mouse subclass annotations


```{r}
library(Seurat)
library(SingleCellExperiment)
library(ggplot2)
library(dplyr)
library(MetaMarkers)
library(ComplexHeatmap)

options(dplyr.summarise.inform = FALSE)


```


```{r}
celltype_color_palette = c("Chandelier"='#F641A8', 
                           "Pvalb"='#D93137', 
                           "Sst"='#FF9900', 
                           "Sst Chodl"='#FFD700', 
                           "Lamp5"='#DA808C',
                           "Sncg"='#DF70FF',
                           "Vip"='#A45FBF', 
                           "Lamp5 Lhx6"='#935F50',
                           "Pax6"='#71238C',
                           'NA' = 'grey90')


```


```{r}

load(file = '/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/mouse_primate_MN_integrated.Rdata')
interneuron.combined$cell_subclass[interneuron.combined$cell_subclass == 'Serpinf1'] = 'Vip'

interneuron.combined
```

Get the whole brain atlas clusters for Leon

```{r}

load(file = '/inkwell03/werner/mouse_brain_consensus_taxonomy/20_3_25_interneuron_mouse_metadata.Rdata')
interneuron_metadata

table(interneuron_metadata$study)


interneuron_metadata %>% filter(study == 'tasic18') %>% group_by(cell_cluster, cell_subclass, primate_subclass) %>%
  summarise(n = n()) %>% arrange(cell_subclass)
```



```{r}


wba_cluster_df = interneuron_metadata %>% filter(study %in% c('macosko_wba','zeng_10xv3_wba','zeng_10xv2_wba')) %>% group_by(study, cell_cluster) %>%
  summarise(n = n())


macosko_df = wba_cluster_df %>% filter(study == 'macosko_wba') %>% select(study, cell_cluster)

write.csv2(macosko_df, file = '/home/werner/projects/mouse_brain_consensus_taxonomy/data/macosko_mouse_wba_kept_clusters.csv', row.names = F)


zeng_df = wba_cluster_df %>% filter(study != 'macosko_wba') %>% select(study, cell_cluster)

write.csv2(zeng_df, file = '/home/werner/projects/mouse_brain_consensus_taxonomy/data/zeng_mouse_wba_kept_clusters.csv', row.names = F)


```


```{r}

zeng_wba_region_mapping = data.table::fread('/vault/lfrench/mouse_brain_consensus_taxonomy_spatial/results/Zeng_mouse_wba_kept_clusters.region_counts.csv')

table(zeng_wba_region_mapping$region_remap)



macosko_wba_region_mapping = data.table::fread('/vault/lfrench/mouse_brain_consensus_taxonomy_spatial/results/Macosko_mouse_wba_kept_clusters.region_counts.csv')

table(macosko_wba_region_mapping$region_remap)

```



For the other datasets, Tasic18 is Primary Visual cortex (VISP) and Anterior Lateral Motor cortex (ALM)
the zeng different technologies datasets are all primate motor cortex (MOp)

```{r}
tasic18 = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/tasic18_gaba_primate_MN.rds')
#zeng_10x_cells = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/zeng_10x_cells_gaba_primate_MN.rds')
#zeng_10x_nuclei = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/zeng_10x_nuclei_gaba_primate_MN.rds')
#zeng_smart_cells = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/zeng_smart_cells_gaba_primate_MN.rds')
#zeng_smart_nuclei = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/zeng_smart_nuclei_gaba_primate_MN.rds')


tasic18 = tasic18[ , tasic18$orthology == 'high_orthology']
```

And the unprocessed tasic18, since this has the region metadata

```{r}

tasic18_full = readRDS(file = '/vault/werner/rotation/jonathan/data/tasic18.rds')
table(as.character(tasic18_full$source.name), tasic18_full$characteristics..cell_subclass)

#Just grab the GABAergic cells
inhib_filt = colData(tasic18_full)$cell_class == 'GABAergic'
tasic18_full = tasic18_full[ ,inhib_filt]

table(colData(tasic18_full)$cell_cluster)
ncol(tasic18_full)


colData(tasic18_full)
colData(tasic18)

index = match(rownames(colData(tasic18)), tasic18_full$title)
tasic18$region = as.character(tasic18_full$source.name[index])

tasic18_region_mapping = as.data.frame(colData(tasic18)) %>% group_by(cell_cluster, region, primate_subclass) %>% summarise(n = n()) %>%
  select(cell_cluster, primate_subclass, n, region)



tasic18_region_mapping$region[tasic18_region_mapping$region == 'Anterior Lateral Motor Cortex (ALM)'] = 'ALM'
tasic18_region_mapping$region[tasic18_region_mapping$region == 'Primary Visual Cortex (VISp)'] = 'VISP'

table(tasic18_region_mapping$region)
```


```{r}

macosko_clust_subclass_mapping = interneuron_metadata %>% filter(study == 'macosko_wba') %>% group_by(cell_cluster, primate_subclass) %>%
  summarise(n = n())

#Add primate subclass labels to the clusters
index = match(macosko_wba_region_mapping$cell_cluster, macosko_clust_subclass_mapping$cell_cluster )
macosko_wba_region_mapping$primate_subclass = macosko_clust_subclass_mapping$primate_subclass[index]

macosko_wba_region_mapping = macosko_wba_region_mapping %>% filter(cell_count_Macosko != 0) %>%
  select(cell_cluster, primate_subclass, cell_count_Macosko, region_remap)
colnames(macosko_wba_region_mapping) = c('cell_cluster','primate_subclass','n','region')



zeng_clust_subclass_mapping = interneuron_metadata %>% filter(study %in% c('zeng_10xv2_wba', 'zeng_10xv3_wba')) %>%
  group_by(cell_cluster, primate_subclass) %>%
  summarise(n = n())


#Add primate subclass labels to the clusters
index = match(zeng_wba_region_mapping$cell_cluster, zeng_clust_subclass_mapping$cell_cluster )
zeng_wba_region_mapping$primate_subclass = zeng_clust_subclass_mapping$primate_subclass[index]

zeng_wba_region_mapping = zeng_wba_region_mapping %>% filter(cell_count_Zeng != 0) %>%
  select(cell_cluster, primate_subclass, cell_count_Zeng, region_remap)
colnames(zeng_wba_region_mapping) = c('cell_cluster','primate_subclass','n','region')




zeng_motor_cortex_df = interneuron_metadata %>% 
  filter(study %in% c('zeng_10x_cells', 'zeng_10x_nuclei', 'zeng_smart_cells', 'zeng_smart_nuclei')) %>%
  group_by(cell_cluster, primate_subclass) %>%
  summarise(n = n()) %>% mutate(region = 'MOp')

```


```{r}

all_region_mapping = do.call('rbind', 
                             list(zeng_wba_region_mapping, macosko_wba_region_mapping, zeng_motor_cortex_df, tasic18_region_mapping))


all_region_mapping


all_region_mapping$primate_subclass = factor(all_region_mapping$primate_subclass,
                                              levels = c('Sst','Sst Chodl','Pvalb','Chandelier',
                                                         'Vip','Sncg','Pax6','Lamp5','Lamp5 Lhx6'))

all_regions = unique(all_region_mapping$region)

region_colors = MetBrewer::met.brewer("Moreau", n=20, type="continuous")
region_colors = region_colors[1:20] 
names(region_colors) =c(all_regions)


# Stacked + percent
ggplot(all_region_mapping, aes(fill=region, y=n, x=primate_subclass)) + 
    geom_bar(position="fill", stat="identity") + ylab('% cell-type') + 
  scale_fill_manual(values = region_colors) +
  ggtitle('Regional contribution to orthologous subclasses') +
  scale_x_discrete(guide = guide_axis(n.dodge=2))

ggsave(filename = 'region_stacked_orth_subclass_mouse.pdf',device = 'pdf',
       path = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/integrated_interneurons/',
       useDingbats = F, height = 6, width = 8)
```

```{r}


all_regions = unique(all_region_mapping$region)

pvalb_regions = unique(all_region_mapping %>% filter(primate_subclass == 'Pvalb') %>% pull(region))
print('Regions with no Pvalb cells...')
all_regions[!all_regions %in% pvalb_regions]

chandelier_regions = unique(all_region_mapping %>% filter(primate_subclass == 'Chandelier') %>% pull(region))
print('Regions with no Chandelier cells...')
all_regions[!all_regions %in% chandelier_regions]

sst_regions = unique(all_region_mapping %>% filter(primate_subclass == 'Sst') %>% pull(region))
print('Regions with no sst cells...')
all_regions[!all_regions %in% sst_regions]

sst_chodl_regions = unique(all_region_mapping %>% filter(primate_subclass == 'Sst Chodl') %>% pull(region))
print('Regions with no sst_chodl cells...')
all_regions[!all_regions %in% sst_chodl_regions]

vip_regions = unique(all_region_mapping %>% filter(primate_subclass == 'Vip') %>% pull(region))
print('Regions with no vip cells...')
all_regions[!all_regions %in% vip_regions]

pax6_regions = unique(all_region_mapping %>% filter(primate_subclass == 'Pax6') %>% pull(region))
print('Regions with no pax6 cells...')
all_regions[!all_regions %in% pax6_regions]


sncg_regions = unique(all_region_mapping %>% filter(primate_subclass == 'Sncg') %>% pull(region))
print('Regions with no sncg cells...')
all_regions[!all_regions %in% sncg_regions]

lamp5_regions = unique(all_region_mapping %>% filter(primate_subclass == 'Lamp5') %>% pull(region))
print('Regions with no lamp5 cells...')
all_regions[!all_regions %in% lamp5_regions]

lamp5_lhx6_regions = unique(all_region_mapping %>% filter(primate_subclass == 'Lamp5 Lhx6') %>% pull(region))
print('Regions with no lamp5_lhx6 cells...')
all_regions[!all_regions %in% lamp5_lhx6_regions]

```

Do a heatmap instead 
Supp. Fig. 7 B

```{r}

all_regions = unique(all_region_mapping$region)

all_region_subclasses = paste(rep(all_regions, 9), c(rep('Pvalb',length(all_regions)), rep('Chandelier',length(all_regions)),rep('Sst',length(all_regions)),
  rep('Sst Chodl',length(all_regions)),rep('Vip',length(all_regions)),rep('Sncg',length(all_regions)),
  rep('Pax6',length(all_regions)),rep('Lamp5',length(all_regions)),rep('Lamp5 Lhx6',length(all_regions))), sep = '_')

all_region_mapping$primate_subclass = as.character(all_region_mapping$primate_subclass)
all_region_mapping$region_subclass = paste(all_region_mapping$region, all_region_mapping$primate_subclass, sep = '_')

#Missing region and subclass intersections, get 0 counts
missing_region_subclasses = all_region_subclasses[!all_region_subclasses %in% all_region_mapping$region_subclass ]

missing_regions = sapply(strsplit(missing_region_subclasses, split = '_', fixed = T), '[[',1)
missing_subclasses = sapply(strsplit(missing_region_subclasses, split = '_', fixed = T), '[[',2)

missing_df = data.frame(cell_cluster = NA, primate_subclass = missing_subclasses, n = 0, 
                        region = missing_regions , region_subclass = missing_region_subclasses )

all_region_mapping = rbind(all_region_mapping, missing_df)


subclass_region_mat = all_region_mapping %>% select(primate_subclass, n, region) %>% group_by(region, primate_subclass) %>%
  summarise(total = sum(n)) %>% tidyr::pivot_wider(names_from = region, values_from = total) %>%
  tibble::column_to_rownames(var = 'primate_subclass')

subclass_region_mat = as.matrix(subclass_region_mat)


col_fun = circlize::colorRamp2(breaks = c(0,log10(2), max(log10(subclass_region_mat+1))), colors = c('grey','white','red'))
ComplexHeatmap::Heatmap(log10(subclass_region_mat+1), col = col_fun, name = 'log10( cell count + 1 )',
                        clustering_method_rows = 'ward.D2', clustering_method_columns = 'ward.D2',
                        show_row_dend = F, show_column_dend = F)


pdf(file = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/integrated_interneurons/region_orth_mouse_subclass_heatmap.pdf',
    useDingbats = F, height = 4, width = 6)
ComplexHeatmap::Heatmap(log10(subclass_region_mat+1), col = col_fun, name = 'log10( cell count + 1 )',
                        clustering_method_rows = 'ward.D2', clustering_method_columns = 'ward.D2',
                        show_row_dend = F, show_column_dend = F)

dev.off()

```


###################################

Get the dot plot enrichments of the primate Vip Pax6 and Sncg markers across the mouse data
Get the average per dataset per subclass


Also get the pareto front markers for the mouse, primate, and cross-mammal markers of just Sncg and Pax6, since that's where the confusion lies

##################################

```{r}
all_mouse_markers = list(
    tasic18 = read_markers("../metamarkers/ortho_markers_tasic18.csv.gz"),
    zeng_10x_cells = read_markers("../metamarkers/ortho_markers_zeng_10x_cells.csv.gz"),
    zeng_10x_nuclei = read_markers("../metamarkers/ortho_markers_zeng_10x_nuclei.csv.gz"),
    zeng_smart_nuclei = read_markers("../metamarkers/ortho_markers_zeng_smart_nuclei.csv.gz"),
    zeng_smart_cells = read_markers("../metamarkers/ortho_markers_zeng_smart_cells.csv.gz"),
    zeng_10xv2_wba = read_markers("../metamarkers/ortho_markers_zeng_10xv2_wba.csv.gz"),
    zeng_10xv3_wba = read_markers("../metamarkers/ortho_markers_zeng_10xv3_wba.csv.gz"),
    macosko_wba = read_markers("../metamarkers/ortho_markers_macosko_wba.csv.gz")
)


mouse_subclass_meta_markers = make_meta_markers(all_mouse_markers, detailed_stats = TRUE)

mouse_subclass_meta_markers %>% group_by(cell_type) %>% filter(rank <= 10)


```

Also get plots demonstrating the homologous DE in mouse, people keep asking questions that amount to How transcriptomically distinct are the CGE homologous subclasses
heatmap of expression, but grabbing the top signal-to-noise markers per subclass


Grab the top 10 recurrent DE markers, get heatmaps of the AUROCs and the fold_change_detection as coverage markers and signal-to-noise markers
Then get dot-line plots of the average AUROC (or average fold_change detection) for each level of recurrence for the subclasses. As we go from consistent DE to dataset specific DE, what's the variability across the subclasses? Are the Vip, Pax6, Sncg subclasses the less distinct globally?


```{r}

top_markers = mouse_subclass_meta_markers %>% filter(rank <= 10)
dup_genes = top_markers$gene[which(duplicated(top_markers$gene))]

#For each duplicate, assign it to the cell-type with the better (minimum) rank
keep_dups = top_markers %>% filter(gene %in% dup_genes) %>% group_by(gene) %>% filter(rank == min(rank))
top_markers = top_markers %>% filter(!gene %in% dup_genes)
top_markers = rbind(top_markers, keep_dups)
top_markers = top_markers %>% arrange(cell_type)

#Set order
top_markers = top_markers %>% arrange(match(cell_type, c('Pvalb','Chandelier','Sst','Sst Chodl','Lamp5','Lamp5 Lhx6','Vip','Sncg','Pax6')))


#AUROC stat
top_auroc_mat = mouse_subclass_meta_markers %>% filter(gene %in% top_markers$gene) %>% select(cell_type, gene, auroc) %>%
  tidyr::pivot_wider(names_from = gene, values_from = auroc) %>% tibble::column_to_rownames(var = 'cell_type')
top_auroc_mat = t(as.matrix(top_auroc_mat))

#Keep the same ordering
index = match(top_markers$gene, rownames(top_auroc_mat))
top_auroc_mat = top_auroc_mat[index, ]
top_auroc_mat = top_auroc_mat[ , c('Pvalb','Chandelier','Sst','Sst Chodl','Lamp5','Lamp5 Lhx6','Vip','Sncg','Pax6')]

marker_annot = rowAnnotation(subclass = top_markers$cell_type,
                                 col = list(subclass = celltype_color_palette))

col_fun = circlize::colorRamp2(breaks = c(0,.5,1), colors = c('blue','white','red'))
Heatmap(top_auroc_mat, name = 'AUROC', col = col_fun,
        cluster_rows = F, cluster_columns = F,
        right_annotation = marker_annot,
        )

#Fold change stat
top_FC_mat = mouse_subclass_meta_markers %>% filter(gene %in% top_markers$gene) %>% select(cell_type, gene, fold_change_detection) %>%
  tidyr::pivot_wider(names_from = gene, values_from = fold_change_detection) %>% tibble::column_to_rownames(var = 'cell_type')
top_FC_mat = t(as.matrix(top_FC_mat))
top_FC_mat = log2(top_FC_mat)

#Keep the same ordering
index = match(top_markers$gene, rownames(top_FC_mat))
top_FC_mat = top_FC_mat[index, ]
top_FC_mat = top_FC_mat[ , c('Pvalb','Chandelier','Sst','Sst Chodl','Lamp5','Lamp5 Lhx6','Vip','Sncg','Pax6')]

marker_annot = rowAnnotation(subclass = top_markers$cell_type,
                                 col = list(subclass = celltype_color_palette))

col_fun_2 = circlize::colorRamp2(breaks = c(-6,0,8), colors = c('blue','white','red'))
Heatmap(top_FC_mat, name = 'log2(FC)', col = col_fun_2,
        cluster_rows = F, cluster_columns = F,
        right_annotation = marker_annot
        )



de_recurrence_summary_df = mouse_subclass_meta_markers %>% filter(recurrence != 0) %>% group_by(cell_type, recurrence) %>% 
  summarise(mean_coverage = mean(auroc), mean_signal = mean(log2(fold_change_detection)))



ggplot(de_recurrence_summary_df %>% filter(cell_type %in% c('Vip','Pax6','Sncg')), aes(x = recurrence, y = mean_coverage, group = cell_type, color = cell_type)) + 
  geom_point(size = 3) + geom_line() +
  scale_color_manual(values = celltype_color_palette) + scale_x_reverse() +
  theme_bw()

ggplot(de_recurrence_summary_df %>% filter(cell_type %in% c('Vip','Pax6','Sncg')), aes(x = recurrence, y = mean_signal, group = cell_type, color = cell_type)) + 
  geom_point(size = 3) + geom_line() +
  scale_color_manual(values = celltype_color_palette) + scale_x_reverse() +
  theme_bw()


high_recur_de = mouse_subclass_meta_markers %>% filter(recurrence %in% c(7,8))
order_vec = high_recur_de %>% group_by(cell_type) %>% summarise(median_auroc = median(auroc)) %>% arrange(desc(median_auroc)) %>% pull(cell_type)
high_recur_de$cell_type = factor(high_recur_de$cell_type , levels = order_vec)
ggplot(high_recur_de, aes(x = cell_type, y = auroc, fill = cell_type)) + ggtitle('Consistent DE genes') + ylab('Coverage (AUROC)') +
  geom_boxplot() + scale_fill_manual(values = celltype_color_palette) + theme_bw()

order_vec = high_recur_de %>% group_by(cell_type) %>% summarise(median_fc = median(fold_change_detection)) %>% arrange(desc(median_fc)) %>% pull(cell_type)
high_recur_de$cell_type = factor(high_recur_de$cell_type , levels = order_vec)
ggplot(high_recur_de, aes(x = cell_type, y = log2(fold_change_detection), fill = cell_type)) + ggtitle('Consistent DE genes') + ylab('Signal-to-noise ( log2(FC detection) )') +
  geom_boxplot() + scale_fill_manual(values = celltype_color_palette) + theme_bw()



ggplot(mouse_subclass_meta_markers %>% filter(recurrence %in% c(1)), aes(x = cell_type, y = auroc, fill = cell_type)) +
  geom_boxplot() + scale_fill_manual(values = celltype_color_palette) + theme_bw()
ggplot(mouse_subclass_meta_markers %>% filter(recurrence %in% c(1)), aes(x = cell_type, y = log2(fold_change_detection), fill = cell_type)) +
  geom_boxplot() + scale_fill_manual(values = celltype_color_palette) + theme_bw()


```


```{r}


p1 = plot_pareto_summary(mouse_subclass_meta_markers %>% filter(cell_type %in% c('Pax6','Sncg')), min_recurrence=4) +
  theme_bw() + scale_color_manual(values = celltype_color_palette) + ggtitle('Homologous Mouse subclass markers') + ylim(.5, 1) + xlim(0,5)

get_pareto_markers(mouse_subclass_meta_markers, "Pax6", min_recurrence=4)

get_pareto_markers(mouse_subclass_meta_markers, "Sncg", min_recurrence=4)


p1


ggsave(plot = p1, filename = 'mouse_Sncg_Pax6_pareto_markers.pdf', path = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/meta_markers/', device = 'pdf', 
       useDingbats = F, family = 'Helvetica', height = 4, width = 6)

```
Save source data for plot
```{r}

save(mouse_subclass_meta_markers, file = '../../source_data/data_for_plots/mouse_meta_markers.Rdata')

```









```{r}
all_primate_markers = list(
    human_10x = read_markers("../metamarkers/markers_human_10x.csv.gz"),
    human_ss = read_markers("../metamarkers/markers_human_ss.csv.gz"),
    chimp_10x = read_markers("../metamarkers/markers_chimp_10x.csv.gz"),
    chimp_ss = read_markers("../metamarkers/markers_chimp_ss.csv.gz"),
    gorilla_10x = read_markers("../metamarkers/markers_gorilla_10x.csv.gz"),
    gorilla_ss = read_markers("../metamarkers/markers_gorilla_ss.csv.gz"),
    macaca_10x = read_markers("../metamarkers/markers_macaca_10x.csv.gz"),
    marmoset_10x= read_markers("../metamarkers/markers_marmoset_10x.csv.gz")
)


all_primate_markers$human_10x$cell_type[all_primate_markers$human_10x$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
all_primate_markers$human_ss$cell_type[all_primate_markers$human_ss$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
all_primate_markers$chimp_10x$cell_type[all_primate_markers$chimp_10x$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
all_primate_markers$chimp_ss$cell_type[all_primate_markers$chimp_ss$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
all_primate_markers$gorilla_10x$cell_type[all_primate_markers$gorilla_10x$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
all_primate_markers$gorilla_ss$cell_type[all_primate_markers$gorilla_ss$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
all_primate_markers$macaca_10x$cell_type[all_primate_markers$macaca_10x$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
all_primate_markers$marmoset_10x$cell_type[all_primate_markers$marmoset_10x$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'


primate_subclass_meta_markers = make_meta_markers(all_primate_markers, detailed_stats = TRUE)

primate_subclass_meta_markers %>% group_by(cell_type) %>% filter(rank <= 10)



```


```{r}

primate_subclass_meta_markers 

human_de_counts = rowSums(primate_subclass_meta_markers  %>% select(human_10x, human_ss))
chimp_de_counts = rowSums(primate_subclass_meta_markers  %>% select(chimp_10x, chimp_ss))
gorilla_de_counts = rowSums(primate_subclass_meta_markers  %>% select(gorilla_10x, gorilla_ss))
macaca_de_counts = primate_subclass_meta_markers$macaca_10x
marmoset_de_counts = primate_subclass_meta_markers$marmoset_10x

#Getting genes DE in at least 1 dataset per species
cross_species_de_index = human_de_counts >= 1 & chimp_de_counts >= 1 & gorilla_de_counts >= 1 & macaca_de_counts >= 1 & marmoset_de_counts >= 1



p2 = plot_pareto_summary(primate_subclass_meta_markers[cross_species_de_index, ] %>% filter(cell_type %in% c('Pax6','Sncg')), min_recurrence=0) +
  theme_bw() + scale_color_manual(values = celltype_color_palette) + ggtitle('Homologous Primate subclass markers') + ylim(.5, 1) + xlim(0,4)

get_pareto_markers(primate_subclass_meta_markers[cross_species_de_index, ], "Pax6", min_recurrence=0)

get_pareto_markers(primate_subclass_meta_markers[cross_species_de_index, ], "Sncg", min_recurrence=0)


p2

ggsave(plot = p2, filename = 'primate_Sncg_Pax6_pareto_markers.pdf', path = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/meta_markers/', device = 'pdf', 
       useDingbats = F, family = 'Helvetica', height = 4, width = 6)

```

Save source data for plot
```{r}

save(primate_subclass_meta_markers , file = '../../source_data/data_for_plots/primate_subclass_meta_markers.Rdata')

```


```{r}

all_species_markers = c(all_mouse_markers, all_primate_markers)

mammal_subclass_meta_markers = make_meta_markers(all_species_markers, detailed_stats = TRUE)

mouse_de_counts = rowSums(mammal_subclass_meta_markers %>% select(tasic18, zeng_10x_cells, zeng_10x_nuclei, zeng_smart_cells, zeng_smart_nuclei, zeng_10xv2_wba, zeng_10xv3_wba, macosko_wba))

human_de_counts = rowSums(mammal_subclass_meta_markers %>% select(human_10x, human_ss))
chimp_de_counts = rowSums(mammal_subclass_meta_markers %>% select(chimp_10x, chimp_ss))
gorilla_de_counts = rowSums(mammal_subclass_meta_markers %>% select(gorilla_10x, gorilla_ss))
macaca_de_counts = mammal_subclass_meta_markers$macaca_10x
marmoset_de_counts = mammal_subclass_meta_markers$marmoset_10x


#Getting genes DE in at least 1 dataset per species
cross_species_de_index = mouse_de_counts >= 1 & human_de_counts >= 1 & chimp_de_counts >= 1 & gorilla_de_counts >= 1 & macaca_de_counts >= 1 & marmoset_de_counts >= 1

p3 = plot_pareto_summary(mammal_subclass_meta_markers[cross_species_de_index, ] %>% filter(cell_type %in% c('Pax6','Sncg')), min_recurrence=6) +
  theme_bw() + scale_color_manual(values = celltype_color_palette) + ggtitle('Homologous Primate-Mouse subclass markers') + ylim(.5, 1) + xlim(0,4)


get_pareto_markers(mammal_subclass_meta_markers[cross_species_de_index, ], 'Pax6', min_recurrence = 6)
get_pareto_markers(mammal_subclass_meta_markers[cross_species_de_index, ], 'Sncg', min_recurrence = 6)

p3

ggsave(plot = p3, filename = 'mammalian_Sncg_Pax6_pareto_markers.pdf', path = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/meta_markers/', device = 'pdf', 
       useDingbats = F, family = 'Helvetica', height = 4, width = 6)

```

Save source data for plot
```{r}

save(mammal_subclass_meta_markers , file = '../../source_data/data_for_plots/mammal_subclass_meta_markers.Rdata')

```


```{r}


mammal_subclass_meta_markers %>% filter(gene %in% c("SORCS3","CCK","CXCL14","FREM1","PAX6" ) & cell_type == 'Pax6')

mammal_subclass_meta_markers %>% filter(gene %in% c("CXCL14","SYT6"  ,"COBLL1") & cell_type == 'Sncg')




```

Get the cross mammal pareto fronts for all the subclasses, put in supplement
Get the aurocs going from human->chimp, to human->mouse



```{r}

p4 = plot_pareto_summary(mammal_subclass_meta_markers[cross_species_de_index, ] %>% filter(cell_type == 'Vip'),  min_recurrence=6) +
  theme_bw() + scale_color_manual(values = celltype_color_palette) + ggtitle('Homologous Primate-Mouse Vip markers') + ylim(.5, 1) + xlim(0,4)
p5 = plot_pareto_summary(mammal_subclass_meta_markers[cross_species_de_index, ] %>% filter(cell_type == 'Lamp5'),  min_recurrence=6) +
  theme_bw() + scale_color_manual(values = celltype_color_palette) + ggtitle('Homologous Primate-Mouse Lamp5 markers') + ylim(.5, 1) + xlim(0,4)
p6 = plot_pareto_summary(mammal_subclass_meta_markers[cross_species_de_index, ] %>% filter(cell_type == 'Lamp5 Lhx6'),  min_recurrence=6) +
  theme_bw() + scale_color_manual(values = celltype_color_palette) + ggtitle('Homologous Primate-Mouse Lamp5 Lhx6 markers') + ylim(.5, 1) + xlim(0,4)

p7 = plot_pareto_summary(mammal_subclass_meta_markers[cross_species_de_index, ] %>% filter(cell_type == 'Sst'),  min_recurrence=6) +
  theme_bw() + scale_color_manual(values = celltype_color_palette) + ggtitle('Homologous Primate-Mouse Sst markers') + ylim(.5, 1) + xlim(0,4)
p8 = plot_pareto_summary(mammal_subclass_meta_markers[cross_species_de_index, ] %>% filter(cell_type == 'Sst Chodl'),  min_recurrence=6) +
  theme_bw() + scale_color_manual(values = celltype_color_palette) + ggtitle('Homologous Primate-Mouse Sst Chodl markers') + ylim(.5, 1) + xlim(0,4)
p9 = plot_pareto_summary(mammal_subclass_meta_markers[cross_species_de_index, ] %>% filter(cell_type == 'Pvalb'),  min_recurrence=6) +
  theme_bw() + scale_color_manual(values = celltype_color_palette) + ggtitle('Homologous Primate-Mouse Pvalb markers') + ylim(.5, 1) + xlim(0,4)
p10 = plot_pareto_summary(mammal_subclass_meta_markers[cross_species_de_index, ] %>% filter(cell_type == 'Chandelier'),  min_recurrence=6) +
  theme_bw() + scale_color_manual(values = celltype_color_palette) + ggtitle('Homologous Primate-Mouse Chandelier markers') + ylim(.5, 1) + xlim(0,4)

p4
p5
p6
p7
p8
p9
p10

get_pareto_markers(mammal_subclass_meta_markers[cross_species_de_index, ], 'Pax6', min_recurrence = 6)
get_pareto_markers(mammal_subclass_meta_markers[cross_species_de_index, ], 'Sncg', min_recurrence = 6)
get_pareto_markers(mammal_subclass_meta_markers[cross_species_de_index, ], 'Vip', min_recurrence = 6)
get_pareto_markers(mammal_subclass_meta_markers[cross_species_de_index, ], 'Lamp5', min_recurrence = 6)
get_pareto_markers(mammal_subclass_meta_markers[cross_species_de_index, ], 'Lamp5 Lhx6', min_recurrence = 6)
get_pareto_markers(mammal_subclass_meta_markers[cross_species_de_index, ], 'Sst', min_recurrence = 6)
get_pareto_markers(mammal_subclass_meta_markers[cross_species_de_index, ], 'Sst Chodl', min_recurrence = 6)
get_pareto_markers(mammal_subclass_meta_markers[cross_species_de_index, ], 'Pvalb', min_recurrence = 6)
get_pareto_markers(mammal_subclass_meta_markers[cross_species_de_index, ], 'Chandelier', min_recurrence = 6)





```


```{r}
#human - chimp
speciesX_subclass_meta_markers = make_meta_markers(all_species_markers[names(all_species_markers) %in% c('human_10x','human_ss','chimp_10x','chimp_ss')], detailed_stats = TRUE)

human_de_counts = rowSums(speciesX_subclass_meta_markers %>% select(human_10x, human_ss))
chimp_de_counts = rowSums(speciesX_subclass_meta_markers %>% select(chimp_10x, chimp_ss))

#Getting genes DE in at least 1 dataset per species
cross_species_de_index =  human_de_counts >= 1 & chimp_de_counts >= 1
human_chimp_de_num_df = speciesX_subclass_meta_markers[cross_species_de_index, ] %>% group_by(cell_type) %>% summarise(n = n()) %>% mutate(label = 'Human-Chimp')

#Human - chimp - gorilla
speciesX_subclass_meta_markers = make_meta_markers(all_species_markers[names(all_species_markers) %in% c('human_10x','human_ss',
                                                                                                         'gorilla_10x','gorilla_ss')], 
                                                   detailed_stats = TRUE)

human_de_counts = rowSums(speciesX_subclass_meta_markers %>% select(human_10x, human_ss))
gorilla_de_counts = rowSums(speciesX_subclass_meta_markers %>% select(gorilla_10x, gorilla_ss))

#Getting genes DE in at least 1 dataset per species
cross_species_de_index =  human_de_counts >= 1 & gorilla_de_counts >= 1
human_gorilla_de_num_df = speciesX_subclass_meta_markers[cross_species_de_index, ] %>% group_by(cell_type) %>% summarise(n = n()) %>% mutate(label = 'Human-Gorilla')



#Human - macaca
speciesX_subclass_meta_markers = make_meta_markers(all_species_markers[names(all_species_markers) %in% c('human_10x','human_ss','macaca_10x')], 
                                                   detailed_stats = TRUE)

human_de_counts = rowSums(speciesX_subclass_meta_markers %>% select(human_10x, human_ss))
macaca_de_counts = speciesX_subclass_meta_markers$macaca_10x

#Getting genes DE in at least 1 dataset per species
cross_species_de_index =  human_de_counts >= 1 & macaca_de_counts >= 1
human_macaca_de_num_df = speciesX_subclass_meta_markers[cross_species_de_index, ] %>% group_by(cell_type) %>% summarise(n = n()) %>% mutate(label = 'Human-Macaca')


#Human - marmoset
speciesX_subclass_meta_markers = make_meta_markers(all_species_markers[names(all_species_markers) %in% c('human_10x','human_ss','marmoset_10x')], 
                                                   detailed_stats = TRUE)

human_de_counts = rowSums(speciesX_subclass_meta_markers %>% select(human_10x, human_ss))

marmoset_de_counts = speciesX_subclass_meta_markers$marmoset_10x

#Getting genes DE in at least 1 dataset per species
cross_species_de_index =  human_de_counts >= 1 & marmoset_de_counts >= 1
human_marmoset_de_num_df = speciesX_subclass_meta_markers[cross_species_de_index, ] %>% group_by(cell_type) %>% summarise(n = n()) %>% mutate(label = 'Human-Marmoset')


#Human - mouse
speciesX_subclass_meta_markers = make_meta_markers(all_species_markers[names(all_species_markers) %in% 
                                                                         c('human_10x','human_ss','tasic18', 'zeng_10x_cells', 'zeng_10x_nuclei', 
                                                                           'zeng_smart_nuclei', 'zeng_smart_cells', 'zeng_10xv2_wba', 'zeng_10xv3_wba', 'macosko_wba')], 
                                                   detailed_stats = TRUE)

human_de_counts = rowSums(speciesX_subclass_meta_markers %>% select(human_10x, human_ss))
mouse_de_counts = rowSums(speciesX_subclass_meta_markers %>% select(tasic18, zeng_10x_cells, zeng_10x_nuclei, zeng_smart_cells, zeng_smart_nuclei, zeng_10xv2_wba, zeng_10xv3_wba, macosko_wba))

#Getting genes DE in at least 1 dataset per species
cross_species_de_index =  human_de_counts >= 1 & mouse_de_counts >= 1
human_mouse_de_num_df = speciesX_subclass_meta_markers[cross_species_de_index, ] %>% group_by(cell_type) %>% summarise(n = n()) %>% mutate(label = 'Human-Mouse')


de_divg_df = do.call('rbind', list(human_chimp_de_num_df, human_gorilla_de_num_df, human_macaca_de_num_df, human_marmoset_de_num_df, human_mouse_de_num_df))


p11 = ggplot(de_divg_df, aes(x = label, y = n)) + geom_boxplot() + 
  geom_point(aes(color = cell_type), size = 3) + 
  geom_line(aes(group = cell_type, color = cell_type), linetype = 'dashed') +
  scale_color_manual(values = celltype_color_palette) + theme_bw() +
  ylab('# of conserved cell-type marker genes')

p11
```




```{r}
#human - chimp
speciesX_subclass_meta_markers = make_meta_markers(all_species_markers[names(all_species_markers) %in% c('human_10x','human_ss','chimp_10x','chimp_ss')], detailed_stats = TRUE)

human_de_counts = rowSums(speciesX_subclass_meta_markers %>% select(human_10x, human_ss))
chimp_de_counts = rowSums(speciesX_subclass_meta_markers %>% select(chimp_10x, chimp_ss))

#Getting genes DE in at least 1 dataset per species
cross_species_de_index =  human_de_counts >= 1 & chimp_de_counts >= 1
human_chimp_de_num_df = speciesX_subclass_meta_markers[cross_species_de_index, ] %>% group_by(cell_type) %>% summarise(n = n()) %>% mutate(label = 'Human-C')

#Human - chimp - gorilla
speciesX_subclass_meta_markers = make_meta_markers(all_species_markers[names(all_species_markers) %in% c('human_10x','human_ss','chimp_10x','chimp_ss',
                                                                                                         'gorilla_10x','gorilla_ss')], 
                                                   detailed_stats = TRUE)

human_de_counts = rowSums(speciesX_subclass_meta_markers %>% select(human_10x, human_ss))
chimp_de_counts = rowSums(speciesX_subclass_meta_markers %>% select(chimp_10x, chimp_ss))
gorilla_de_counts = rowSums(speciesX_subclass_meta_markers %>% select(gorilla_10x, gorilla_ss))

#Getting genes DE in at least 1 dataset per species
cross_species_de_index =  human_de_counts >= 1 & chimp_de_counts >= 1 & gorilla_de_counts >= 1
human_gorilla_de_num_df = speciesX_subclass_meta_markers[cross_species_de_index, ] %>% group_by(cell_type) %>% summarise(n = n()) %>% mutate(label = 'Human-C-G')



#Human - macaca
speciesX_subclass_meta_markers = make_meta_markers(all_species_markers[names(all_species_markers) %in% c('human_10x','human_ss','chimp_10x','chimp_ss',
                                                                                                         'gorilla_10x','gorilla_ss','macaca_10x')], 
                                                   detailed_stats = TRUE)

human_de_counts = rowSums(speciesX_subclass_meta_markers %>% select(human_10x, human_ss))
chimp_de_counts = rowSums(speciesX_subclass_meta_markers %>% select(chimp_10x, chimp_ss))
gorilla_de_counts = rowSums(speciesX_subclass_meta_markers %>% select(gorilla_10x, gorilla_ss))
macaca_de_counts = speciesX_subclass_meta_markers$macaca_10x

#Getting genes DE in at least 1 dataset per species
cross_species_de_index =  human_de_counts >= 1 & chimp_de_counts >= 1 & gorilla_de_counts >= 1 & macaca_de_counts >= 1
human_macaca_de_num_df = speciesX_subclass_meta_markers[cross_species_de_index, ] %>% group_by(cell_type) %>% summarise(n = n()) %>% mutate(label = 'Human-C-G-Mac')


#Human - marmoset
speciesX_subclass_meta_markers = make_meta_markers(all_species_markers[names(all_species_markers) %in% c('human_10x','human_ss','chimp_10x','chimp_ss',
                                                                                                         'gorilla_10x','gorilla_ss','macaca_10x',
                                                                                                         'marmoset_10x')], 
                                                   detailed_stats = TRUE)

human_de_counts = rowSums(speciesX_subclass_meta_markers %>% select(human_10x, human_ss))
chimp_de_counts = rowSums(speciesX_subclass_meta_markers %>% select(chimp_10x, chimp_ss))
gorilla_de_counts = rowSums(speciesX_subclass_meta_markers %>% select(gorilla_10x, gorilla_ss))
macaca_de_counts = speciesX_subclass_meta_markers$macaca_10x
marmoset_de_counts = speciesX_subclass_meta_markers$marmoset_10x

#Getting genes DE in at least 1 dataset per species
cross_species_de_index =  human_de_counts >= 1 & chimp_de_counts >= 1 & gorilla_de_counts >= 1 & macaca_de_counts >= 1 & marmoset_de_counts >= 1
human_marmoset_de_num_df = speciesX_subclass_meta_markers[cross_species_de_index, ] %>% group_by(cell_type) %>% summarise(n = n()) %>% mutate(label = 'Human-C-G-Mac-Mar')


#Human - mouse
speciesX_subclass_meta_markers = make_meta_markers(all_species_markers, 
                                                   detailed_stats = TRUE)

human_de_counts = rowSums(speciesX_subclass_meta_markers %>% select(human_10x, human_ss))
chimp_de_counts = rowSums(speciesX_subclass_meta_markers %>% select(chimp_10x, chimp_ss))
gorilla_de_counts = rowSums(speciesX_subclass_meta_markers %>% select(gorilla_10x, gorilla_ss))
macaca_de_counts = speciesX_subclass_meta_markers$macaca_10x
marmoset_de_counts = speciesX_subclass_meta_markers$marmoset_10x
mouse_de_counts = rowSums(speciesX_subclass_meta_markers %>% select(tasic18, zeng_10x_cells, zeng_10x_nuclei, zeng_smart_cells, zeng_smart_nuclei, zeng_10xv2_wba, zeng_10xv3_wba, macosko_wba))

#Getting genes DE in at least 1 dataset per species
cross_species_de_index =  human_de_counts >= 1 & chimp_de_counts >= 1 & gorilla_de_counts >= 1 & macaca_de_counts >= 1 & marmoset_de_counts >= 1 & mouse_de_counts >= 1
human_mouse_de_num_df = speciesX_subclass_meta_markers[cross_species_de_index, ] %>% group_by(cell_type) %>% summarise(n = n()) %>% mutate(label = 'Human-C-G-Mac-Mar-Mou')


de_divg_df = do.call('rbind', list(human_chimp_de_num_df, human_gorilla_de_num_df, human_macaca_de_num_df, human_marmoset_de_num_df, human_mouse_de_num_df))

de_divg_df

p12 = ggplot(de_divg_df, aes(x = label, y = n)) + geom_boxplot() + 
  geom_point(aes(color = cell_type), size = 3) + 
  geom_line(aes(group = cell_type, color = cell_type), linetype = 'dashed') +
  scale_color_manual(values = celltype_color_palette) + theme_bw() +
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +
  ylab('# of conserved cell-type marker genes')

p12

```


ggsave(plot = p3, filename = 'mammalian_Sncg_Pax6_pareto_markers.pdf', path = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/meta_markers/', device = 'pdf', 
       useDingbats = F, family = 'Helvetica', height = 4, width = 6)

```{r}


ggsave(plot = p4, filename = 'mammalian_Vip_pareto_markers.pdf', path = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/meta_markers/', device = 'pdf', 
       useDingbats = F, family = 'Helvetica', height = 4, width = 6)
ggsave(plot = p5, filename = 'mammalian_Lamp5_pareto_markers.pdf', path = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/meta_markers/', device = 'pdf', 
       useDingbats = F, family = 'Helvetica', height = 4, width = 6)
ggsave(plot = p6, filename = 'mammalian_Lamp5_Lhx6_pareto_markers.pdf', path = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/meta_markers/', device = 'pdf', 
       useDingbats = F, family = 'Helvetica', height = 4, width = 6)
ggsave(plot = p7, filename = 'mammalian_Sst_pareto_markers.pdf', path = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/meta_markers/', device = 'pdf', 
       useDingbats = F, family = 'Helvetica', height = 4, width = 6)
ggsave(plot = p8, filename = 'mammalian_Sst_Chodl_pareto_markers.pdf', path = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/meta_markers/', device = 'pdf', 
       useDingbats = F, family = 'Helvetica', height = 4, width = 6)
ggsave(plot = p9, filename = 'mammalian_Pvalb_pareto_markers.pdf', path = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/meta_markers/', device = 'pdf', 
       useDingbats = F, family = 'Helvetica', height = 4, width = 6)
ggsave(plot = p10, filename = 'mammalian_Chandelier_pareto_markers.pdf', path = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/meta_markers/', device = 'pdf', 
       useDingbats = F, family = 'Helvetica', height = 4, width = 6)

ggsave(plot = p11, filename = 'conserved_de_human_species_pair.pdf', path = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/meta_markers/', device = 'pdf', 
       useDingbats = F, family = 'Helvetica', height = 4, width = 6)
ggsave(plot = p12, filename = 'conserved_de_human_species_cuml.pdf', path = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/meta_markers/', device = 'pdf', 
       useDingbats = F, family = 'Helvetica', height = 4, width = 6)

```


```{r}
DefaultAssay(interneuron.combined) = 'originalexp'

all_mouse_interneurons = as.SingleCellExperiment(interneuron.combined)
names(assays(all_mouse_interneurons )) = c('counts','cpm')

#get the top primate and mouse MetaMarkers
top_primate_markers = primate_subclass_meta_markers %>% filter(rank <= 50)


#Get the primate enrichments, summarise by the mean enrichment per de novo cluster
ct_scores_mouse = score_cells(log1p(cpm(all_mouse_interneurons)), top_primate_markers)
ct_enrichment_mouse = compute_marker_enrichment(ct_scores_mouse)


#Add a dataset and primate subclass label
study_primate_subclass = paste(all_mouse_interneurons$study, all_mouse_interneurons$primate_subclass, sep = '|')
colnames(ct_enrichment_mouse) = study_primate_subclass

ct_enrichment_mouse_df = reshape2::melt(t(ct_enrichment_mouse), varnames = c('study_primate_subclass','primate_markers'), value.name = c('enrichment'))
ct_enrichment_mouse_df = ct_enrichment_mouse_df %>% group_by(study_primate_subclass, primate_markers) %>% 
  summarise(mean_enrichment = mean(enrichment), sd_enrichment = sd(enrichment))




ct_enrichment_mouse_df$mapped_primate_subclass = sapply(strsplit(as.character(ct_enrichment_mouse_df$study_primate_subclass), split = '|', fixed = T), '[[', 2)
ct_enrichment_mouse_df$mapped_primate_subclass = factor(ct_enrichment_mouse_df$mapped_primate_subclass, 
                                                          levels = c('Sst','Sst Chodl','Pvalb','Chandelier','Lamp5 Lhx6','Lamp5','Vip','Pax6','Sncg'))


plotting_df = ct_enrichment_mouse_df %>% filter(primate_markers == 'all|Vip')

order_vec = plotting_df %>% arrange(match(mapped_primate_subclass, c('Sst','Sst Chodl','Pvalb','Chandelier','Lamp5 Lhx6','Lamp5','Vip','Pax6','Sncg'))) %>%
  pull(study_primate_subclass)

plotting_df$study_primate_subclass = factor(plotting_df$study_primate_subclass, levels = order_vec)

mouse_vip_enrich_dot_plot = ggplot(plotting_df, aes(x = study_primate_subclass, y = mean_enrichment, color = mapped_primate_subclass)) +
  geom_point(size = 3, show.legend = F) + 
  geom_linerange(aes(ymin = mean_enrichment - sd_enrichment, ymax = mean_enrichment + sd_enrichment),
                 show.legend = F) +
  geom_hline(yintercept = 1, linetype = 'dashed', color = 'black') +
  ylab('Mean Primate Vip enrichment') + xlab('Mouse dataset Primate subclasses') + ylim(0,3) +
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +
  scale_color_manual(values = celltype_color_palette) +
  theme_bw() + theme(panel.border = element_blank(), axis.line = element_line(),
                     panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(),axis.text.x=element_blank())

mouse_vip_enrich_dot_plot
#save source data 
save(plotting_df, file = '../../source_data/data_for_plots/mouse_Vip_dot_plot_primate_enrich.Rdata')


plotting_df = ct_enrichment_mouse_df %>% filter(primate_markers == 'all|Pax6')

order_vec = plotting_df %>% arrange(match(mapped_primate_subclass, c('Sst','Sst Chodl','Pvalb','Chandelier','Lamp5 Lhx6','Lamp5','Vip','Pax6','Sncg'))) %>%
  pull(study_primate_subclass)

plotting_df$study_primate_subclass = factor(plotting_df$study_primate_subclass, levels = order_vec)

mouse_pax6_enrich_dot_plot = ggplot(plotting_df, aes(x = study_primate_subclass, y = mean_enrichment, color = mapped_primate_subclass)) +
  geom_point(size = 3, show.legend = F) + 
  geom_linerange(aes(ymin = mean_enrichment - sd_enrichment, ymax = mean_enrichment + sd_enrichment),
                 show.legend = F) +
  geom_hline(yintercept = 1, linetype = 'dashed', color = 'black') +
  ylab('Mean Primate Pax6 enrichment') + xlab('Mouse dataset Primate subclasses') + ylim(0,3) +
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +
  scale_color_manual(values = celltype_color_palette) +
  theme_bw() + theme(panel.border = element_blank(), axis.line = element_line(),
                     panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(),axis.text.x=element_blank())

mouse_pax6_enrich_dot_plot
#save source data 
save(plotting_df, file = '../../source_data/data_for_plots/mouse_Pax6_dot_plot_primate_enrich.Rdata')

plotting_df = ct_enrichment_mouse_df %>% filter(primate_markers == 'all|Sncg')

order_vec = plotting_df %>% arrange(match(mapped_primate_subclass, c('Sst','Sst Chodl','Pvalb','Chandelier','Lamp5 Lhx6','Lamp5','Vip','Pax6','Sncg'))) %>%
  pull(study_primate_subclass)

plotting_df$study_primate_subclass = factor(plotting_df$study_primate_subclass, levels = order_vec)

mouse_sncg_enrich_dot_plot = ggplot(plotting_df, aes(x = study_primate_subclass, y = mean_enrichment, color = mapped_primate_subclass)) +
  geom_point(size = 3, show.legend = F) + 
  geom_linerange(aes(ymin = mean_enrichment - sd_enrichment, ymax = mean_enrichment + sd_enrichment),
                 show.legend = F) +
  geom_hline(yintercept = 1, linetype = 'dashed', color = 'black') +
  ylab('Mean Primate Sncg enrichment') + xlab('Mouse dataset Primate subclasses') + ylim(0,3) +
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +
  scale_color_manual(values = celltype_color_palette) +
  theme_bw() + theme(panel.border = element_blank(), axis.line = element_line(),
                     panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(),axis.text.x=element_blank())

#save source data 
save(plotting_df, file = '../../source_data/data_for_plots/mouse_Sncg_dot_plot_primate_enrich.Rdata')

mouse_sncg_enrich_dot_plot
mouse_pax6_enrich_dot_plot
mouse_vip_enrich_dot_plot


ggsave(plot = mouse_sncg_enrich_dot_plot, filename = 'dot_mouse_sncg_mean_primate_enrich.pdf', path = '../graphs/non_allen_data/', device = 'pdf',
       useDingbats = F, height = 2, width = 6)
ggsave(plot = mouse_pax6_enrich_dot_plot, filename = 'dot_mouse_pax6_mean_primate_enrich.pdf', path = '../graphs/non_allen_data/', device = 'pdf',
       useDingbats = F, height = 2, width = 6)
ggsave(plot = mouse_vip_enrich_dot_plot, filename = 'dot_mouse_vip_mean_primate_enrich.pdf', path = '../graphs/non_allen_data/', device = 'pdf',
       useDingbats = F, height = 2, width = 6)



#And the mouse Sncg markers
top_mouse_markers = mouse_subclass_meta_markers %>% filter(rank <= 50)

#Get the primate enrichments, summarise by the mean enrichment per de novo cluster
ct_scores_mouse = score_cells(log1p(cpm(all_mouse_interneurons)), top_mouse_markers)
ct_enrichment_mouse = compute_marker_enrichment(ct_scores_mouse)

#Add a dataset and primate subclass label
study_primate_subclass = paste(all_mouse_interneurons$study, all_mouse_interneurons$primate_subclass, sep = '|')
colnames(ct_enrichment_mouse) = study_primate_subclass

ct_enrichment_mouse_df = reshape2::melt(t(ct_enrichment_mouse), varnames = c('study_primate_subclass','mouse_markers'), value.name = c('enrichment'))
ct_enrichment_mouse_df = ct_enrichment_mouse_df %>% group_by(study_primate_subclass, mouse_markers) %>% 
  summarise(mean_enrichment = mean(enrichment), sd_enrichment = sd(enrichment))


ct_enrichment_mouse_df$mapped_primate_subclass = sapply(strsplit(as.character(ct_enrichment_mouse_df$study_primate_subclass), split = '|', fixed = T), '[[', 2)
ct_enrichment_mouse_df$mapped_primate_subclass = factor(ct_enrichment_mouse_df$mapped_primate_subclass, 
                                                          levels = c('Sst','Sst Chodl','Pvalb','Chandelier','Lamp5 Lhx6','Lamp5','Vip','Pax6','Sncg'))

plotting_df = ct_enrichment_mouse_df %>% filter(mouse_markers == 'all|Sncg')

order_vec = plotting_df %>% arrange(match(mapped_primate_subclass, c('Sst','Sst Chodl','Pvalb','Chandelier','Lamp5 Lhx6','Lamp5','Vip','Pax6','Sncg'))) %>%
  pull(study_primate_subclass)

plotting_df$study_primate_subclass = factor(plotting_df$study_primate_subclass, levels = order_vec)

mouse_mouse_sncg_enrich_dot_plot = ggplot(plotting_df, aes(x = study_primate_subclass, y = mean_enrichment, color = mapped_primate_subclass)) +
  geom_point(size = 3, show.legend = F) + 
  geom_linerange(aes(ymin = mean_enrichment - sd_enrichment, ymax = mean_enrichment + sd_enrichment),
                 show.legend = F) +
  geom_hline(yintercept = 1, linetype = 'dashed', color = 'black') +
  ylab('Mean Mouse Sncg enrichment') + xlab('Mouse dataset Primate subclasses') + ylim(0,3.7) +
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +
  scale_color_manual(values = celltype_color_palette) +
  theme_bw() + theme(panel.border = element_blank(), axis.line = element_line(),
                     panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(),axis.text.x=element_blank())

mouse_mouse_sncg_enrich_dot_plot

#save source data 
save(plotting_df, file = '../../source_data/data_for_plots/mouse_Sncg_dot_plot_mouse_enrich.Rdata')

ggsave(plot = mouse_mouse_sncg_enrich_dot_plot, filename = 'dot_mouse_sncg_mean_mouse_enrich.pdf', path = '../graphs/non_allen_data/', device = 'pdf',
       useDingbats = F, height = 2, width = 6)


plotting_df %>% filter(mapped_primate_subclass == 'Sncg')


```





################################

Comparing markers across the original mouse subclasses and the primate subclasses

Just start with the mouse original metamarkers and the primate metamarkers

Aiming to show what does and what does not overlap between the subclass labels

##################################


Need to compute mouse subclass markers using the non-orthologous filtered data, since we are comparing marker transferability before the mapping to the primate subclasses.

No filtering for the high-orthology clusters

```{r}
tasic18 = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/tasic18_gaba_primate_MN.rds')
zeng_10x_cells = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/zeng_10x_cells_gaba_primate_MN.rds')
zeng_10x_nuclei = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/zeng_10x_nuclei_gaba_primate_MN.rds')
zeng_smart_cells = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/zeng_smart_cells_gaba_primate_MN.rds')
zeng_smart_nuclei = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/zeng_smart_nuclei_gaba_primate_MN.rds')
zeng_10xv2_wba = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/zeng_wb_atlas_10xv2_gaba_primate_MN.rds')
zeng_10xv3_wba = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/zeng_wb_atlas_10xv3_gaba_primate_MN.rds')
macosko_wba = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/macosko_wb_atlas_gaba_primate_MN.rds')
```



Rename the old Allen institute subclasses

```{r}

zeng_10x_cells$cell_subclass = as.character(zeng_10x_cells$cell_subclass)
zeng_10x_cells$cell_cluster = as.character(zeng_10x_cells$cell_cluster)

zeng_smart_cells$cell_subclass = as.character(zeng_smart_cells$cell_subclass)
zeng_smart_cells$cell_cluster = as.character(zeng_smart_cells$cell_cluster)

zeng_smart_nuclei$cell_subclass = as.character(zeng_smart_nuclei$cell_subclass)
zeng_smart_nuclei$cell_cluster = as.character(zeng_smart_nuclei$cell_cluster)

zeng_10x_nuclei$cell_subclass = as.character(zeng_10x_nuclei$cell_subclass)
zeng_10x_nuclei$cell_cluster = as.character(zeng_10x_nuclei$cell_cluster)

zeng_10xv2_wba$cell_cluster = as.character(zeng_10xv2_wba$cell_cluster)
zeng_10xv3_wba$cell_cluster = as.character(zeng_10xv3_wba$cell_cluster)

tasic18$cell_subclass[tasic18$cell_cluster == 'Lamp5 Lhx6'] = 'Lamp5 Lhx6'
tasic18$cell_subclass[tasic18$cell_cluster == 'Sst Chodl'] = 'Sst Chodl'
tasic18$cell_subclass[tasic18$cell_cluster == 'Pvalb Vipr2'] = 'Chandelier'
tasic18$cell_subclass[tasic18$cell_subclass == 'Serpinf1'] = 'Vip'

zeng_10x_cells$cell_subclass[zeng_10x_cells$cell_cluster == 'Lamp5 Lhx6'] = 'Lamp5 Lhx6'
zeng_10x_cells$cell_subclass[zeng_10x_cells$cell_cluster == 'Sst Chodl'] = 'Sst Chodl'
zeng_10x_cells$cell_subclass[zeng_10x_cells$cell_cluster == 'Pvalb Vipr2'] = 'Chandelier'

zeng_10x_nuclei$cell_subclass[zeng_10x_nuclei$cell_cluster == 'Lamp5 Lhx6'] = 'Lamp5 Lhx6'
zeng_10x_nuclei$cell_subclass[zeng_10x_nuclei$cell_cluster == 'Sst Chodl'] = 'Sst Chodl'
zeng_10x_nuclei$cell_subclass[zeng_10x_nuclei$cell_cluster == 'Pvalb Vipr2'] = 'Chandelier'

zeng_smart_nuclei$cell_subclass[zeng_smart_nuclei$cell_cluster == 'Lamp5 Lhx6'] = 'Lamp5 Lhx6'
zeng_smart_nuclei$cell_subclass[zeng_smart_nuclei$cell_cluster == 'Sst Chodl'] = 'Sst Chodl'
zeng_smart_nuclei$cell_subclass[zeng_smart_nuclei$cell_cluster == 'Pvalb Vipr2'] = 'Chandelier'

zeng_smart_cells$cell_subclass[zeng_smart_cells$cell_cluster == 'Lamp5 Lhx6'] = 'Lamp5 Lhx6'
zeng_smart_cells$cell_subclass[zeng_smart_cells$cell_cluster == 'Sst Chodl'] = 'Sst Chodl'
zeng_smart_cells$cell_subclass[zeng_smart_cells$cell_cluster == 'Pvalb Vipr2'] = 'Chandelier'


zeng_10xv2_wba$cell_subclass[zeng_10xv2_wba$cell_subclass == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
zeng_10xv3_wba$cell_subclass[zeng_10xv3_wba$cell_subclass == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'


```


Compute the mouse subclass metamarkers


```{r}

markers_tasic18 = compute_markers(assay(tasic18, "cpm"), tasic18$cell_subclass)
markers_zeng_10x_cells = compute_markers(assay(zeng_10x_cells, "cpm"), zeng_10x_cells$cell_subclass)
markers_zeng_10x_nuclei = compute_markers(assay(zeng_10x_nuclei, "cpm"), zeng_10x_nuclei$cell_subclass)
markers_zeng_smart_nuclei = compute_markers(assay(zeng_smart_nuclei, "cpm"), zeng_smart_nuclei$cell_subclass)
markers_zeng_smart_cells = compute_markers(assay(zeng_smart_cells, "cpm"), zeng_smart_cells$cell_subclass)
markers_zeng_10xv2_wba = compute_markers(assay(zeng_10xv2_wba, "cpm"), zeng_10xv2_wba$cell_subclass)
markers_zeng_10xv3_wba = compute_markers(assay(zeng_10xv3_wba, "cpm"), zeng_10xv3_wba$cell_subclass)
markers_macosko_wba = compute_markers(assay(macosko_wba, "cpm"), macosko_wba$cell_subclass)


original_mouse_markers = list(
    tasic18 = markers_tasic18,
    zeng_10x_cells = markers_zeng_10x_cells,
    zeng_10x_nuclei = markers_zeng_10x_nuclei ,
    zeng_smart_nuclei = markers_zeng_smart_nuclei,
    zeng_smart_cells = markers_zeng_smart_cells ,
    zeng_10xv2_wba = markers_zeng_10xv2_wba,
    zeng_10xv3_wba = markers_zeng_10xv3_wba,
    macosko_wba = markers_macosko_wba
)


original_mouse_subclass_meta_markers = make_meta_markers(original_mouse_markers, detailed_stats = TRUE)
original_mouse_subclass_meta_markers %>% group_by(cell_type) %>% filter(rank <= 10)

original_mouse_subclass_meta_markers %>% filter(gene == 'SNCG')
```



And the primate markers

```{r}
all_primate_markers = list(
    human_10x = read_markers("../metamarkers/markers_human_10x.csv.gz"),
    human_ss = read_markers("../metamarkers/markers_human_ss.csv.gz"),
    chimp_10x = read_markers("../metamarkers/markers_chimp_10x.csv.gz"),
    chimp_ss = read_markers("../metamarkers/markers_chimp_ss.csv.gz"),
    gorilla_10x = read_markers("../metamarkers/markers_gorilla_10x.csv.gz"),
    gorilla_ss = read_markers("../metamarkers/markers_gorilla_ss.csv.gz"),
    macaca_10x = read_markers("../metamarkers/markers_macaca_10x.csv.gz"),
    marmoset_10x= read_markers("../metamarkers/markers_marmoset_10x.csv.gz")
)


all_primate_markers$human_10x$cell_type[all_primate_markers$human_10x$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
all_primate_markers$human_ss$cell_type[all_primate_markers$human_ss$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
all_primate_markers$chimp_10x$cell_type[all_primate_markers$chimp_10x$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
all_primate_markers$chimp_ss$cell_type[all_primate_markers$chimp_ss$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
all_primate_markers$gorilla_10x$cell_type[all_primate_markers$gorilla_10x$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
all_primate_markers$gorilla_ss$cell_type[all_primate_markers$gorilla_ss$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
all_primate_markers$macaca_10x$cell_type[all_primate_markers$macaca_10x$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
all_primate_markers$marmoset_10x$cell_type[all_primate_markers$marmoset_10x$cell_type == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'


primate_subclass_meta_markers = make_meta_markers(all_primate_markers, detailed_stats = TRUE)

primate_subclass_meta_markers %>% group_by(cell_type) %>% filter(rank <= 10)



```



take the top 10 primate markers for a subclass and just see where they are in the mouse DE across all subclasses


```{r}

top_primate_markers = primate_subclass_meta_markers %>% filter(cell_type == 'Pax6' & rank <= 10) %>% pull(gene) 


test_df = original_mouse_subclass_meta_markers %>% filter(gene %in% top_primate_markers) #%>% group_by(cell_type) %>%
  #summarise(mean_auroc = mean(auroc), mean_fc = mean(fold_change)) %>% arrange(desc(mean_auroc))


p1 = ggplot(test_df , aes(x = log2(fold_change), y = auroc, color = cell_type)) + geom_point() +
  scale_color_manual(values = celltype_color_palette) 
  


ggExtra::ggMarginal(p1, groupColour = TRUE, groupFill = TRUE)


#VIPR2 or NKX2-1
primate_subclass_meta_markers %>% filter(gene == 'VIPR2')

original_mouse_subclass_meta_markers %>% filter(gene == 'VIPR2')
```



Try the name sake genes, comparing the primate and the mouse aurocs

```{r}

#Get the primate name sake gene DE
primate_name_sake_genes = primate_subclass_meta_markers %>% 
  filter(gene %in% c('PVALB','SST','CHODL','NKX2-1','LHX6','LAMP5','VIP','SNCG','PAX6') & cell_type != 'Pax6') %>%
  mutate(primate_auroc = auroc) %>% select(cell_type, gene, primate_auroc)

primate_name_sake_genes$full_label = paste(primate_name_sake_genes$cell_type, primate_name_sake_genes$gene, sep = '_' )

#Get the mouse name sake gene DE
mouse_name_sake_genes = original_mouse_subclass_meta_markers %>% filter(gene %in% c('PVALB','SST','CHODL','NKX2-1','LHX6','LAMP5','VIP','SNCG','PAX6')) %>%
  mutate(mouse_auroc = auroc) %>% select(cell_type, gene, mouse_auroc)

mouse_name_sake_genes$full_label = paste(mouse_name_sake_genes$cell_type, mouse_name_sake_genes$gene, sep = '_' )

#Align and add
index = match(mouse_name_sake_genes$full_label, primate_name_sake_genes$full_label)
primate_name_sake_genes = primate_name_sake_genes[index, ]

mouse_name_sake_genes$primate_auroc = primate_name_sake_genes$primate_auroc

#Set the order for plotting
mouse_name_sake_genes$gene = factor(mouse_name_sake_genes$gene, levels = c('SST','CHODL','PVALB','NKX2-1','LHX6','LAMP5','VIP','SNCG', 'PAX6'))
#Add a mouse subclass label
mouse_name_sake_genes$plot_label = mouse_name_sake_genes$full_label
mouse_name_sake_genes$plot_label[!mouse_name_sake_genes$plot_label %in% c('Chandelier_NKX2-1','Pvalb_PVALB','Sst_SST', 'Sst Chodl_CHODL',
                                                                          'Lamp5_LAMP5','Lamp5 Lhx6_LHX6','Sncg_SNCG','Vip_VIP','Sncg_PAX6')] = NA

mouse_name_sake_genes$plot_label = sapply(strsplit(mouse_name_sake_genes$plot_label, split = '_', fixed = T), '[[', 1)


primate_pax6_PAX6_auroc = primate_subclass_meta_markers %>% filter(cell_type == 'Pax6' & gene == 'PAX6') %>% pull(auroc)


library(ggrepel)

p1 = ggplot(mouse_name_sake_genes, aes(x = primate_auroc, y = mouse_auroc, color = cell_type, label = plot_label)) +
  geom_point(size = 3, alpha = .75) + ylim(0,1) + xlim(0,1) +
  geom_label_repel(min.segment.length = 0, max.overlaps = Inf, hjust = 1, direction = 'y', nudge_x = -.4,box.padding = .1, show.legend = F) +
  geom_vline(xintercept = .5, color = 'black', linetype = 'dashed', alpha = .5) +
  geom_hline(yintercept = .5, color = 'black', linetype = 'dashed', alpha = .5) +
  scale_color_manual(values = celltype_color_palette, name = 'Subclass', 
                     breaks = c('Sst','Sst Chodl','Pvalb','Chandelier','Lamp5 Lhx6','Lamp5','Vip','Sncg')) +
  ylab('Mouse AUROC') + xlab('Primate AUROC') +
  theme_bw()+ ggtitle('Subclass namesake gene DE') +
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank()) +
  facet_wrap(~gene, nrow = 3) +
  geom_vline(data = subset(mouse_name_sake_genes, gene == "PAX6"), aes(xintercept = primate_pax6_PAX6_auroc), 
             color = 'red', linetype = 'longdash')

p1
ggsave(plot = p1, filename = 'namesake_gene_de.pdf', path = '../graphs/mouse_primate_marker_comp/', device = 'pdf', useDingbats = F,
       height = 6, width = 8)


```

#save source data
```{r}
save(mouse_name_sake_genes, file = '../../source_data/data_for_plots/namesake_marker_de.Rdata')
```

######################################

And now try aggregating increasing primate markers in the mouse data

Given a marker set, compute the subclass specificity of those markers expression within the mouse subclasses
Do for the 9 primate markers sets, increasing number of markers


Could also just do the MetaMarker enrichment score as well, probably start with that

####################################




Using the original subclass labels for the mouse data, cell_subclass, get the marker enrichments of the primate metamarkers


```{r}

get_avg_marker_enrichment = function(sce_obj, meta_markers, num_markers){

  top_markers = meta_markers %>% filter(rank <= num_markers)
  
  #Compute enrichments
  ct_scores = score_cells(log1p(cpm(sce_obj)), top_markers)
  ct_enrichment = compute_marker_enrichment(ct_scores)
  #Add the mouse original subclass label to the cells
  dataset_cell_marker_enrichs_df = as.data.frame(t(ct_enrichment))
  dataset_cell_marker_enrichs_df$mouse_subclass = sce_obj$cell_subclass
  #Get the mean enrichment for all the marker sets across the mouse subclass labels
  summary_marker_enrich_df = dataset_cell_marker_enrichs_df %>% group_by(mouse_subclass) %>% 
    summarise(`Primate Sst` = mean(`all|Sst`), `Primate Sst Chodl` = mean(`all|Sst Chodl`),
              `Primate Pvalb` = mean(`all|Pvalb`), `Primate Chandelier` = mean(`all|Chandelier`),
              `Primate Lamp5` = mean(`all|Lamp5`), `Primate Lamp5 Lhx6` = mean(`all|Lamp5 Lhx6`),
              `Primate Vip` = mean(`all|Vip`), `Primate Sncg` = mean(`all|Sncg`),
              `Primate Pax6` = mean(`all|Pax6`)) %>% tibble::column_to_rownames(var = "mouse_subclass")
  summary_marker_enrich_df = as.matrix(summary_marker_enrich_df)
    #tidyr::pivot_longer(!mouse_subclass, names_to = 'primate_markers', values_to = 'mean_mouse_enrich') %>%
    #mutate(num_markers = current_markers)
  
  #all_summary_dfs[[i]] = summary_marker_enrich_df
  
  return(summary_marker_enrich_df)

}


num_total_markers = c(1, 10, 25, 50, 100)

all_enrich_dfs = list()

for(i in 1:length(num_total_markers)){
  

  current_markers = num_total_markers[i]
  
  all_enrich_mats = vector(mode = 'list', length = 8)
  
  all_enrich_mats[[1]] = get_avg_marker_enrichment(tasic18, primate_subclass_meta_markers, num_markers = current_markers)
  all_enrich_mats[[2]] = get_avg_marker_enrichment(zeng_10x_cells, primate_subclass_meta_markers, num_markers = current_markers)
  all_enrich_mats[[3]] = get_avg_marker_enrichment(zeng_10x_nuclei, primate_subclass_meta_markers, num_markers = current_markers)
  all_enrich_mats[[4]] = get_avg_marker_enrichment(zeng_smart_cells, primate_subclass_meta_markers, num_markers = current_markers)
  all_enrich_mats[[5]] = get_avg_marker_enrichment(zeng_smart_nuclei, primate_subclass_meta_markers, num_markers = current_markers)
  all_enrich_mats[[6]] = get_avg_marker_enrichment(zeng_10xv2_wba, primate_subclass_meta_markers, num_markers = current_markers)
  all_enrich_mats[[7]] = get_avg_marker_enrichment(zeng_10xv3_wba, primate_subclass_meta_markers, num_markers = current_markers)
  all_enrich_mats[[8]] = get_avg_marker_enrichment(macosko_wba, primate_subclass_meta_markers, num_markers = current_markers)
  
  dataset_avg_enrich_mat = Reduce('+', all_enrich_mats ) / 8
  
  
  avg_enrich_df = reshape2::melt(dataset_avg_enrich_mat, varnames = c('Mouse Subclass','Primate Subclass markers'), value.name = 'avg_enrichment')
  avg_enrich_df$num_markers = current_markers
  
  all_enrich_dfs[[i]] = avg_enrich_df


}

temp_enrich_dfs = do.call('rbind', all_enrich_dfs)


ggplot(temp_enrich_dfs, aes(x = num_markers, y = avg_enrichment, color = `Mouse Subclass`, group = `Mouse Subclass`)) + 
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = num_total_markers) +
  scale_color_manual(values = celltype_color_palette, name = 'Mouse Subclass') +
  theme_bw() + ylab('Average expression enrichment') + xlab('Number of Primate subclass markers') +
  facet_wrap(~`Primate Subclass markers`, scales = 'free_y') + ggtitle('Primate marker enrichment in Mouse subclasses')


```


Trying some type of signal to noise ratio

given a gene set, number of cells with detectable expression / total cells.
then that ratio divided by the same ratio computed from all background cells


```{r}

compute_avg_dectection = function(sce_obj, meta_markers, num_markers){

  #Grab the top markers
  top_markers = meta_markers %>% filter(rank <= num_markers)
  #Set up the marker gene annotation matrix
  gene_annotations = as.matrix(table(top_markers$cell_type, top_markers$gene))
  #And the mouse subclass annotation matrix
  #Make sure the annotations line up with the actual expression data order, doing table() sorts the data
  celltype_annotations = as.matrix(table(colnames(sce_obj), sce_obj$cell_subclass))
  index = match(colnames(sce_obj), rownames(celltype_annotations))
  celltype_annotations = celltype_annotations[index, ]
  
  #Just need the expression of the primate markers
  gene_index = rownames(sce_obj) %in% top_markers$gene
  marker_exp = cpm(sce_obj)[gene_index, ]
  #Detected or not
  marker_exp = as.matrix(marker_exp > 0) *1
  
  #This computes the detection rate of each gene in each subclass  
  cell_totals = colSums(celltype_annotations)
  detect_ratios = t(t(marker_exp %*% celltype_annotations) / cell_totals)
  
  #And this computes the average detection rate for each gene set across all the mouse subclasses
  index = match(colnames(gene_annotations), rownames(detect_ratios))
  detect_ratios = detect_ratios[index, ]
  
  avg_detection_rates = (gene_annotations %*% detect_ratios) / num_markers
  return(avg_detection_rates)
}



```




```{r}

num_total_markers = c(1, 10, 25, 50, 100)

all_avg_detect_dfs = list()

for(i in 1:length(num_total_markers)){
  
current_markers = num_total_markers[i]

all_avg_detect_mats = vector(mode = 'list', length = 8)

all_avg_detect_mats[[1]] = compute_avg_dectection(tasic18, primate_subclass_meta_markers, num_markers = current_markers)
all_avg_detect_mats[[2]] = compute_avg_dectection(zeng_10x_cells, primate_subclass_meta_markers, num_markers = current_markers)
all_avg_detect_mats[[3]] = compute_avg_dectection(zeng_10x_nuclei, primate_subclass_meta_markers, num_markers = current_markers)
all_avg_detect_mats[[4]] = compute_avg_dectection(zeng_smart_cells, primate_subclass_meta_markers, num_markers = current_markers)
all_avg_detect_mats[[5]] = compute_avg_dectection(zeng_smart_nuclei, primate_subclass_meta_markers, num_markers = current_markers)
all_avg_detect_mats[[6]] = compute_avg_dectection(zeng_10xv2_wba, primate_subclass_meta_markers, num_markers = current_markers)
all_avg_detect_mats[[7]] = compute_avg_dectection(zeng_10xv3_wba, primate_subclass_meta_markers, num_markers = current_markers)
all_avg_detect_mats[[8]] = compute_avg_dectection(macosko_wba, primate_subclass_meta_markers, num_markers = current_markers)

dataset_avg_detection_mat = Reduce('+', all_avg_detect_mats ) / 8


avg_detection_df = reshape2::melt(dataset_avg_detection_mat, varnames = c('Primate Subclass Markers','Mouse Subclass'), value.name = 'avg_detection_rate')
avg_detection_df$num_markers = current_markers

all_avg_detect_dfs[[i]] = avg_detection_df


}

temp_detect_dfs = do.call('rbind', all_avg_detect_dfs )


temp_detect_dfs$full_label = paste(temp_detect_dfs$`Primate Subclass Markers`, temp_detect_dfs$`Mouse Subclass`, sep = '_')
table(temp_detect_dfs$full_label)
keep_index = temp_detect_dfs$full_label %in% c('Chandelier_Chandelier','Chandelier_Pvalb',
                                               'Pvalb_Pvalb','Pvalb_Chandelier',
                                               'Sst_Sst','Sst_Sst Chodl',
                                               'Sst Chodl_Sst Chodl','Sst Chodl_Sst',
                                               'Vip_Vip','Vip_Sncg',
                                               'Sncg_Sncg','Sncg_Vip',
                                               'Lamp5_Lamp5','Lamp5_Lamp5 Lhx6',
                                               'Lamp5 Lhx6_Lamp5 Lhx6','Lamp5 Lhx6_Lamp5',
                                               'Pax6_Sncg','Pax6_Vip')
temp_detect_dfs$color_label = 'NA'
temp_detect_dfs$color_label[keep_index] = as.character(temp_detect_dfs$`Mouse Subclass`)[keep_index]
temp_detect_dfs
```


```{r}

temp_detect_dfs$`Primate Subclass Markers` = factor(temp_detect_dfs$`Primate Subclass Markers`, levels = c('Sst','Sst Chodl','Pvalb','Chandelier',
                                                                                                           'Lamp5 Lhx6','Lamp5','Vip','Pax6','Sncg'))

temp_detect_dfs$color_label = factor(temp_detect_dfs$color_label, levels = c('Sst','Sst Chodl','Pvalb','Chandelier',
                                                                                                           'Lamp5 Lhx6','Lamp5','Vip','Pax6','Sncg', 'NA'))

p2 = ggplot(temp_detect_dfs, aes(x = num_markers, y = avg_detection_rate, color = color_label, group = `Mouse Subclass`)) + 
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = num_total_markers) +
  scale_color_manual(values = celltype_color_palette, name = 'Mouse Subclass') +
  theme_bw() + ylab('Average gene detection rate') + xlab('Number of Primate subclass markers') +
  facet_wrap(~`Primate Subclass Markers`, scales = 'free_y') + ggtitle('Detection of Primate marker expression in Mouse subclasses')

p2
ggsave(plot = p2, filename = 'avg_detection_rate_primate_in_mouse.pdf', path = '../graphs/mouse_primate_marker_comp/', device = 'pdf', useDingbats = F,
       height = 6, width = 8)


```


And try with aggregated expression.
ratio of avg(agg_in_celltype) / avg(agg_in_all_other_cells)


```{r}


compute_signal_to_noise = function(sce_obj, meta_markers, num_markers){

  top_markers = meta_markers %>% filter(rank <= num_markers)
  
  gene_annotations = as.matrix(table(top_markers$cell_type, top_markers$gene))
  
  #Just need the expression of the primate markers
  gene_index = rownames(sce_obj) %in% top_markers$gene
  marker_exp = as.matrix(log1p(cpm(sce_obj))[gene_index, ])
  
  index = match(rownames(marker_exp), colnames(gene_annotations))
  gene_annotations = gene_annotations[ , index]
  
  #Computes the aggregate expression for each marker set in each cell
  agg_marker_exp_mat = t(marker_exp) %*% t(gene_annotations)
  
  
  #Make sure the annotations line up with the actual expression data order, doing table() sorts the data
  celltype_annotations = as.matrix(table(colnames(sce_obj), sce_obj$cell_subclass))
  index = match(rownames(agg_marker_exp_mat), rownames(celltype_annotations))
  celltype_annotations = celltype_annotations[index, ]
  
  cell_totals = colSums(celltype_annotations)
  cell_totals_background = colSums(!celltype_annotations)
  
  
  #This computes the average aggregate expression per cell-type
  sum_aggs = t(agg_marker_exp_mat ) %*% celltype_annotations
  avg_agg_exp = t(t(sum_aggs) / cell_totals)
  
  #This computes the average aggregate expression in the background cells
  sum_aggs_background = t(agg_marker_exp_mat ) %*% !celltype_annotations
  avg_agg_exp_background = t(t(sum_aggs_background) / cell_totals_background)
  
  #Divide to get a signal-to-noise ratio
  signal_to_noise_mat = avg_agg_exp / avg_agg_exp_background

  
  return(list('agg_exp' = avg_agg_exp , 'signal_noise' = signal_to_noise_mat))
}


```


```{r}

num_total_markers = c(1, 10, 25, 50, 100)

all_signal_noise_dfs = list()
all_exp_dfs = list()

for(i in 1:length(num_total_markers)){

  current_markers = num_total_markers[i]

  all_signal_noise_mats = vector(mode = 'list', length = 8)
  all_exp_mats = vector(mode = 'list', length = 8)
  
  temp_results = compute_signal_to_noise(tasic18, primate_subclass_meta_markers, num_markers = current_markers)
  all_exp_mats[[1]] = temp_results[['agg_exp']]
  all_signal_noise_mats[[1]] = temp_results[['signal_noise']]
  
  temp_results = compute_signal_to_noise(zeng_10x_cells, primate_subclass_meta_markers, num_markers = current_markers)
  all_exp_mats[[2]] = temp_results[['agg_exp']]
  all_signal_noise_mats[[2]] = temp_results[['signal_noise']]  

  temp_results = compute_signal_to_noise(zeng_10x_nuclei, primate_subclass_meta_markers, num_markers = current_markers)
  all_exp_mats[[3]] = temp_results[['agg_exp']]
  all_signal_noise_mats[[3]] = temp_results[['signal_noise']] 
  
  temp_results = compute_signal_to_noise(zeng_smart_cells, primate_subclass_meta_markers, num_markers = current_markers)
  all_exp_mats[[4]] = temp_results[['agg_exp']]
  all_signal_noise_mats[[4]] = temp_results[['signal_noise']]  
  
  temp_results = compute_signal_to_noise(zeng_smart_nuclei, primate_subclass_meta_markers, num_markers = current_markers)
  all_exp_mats[[5]] = temp_results[['agg_exp']]
  all_signal_noise_mats[[5]] = temp_results[['signal_noise']]  
  
  temp_results = compute_signal_to_noise(zeng_10xv2_wba, primate_subclass_meta_markers, num_markers = current_markers)
  all_exp_mats[[6]] = temp_results[['agg_exp']]
  all_signal_noise_mats[[6]] = temp_results[['signal_noise']]  
  
  temp_results = compute_signal_to_noise(zeng_10xv3_wba, primate_subclass_meta_markers, num_markers = current_markers)
  all_exp_mats[[7]] = temp_results[['agg_exp']]
  all_signal_noise_mats[[7]] = temp_results[['signal_noise']]  
  
  temp_results = compute_signal_to_noise(macosko_wba, primate_subclass_meta_markers, num_markers = current_markers)
  all_exp_mats[[8]] = temp_results[['agg_exp']]
  all_signal_noise_mats[[8]] = temp_results[['signal_noise']]
  
  dataset_avg_signal_noise_mat = Reduce('+', all_signal_noise_mats ) / 8
  dataset_avg_exp_mat = Reduce('+', all_exp_mats ) / 8
  
  avg_signal_noise_df = reshape2::melt(dataset_avg_signal_noise_mat, varnames = c('Primate Subclass Markers','Mouse Subclass'), value.name = 'avg_signal_noise')
  avg_signal_noise_df$num_markers = current_markers
  
  all_signal_noise_dfs[[i]] = avg_signal_noise_df
  
  avg_exp_df = reshape2::melt(dataset_avg_exp_mat, varnames = c('Primate Subclass Markers','Mouse Subclass'), value.name = 'avg_agg_exp')
  avg_exp_df$num_markers = current_markers
  
  all_exp_dfs[[i]] = avg_exp_df


}

temp_signal_noise_dfs = do.call('rbind', all_signal_noise_dfs )
temp_exp_dfs = do.call('rbind', all_exp_dfs )

temp_signal_noise_dfs$full_label = paste(temp_signal_noise_dfs$`Primate Subclass Markers`, temp_signal_noise_dfs$`Mouse Subclass`, sep = '_')
table(temp_signal_noise_dfs$full_label)
keep_index = temp_signal_noise_dfs$full_label %in% c('Chandelier_Chandelier','Chandelier_Pvalb',
                                               'Pvalb_Pvalb','Pvalb_Chandelier',
                                               'Sst_Sst','Sst_Sst Chodl',
                                               'Sst Chodl_Sst Chodl','Sst Chodl_Sst',
                                               'Vip_Vip','Vip_Sncg',
                                               'Sncg_Sncg','Sncg_Vip',
                                               'Lamp5_Lamp5','Lamp5_Lamp5 Lhx6',
                                               'Lamp5 Lhx6_Lamp5 Lhx6','Lamp5 Lhx6_Lamp5',
                                               'Pax6_Sncg', 'Pax6_Vip')
temp_signal_noise_dfs$color_label = 'NA'
temp_signal_noise_dfs$color_label[keep_index] = as.character(temp_signal_noise_dfs$`Mouse Subclass`)[keep_index]
temp_signal_noise_dfs
```




```{r}

temp_signal_noise_dfs$`Primate Subclass Markers` = paste0('Primate ', temp_signal_noise_dfs$`Primate Subclass Markers` )

temp_signal_noise_dfs$`Primate Subclass Markers` = factor(temp_signal_noise_dfs$`Primate Subclass Markers`, 
                                                          levels = c('Primate Sst','Primate Sst Chodl','Primate Pvalb','Primate Chandelier',
                                                                     'Primate Lamp5 Lhx6','Primate Lamp5','Primate Vip','Primate Sncg','Primate Pax6'))

temp_exp_dfs$`Primate Subclass Markers` = paste0('Primate ', temp_exp_dfs$`Primate Subclass Markers` )

temp_exp_dfs$`Primate Subclass Markers` = factor(temp_exp_dfs$`Primate Subclass Markers`, 
                                                          levels = c('Primate Sst','Primate Sst Chodl','Primate Pvalb','Primate Chandelier',
                                                                     'Primate Lamp5 Lhx6','Primate Lamp5','Primate Vip','Primate Sncg','Primate Pax6'))


temp_signal_noise_dfs$color_label = factor(temp_signal_noise_dfs$color_label, levels = c('Sst','Sst Chodl','Pvalb','Chandelier',
                                                                                                           'Lamp5 Lhx6','Lamp5','Vip','Sncg','Pax6', 'NA'))

#temp_exp_dfs$`Primate Subclass Markers` = factor(temp_exp_dfs$`Primate Subclass Markers`, levels = c('Sst','Sst Chodl','Pvalb','Chandelier',
#                                                                                                           'Lamp5 Lhx6','Lamp5','Vip','Pax6','Sncg'))

#temp_exp_dfs$color_label = factor(temp_exp_dfs$color_label, levels = c('Sst','Sst Chodl','Pvalb','Chandelier',
#                                                                                                           'Lamp5 Lhx6','Lamp5','Vip','Pax6','Sncg', 'NA'))





p3 = ggplot(temp_signal_noise_dfs, aes(x = num_markers, y = avg_signal_noise, color =`Mouse Subclass`, group = `Mouse Subclass`)) + 
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = num_total_markers) +
  scale_color_manual(values = celltype_color_palette, name = 'Mouse Subclass') +
  theme_bw() + ylab('Average signal-to-noise ratio') + xlab('Number of Primate subclass markers') +
  facet_wrap(~`Primate Subclass Markers`, scales = 'free_y') + ggtitle('Signal-to-Noise of Primate marker expression in Mouse subclasses')

p3
ggsave(plot = p3, filename = 'avg_signal_noise_primate_in_mouse.pdf', path = '../graphs/mouse_primate_marker_comp/', device = 'pdf', useDingbats = F,
       height = 6, width = 8)


temp_exp_dfs = temp_exp_dfs %>% group_by(`Primate Subclass Markers`, num_markers) %>% mutate(zscore_exp = scale(avg_agg_exp))



p4 = ggplot(temp_exp_dfs, aes(x = num_markers, y = zscore_exp, color =`Mouse Subclass`, group = `Mouse Subclass`)) + 
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = num_total_markers) +
  scale_color_manual(values = celltype_color_palette, name = 'Mouse Subclass') +
  theme_bw() + ylab('Average aggregate expression') + xlab('Number of Primate subclass markers') +
  facet_wrap(~`Primate Subclass Markers`, scales = 'free_y') + ggtitle('Aggregate expression of Primate markers in Mouse subclasses')

p4
ggsave(plot = p4, filename = 'avg_exp_primate_in_mouse.pdf', path = '../graphs/mouse_primate_marker_comp/', device = 'pdf', useDingbats = F,
       height = 6, width = 8)


```



Re order the enrichment df to match the other stat dfs
And add a scaling of the enrichment values

```{r}

temp_enrich_dfs = temp_enrich_dfs %>% group_by(num_markers, `Mouse Subclass`) %>% arrange( as.character(`Primate Subclass markers`), .by_group = T)

table(temp_signal_noise_dfs$`Mouse Subclass` == temp_enrich_dfs$`Mouse Subclass`)
table(temp_signal_noise_dfs$`Primate Subclass Markers` == temp_enrich_dfs$`Primate Subclass markers`)

table(temp_exp_dfs$`Mouse Subclass` == temp_enrich_dfs$`Mouse Subclass`)
table(temp_exp_dfs$`Primate Subclass Markers` == temp_enrich_dfs$`Primate Subclass markers`)



temp_enrich_dfs = temp_enrich_dfs %>% group_by(`Primate Subclass markers`, num_markers) %>% mutate(zscore_enrich = scale(avg_enrichment))


```




```{r}

all_stats_df = temp_signal_noise_dfs
all_stats_df$avg_agg_exp = temp_exp_dfs$avg_agg_exp
all_stats_df$zscore_exp = temp_exp_dfs$zscore_exp
all_stats_df$avg_enrich_exp = temp_enrich_dfs$avg_enrichment
all_stats_df$zscore_enrich = temp_enrich_dfs$zscore_enrich

marker_colors = c('white','grey70','grey50','grey30','black')
names(marker_colors) = as.character(num_total_markers)

all_stats_df$num_markers = as.character(all_stats_df$num_markers)
all_stats_df$num_markers = factor(all_stats_df$num_markers, levels = c(as.character(num_total_markers)))


p5 = ggplot(all_stats_df, aes( y = avg_signal_noise, x = avg_agg_exp , color =color_label, group = `Mouse Subclass`)) + 
  geom_point(shape = 21,size = 2, aes(fill = num_markers)) +
  geom_line() +
  scale_color_manual(values = celltype_color_palette, name = 'Mouse Subclass') +
  scale_fill_manual(values = marker_colors, name = '# markers') +
  theme_bw() + ylab('Average signal-to-noise ratio') + xlab('Average aggregate expression') +
  facet_wrap(~`Primate Subclass Markers`, scales = 'free_y') + ggtitle('Primate marker expression in Mouse subclasses')

p5

ggsave(plot = p5, filename = 'exp_vs_signal_noise_primate_in_mouse.pdf', path = '../graphs/mouse_primate_marker_comp/', device = 'pdf', useDingbats = F,
       height = 6, width = 8)


p6 = ggplot(all_stats_df, aes( x = avg_agg_exp, y =avg_enrich_exp , color =color_label, group = `Mouse Subclass`)) + 
  geom_point(shape = 21,size = 2, aes(fill = num_markers)) +
  geom_line() + 
  #coord_cartesian(ylim=c(0, 3.5)) +
  scale_color_manual(values = celltype_color_palette, name = 'Mouse Subclass') +
  scale_fill_manual(values = marker_colors, name = '# markers') +
  theme_bw() + ylab('Expression enrichment') + xlab('Aggregate expression') +
  facet_wrap(~`Primate Subclass Markers`, scales = 'free') + ggtitle('Primate marker expression in Mouse subclasses')

p6

ggsave(plot = p6, filename = 'enrichment_vs_agg_exp_primate_in_mouse.pdf', path = '../graphs/mouse_primate_marker_comp/', device = 'pdf', useDingbats = F,
       height = 6, width = 8)

```

```{r}


all_stats_df %>% group_by(`Primate Subclass Markers`, `Mouse Subclass`) %>% summarise(mean_enrichment = mean(avg_enrich_exp)) %>%
  group_by(`Primate Subclass Markers`) %>% arrange(desc(mean_enrichment), .by_group = T)

```

And plot the Primate Sncg and Primate Pax6 enrichment against each other for the Mouse Sncg subclass

```{r}

snapshot_df = all_stats_df %>% filter(`Primate Subclass Markers` %in% c('Primate Sncg','Primate Pax6') & `Mouse Subclass` == 'Sncg') %>%
  select(`Primate Subclass Markers`, `Mouse Subclass`, avg_enrich_exp, num_markers) %>% 
  tidyr::pivot_wider(names_from = `Primate Subclass Markers`, values_from = avg_enrich_exp)


p7 = ggplot(snapshot_df , aes(x = `Primate Sncg`, y = `Primate Pax6`, color = `Mouse Subclass`)) +
  geom_point(shape = 21,size = 5, aes(fill = num_markers)) +
  geom_abline(slope = 1, intercept = 0, color = 'red') +
  ylim(0, 2.8) + xlim(0, 2) +
  scale_color_manual(values = celltype_color_palette, name = 'Mouse Subclass') +
  scale_fill_manual(values = marker_colors, name = '# markers') +
  ggtitle('Mouse Sncg Subclass') + ylab('Primate Pax6 marker enrichment') + xlab('Primate Sncg marker enrichment') +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())

p7

ggsave(plot = p7, filename = 'mouse_snch_primate_sncg_pax6_enrichment.pdf', path = '../graphs/mouse_primate_marker_comp/', device = 'pdf', useDingbats = F,
       height = 2.5, width = 4)


```


```{r}

saveRDS(all_stats_df, file = '/home/werner/projects/mouse_brain_consensus_taxonomy/data/mouse_primate_marker_comp/all_stats_df.rds')

```


```{r}

all_stats_df = readRDS('/home/werner/projects/mouse_brain_consensus_taxonomy/data/mouse_primate_marker_comp/all_stats_df.rds')


all_stats_df %>% filter(`Mouse Subclass` == 'Pvalb' & num_markers == 50) %>% arrange(desc(avg_enrich_exp))

```

```{r}
num_total_markers = c(1, 10, 25, 50, 100)
marker_colors = c('white','grey70','grey50','grey30','black')
names(marker_colors) = as.character(num_total_markers)



ggplot(all_stats_df, aes( y = avg_signal_noise, x= avg_enrich_exp, color =color_label, group = `Mouse Subclass`)) + 
  geom_point(shape = 21,size = 2, aes(fill = num_markers)) +
  geom_line() + 
  #coord_cartesian(ylim=c(0, 3.5)) +
  scale_color_manual(values = celltype_color_palette, name = 'Mouse Subclass') +
  scale_fill_manual(values = marker_colors, name = '# markers') +
  theme_bw() + ylab('Expression enrichment') + xlab('Aggregate expression') +
  facet_wrap(~`Primate Subclass Markers`, scales = 'free') + ggtitle('Primate marker expression in Mouse subclasses')


```


```{r}

p6 = ggplot(all_stats_df, aes( x = avg_agg_exp, y =avg_enrich_exp , color =color_label, group = `Mouse Subclass`)) + 
  geom_point(shape = 21,size = 3, aes(fill = num_markers)) +
  geom_line(linewidth = 1) + geom_hline(yintercept = 1, color = 'black', linetype = 'dashed') +
  scale_color_manual(values = celltype_color_palette, name = 'Mouse Subclass') +
  scale_fill_manual(values = marker_colors, name = '# markers') +
  theme_bw() + ylab('Mouse expression enrichment') + xlab('Aggregate mouse expression') +
  facet_wrap(~`Primate Subclass Markers`, scales = 'free') + ggtitle('Primate marker expression in Mouse subclasses')

p6

ggsave(plot = p6, filename = 'enrichment_vs_agg_exp_primate_in_mouse.pdf', path = '../graphs/mouse_primate_marker_comp/', device = 'pdf', useDingbats = F,
       height = 6, width = 8)


```


And now look at the primate marker gene enrichment in the orthologous mapped mouse subclasses.
Specifically interested in these

Mouse Lamp5 - Primate Lamp5
Mouse Lamp5 - Primate Pax6
Mouse Vip - Primate Sncg
Mouse Sncg - Primate Pax6
Mouse Vip - Primate Vip


```{r}


get_avg_marker_enrichment_2 = function(sce_obj, study_label, meta_markers, num_markers){

  #Get an annotation for the previous mouse and now orthologous primate subclass labels
  mouse_subclass = paste0('Mouse_', sce_obj$cell_subclass)
  primate_subclass = paste0('Primate_', sce_obj$primate_subclass)
  sce_obj$combined_subclass_label = paste(mouse_subclass, primate_subclass, sep = '|')

  orth_filt = sce_obj$orthology == 'high_orthology'
  sce_obj = sce_obj[ ,orth_filt ]
  
  top_markers = meta_markers %>% filter(rank <= num_markers)
  
  #Compute enrichments
  ct_scores = score_cells(log1p(cpm(sce_obj)), top_markers)
  ct_enrichment = compute_marker_enrichment(ct_scores)
  
  #Get the average enrichment of the marker sets by that combined subclass label
  colnames(ct_enrichment) = sce_obj$combined_subclass_label
  marker_enrich_df = reshape2::melt(t(ct_enrichment), varnames = c('combined_subclass', 'primate_marker_set'), value.name = 'enrichment')

  marker_enrich_df = marker_enrich_df %>% group_by(combined_subclass, primate_marker_set) %>% summarise(mean_enrichment = mean(enrichment))
  marker_enrich_df$study = study_label
  
  return(marker_enrich_df)

}

num_total_markers = c(1, 10, 25, 50, 100)
total_enrich_dfs_list = vector(mode = 'list', length = length(num_total_markers))

for(i in 1:length(num_total_markers)){

  current_num_markers = num_total_markers[i]
  
  temp_df_1 = get_avg_marker_enrichment_2(tasic18, 'tasic18', primate_subclass_meta_markers, current_num_markers)
  temp_df_2 = get_avg_marker_enrichment_2(zeng_10x_cells, 'zeng_10x_cells', primate_subclass_meta_markers, current_num_markers)
  temp_df_3 = get_avg_marker_enrichment_2(zeng_10x_nuclei, 'zeng_10x_nuclei', primate_subclass_meta_markers, current_num_markers)
  temp_df_4 = get_avg_marker_enrichment_2(zeng_smart_cells, 'zeng_smart_cells', primate_subclass_meta_markers, current_num_markers)
  temp_df_5 = get_avg_marker_enrichment_2(zeng_smart_nuclei, 'zeng_smart_nuclei', primate_subclass_meta_markers, current_num_markers)
  temp_df_6 = get_avg_marker_enrichment_2(zeng_10xv2_wba, 'zeng_10xv2_wba', primate_subclass_meta_markers, current_num_markers)
  temp_df_7 = get_avg_marker_enrichment_2(zeng_10xv3_wba, 'zeng_10xv3_wba', primate_subclass_meta_markers, current_num_markers)
  temp_df_8 = get_avg_marker_enrichment_2(macosko_wba, 'macosko_wba', primate_subclass_meta_markers, current_num_markers)
  
  
  all_enrich_df_list = list(temp_df_1, temp_df_2, temp_df_3, temp_df_4, temp_df_5, temp_df_6, temp_df_7, temp_df_8)
  
  all_enrich_dfs = do.call('rbind', all_enrich_df_list)
  
  all_enrich_dfs = all_enrich_dfs %>% group_by(combined_subclass, primate_marker_set) %>% 
    summarise(overall_mean_enrichment = mean(mean_enrichment), sd_mean_enrichment = sd(mean_enrichment)) %>%
    mutate(num_markers = current_num_markers)
  
  total_enrich_dfs_list[[i]] = all_enrich_dfs

}

mouse_primate_label_enrich_df = do.call('rbind', total_enrich_dfs_list)


mouse_primate_label_enrich_df


```





```{r}

mouse_primate_label_enrich_df$primate_subclass_label = sapply(strsplit(as.character(mouse_primate_label_enrich_df$primate_marker_set), 
                                                                       split = '|', fixed = T), '[[', 2)
mouse_primate_label_enrich_df$num_markers = as.character(mouse_primate_label_enrich_df$num_markers)
mouse_primate_label_enrich_df$num_markers = factor(mouse_primate_label_enrich_df$num_markers , levels = c('1','10','25','50','100'))
```


```{r}

plotting_df = mouse_primate_label_enrich_df %>% filter(combined_subclass %in% 
                                                         c('Mouse_Sncg|Primate_Pax6', 'Mouse_Vip|Primate_Pax6', 'Mouse_Lamp5|Primate_Pax6'))

plotting_df$label_color = plotting_df$primate_subclass_label

index = plotting_df$combined_subclass == 'Mouse_Lamp5|Primate_Pax6' & !plotting_df$primate_subclass_label %in% c('Lamp5','Pax6')
plotting_df$label_color[index] = 'NA'
plotting_df$overall_mean_enrichment[index] = NA
plotting_df$sd_mean_enrichment[index] = NA
index = plotting_df$combined_subclass == 'Mouse_Vip|Primate_Pax6' & !plotting_df$primate_subclass_label %in% c('Vip','Pax6')
plotting_df$label_color[index] = 'NA'
plotting_df$overall_mean_enrichment[index] = NA
plotting_df$sd_mean_enrichment[index] = NA
index = plotting_df$combined_subclass == 'Mouse_Sncg|Primate_Pax6' & !plotting_df$primate_subclass_label %in% c('Sncg','Pax6')
plotting_df$label_color[index] = 'NA'
plotting_df$overall_mean_enrichment[index] = NA
plotting_df$sd_mean_enrichment[index] = NA


plotting_df$combined_subclass = factor(plotting_df$combined_subclass, levels = c('Mouse_Vip|Primate_Pax6','Mouse_Lamp5|Primate_Pax6','Mouse_Sncg|Primate_Pax6'))

p8 = ggplot(plotting_df, aes(x = num_markers, y = overall_mean_enrichment, color = label_color, group = primate_subclass_label)) +
  geom_point(size = 3) + geom_line(show.legend = F) + 
  geom_pointrange(aes(ymax = overall_mean_enrichment + sd_mean_enrichment, ymin = overall_mean_enrichment - sd_mean_enrichment ),
                position=position_dodge(0.05), show.legend = F) +
  ggtitle('Previous Mouse Sncg/Vip/Lamp5 -> Orthologous Pax6') +
  geom_hline(yintercept = 1, linetype = 'dashed', color = 'black') +
  xlab('Number of Primate Markers') + ylab('Mouse Expression enrichment') +
  scale_color_manual(values = celltype_color_palette, name = 'Primate markers') +
  facet_wrap(~combined_subclass)+
  theme_bw()
p8
ggsave(plot = p8, filename = 'mouse_vip_lamp5_sncg_primate_pax6_marker_exp.pdf', path = '../graphs/mouse_primate_marker_comp/', device = 'pdf', useDingbats = F,
       height = 4, width = 7)



plotting_df = mouse_primate_label_enrich_df %>% filter(combined_subclass %in% 
                                                         c('Mouse_Vip|Primate_Sncg', 'Mouse_Sncg|Primate_Sncg', 'Mouse_Lamp5|Primate_Sncg'))

plotting_df$label_color = plotting_df$primate_subclass_label

 index = plotting_df$combined_subclass == 'Mouse_Vip|Primate_Sncg' & !plotting_df$primate_subclass_label %in% c('Vip','Sncg')
 plotting_df$label_color[index] = 'NA'
 plotting_df$overall_mean_enrichment[index] = NA
 plotting_df$sd_mean_enrichment[index] = NA
 index = plotting_df$combined_subclass == 'Mouse_Lamp5|Primate_Sncg' & !plotting_df$primate_subclass_label %in% c('Lamp5','Sncg')
 plotting_df$label_color[index] = 'NA'
 plotting_df$overall_mean_enrichment[index] = NA
 plotting_df$sd_mean_enrichment[index] = NA
 index = plotting_df$combined_subclass == 'Mouse_Sncg|Primate_Sncg' & !plotting_df$primate_subclass_label %in% c('Pax6','Sncg')
 plotting_df$label_color[index] = 'NA'
 plotting_df$overall_mean_enrichment[index] = NA
 plotting_df$sd_mean_enrichment[index] = NA

plotting_df$combined_subclass = factor(plotting_df$combined_subclass, levels = c('Mouse_Vip|Primate_Sncg','Mouse_Lamp5|Primate_Sncg','Mouse_Sncg|Primate_Sncg'))
 
p9 = ggplot(plotting_df, aes(x = num_markers, y = overall_mean_enrichment, color = label_color, group = primate_subclass_label)) +
  geom_point(size = 3) + geom_line(show.legend = F) + 
  geom_pointrange(aes(ymax = overall_mean_enrichment + sd_mean_enrichment, ymin = overall_mean_enrichment - sd_mean_enrichment ),
                position=position_dodge(0.05), show.legend = F) +
  ggtitle('Previous Mouse Sncg/Vip/Lamp5 -> Orthologous Sncg') +
  geom_hline(yintercept = 1, linetype = 'dashed', color = 'black') +
  xlab('Number of Primate Markers') + ylab('Mouse Expression enrichment') +
  scale_color_manual(values = celltype_color_palette, name = 'Primate markers') +
  facet_wrap(~combined_subclass)+
  theme_bw()
p9
ggsave(plot = p9, filename = 'mouse_vip_lamp5_sncg_primate_sncg_marker_exp.pdf', path = '../graphs/mouse_primate_marker_comp/', device = 'pdf', useDingbats = F,
       height = 4, width = 7)



```

```{r}

plotting_df = mouse_primate_label_enrich_df %>% filter(combined_subclass == 'Mouse_Vip|Primate_Sncg')

ggplot(plotting_df, aes(x = num_markers, y = overall_mean_enrichment, color = primate_subclass_label, group = primate_subclass_label)) +
  geom_point(size = 3) + geom_line(show.legend = F) + ggtitle('Previous Mouse Vip -> Orthologous Sncg') +
  geom_hline(yintercept = 1, linetype = 'dashed', color = 'black') +
  xlab('Number of Primate Markers') + ylab('Mouse Expression enrichment') +
  scale_color_manual(values = celltype_color_palette, name = 'Primate markers') +
  theme_bw()


```




###################################

For those same comparisons, plot the full metaNeighbor AUROCs.
Highlighting that while metaneighbor is identifying the best matches, at least for the orthologous Sncg subclass, it is just barely the best match


####################################

mouse vs primate subclass full aurocs



```{r}

#contains mouse_clust_primate_sub_aurocs, mouse_clust_primate_sub_best_hits, mouseMC_primate_sub_best_hits, mouseMC_primate_clust_best_hits
#looking for the mouse_clust_primate_sub_aurocs
load('/home/werner/projects/mouse_brain_consensus_taxonomy/data/metaneighbor/cross_mammal_MN_results.Rdata')

#Has the combined mouse metadata
#interneuron_metadata
load(file = '/inkwell03/werner/mouse_brain_consensus_taxonomy/20_3_25_interneuron_mouse_metadata.Rdata')

interneuron_metadata %>% filter(cell_subclass == 'Sncg' & primate_subclass == 'Sncg') %>% group_by(study, cell_cluster) %>%
  summarise(n = n())

```

From the combined mouse metadata, grab the clusters of interest

first comparison is

Mouse Vip/Lamp5/Sncg -> Primate Pax6
```{r}

get_MN_mouse_primate_comp = function(target_mouse_subclass, ref_primate_subclass, mouse_metadata, MN_aurocs){

  comp_df = mouse_metadata %>% filter(cell_subclass == target_mouse_subclass & primate_subclass == ref_primate_subclass) %>% 
    group_by(study, cell_cluster) %>%
    summarise(n = n())
  comp_df$full_cluster = paste(comp_df$study, comp_df$cell_cluster, sep = '|')
  
  c_index = colnames(MN_aurocs) %in% comp_df$full_cluster
  r_index = grepl('human', rownames(MN_aurocs)) | grepl('chimp', rownames(MN_aurocs)) |
    grepl('gorilla', rownames(MN_aurocs)) | grepl('macaca', rownames(MN_aurocs)) | 
    grepl('marmoset', rownames(MN_aurocs)) 
  
  
  comp_auroc_df = reshape2::melt(MN_aurocs[r_index, c_index], varnames = c('ref_primate_cluster','target_mouse_cluster'), value.name = 'auroc')
  comp_auroc_df$primate_subclass = sapply(strsplit(as.character(comp_auroc_df$ref_primate_cluster), split = '|', fixed = T), '[[', 2)
  comp_auroc_df$primate_subclass[comp_auroc_df$primate_subclass == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
  comp_auroc_df = comp_auroc_df %>% group_by(target_mouse_cluster, primate_subclass) %>% summarise(primate_auroc = mean(auroc))
  order_vec = comp_auroc_df %>% group_by(primate_subclass) %>% summarise(median = median(primate_auroc)) %>% arrange(median) %>% pull(primate_subclass)
  
  comp_auroc_df$primate_subclass = factor(comp_auroc_df$primate_subclass, levels = order_vec)
  
  auroc_plot = ggplot(comp_auroc_df, aes(x = primate_subclass, y = primate_auroc, fill = primate_subclass)) + geom_boxplot() +
    geom_hline(yintercept = .5, linetype = 'dashed', color = 'black') +
    xlab('Primate Subclass') + ylab('Mean Primate MetaNeighbor AUROC') + 
    scale_x_discrete(guide = guide_axis(n.dodge=2)) +
    ggtitle(sprintf('Previous Mouse %s -> Orthologous %s',target_mouse_subclass ,ref_primate_subclass )) +
    scale_fill_manual(values = celltype_color_palette, name = 'Primate Subclass') + ylim(0,1) + theme_bw()
  
  return(auroc_plot)

}


p_vip_pax6 = get_MN_mouse_primate_comp('Vip','Pax6', interneuron_metadata, mouse_clust_primate_sub_aurocs)
p_vip_pax6

p_lamp5_pax6 = get_MN_mouse_primate_comp('Lamp5','Pax6', interneuron_metadata, mouse_clust_primate_sub_aurocs)
p_lamp5_pax6

p_sncg_pax6 = get_MN_mouse_primate_comp('Sncg','Pax6', interneuron_metadata, mouse_clust_primate_sub_aurocs)
p_sncg_pax6


p_vip_sncg = get_MN_mouse_primate_comp('Vip','Sncg', interneuron_metadata, mouse_clust_primate_sub_aurocs)
p_vip_sncg


```


```{r}

mouse_metadata = interneuron_metadata
target_mouse_subclass = 'Sncg'
ref_primate_subclass = 'Sncg'
MN_aurocs = mouse_clust_primate_sub_aurocs

comp_df = mouse_metadata %>% filter(cell_subclass == target_mouse_subclass & primate_subclass == ref_primate_subclass) %>% 
  group_by(study, cell_cluster) %>%
  summarise(n = n())
comp_df$full_cluster = paste(comp_df$study, comp_df$cell_cluster, sep = '|')

c_index = colnames(MN_aurocs) %in% comp_df$full_cluster
r_index = grepl('human', rownames(MN_aurocs)) | grepl('chimp', rownames(MN_aurocs)) |
  grepl('gorilla', rownames(MN_aurocs)) | grepl('macaca', rownames(MN_aurocs)) | 
  grepl('marmoset', rownames(MN_aurocs)) 

comp_auroc_df = data.frame(auroc = MN_aurocs[r_index, c_index], 
                           ref_primate_cluster = names(MN_aurocs[r_index, c_index]),
                           target_mouse_cluster = comp_df$full_cluster[1])

#comp_auroc_df = reshape2::melt(MN_aurocs[r_index, c_index], varnames = c('ref_primate_cluster','target_mouse_cluster'), value.name = 'auroc')
comp_auroc_df$primate_subclass = sapply(strsplit(as.character(comp_auroc_df$ref_primate_cluster), split = '|', fixed = T), '[[', 2)
comp_auroc_df$primate_subclass[comp_auroc_df$primate_subclass == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
comp_auroc_df = comp_auroc_df %>% group_by(target_mouse_cluster, primate_subclass) %>% summarise(primate_auroc = mean(auroc))
order_vec = comp_auroc_df %>% arrange(primate_auroc) %>% pull(primate_subclass)

comp_auroc_df$primate_subclass = factor(comp_auroc_df$primate_subclass, levels = order_vec)

p_sncg_sncg = ggplot(comp_auroc_df, aes(x = primate_subclass, y = primate_auroc, color = primate_subclass)) + geom_point(size = 5) +
  geom_hline(yintercept = .5, linetype = 'dashed', color = 'black') +
  xlab('Primate Subclass') + ylab('Mean Primate MetaNeighbor AUROC') + 
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +
  ggtitle(sprintf('Previous Mouse %s -> Orthologous %s',target_mouse_subclass ,ref_primate_subclass )) +
  scale_color_manual(values = celltype_color_palette, name = 'Primate Subclass') + ylim(0,1) + theme_bw()
p_sncg_sncg




target_mouse_subclass = 'Lamp5'
ref_primate_subclass = 'Sncg'

comp_df = mouse_metadata %>% filter(cell_subclass == target_mouse_subclass & primate_subclass == ref_primate_subclass) %>% 
  group_by(study, cell_cluster) %>%
  summarise(n = n())
comp_df$full_cluster = paste(comp_df$study, comp_df$cell_cluster, sep = '|')

c_index = colnames(MN_aurocs) %in% comp_df$full_cluster
r_index = grepl('human', rownames(MN_aurocs)) | grepl('chimp', rownames(MN_aurocs)) |
  grepl('gorilla', rownames(MN_aurocs)) | grepl('macaca', rownames(MN_aurocs)) | 
  grepl('marmoset', rownames(MN_aurocs)) 

comp_auroc_df = data.frame(auroc = MN_aurocs[r_index, c_index], 
                           ref_primate_cluster = names(MN_aurocs[r_index, c_index]),
                           target_mouse_cluster = comp_df$full_cluster[1])

#comp_auroc_df = reshape2::melt(MN_aurocs[r_index, c_index], varnames = c('ref_primate_cluster','target_mouse_cluster'), value.name = 'auroc')
comp_auroc_df$primate_subclass = sapply(strsplit(as.character(comp_auroc_df$ref_primate_cluster), split = '|', fixed = T), '[[', 2)
comp_auroc_df$primate_subclass[comp_auroc_df$primate_subclass == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
comp_auroc_df = comp_auroc_df %>% group_by(target_mouse_cluster, primate_subclass) %>% summarise(primate_auroc = mean(auroc))
order_vec = comp_auroc_df %>% arrange(primate_auroc) %>% pull(primate_subclass)

comp_auroc_df$primate_subclass = factor(comp_auroc_df$primate_subclass, levels = order_vec)

p_lamp5_sncg = ggplot(comp_auroc_df, aes(x = primate_subclass, y = primate_auroc, color = primate_subclass)) + geom_point(size = 5) +
  geom_hline(yintercept = .5, linetype = 'dashed', color = 'black') +
  xlab('Primate Subclass') + ylab('Mean Primate MetaNeighbor AUROC') + 
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +
  ggtitle(sprintf('Previous Mouse %s -> Orthologous %s',target_mouse_subclass ,ref_primate_subclass )) +
  scale_color_manual(values = celltype_color_palette, name = 'Primate Subclass') + ylim(0,1) + theme_bw()
p_lamp5_sncg


```

```{r}

ggsave(plot = p_vip_pax6, filename = 'full_MNaurocs_mouse_vip_primate_pax6.pdf', 
       path = '../graphs/mouse_primate_marker_comp/', device = 'pdf', useDingbats = F,
       height = 3, width = 5)

ggsave(plot = p_lamp5_pax6, filename = 'full_MNaurocs_mouse_lamp5_primate_pax6.pdf', 
       path = '../graphs/mouse_primate_marker_comp/', device = 'pdf', useDingbats = F,
       height = 3, width = 5)

ggsave(plot = p_sncg_pax6, filename = 'full_MNaurocs_mouse_sncg_primate_pax6.pdf', 
       path = '../graphs/mouse_primate_marker_comp/', device = 'pdf', useDingbats = F,
       height = 3, width = 5)

ggsave(plot = p_vip_sncg, filename = 'full_MNaurocs_mouse_vip_primate_sncg.pdf', 
       path = '../graphs/mouse_primate_marker_comp/', device = 'pdf', useDingbats = F,
       height = 3, width = 5)

ggsave(plot = p_lamp5_sncg, filename = 'full_MNaurocs_mouse_lamp5_primate_sncg.pdf', 
       path = '../graphs/mouse_primate_marker_comp/', device = 'pdf', useDingbats = F,
       height = 3, width = 5)

ggsave(plot = p_sncg_sncg, filename = 'full_MNaurocs_mouse_sncg_primate_sncg.pdf', 
       path = '../graphs/mouse_primate_marker_comp/', device = 'pdf', useDingbats = F,
       height = 3, width = 5)


```



And now get the difference between the top and next best match in the full aurocs, between mouse and primates, 
show distributions across the mouse orthologous subclasses


```{r}

orth_color_palette = c("Chandelier"='#F641A8', 
                           "Pvalb"='#D93137', 
                           "Sst"='#FF9900', 
                           "Sst Chodl"='gold', 
                           "Lamp5"='#DA808C',
                           "Sncg"='#DF70FF',
                           "Vip"='#A45FBF', 
                           "Lamp5 Lhx6"='#935F50',
                           "Pax6"='#71238C',
                           'low orthology' = 'grey90')


MN_aurocs = mouse_clust_primate_sub_aurocs

r_index = grepl('human', rownames(MN_aurocs)) | grepl('chimp', rownames(MN_aurocs)) |
  grepl('gorilla', rownames(MN_aurocs)) | grepl('macaca', rownames(MN_aurocs)) | 
  grepl('marmoset', rownames(MN_aurocs)) 

c_index = !r_index

comp_auroc_df = reshape2::melt(MN_aurocs[r_index, c_index], varnames = c('ref_primate_cluster','target_mouse_cluster'), value.name = 'auroc')
comp_auroc_df$primate_subclass = sapply(strsplit(as.character(comp_auroc_df$ref_primate_cluster), split = '|', fixed = T), '[[', 2)
comp_auroc_df$primate_subclass[comp_auroc_df$primate_subclass == 'Lamp5_Lhx6'] = 'Lamp5 Lhx6'
comp_auroc_df = comp_auroc_df %>% group_by(target_mouse_cluster, primate_subclass) %>% summarise(primate_auroc = mean(auroc))
  
comp_auroc_df$row_id = 1:nrow(comp_auroc_df)


#Get the max auroc per cluster
max_aurocs_df = comp_auroc_df %>% group_by(target_mouse_cluster) %>% mutate(max_auroc = max(primate_auroc)) %>% filter(primate_auroc == max_auroc)

#Filter out the max and get the next best
second_best_aurocs_df = comp_auroc_df %>% filter(!row_id %in% max_aurocs_df$row_id) %>% group_by(target_mouse_cluster) %>%
  mutate(second_best_auroc = max(primate_auroc)) %>% filter(primate_auroc == second_best_auroc)


table(max_aurocs_df$target_mouse_cluster == second_best_aurocs_df$target_mouse_cluster)
max_aurocs_df$second_best_auroc = second_best_aurocs_df$second_best_auroc

#This will have all the mouse clusters, including the ones with low orthology
max_aurocs_df = max_aurocs_df %>% mutate(best_next_best_diff = max_auroc - second_best_auroc)


#The interneuron metadata only has data for the high orthologt clusters
interneuron_metadata$full_cluster = paste(interneuron_metadata$study, interneuron_metadata$cell_cluster, sep = '|')
primate_subclass_mouse_cluster_mapping = interneuron_metadata %>% group_by(full_cluster, primate_subclass) %>% summarise(n = n())


index = match(max_aurocs_df$target_mouse_cluster, primate_subclass_mouse_cluster_mapping$full_cluster )
primate_subclass_mouse_cluster_mapping[index, ]

max_aurocs_df$orthologous_subclass = primate_subclass_mouse_cluster_mapping$primate_subclass[index]
max_aurocs_df$orthologous_subclass[is.na(max_aurocs_df$orthologous_subclass)] = 'low orthology'

order_vec = max_aurocs_df %>% group_by(orthologous_subclass) %>% summarise(median = median(best_next_best_diff)) %>% arrange(desc(median)) %>%
  pull(orthologous_subclass)

max_aurocs_df$orthologous_subclass = factor(max_aurocs_df$orthologous_subclass, levels = order_vec)

wide_discrim_plot = ggplot(max_aurocs_df, aes(x = orthologous_subclass, y = best_next_best_diff, fill = orthologous_subclass)) + geom_boxplot(show.legend = F) +
  ylab('Mouse cluster discriminability (Best - Next best AUROC)') + xlab('Mouse orthologous subclasses') +
  scale_fill_manual(values = orth_color_palette) +
  theme_bw() + theme(aspect.ratio = 1)

long_discrim_plot = ggplot(max_aurocs_df, aes(y = orthologous_subclass, x = best_next_best_diff, fill = orthologous_subclass)) + geom_boxplot(show.legend = F) +
  xlab('Mouse cluster discriminability (Best - Next best AUROC)') + ylab('Mouse orthologous subclasses') +
  scale_fill_manual(values = orth_color_palette) + scale_y_discrete(limits=rev) +
  theme_bw() + theme(aspect.ratio = 1)

wide_discrim_plot 
long_discrim_plot 

ggsave(plot = wide_discrim_plot , filename = 'wide_mouse_cluster_discriminability_boxplots.pdf', 
       path = '../graphs/mouse_primate_marker_comp/', device = 'pdf', useDingbats = F,
       height = 3, width = 5)
ggsave(plot = long_discrim_plot , filename = 'long_mouse_cluster_discriminability_boxplots.pdf', 
       path = '../graphs/mouse_primate_marker_comp/', device = 'pdf', useDingbats = F,
       height = 5, width = 3)


```






