

Exploring some human and mouse published patch-seq data

At least for some mouse data, the cells are already mapped to the Tasic18 clusters, can assign them to the homologous primate subclasses.


```{r}
library(ggplot2)
library(dplyr)
library(SingleCellExperiment)
library(ggExtra)
library(ComplexHeatmap)

```

```{r}
celltype_color_palette = c("Chandelier"='#F641A8', 
                           "Pvalb"='#D93137', 
                           "Sst"='#FF9900', 
                           "Sst Chodl"='gold', 
                           "Lamp5"='#DA808C',
                           "Sncg"='#DF70FF',
                           "Vip"='#A45FBF', 
                           "Lamp5 Lhx6"='#935F50',
                           "Pax6"='#71238C')


```


```{r}

tasic18 = readRDS('/inkwell03/werner/mouse_brain_consensus_taxonomy/datasets/ortho_filt/mouse/tasic18_gaba_primate_MN.rds')
tasic18 = tasic18[ , tasic18$orthology == 'high_orthology']

tasic_metadata = as.data.frame(colData(tasic18))

table(tasic_metadata$orthology)



tasic_clust_to_subclass_df = tasic_metadata %>% group_by(cell_cluster, cell_subclass, primate_subclass, orthology) %>% summarise(n = n()) %>% arrange(orthology)


clust_to_primate_subclass = c(tasic_clust_to_subclass_df$primate_subclass)
names(clust_to_primate_subclass) = tasic_clust_to_subclass_df$cell_cluster
clust_to_primate_subclass

clust_to_mouse_subclass = c(tasic_clust_to_subclass_df$cell_subclass)
names(clust_to_mouse_subclass) = tasic_clust_to_subclass_df$cell_cluster
clust_to_mouse_subclass


```





Tasic18 pax6 clusters
c('Lamp5 Fam19a1 Pax6','Lamp5 Krt73','Serpinf1 Clrn1','Sncg Slc17a8')

Tasic18 sncg cluster
c('Vip Igfbp6 Pltp')


```{r}

mouse_file_path = '/home/werner/projects/mouse_brain_consensus_taxonomy/data/patch_seq/mouse/'

#Has cell IDs and their cluster annotation
mouse_patch_meta = data.table::fread(paste0(mouse_file_path, '20200625_patchseq_metadata_mouse.csv'))

table(mouse_patch_meta$corresponding_AIT2.3.1_alias)


mouse_patch_meta %>% filter(corresponding_AIT2.3.1_alias %in% c('Lamp5 Fam19a1 Pax6','Lamp5 Krt73','Serpinf1 Clrn1','Sncg Slc17a8') )

#Add the mouse and primate subclass labels to the clusters
mouse_patch_meta$primate_subclass = clust_to_primate_subclass[mouse_patch_meta$corresponding_AIT2.3.1_alias]
mouse_patch_meta$mouse_subclass = clust_to_mouse_subclass[mouse_patch_meta$corresponding_AIT2.3.1_alias]


```



Add the cluster annotations from the metadata and then add the mouse subclass and homologous subclass annotations from the Tasic data



```{r}

mouse_patch_data = data.table::fread(paste0(mouse_file_path, 'ephys_features_gouwens_w_QC.csv'))

index = match(mouse_patch_data$cell_specimen_id,mouse_patch_meta$cell_specimen_id )
table(mouse_patch_data$cell_specimen_id == mouse_patch_meta$cell_specimen_id[index] )

mouse_patch_data$cell_cluster = mouse_patch_meta$corresponding_AIT2.3.1_alias[index]
mouse_patch_data$mouse_subclass = mouse_patch_meta$mouse_subclass[index]
mouse_patch_data$primate_subclass = mouse_patch_meta$primate_subclass[index]





```


```{r}

ggplot(filter(mouse_patch_data, mouse_subclass %in% c('Vip','Sncg','Lamp5')), 
       aes(x = qc_initial_access_resistance_mohm , y = input_resistance, color = mouse_subclass )) + 
  geom_point(alpha = .5) +
  scale_color_manual(values = celltype_color_palette) 


ggplot(filter(mouse_patch_data, mouse_subclass %in% c('Vip','Sncg','Lamp5')), 
       aes(x = rheo_fast_trough_t , y = input_resistance, color = primate_subclass )) + 
  geom_point(alpha = .5) +
  scale_color_manual(values = celltype_color_palette) 

table(mouse_patch_data$mouse_subclass)
table(mouse_patch_data$primate_subclass)
```


Trying PCA just to see 
Filtering out the variables that have mostly NAs and the QC filters
filt_vars = c('file_id','file_name','cell_cluster','mouse_subclass','primate_subclass','qc_blowout_mv','rheo_slow_trough_t','rheo_adp_v','rheo_slow_trough_v',
'hero_adapt','hero_first_isi','hero_isi_cv','hearo_mean_isi')

```{r}

#Denote the Lamp5 Lhx6 as a subclass in the original mouse annotations
mouse_patch_data$mouse_subclass[mouse_patch_data$cell_cluster == 'Lamp5 Lhx6'] = 'Lamp5 Lhx6'
#Denote the Serpin mouse subclass as Vip
mouse_patch_data$mouse_subclass[mouse_patch_data$mouse_subclass == 'Serpinf1'] = 'Vip'



filt_vars = c('file_id','file_name','cell_cluster','mouse_subclass','primate_subclass','qc_blowout_mv','qc_electrode_0_pa','qc_seal_gohm',
              'qc_input_resistance_mohm','qc_initial_access_resistance_mohm','qc_input_access_resistance_ratio',
              'rheo_slow_trough_t','rheo_adp_v','rheo_slow_trough_v',
              'hero_adapt','hero_first_isi','hero_isi_cv','hero_mean_isi', 'qc_recording_date')

filt_mouse_patch_data = mouse_patch_data %>% filter(mouse_subclass %in% c('Vip','Lamp5','Sncg', 'Lamp5 Lhx6')) %>% select(-filt_vars)

#Filter out cells with NA values
row_filt = rowSums(is.na(filt_mouse_patch_data)) == 0 
filt_mouse_patch_data = filt_mouse_patch_data[row_filt, ]


filt_mouse_patch_data = as.matrix(filt_mouse_patch_data %>% tibble::column_to_rownames(var = 'cell_specimen_id'))


scaled_filt_mouse_patch_data = scale(filt_mouse_patch_data)



data.pca <- princomp(scaled_filt_mouse_patch_data)
summary(data.pca)

pca_loadings_df = as.data.frame(data.pca$scores)


index = match(rownames(pca_loadings_df), mouse_patch_data$cell_specimen_id )
table(rownames(pca_loadings_df) == mouse_patch_data$cell_specimen_id[index] )

pca_loadings_df$cell_cluster = mouse_patch_data$cell_cluster[index]
pca_loadings_df$mouse_subclass = mouse_patch_data$mouse_subclass[index]
pca_loadings_df$primate_subclass = mouse_patch_data$primate_subclass[index]


pca_loadings_df

```


Generally very subtle differences across the ephys differences for the different subtypes, largely looks like a Lamp5 vs Vip spectrum, Sncg and Pax6 are inbetween, but very continuous.


```{r}

ggplot(pca_loadings_df, aes(x = Comp.1, y = Comp.2, color = mouse_subclass)) + geom_point(alpha = .75) +
  scale_color_manual(values = celltype_color_palette)


ggplot(pca_loadings_df, aes(x = Comp.1, y = Comp.2, color = primate_subclass)) + geom_point(alpha = .75) +
  scale_color_manual(values = celltype_color_palette)

```

And just try umap

```{r}

#mouse_patch_umap = umap::umap(scaled_filt_mouse_patch_data)

mouse_patch_umap = umap::umap(as.matrix(pca_loadings_df[ , 1:10]))


umap_df = as.data.frame(mouse_patch_umap$layout)

index = match(rownames(umap_df), mouse_patch_data$cell_specimen_id )
table(rownames(umap_df) == mouse_patch_data$cell_specimen_id[index] )

umap_df$cell_cluster = mouse_patch_data$cell_cluster[index]
umap_df$mouse_subclass = mouse_patch_data$mouse_subclass[index]
umap_df$primate_subclass = mouse_patch_data$primate_subclass[index]

ggplot(umap_df, aes(x = V1, y = V2, color = mouse_subclass)) + geom_point(size = 2) +
  scale_color_manual(values = celltype_color_palette) + theme_bw() + ylab("UMAP 2") + xlab('UMAP 1') + ggtitle('Mouse CGE ephys UMAP')

ggplot(umap_df, aes(x = V1, y = V2, color = primate_subclass)) + geom_point(size = 2) +
  scale_color_manual(values = celltype_color_palette) + theme_bw() + ylab("UMAP 2") + xlab('UMAP 1') + ggtitle('Mouse CGE ephys UMAP')

ggsave(filename = 'mouse_patchseq_ephys_CGE_umap.pdf',path = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/mouse_patchSeq/', device = 'pdf',
       useDingbats = F, height = 4, width = 5, family = 'Helvetica')



```






PUlling together a standard variance decomposition approach to compare across the different annotations.
Starting simply with a one-way ANOVA, univariate decomposition

The goal is to compare the variance explained by the original annotations to the variance explained by the homologous annotations.
Use ANOVA decomposition to get the coefficient of determination for labelsA (mouse subclasses) and then labelsB (homologous subclasses). 
Can get a p-value for each coefficient/feature and can also compare the distribution of coefficients across all the features for the two labels.

Since the homologous labels have 1 more annotation, need to use either adjustedR^2 or omega^2 as the coefficient. Empirical p-values will also be useful due to the varying sample size of each annotation group.

Z-score the ephys features to start, since all in different units.

Don't necessarily have to z-score the genes, since all CPM (may try logCPM), but if I want to compare the omegas across the modalities, will need to zscore anyhow.


Code is from chatGPT, double checking implementation and output

```{r}

library(parallel)

# ---------- omega^2 estimator ----------
omega2 <- function(y, grp) {
  grp <- factor(grp)
  n <- length(y)
  k <- nlevels(grp)
  if (k < 2) return(0)
  grand_mean <- mean(y)
  group_means <- tapply(y, grp, mean)
  n_k <- as.numeric(table(grp))
  ss_b <- sum(n_k * (group_means - grand_mean)^2)
  ss_t <- sum((y - grand_mean)^2)
  ms_w <- (ss_t - ss_b) / (n - k)
  num <- ss_b - (k - 1) * ms_w
  denom <- ss_t + ms_w
  if (denom <= 0 || num <= 0) return(0)
  num / denom
}

# ---------- permutation test per feature ----------
omega_perm_test <- function(y, grp, n_perm = 1000) {
  obs <- omega2(y, grp)
  # permute while preserving group sizes
  n_k <- table(grp)
  perm_fun <- function() {
    shuffled <- sample(y)
    shuffled_grp <- rep(names(n_k), n_k)
    omega2(shuffled, shuffled_grp)
  }
  perms <- replicate(n_perm, perm_fun())
  pval <- (1 + sum(perms >= obs)) / (1 + n_perm)
  list(obs = obs, pval = pval)
}

# ---------- wrapper across many features ----------
run_omega_perm <- function(mat, grp, n_perm = 1000, n_cores = 4) {
  stopifnot(is.matrix(mat))
  cl <- makeCluster(n_cores)
  clusterExport(cl, c("omega2", "omega_perm_test", "grp", "n_perm"), envir = environment())
  res <- parApply(cl, mat, 1, function(y) omega_perm_test(y, grp, n_perm))
  stopCluster(cl)
  
  # Convert list output to data.frame
  obs <- sapply(res, function(x) x$obs)
  pval <- sapply(res, function(x) x$pval)
  data.frame(feature = rownames(mat), omega2 = obs, pval = pval)
}

# ---------- example usage ----------
# expr: matrix of features (genes/ephys vars) x samples (cells)
# labels_A and labels_B: factor vectors of length ncol(expr)

#res_A <- run_omega_perm(expr, labels_A, n_perm = 1000, n_cores = 8)
#res_B <- run_omega_perm(expr, labels_B, n_perm = 1000, n_cores = 8)

# Add FDR correction
#res_A$padj <- p.adjust(res_A$pval, method = "BH")
#res_B$padj <- p.adjust(res_B$pval, method = "BH")

```





```{r}
#Rows are cells, columns are ephys features
filt_mouse_patch_data 

scaled_filt_mouse_patch_data = scale(filt_mouse_patch_data)

#Get the two different annotations
index = match(rownames(filt_mouse_patch_data ), mouse_patch_data$cell_specimen_id )
mouse_subclass = mouse_patch_data$mouse_subclass[index]
primate_subclass = mouse_patch_data$primate_subclass[index]

table(mouse_subclass, primate_subclass)
table(primate_subclass)


```

```{r}

# ---------- example usage ----------
# expr: matrix of features (genes/ephys vars) x samples (cells)
# labels_A and labels_B: factor vectors of length ncol(expr)

ephys_feature_mat = t(scaled_filt_mouse_patch_data )

res_mouse <- run_omega_perm(ephys_feature_mat, mouse_subclass, n_perm = 1000, n_cores = 8)
res_primate <- run_omega_perm(ephys_feature_mat, primate_subclass, n_perm = 1000, n_cores = 8)

# Add FDR correction
res_mouse$padj <- p.adjust(res_mouse$pval, method = "BH")
res_primate$padj <- p.adjust(res_primate$pval, method = "BH")

par(pty= 's')
plot(res_mouse$omega2, res_primate$omega2, main = 'Electrophysiology features', 
     ylab = 'Variance explained by homologous subclasses', 
     xlab = 'Variance explained by original subclasses', xlim = c(0,1), ylim = c(0,1) )
abline(a = 0, b = 1, col = 'red', lty = 'dashed')


pdf(file = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/mouse_patchSeq/omega2_ephys_features.pdf', useDingbats = F, height = 4, width = 4)
par(pty= 's')
plot(res_mouse$omega2, res_primate$omega2, main = 'Electrophysiology features', 
     ylab = 'Variance explained by homologous subclasses', 
     xlab = 'Variance explained by original subclasses', xlim = c(0,1), ylim = c(0,1) )
abline(a = 0, b = 1, col = 'red', lty = 'dashed')
dev.off()



combined_res = res_mouse
combined_res$primate_omega2 = res_primate$omega2
combined_res$primate_pval = res_primate$pval
combined_res$primate_padj = res_primate$padj

combined_res = combined_res %>% mutate(omega_diff = abs(omega2 - primate_omega2)) %>% arrange(desc(omega_diff))
combined_res %>% arrange(desc(primate_omega2))

```

#save source data
```{r}
save(res_mouse, res_primate , file = '../../source_data/data_for_plots/omega_ephys_scatter.Rdata')
```




```{r}

combined_res %>% arrange(desc(primate_omega2))

plotting_omega_df = combined_res %>% select(feature, omega2, primate_omega2) %>% tidyr::pivot_longer(!feature, names_to = 'annotation', values_to = 'omega-squared')

ggplot(plotting_omega_df, aes(x = annotation, y = `omega-squared`)) + geom_boxplot() +
  theme_bw() + ylim(0,1) + ggtitle('Electrophysiological feature variance')


wilcox.test(combined_res$omega2, combined_res$primate_omega2, paired = T, alternative = 'less')


```



good features for separability
tau
rheo_width
rheobase_i
qc_input_resistance_mohm



```{r}

library(ggExtra)

```


```{r}

p_mouse = ggplot(filter(mouse_patch_data, primate_subclass %in% c('Vip','Sncg','Lamp5', 'Lamp5 Lhx6','Pax6')), 
       aes(x = rheo_peak_t, y = sag, color = mouse_subclass )) + 
  geom_point(alpha = .75, size = 2) +
  scale_color_manual(values = celltype_color_palette) + theme_bw()

ggMarginal(p_mouse, type="density",  groupColour = TRUE, groupFill = TRUE)

p_primate = ggplot(filter(mouse_patch_data, primate_subclass %in% c('Vip','Sncg','Lamp5', 'Lamp5 Lhx6','Pax6')), 
       aes(x = rheo_peak_t , y = sag, color = primate_subclass )) + 
  geom_point(alpha = .75, size = 2) +
  scale_color_manual(values = celltype_color_palette) + theme_bw()

ggMarginal(p_primate, type="density",  groupColour = TRUE, groupFill = TRUE)

```

The features with the biggest differn

```{r}

p_mouse = ggplot(filter(mouse_patch_data, mouse_subclass %in% c('Lamp5', 'Sncg')), 
       aes(x = sag, y = rheo_width, color = mouse_subclass )) + 
  geom_point(alpha = .75, size = 2) +
  scale_color_manual(values = celltype_color_palette) + theme_bw()

ggMarginal(p_mouse, type="density",  groupColour = TRUE, groupFill = TRUE)

p_primate = ggplot(filter(mouse_patch_data, mouse_subclass %in% c('Lamp5', 'Sncg')), 
       aes(x = sag , y = rheo_width, color = primate_subclass )) + 
  geom_point(alpha = .75, size = 2) +
  scale_color_manual(values = celltype_color_palette) + theme_bw()

ggMarginal(p_primate, type="density",  groupColour = TRUE, groupFill = TRUE)



p_mouse = ggplot(filter(mouse_patch_data, mouse_subclass %in% c('Vip')), 
       aes(x = sag, y = rheo_width, color = mouse_subclass )) + 
  geom_point(alpha = .75, size = 2) +
  scale_color_manual(values = celltype_color_palette) + theme_bw()

ggMarginal(p_mouse, type="density",  groupColour = TRUE, groupFill = TRUE)

p_primate = ggplot(filter(mouse_patch_data, mouse_subclass %in% c('Vip')), 
       aes(x = sag , y = rheo_width, color = primate_subclass )) + 
  geom_point( size = 2) +
  scale_color_manual(values = celltype_color_palette) + theme_bw()

ggMarginal(p_primate, type="density",  groupColour = TRUE, groupFill = TRUE)



p_mouse = ggplot(filter(mouse_patch_data, primate_subclass %in% c('Pax6')), 
       aes(x = sag, y = rheo_width, color = mouse_subclass )) + 
  geom_point(alpha = .75, size = 2) +
  scale_color_manual(values = celltype_color_palette) + theme_bw()

ggMarginal(p_mouse, type="density",  groupColour = TRUE, groupFill = TRUE)

p_primate = ggplot(filter(mouse_patch_data, primate_subclass %in% c('Pax6')), 
       aes(x = sag , y = rheo_width, color = primate_subclass )) + 
  geom_point(alpha = .75, size = 2) +
  scale_color_manual(values = celltype_color_palette) + theme_bw()

ggMarginal(p_primate, type="density",  groupColour = TRUE, groupFill = TRUE)




```



```{r}


ephys_correlation_mat = cor(ephys_feature_mat, method = 'spearman')
ephys_dist_mat = as.dist(1- ephys_correlation_mat)

index = match(colnames(ephys_correlation_mat), mouse_patch_data$cell_specimen_id)

corr_mouse_subclass = mouse_patch_data$mouse_subclass[index]
corr_primate_subclass = mouse_patch_data$primate_subclass[index]

top_annot = HeatmapAnnotation(mouse_subclass = corr_mouse_subclass, homologous_subclass = corr_primate_subclass,
                              col = list(mouse_subclass = celltype_color_palette, homologous_subclass = celltype_color_palette))



corr_cols <- rev(grDevices::colorRampPalette(RColorBrewer::brewer.pal(11,"RdYlBu"))(100))
Heatmap(ephys_correlation_mat , name = 'spearman', col = corr_cols,
        clustering_distance_rows = ephys_dist_mat, use_raster = T,
        clustering_distance_columns = ephys_dist_mat,
        clustering_method_rows = 'ward.D2', clustering_method_columns = 'ward.D2',
        show_row_names = F, show_column_names = F,
        top_annotation = top_annot, column_title = 'Cell similarity over electrophysiology features')

pdf(file = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/mouse_patchSeq/cell_siml_of_ephys_features.pdf', 
    useDingbats = F, height = 4, width = 6, family = 'sans')
Heatmap(ephys_correlation_mat , name = 'spearman', use_raster = T,col = corr_cols, 
        clustering_distance_rows = ephys_dist_mat,
        clustering_distance_columns = ephys_dist_mat,
        clustering_method_rows = 'ward.D2', clustering_method_columns = 'ward.D2',
        show_row_names = F, show_column_names = F,
        top_annotation = top_annot, column_title = 'Cell similarity over electrophysiology features')

dev.off()

pdf(file = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/mouse_patchSeq/cell_siml_of_ephys_features_nonRaster.pdf', 
    useDingbats = F, height = 4, width = 6, family = 'sans')
Heatmap(ephys_correlation_mat , name = 'spearman', use_raster =F,col = corr_cols, 
        clustering_distance_rows = ephys_dist_mat,
        clustering_distance_columns = ephys_dist_mat,
        clustering_method_rows = 'ward.D2', clustering_method_columns = 'ward.D2',
        show_row_names = F, show_column_names = F,
        top_annotation = top_annot, column_title = 'Cell similarity over electrophysiology features')

dev.off()



```



#save source data
```{r}
save(mouse_patch_data, ephys_correlation_mat , file = '../../source_data/data_for_plots/mouse_ephys_corr_heatmap.Rdata')
```





Globally it's a bit difficult to see any signal other than the large Lamp5 and Vip difference.
For a targeted analysis, focus on just the Lamp5s and the subset we are claiming are homologous to Pax6. (And Vip and Pax6)

Using the Lamp5 clusters that stay Lamp5 across species, compute empirical p-values for each cell for each ephys feature
Are the cells we're claiming are homologous Pax6 cells consistent outliers in ephys features to the Lamp5 cells? Which features?
Can combine p-values for an overall statement, or keep it at the feature level

Also do this for all newly labeled Pax6 cells versus all other CGE cell-types

```{r}

#Filtering out the QC variables and the variables with vast majority NAs

mouse_lamp5_lamp5_patch_df = mouse_patch_data %>% filter(mouse_subclass == 'Lamp5' & primate_subclass == 'Lamp5') %>%
  select(-c(rheo_adp_v, rheo_slow_trough_t, rheo_slow_trough_v, 
            qc_blowout_mv, qc_recording_date, qc_electrode_0_pa, qc_seal_gohm, qc_input_resistance_mohm, qc_initial_access_resistance_mohm, 
            qc_input_access_resistance_ratio,   
            file_id, file_name, cell_cluster, mouse_subclass, primate_subclass)) %>%
  tibble::column_to_rownames(var = 'cell_specimen_id')

mouse_lamp5_pax6_patch_df = mouse_patch_data %>% filter(mouse_subclass == 'Lamp5' & primate_subclass == 'Pax6') %>%
  select(-c(rheo_adp_v, rheo_slow_trough_t, rheo_slow_trough_v, 
            qc_blowout_mv, qc_recording_date, qc_electrode_0_pa, qc_seal_gohm, qc_input_resistance_mohm, qc_initial_access_resistance_mohm, 
            qc_input_access_resistance_ratio,   
            file_id, file_name, cell_cluster, mouse_subclass, primate_subclass)) %>%
  tibble::column_to_rownames(var = 'cell_specimen_id')


#Getting the pvalues of the ephys features for all the Lamp5 cells

# Rank-based empirical two-sided p-values
pvals <- apply(mouse_lamp5_lamp5_patch_df, 2, function(col) {
  # rank of each value in its column
  #Dropping NA values
  ranks <- rank(col, ties.method = "average", na.last = NA)
  N = length(ranks)
  # convert ranks to one-sided p-values
  p_lower <- ranks / N 
  p_upper <-  1 - (ranks - 1) / N
  # two-sided empirical p-value
  p_two <- 2 * pmin(p_lower, p_upper)
  #Cap at 1, for the cases of median values in odd numbered sample sizes
  pmin(p_two, 1)
})



for(i in 1:length(pvals)){
  
  hist(pvals[[i]],breaks = 20, main = i)
  
}


#And now a function to get the empirical pvalues of test samples
empirical_pvals_rank <- function(test_mat, ref_mat) {
  stopifnot(ncol(test_mat) == ncol(ref_mat))
  
  N <- nrow(ref_mat)

  pvals <- sapply(seq_len(ncol(ref_mat)), function(j) {
    ref_col <- ref_mat[, j]
    test_col <- test_mat[, j]
    
    # Rank reference distribution
    sorted_ref <- sort(ref_col)

    # For each test value, find how many ref values are <= that value
    ranks_lower <- findInterval(test_col, sorted_ref)
    p_lower <- ranks_lower / N

    # upper tail: how many are >=
    ranks_upper <- N - ranks_lower + 1
    p_upper <- ranks_upper / N

    # two-sided p-values
    p_two <- 2 * pmin(p_lower, p_upper)
    pmin(p_two, 1)  # cap at 1
  })

  df= as.data.frame(pvals)
  colnames(df) = colnames(ref_mat)
  return(df)
}

test_lamp5_pax6_pval_df = empirical_pvals_rank(mouse_lamp5_pax6_patch_df, mouse_lamp5_lamp5_patch_df)


#And combine for a single p-value per feature

# df_pvals: rows = samples, columns = features
fisher_combine <- function(pvals) {
  # Ensure no zero
  pvals <- pmax(pvals, 1e-300)
  k <- length(pvals)
  stat <- -2 * sum(log(pvals))
  p_combined <- pchisq(stat, df = 2 * k, lower.tail = FALSE)
  return(p_combined)
}

# Apply to each column (feature)
combined_pvals_lamp5_pax6 <- apply(test_lamp5_pax6_pval_df, 2, fisher_combine)
combined_pvals_lamp5_pax6 <- p.adjust(combined_pvals_lamp5_pax6, method = 'BH')
combined_pvals_lamp5_pax6

plotting_feature_pval = combined_pvals_lamp5_pax6[!is.na(combined_pvals_lamp5_pax6)]
plotting_feature_pval = plotting_feature_pval[order(plotting_feature_pval, decreasing = T)]

par(mar = c(4, 10, 2, 1))
barplot(-1*log10(plotting_feature_pval+1e-10), 
        names.arg = names(plotting_feature_pval),
        las = 2, horiz = T, cex.names = .75, 
        main = 'Mouse Lamp5 -> Homologous Pax6', xlab = '-log10(meta p-value)' )
abline(v = -1*log10(0.05), col = 'red')

pdf(file = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/mouse_patchSeq/lamp5_to_pax6_sig_features_barplot.pdf', 
    useDingbats = F, family = 'sans', height = 6, width = 7)
par(mar = c(4, 10, 2, 1))
barplot(-1*log10(plotting_feature_pval+1e-10), 
        names.arg = names(plotting_feature_pval),
        las = 2, horiz = T, cex.names = .75, 
        main = 'Mouse Lamp5 -> Homologous Pax6', xlab = '-log10(meta p-value)' )
abline(v = -1*log10(0.05), col = 'red')
dev.off()



#Plot the distributions of values for the lamp5 vs pax6 for all of the features we pull out as significant

mouse_lamp5_lamp5_patch_df$label = 'Lamp5'
mouse_lamp5_pax6_patch_df$label = 'Pax6'

all_lamp5_pax6_df = rbind(mouse_lamp5_lamp5_patch_df,mouse_lamp5_pax6_patch_df )

combined_pvals_lamp5_pax6 = combined_pvals_lamp5_pax6[order(combined_pvals_lamp5_pax6)]
sig_features = names(combined_pvals_lamp5_pax6[combined_pvals_lamp5_pax6 <= 0.05 & !is.na(combined_pvals_lamp5_pax6)])

#Filter for the significant features and plot
sig_subset = all_lamp5_pax6_df[ , colnames(all_lamp5_pax6_df) %in% c(sig_features, 'label')]

sig_subset = sig_subset %>% tidyr::pivot_longer(!label, names_to = 'feature', values_to = 'value')

sig_subset$feature = factor(sig_subset$feature, levels = sig_features)

sig_subset

ggplot(sig_subset, aes(x = feature, y = value, color = label)) + geom_boxplot() + facet_wrap(~feature, scales = 'free') +
  scale_color_manual(values = celltype_color_palette) + theme_bw()
ggsave(filename = 'all_sig_lamp5_pax6_ephy_features_boxplots.pdf', path = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/mouse_patchSeq/', 
       device = 'pdf', family = 'Helvetica',
       useDingbats = F, height = 4, width = 6)

hist(pvals$tau, density = T, col = 'blue')
hist(test_lamp5_pax6_pval_df$tau, add = T, density = T, col = 'red' )


pval_df = data.frame(pvals = c(pvals$tau, test_lamp5_pax6_pval_df$tau ), 
                     labels = c(rep('Lamp5', length = length(pvals$tau)), rep('Pax6', length = length(test_lamp5_pax6_pval_df$tau)) ) )

ggplot(pval_df, aes(x = pvals, fill = labels)) +
  geom_histogram(alpha = 0.5, position = "identity", bins = 10) +
  labs(title = "Tau", x = "P-values", y = "Frequency") +
  theme_minimal()


```


```{r}

p_mouse = ggplot(filter(mouse_patch_data, mouse_subclass %in% c('Lamp5')), 
       aes(x = rheo_width, y = tau, color = mouse_subclass )) + 
  geom_point(alpha = .75, size = 2, show.legend = F) +
  scale_color_manual(values = celltype_color_palette) + theme_bw()

ggMarginal(p_mouse, type="density",  groupColour = TRUE, groupFill = TRUE)

ggsave(filename = 'taue_and_rheo_width_lamp5_scatter.pdf', path = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/mouse_patchSeq/', 
       device = 'pdf', family = 'Helvetica',
       useDingbats = F, height = 4, width = 6)


p_primate = ggplot(filter(mouse_patch_data, mouse_subclass %in% c('Lamp5')), 
       aes(x = rheo_width, y =tau , color = primate_subclass )) + 
  geom_point(alpha = .75, size = 2, show.legend = F) +
  scale_color_manual(values = celltype_color_palette) + theme_bw()

ggMarginal(p_primate, type="density",  groupColour = TRUE, groupFill = TRUE)

ggsave(filename = 'taue_and_rheo_width_lamp5_pax6_scatter.pdf', path = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/mouse_patchSeq/', 
       device = 'pdf', family = 'Helvetica',
       useDingbats = F, height = 4, width = 6)

```


```{r}


mouse_vip_vip_patch_df = mouse_patch_data %>% filter(mouse_subclass == 'Vip' & primate_subclass == 'Vip') %>%
  select(-c(rheo_adp_v, rheo_slow_trough_t, rheo_slow_trough_v, 
            qc_blowout_mv, qc_recording_date, qc_electrode_0_pa, qc_seal_gohm, qc_input_resistance_mohm, qc_initial_access_resistance_mohm, 
            qc_input_access_resistance_ratio,   
            file_id, file_name, cell_cluster, mouse_subclass, primate_subclass)) %>%
  tibble::column_to_rownames(var = 'cell_specimen_id')

mouse_vip_pax6_patch_df = mouse_patch_data %>% filter(mouse_subclass == 'Vip' & primate_subclass == 'Pax6') %>%
  select(-c(rheo_adp_v, rheo_slow_trough_t, rheo_slow_trough_v, 
            qc_blowout_mv, qc_recording_date, qc_electrode_0_pa, qc_seal_gohm, qc_input_resistance_mohm, qc_initial_access_resistance_mohm, 
            qc_input_access_resistance_ratio,   
            file_id, file_name, cell_cluster, mouse_subclass, primate_subclass)) %>%
  tibble::column_to_rownames(var = 'cell_specimen_id')




test_vip_pax6_pval_df = empirical_pvals_rank(mouse_vip_pax6_patch_df, mouse_vip_vip_patch_df)


#And combine for a single p-value per feature

# Apply to each column (feature)
combined_pvals_vip_pax6 <- apply(test_vip_pax6_pval_df, 2, fisher_combine)
combined_pvals_vip_pax6 <- p.adjust(combined_pvals_vip_pax6, method = 'BH')
combined_pvals_vip_pax6

plotting_feature_pval = combined_pvals_vip_pax6[!is.na(combined_pvals_vip_pax6)]
plotting_feature_pval = plotting_feature_pval[order(plotting_feature_pval, decreasing = T)]

par(mar = c(4, 10, 2, 1))
barplot(-1*log10(plotting_feature_pval+1e-10), 
        names.arg = names(plotting_feature_pval),
        las = 2, horiz = T, cex.names = .75, main = 'Mouse Vip -> Homologous Pax6')
abline(v = -1*log10(0.05), col = 'red')


pdf(file = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/mouse_patchSeq/vip_to_pax6_sig_features_barplot.pdf', 
    useDingbats = F, family = 'sans', height = 6, width = 7)
par(mar = c(4, 10, 2, 1))
barplot(-1*log10(plotting_feature_pval+1e-10), 
        names.arg = names(plotting_feature_pval),
        las = 2, horiz = T, cex.names = .75, main = 'Mouse Vip -> Homologous Pax6',xlab = '-log10(meta p-value)' )
abline(v = -1*log10(0.05), col = 'red')
dev.off()


#Plot the distributions of values for the Vip vs pax6 for all of the features we pull out as significant

mouse_vip_vip_patch_df$label = 'Vip'
mouse_vip_pax6_patch_df$label = 'Pax6'

all_Vip_pax6_df = rbind(mouse_vip_vip_patch_df,mouse_vip_pax6_patch_df )

combined_pvals_vip_pax6 = combined_pvals_vip_pax6[order(combined_pvals_vip_pax6)]
sig_features = names(combined_pvals_vip_pax6[combined_pvals_vip_pax6 <= 0.05 & !is.na(combined_pvals_vip_pax6)])

#Filter for the significant features and plot
sig_subset = all_Vip_pax6_df[ , colnames(all_Vip_pax6_df) %in% c(sig_features, 'label')]

sig_subset = sig_subset %>% tidyr::pivot_longer(!label, names_to = 'feature', values_to = 'value')

sig_subset$feature = factor(sig_subset$feature, levels = sig_features)

sig_subset

ggplot(sig_subset, aes(x = feature, y = value, color = label)) + geom_boxplot() + facet_wrap(~feature, scales = 'free') +
  scale_color_manual(values = celltype_color_palette) + theme_bw()
ggsave(filename = 'all_sig_Vip_pax6_ephy_features_boxplots.pdf', path = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/mouse_patchSeq/', 
       device = 'pdf', family = 'Helvetica',
       useDingbats = F, height = 4, width = 6)




```


```{r}

p_mouse = ggplot(filter(mouse_patch_data, mouse_subclass %in% c('Vip') & primate_subclass != 'Sncg'), 
       aes(x = rheo_width, y = tau, color = mouse_subclass )) + 
  geom_point(alpha = .75, size = 2, show.legend = F) +
  scale_color_manual(values = celltype_color_palette) + theme_bw()

ggMarginal(p_mouse, type="density",  groupColour = TRUE, groupFill = TRUE)

ggsave(filename = 'taue_and_rheo_width_vip_scatter.pdf', path = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/mouse_patchSeq/', 
       device = 'pdf', family = 'Helvetica',
       useDingbats = F, height = 4, width = 6)


p_primate = ggplot(filter(mouse_patch_data, mouse_subclass %in% c('Vip') & primate_subclass != 'Sncg'), 
       aes(y = tau, x =rheo_width , color = primate_subclass )) + 
  geom_point(alpha = .75, size = 2, show.legend = F) +
  scale_color_manual(values = celltype_color_palette) + theme_bw()

ggMarginal(p_primate, type="density",  groupColour = TRUE, groupFill = TRUE)
ggsave(filename = 'taue_and_rheo_width_vip_pax6_scatter.pdf', path = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/mouse_patchSeq/', 
       device = 'pdf', family = 'Helvetica',
       useDingbats = F, height = 4, width = 6)
```


And now all the pax6 cells



```{r}

 table(mouse_patch_data$primate_subclass)


mouse_cge_patch_df = mouse_patch_data %>% filter( primate_subclass %in% c('Vip','Sncg','Lamp5', 'Lamp5 Lhx6')) %>%
  select(-c(rheo_adp_v, rheo_slow_trough_t, rheo_slow_trough_v, 
            qc_blowout_mv, qc_recording_date, qc_electrode_0_pa, qc_seal_gohm, qc_input_resistance_mohm, qc_initial_access_resistance_mohm, 
            qc_input_access_resistance_ratio,   
            file_id, file_name, cell_cluster, mouse_subclass, primate_subclass)) %>%
  tibble::column_to_rownames(var = 'cell_specimen_id')

mouse_mge_patch_df = mouse_patch_data %>% filter( primate_subclass %in% c('Pvalb','Chandelier','Sst', 'Sst Chodl')) %>%
  select(-c(rheo_adp_v, rheo_slow_trough_t, rheo_slow_trough_v, 
            qc_blowout_mv, qc_recording_date, qc_electrode_0_pa, qc_seal_gohm, qc_input_resistance_mohm, qc_initial_access_resistance_mohm, 
            qc_input_access_resistance_ratio,   
            file_id, file_name, cell_cluster, mouse_subclass, primate_subclass)) %>%
  tibble::column_to_rownames(var = 'cell_specimen_id')


mouse_pax6_patch_df = mouse_patch_data %>% filter(primate_subclass == 'Pax6') %>%
  select(-c(rheo_adp_v, rheo_slow_trough_t, rheo_slow_trough_v, 
            qc_blowout_mv, qc_recording_date, qc_electrode_0_pa, qc_seal_gohm, qc_input_resistance_mohm, qc_initial_access_resistance_mohm, 
            qc_input_access_resistance_ratio,   
            file_id, file_name, cell_cluster, mouse_subclass, primate_subclass)) %>%
  tibble::column_to_rownames(var = 'cell_specimen_id')




test_cge_pax6_pval_df = empirical_pvals_rank(mouse_pax6_patch_df, mouse_cge_patch_df)

test_mge_pax6_pval_df = empirical_pvals_rank(mouse_pax6_patch_df, mouse_mge_patch_df)



#And combine for a single p-value per feature

# Apply to each column (feature)
combined_pvals_pax6 <- apply(test_cge_pax6_pval_df, 2, fisher_combine)
combined_pvals_pax6 <- p.adjust(combined_pvals_pax6, method = 'BH')
combined_pvals_pax6

plotting_feature_pval = combined_pvals_pax6[!is.na(combined_pvals_pax6)]
plotting_feature_pval = plotting_feature_pval[order(plotting_feature_pval, decreasing = T)]

par(mar = c(4, 10, 2, 1))
barplot(-1*log10(plotting_feature_pval+1e-10), 
        names.arg = names(plotting_feature_pval),
        las = 2, horiz = T, cex.names = .75, main = 'All CGE vs Homologous Pax6')
abline(v = -1*log10(0.05), col = 'red')


pdf(file = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/mouse_patchSeq/CGE_vs_pax6_sig_features_barplot.pdf', 
    useDingbats = F, family = 'sans', height = 6, width = 7)
par(mar = c(4, 10, 2, 1))
barplot(-1*log10(plotting_feature_pval+1e-10), 
        names.arg = names(plotting_feature_pval),
        las = 2, horiz = T, cex.names = .75, main = 'All CGE vs Homologous Pax6',xlab = '-log10(meta p-value)' )
abline(v = -1*log10(0.05), col = 'red')
dev.off()


#Plot the distributions of values for the Vip vs pax6 for all of the features we pull out as significant

mouse_cge_patch_df$label = 'non-Pax6 CGE'
mouse_pax6_patch_df$label = 'Pax6'

all_cge_pax6_df = rbind(mouse_cge_patch_df,mouse_pax6_patch_df )

combined_pvals_pax6 = combined_pvals_pax6[order(combined_pvals_pax6)]
sig_features = names(combined_pvals_pax6[combined_pvals_pax6 <= 0.05 & !is.na(combined_pvals_pax6)])

#Filter for the significant features and plot
sig_subset = all_cge_pax6_df[ , colnames(all_cge_pax6_df) %in% c(sig_features, 'label')]

sig_subset = sig_subset %>% tidyr::pivot_longer(!label, names_to = 'feature', values_to = 'value')

sig_subset$feature = factor(sig_subset$feature, levels = sig_features)

sig_subset

ggplot(sig_subset, aes(x = feature, y = value, color = label)) + geom_boxplot() + facet_wrap(~feature, scales = 'free') +
  scale_color_manual(values = c('Pax6' = celltype_color_palette[['Pax6']], 'non-Pax6 CGE' = 'grey20')) + theme_bw()
ggsave(filename = 'all_sig_CGE_pax6_ephy_features_boxplots.pdf', path = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/mouse_patchSeq/', 
       device = 'pdf', family = 'Helvetica',
       useDingbats = F, height = 4, width = 6)


```

save source data

```{r}

save(plotting_feature_pval,sig_subset,  file = '../../source_data/data_for_plots/cge_pax6_distinguishing_features.Rdata')

```



```{r}

#and check out the MGEs too
#And combine for a single p-value per feature

# Apply to each column (feature)
combined_pvals_pax6 <- apply(test_mge_pax6_pval_df, 2, fisher_combine)
combined_pvals_pax6 <- p.adjust(combined_pvals_pax6, method = 'BH')
combined_pvals_pax6

plotting_feature_pval = combined_pvals_pax6[!is.na(combined_pvals_pax6)]
plotting_feature_pval = plotting_feature_pval[order(plotting_feature_pval, decreasing = T)]

par(mar = c(4, 10, 2, 1))
barplot(-1*log10(plotting_feature_pval+1e-10), 
        names.arg = names(plotting_feature_pval),
        las = 2, horiz = T, cex.names = .75, main = 'All mge vs Homologous Pax6')
abline(v = -1*log10(0.05), col = 'red')


pdf(file = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/mouse_patchSeq/mge_vs_pax6_sig_features_barplot.pdf', 
    useDingbats = F, family = 'sans', height = 6, width = 7)
par(mar = c(4, 10, 2, 1))
barplot(-1*log10(plotting_feature_pval+1e-10), 
        names.arg = names(plotting_feature_pval),
        las = 2, horiz = T, cex.names = .75, main = 'All mge vs Homologous Pax6',xlab = '-log10(meta p-value)' )
abline(v = -1*log10(0.05), col = 'red')
dev.off()


#Plot the distributions of values for the Vip vs pax6 for all of the features we pull out as significant

mouse_mge_patch_df$label = 'MGE'
mouse_pax6_patch_df$label = 'Pax6'

all_mge_pax6_df = rbind(mouse_mge_patch_df,mouse_pax6_patch_df )

combined_pvals_pax6 = combined_pvals_pax6[order(combined_pvals_pax6)]
sig_features = names(combined_pvals_pax6[combined_pvals_pax6 <= 0.05 & !is.na(combined_pvals_pax6)])

#Filter for the significant features and plot
sig_subset = all_mge_pax6_df[ , colnames(all_mge_pax6_df) %in% c(sig_features, 'label')]

sig_subset = sig_subset %>% tidyr::pivot_longer(!label, names_to = 'feature', values_to = 'value')

sig_subset$feature = factor(sig_subset$feature, levels = sig_features)

sig_subset

ggplot(sig_subset, aes(x = feature, y = value, color = label)) + geom_boxplot() + facet_wrap(~feature, scales = 'free') +
  scale_color_manual(values = c('Pax6' = celltype_color_palette[['Pax6']], 'MGE' = 'grey20')) + theme_bw()
ggsave(filename = 'all_sig_MGE_pax6_ephy_features_boxplots.pdf', path = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/mouse_patchSeq/', 
       device = 'pdf', family = 'Helvetica',
       useDingbats = F, height = 4, width = 6)






```


save source data

```{r}

save(sig_subset,  file = '../../source_data/data_for_plots/mge_pax6_distinguishing_features.Rdata')

```


Try out a UMAP with the human CGE ephys data Shreejoy pointed us towards

```{r}


mouse_patch_data

human_patch_data = data.table::fread('/home/werner/projects/mouse_brain_consensus_taxonomy/data/patch_seq/human/stripathy_human_int/patchseq_all_features.csv')
human_patch_data

table(human_patch_data$revised_subclass_label)
table(human_patch_data$subclass_scANVI)


cell_type_filt = human_patch_data$subclass_scANVI %in% c('Lamp5','Lamp5_Lhx6','Pax6','Sncg','Vip')

colnames(human_patch_data)[ c(35:50, 63:131)]

```

```{r}

human_to_mouse_label = c('v_baseline' = 'v_baseline', 'rheobase_i' = 'rheobase_i', 'fi_fit_slope' = 'fi_fit_slope', 'adapt_hero' = 'hero_adapt',
                         'sag' = 'sag', 'vm_for_sag' = 'vm_for_sag', 'input_resistance' = 'input_resistance', 'tau' = 'tau', 'avg_rate_hero' = 'hero_avg_rate',
                         'isi_cv_hero' = 'hero_isi_cv', 'latency_hero' = 'hero_latency', 'mean_isi_hero' = 'hero_mean_isi', 
                         'median_isi_hero' = 'hero_median_isi', 'threshold_v_rheo' = 'rheo_threshold_v',  'width_rheo' = 'rheo_width', 
                         'upstroke_downstroke_ratio_rheo' = 'rheo_upstroke_downstroke_ratio', 'subclass_scANVI' = 'primate_subclass')


filt_human_patch_data = human_patch_data %>% 
  filter(subclass_scANVI %in% c('Lamp5','Lamp5_Lhx6','Pax6','Sncg','Vip')) %>% 
  select(c(names(human_to_mouse_label)))



colnames(filt_human_patch_data) = human_to_mouse_label[colnames(filt_human_patch_data)]

#Now filter the mouse to match
index = which(colnames(mouse_patch_data) %in% colnames(filt_human_patch_data))
filt_mouse_patch_data = mouse_patch_data[ , ..index]

filt_mouse_patch_data  = filt_mouse_patch_data %>% filter(primate_subclass %in% c('Lamp5','Lamp5 Lhx6','Vip','Sncg','Pax6'))

#Align the columns
index = match(colnames(filt_mouse_patch_data), colnames(filt_human_patch_data))
filt_human_patch_data = filt_human_patch_data[ , ..index]

#Filter out human cells with no data
filt_human_patch_data = filt_human_patch_data[ rowSums(is.na(filt_human_patch_data)) != 16, ]



filt_human_patch_data$species = 'Human'
filt_mouse_patch_data$species = 'Mouse'

cross_species_ephys_df = rbind(filt_human_patch_data, filt_mouse_patch_data)
cross_species_ephys_df$row_index = 1:nrow(cross_species_ephys_df)

```

Final cell count 
Human Mouse 
  337   735 


```{r}

cross_species_ephys_mat = as.matrix(cross_species_ephys_df %>% select(-c(primate_subclass, species)) %>% tibble::column_to_rownames(var = 'row_index'))

#Filter out cells with NA values
row_filt = rowSums(is.na(cross_species_ephys_mat)) == 0 
cross_species_ephys_mat = cross_species_ephys_mat[row_filt, ]

length(as.numeric(rownames(cross_species_ephys_mat)))
scaled_cross_species_ephys_mat = scale(cross_species_ephys_mat)

data.pca <- princomp(scaled_cross_species_ephys_mat)
summary(data.pca)

pca_loadings_df = as.data.frame(data.pca$scores)

pca_loadings_df$primate_subclass = cross_species_ephys_df$primate_subclass[as.numeric(rownames(cross_species_ephys_mat))]
pca_loadings_df$species = cross_species_ephys_df$species[as.numeric(rownames(cross_species_ephys_mat))]


pca_loadings_df


ggplot(pca_loadings_df, aes(x = Comp.1, y = Comp.2, color = species)) + geom_point(alpha = .75)


ggplot(pca_loadings_df, aes(x = Comp.1, y = Comp.2, color = primate_subclass)) + geom_point(alpha = .75) +
  scale_color_manual(values = celltype_color_palette)


#mouse_patch_umap = umap::umap(scaled_filt_mouse_patch_data)

mouse_patch_umap = umap::umap(as.matrix(pca_loadings_df[ , 1:10]))


umap_df = as.data.frame(mouse_patch_umap$layout)

umap_df$primate_subclass = cross_species_ephys_df$primate_subclass[as.numeric(rownames(cross_species_ephys_mat))]
umap_df$species = cross_species_ephys_df$species[as.numeric(rownames(cross_species_ephys_mat))]

ggplot(umap_df, aes(x = V1, y = V2, color = species)) + geom_point(size = 2) +
  theme_bw() + ylab("UMAP 2") + xlab('UMAP 1') + ggtitle('Human/Mouse CGE ephys UMAP')

ggsave(filename = 'by_species_human_mouse_patchseq_ephys_CGE_umap.pdf',
       path = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/mouse_patchSeq/', device = 'pdf',
       useDingbats = F, height = 4, width = 5, family = 'Helvetica')



ggplot(umap_df, aes(x = V1, y = V2, color = primate_subclass)) + geom_point(size = 2) +
  scale_color_manual(values = celltype_color_palette) + theme_bw() + ylab("UMAP 2") + xlab('UMAP 1') + ggtitle('Human/Mouse CGE ephys UMAP')

ggsave(filename = 'by_subclass_human_mouse_patchseq_ephys_CGE_umap.pdf',
       path = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/mouse_patchSeq/', device = 'pdf',
       useDingbats = F, height = 4, width = 5, family = 'Helvetica')


```

```{r}

human_cge_patch_df = cross_species_ephys_df %>% filter( primate_subclass %in% c('Vip','Sncg','Lamp5', 'Lamp5_Lhx6') & species == 'Human') %>%
  select(-c( species, primate_subclass)) %>% as.data.frame()

human_pax6_patch_df = cross_species_ephys_df %>% filter( primate_subclass %in% c('Pax6') & species == 'Human') %>%
  select(-c( species, primate_subclass)) %>% as.data.frame()



test_cge_pax6_pval_df = empirical_pvals_rank(human_pax6_patch_df, human_cge_patch_df)


#And combine for a single p-value per feature

# Apply to each column (feature)
combined_pvals_pax6 <- apply(test_cge_pax6_pval_df, 2, fisher_combine)
combined_pvals_pax6 <- p.adjust(combined_pvals_pax6, method = 'BH')
combined_pvals_pax6

plotting_feature_pval = combined_pvals_pax6[!is.na(combined_pvals_pax6)]
plotting_feature_pval = plotting_feature_pval[order(plotting_feature_pval, decreasing = T)]

par(mar = c(4, 10, 2, 1))
barplot(-1*log10(plotting_feature_pval+1e-10), 
        names.arg = names(plotting_feature_pval),
        las = 2, horiz = T, cex.names = .75, main = 'All Human CGE vs Human Pax6')
abline(v = -1*log10(0.05), col = 'red')


pdf(file = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/mouse_patchSeq/human_CGE_vs_pax6_sig_features_barplot.pdf', 
    useDingbats = F, family = 'sans', height = 6, width = 7)
par(mar = c(4, 10, 2, 1))
barplot(-1*log10(plotting_feature_pval+1e-10), 
        names.arg = names(plotting_feature_pval),
        las = 2, horiz = T, cex.names = .75, main = 'All Human CGE vs Human Pax6',xlab = '-log10(meta p-value)' )
abline(v = -1*log10(0.05), col = 'red')
dev.off()


#Plot the distributions of values for the Vip vs pax6 for all of the features we pull out as significant

human_cge_patch_df$label = 'non-Pax6 CGE'
human_pax6_patch_df$label = 'Pax6'

all_cge_pax6_df = rbind(human_cge_patch_df,human_pax6_patch_df )

combined_pvals_pax6 = combined_pvals_pax6[order(combined_pvals_pax6)]
sig_features = names(combined_pvals_pax6[combined_pvals_pax6 <= 0.05 & !is.na(combined_pvals_pax6)])

#Filter for the significant features and plot
sig_subset = all_cge_pax6_df[ , colnames(all_cge_pax6_df) %in% c(sig_features, 'label')]

sig_subset = sig_subset %>% tidyr::pivot_longer(!label, names_to = 'feature', values_to = 'value')

sig_subset$feature = factor(sig_subset$feature, levels = sig_features)

sig_subset

ggplot(sig_subset, aes(x = feature, y = value, color = label)) + geom_boxplot() + facet_wrap(~feature, scales = 'free') +
  scale_color_manual(values = c('Pax6' = celltype_color_palette[['Pax6']], 'non-Pax6 CGE' = 'grey20')) + theme_bw()
ggsave(filename = 'human_all_sig_CGE_pax6_ephy_features_boxplots.pdf', path = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/mouse_patchSeq/', 
       device = 'pdf', family = 'Helvetica',
       useDingbats = F, height = 4, width = 6)





```






And do the same univariate decomp with the tasic expression data
Only use highly variable genes


```{r}

tasic18$cell_subclass[tasic18$cell_cluster == 'Lamp5 Lhx6'] = 'Lamp5 Lhx6'
tasic18$cell_subclass[tasic18$cell_subclass == 'Serpinf1'] = 'Vip'

cell_filt = tasic18$cell_subclass %in% c('Lamp5','Lamp5 Lhx6','Vip','Sncg')
tasic18_cge = tasic18[ , cell_filt]

table(tasic18_cge$cell_subclass)
table(tasic18_cge$primate_subclass)




#Get highly variable genes
logcounts(tasic18_cge) = log1p(cpm(tasic18_cge))
stats = scran::modelGeneVar(tasic18_cge)
dataset_hvg = scran::getTopHVGs(stats, n = 2000)

#Get the logcounts of the Hvgs
gene_filt = rownames(tasic18_cge) %in% dataset_hvg
feature_exp = logcounts(tasic18_cge)[gene_filt, ]
scale_feature_exp = t(scale(t(feature_exp)))

#And the subclass annotations
exp_mouse_subclass = tasic18_cge$cell_subclass
exp_primate_subclass = tasic18_cge$primate_subclass

table(exp_mouse_subclass)
table(exp_primate_subclass)
```


```{r}
res_exp_mouse <- run_omega_perm(scale_feature_exp, exp_mouse_subclass, n_perm = 1000, n_cores = 8)
res_exp_primate <- run_omega_perm(scale_feature_exp, exp_primate_subclass, n_perm = 1000, n_cores = 8)

# Add FDR correction
res_exp_mouse$padj <- p.adjust(res_exp_mouse$pval, method = "BH")
res_exp_primate$padj <- p.adjust(res_exp_primate$pval, method = "BH")

par(pty= 's')
plot(res_exp_mouse$omega2, res_exp_primate$omega2, main = 'Transcriptomic features', 
     ylab = 'Variance explained by homologous subclasses', 
     xlab = 'Variance explained by original subclasses', xlim = c(0,1), ylim = c(0,1) )
abline(a = 0, b = 1, col = 'red', lty = 'dashed')


pdf(file = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/mouse_patchSeq/omega2_transcriptomic_features.pdf', 
    useDingbats = F, height = 4, width = 4, family = 'sans')
par(pty= 's')
plot(res_exp_mouse$omega2, res_exp_primate$omega2, main = 'Transcriptomic features', 
     ylab = 'Variance explained by homologous subclasses', 
     xlab = 'Variance explained by original subclasses', xlim = c(0,1), ylim = c(0,1) )
abline(a = 0, b = 1, col = 'red', lty = 'dashed')
dev.off()


combined_res_exp = res_exp_mouse
combined_res_exp$primate_omega2 = res_exp_primate$omega2
combined_res_exp$primate_pval = res_exp_primate$pval
combined_res_exp$primate_padj = res_exp_primate$padj

combined_res_exp = combined_res_exp %>% mutate(omega_diff = primate_omega2 - omega2) %>% arrange(desc(omega_diff))


combined_res_exp %>% arrange(desc(primate_omega2))

```



save source data

```{r}

save(res_exp_mouse, res_exp_primate,  file = '../../source_data/data_for_plots/omega_expression_scatter.Rdata')

```





```{r}

plotting_omega_df = combined_res_exp %>% select(feature, omega2, primate_omega2) %>% tidyr::pivot_longer(!feature, names_to = 'annotation', values_to = 'omega-squared')

ggplot(plotting_omega_df, aes(x = annotation, y = `omega-squared`)) + geom_boxplot() +
  theme_bw() + ylim(0,1) + ggtitle('Transcriptomic feature variance')


hist(combined_res_exp$omega_diff, breaks = seq(-.3, .8, .01))


wilcox.test(combined_res_exp$omega2, combined_res_exp$primate_omega2, paired = T, alternative = 'two.sided')
#wilcox.test(combined_res_exp$omega2, combined_res_exp$primate_omega2, paired = T, alternative = 'greater')
wilcox.test(combined_res_exp$omega2, combined_res_exp$primate_omega2, paired = T, alternative = 'less')
```

VIP
LAMP5
MPPED1
NXPH1
PTPRM
CPLX3
SYNPR


```{r}



current_metadata = as.data.frame(colData(tasic18_cge))


current_metadata$gene_1_exp = feature_exp['VIP' ,]
current_metadata$gene_2_exp = feature_exp['LAMP5' ,]

ggplot(current_metadata, aes(x = gene_1_exp, y = gene_2_exp, color = cell_subclass)) +
  geom_point() + scale_color_manual(values = celltype_color_palette) + theme_bw()

ggplot(current_metadata, aes(x = gene_1_exp, y = gene_2_exp, color = primate_subclass)) +
  geom_point() + scale_color_manual(values = celltype_color_palette) + theme_bw()



```


And same correlation heatmap, using the top 24 hvg genes

```{r}

top_hvg = scran::getTopHVGs(stats, n = 25)
top_hvg

exp_cor_mat = cor(scale_feature_exp[top_hvg, ], method = 'spearman')
dist_exp_cor_mat = as.dist(1- exp_cor_mat)


current_metadata = as.data.frame(colData(tasic18_cge))

index = match(colnames(exp_cor_mat), rownames(current_metadata))
exp_mouse_subclass = current_metadata$cell_subclass[index]
exp_primate_subclass = current_metadata$primate_subclass[index]


top_annot_exp = HeatmapAnnotation(mouse_subclass = exp_mouse_subclass, homologous_subclass = exp_primate_subclass,
                                   col = list(mouse_subclass = celltype_color_palette, homologous_subclass = celltype_color_palette))

corr_cols <- rev(grDevices::colorRampPalette(RColorBrewer::brewer.pal(11,"RdYlBu"))(100))
Heatmap(exp_cor_mat , name = 'spearman',use_raster = F, col = corr_cols,
        clustering_distance_rows = dist_exp_cor_mat,
        clustering_distance_columns = dist_exp_cor_mat,
        clustering_method_rows = 'ward.D2', clustering_method_columns = 'ward.D2',
        show_row_names = F, show_column_names = F,
        top_annotation = top_annot_exp, column_title = 'Cell similarity over transcriptomic features (top 25 hvgs)')


pdf(file = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/mouse_patchSeq/cell_siml_of_trans_features.pdf', 
    useDingbats = F, height = 4, width = 6, family = 'sans')
Heatmap(exp_cor_mat , name = 'spearman', use_raster = F, col = corr_cols,
        clustering_distance_rows = dist_exp_cor_mat,
        clustering_distance_columns = dist_exp_cor_mat,
        clustering_method_rows = 'ward.D2', clustering_method_columns = 'ward.D2',
        show_row_names = F, show_column_names = F,
        top_annotation = top_annot_exp, column_title = 'Cell similarity over transcriptomic features (top 25 hvgs)')
dev.off()


pdf(file = '/home/werner/projects/mouse_brain_consensus_taxonomy/graphs/mouse_patchSeq/raster_cell_siml_of_trans_features.pdf', useDingbats = F, 
    height = 4, width = 6, family = 'sans')
Heatmap(exp_cor_mat , name = 'spearman', use_raster = T, col = corr_cols,
        clustering_distance_rows = dist_exp_cor_mat,
        clustering_distance_columns = dist_exp_cor_mat,
        clustering_method_rows = 'ward.D2', clustering_method_columns = 'ward.D2',
        show_row_names = F, show_column_names = F,
        top_annotation = top_annot_exp, column_title = 'Cell similarity over transcriptomic features (top 25 hvgs)')
dev.off()

```


save source data

```{r}

save(exp_cor_mat, current_metadata,  file = '../../source_data/data_for_plots/exp_corr_heatmap.Rdata')

```




